% Local IspellPersDict: aspell.dunari.pws

\chapter{2}

\noindent Chark wrapped his trunk around the the heavy door handle and then used the three
tentacles at the end of his trunk to tap the access code on the lock panel. The lock clicked
and, leaning onto his back legs, he slid the door aside.

Inside, the room was packed with state of the art electronics, and smelled vaguely of solder and
over heated transformers. The noise of the air cooling unit was a constant drone. The only
windows were uncomfortably small, placed high on the wall, and partially covered by equipment.
Several artificial lights supplemented the meager natural illumination.

Tarmala, the Aerosights engineer who had called Chark, was studying an oscilloscope, one of his
rear legs hooked slightly behind the other. His tentacles operated the slider controls on the
side of the instrument. Fanlon, a young military engineer, was looking over Tarmala's shoulder,
but he twisted his head around and slid his trunk over Tarmala's back when he heard the door
open.

``The Captain is here.'' He motioned excitedly to Chark with his trunk, extending one short
tentacle in the direction of the main display grid of the oscilloscope. ``Take a look!''

Tarmala stepped aside, turning his head slightly so his right eye was toward Chark in invitation
and deference. Chark stepped forward, pushing his front shoulders between them. Sure enough
there was a return signal. It was distinct. It was steady.

``Are you tracking this?'' Chark asked.

``Absolutely, we have the Ear slewing right now,'' Fanlon said.

Chark took in the instruments, quickly scanning over the myriad displays by tilting his head
first to one side and then the next with a well practiced motion. Putting his trunk over his
shoulder he turned his more acute left eye toward the screen. ``That almost looks like
ping-back,'' he said after a moment, ``but you're slewing the antenna, yes? Do you lose the
signal if you don't?''

``We do.''

``So maybe it's a real target.''

``Impossible,'' said Tarmala. ``Look at the return delay.'' He twisted around and raised his
trunk toward a unit perched precariously on a high self above the receivers. The side panel was
open and hanging by one clip. ``If it's a real target it's really far away---and big.''

Chark stepped back to study the display for a moment, flapping his ears slowly in contemplation.
He turned to the plotter, tilting his head to look at the historical data scratched onto the
paper in broken lines. ``This doesn't make sense,'' he said finally. ``According to this you
earlier had the antenna pointed even higher than it is now.''

``Right,'' said Tarmala, ``and the return delay was unusually long even then. See\ldots''
extending his trunk he traced the green line with his second tentacle.

``It can't be real,'' said Chark. ``It would be hundreds of karnons away and at that elevation
it would be in space.''

Later, as was his habit when we went home, Chark walked to the lifter rather than take the
base's shuttle. He needed what exercise he could get, and he enjoyed the outdoor air. On clear
days, when the Nermella wasn't blowing much sand, he could see the Sarnowledge Mountains rising
in the distance as a darkening against the yellow-green sky. Sometimes he liked to look at the
mountains and reflect on the relaxed days when Junella's old seleck took camping trips there. Of
course that was all before the project.

The purpose of the RRRF system was to track aerocars in flight. It was a simple idea, really.
The antenna emitted a powerful pulse of radio energy and some of that energy reflected off the
target to return back to the source. By timing the return delay the distance to the target could
be determined. The orientation of the antenna provided the direction.

It sounded simple but making it work wasn't. They had to generate powerful bursts of radio
energy. They had to measure short time intervals. They had to accurately control a large
antenna. There were many places where things could go wrong and they often did.

When the lifter came, Chark climbed aboard, his military cloak making it easy for him to find a
good seat. He slid his rear legs onto the back cushion and folded his front legs beneath him. He
tugged briefly at his cloak with his three tentacles to straighten it. As the lifter started
again he tilted his head to look out the window.

The muggy, tropical air of Lungast fostered lush fields of harsna grass lined with rigat trees.
It was so different than the desert land on the Northern Plains where he and Junella had lived
before. The large, rubinum-red Rujaran sun hung low over the western horizon. Its face was
covered with several spot complexes. Its subdued, golden-red light filled the muggy sky and lit
the yellow harsna with ghostly highlights.

Chark eyed the sun suspiciously. He was still not used to seeing it so low or so gloomy. The
project was a great opportunity for him, but Junella had left her seleck to make the move. He
found himself wondering how happy they could really be in such a dank and dreary place.

The lifter finally reached the large garjost tree that marked Chark's stop. It was the sight of
its short, thick trunk and graceful horizontal branches more than the cry of the driver that
alerted him to take notice and gather his belongings. From the stop he walked down the lane and
into a circle of seven modest, interconnected homes. Their low greenstone walls blended into the
rich tropical soil. A young woman new to the seleck was tending her garden. She greeted him
warmly with a soft bleat. Chark went to her and they crossed trunks briefly, entwining them,
tentacles brushing against leathery skin.

Once inside his house Chark automatically adjusted the picture of Chartik sitting unobtrusively
on a small shelf. It never really needed adjusting, of course, but he liked to touch it. He
called out for Junella with a low, growling moan but she did not come immediately. He was about
to get his pad to see if she had left a message for him when he heard her enter through the
connectway to the neighbor's house.

``Oh, Chark,'' she said, ``I had a feeling you were back.''

She was wearing a light, rubinum cloak, loosely tied about each leg. Her ears were clipped to
the fabric using the copper snaps Chark had given her when they were courting. They crossed
trunks, lingering slightly in their embrace, and then Chark gently tapped twice behind her head
where her trunk ridge merged into her spine.

``It didn't go well,'' she said more as a statement than a question. Somehow she always seemed
to know.

``Things went fine,'' Chark replied. He paused and affectionately touched her ear with his first
tentacle. The secret nature of the RRRF project made it impossible for him to share the details
of his work. Chark took his oath seriously. She was understanding, of course.

``A quirk came up before I left,'' Chark said, ``that's all.''

They shared a light meal and then relaxed for a while playing a game of conifi with two of the
seleckia and their husbands. Jiapita was the eldest seleckia and Jijar was the young gardener
who had greeted Chark earlier. Junella and Jiapita's husband were experts and dominated the
game, as usual, but Chark didn't mind. It was nice to have something to take his mind off his
research. The work was critical to the safety of his country and the pressure was sometimes hard
to bear.

``Chark forgot to tap me earlier,'' Jijar said playfully at one point. Chark suppressed a
flutter, uncertain of how to react to her forward comment.

``He's been distracted in many ways,'' Junella said clicking her teeth. She twisted her head
slightly to glance at her husband.

After the others had left, Chark got ready to sleep. He first took a few moments to look out
over the back gardens. The vines his wife had planted around the base of the house were finally
starting to bloom. In due course there would be flashes to provoke and ripen the delicious
seeds. Chark winced slightly when he realized he had yet to fix the flash shades in their main
living space. The color of their furniture was already starting to fade. Yet the distraction of
his work made it hard to find time for everyday things.

``I'll be at the hospital for a while,'' Junella said from behind him, ``but I should be back
when you awake.''

Chark made a few notes in his diary and then went up the ramp to the beds overlooking the main
floor. The light of the sun filled the room with an orange, fiery glow. Chark changed into his
sleeping cloak and climbed onto his bed. As he folded his legs beneath him he tugged at his
cloak to loosen it a bit. Soon he entered farth-sleep and meditated about the strange signal
found earlier. Yet his mind drifted and he found himself remembering their old home on the
Northern Plains and the happy times there with Chartik. In due course paren-sleep took him and
his mind wandered no more.

\spacebreak

It seemed like only moments when Chark was awakened from paren-sleep by the sound of the phone
chiming. With a forced shake he roused his body and stumbled out of bed. Junella was still gone
and none of the other seleckia were in the house. In a daze Chark made his way down the ramp to
answer the insistent phone. The right half of his body was still partly asleep and it made him
lurch awkwardly as he moved.

``Yeah?'' he said ungraciously into the transceiver.

``I'm so sorry to call you at home, Captain.'' It was Tarmala. ``It's just that we've been
tracking that target we found earlier.''

``Tracking?'' Chark absent mindedly scratched his front left shoulder against the edge of the
door frame.

``Yeah,'' Tarmala said. ``Fanlon did some calculations and it's in orbit. We predicted that we'd
be able to track it again periodically, and sure enough we can. We have more data on it now. We
have some nice plots with very good return delay measurements. We want\ldots''

``Wait, wait, wait\ldots'' Chark said. ``Wait.'' Tarmala stopped abruptly. Chark shifted a
little and shook his right front leg. ``What do you mean by orbit?''

``It's in orbit around Rujar. It's about 500 karnons up.''

\spacebreak

When Chark next awoke he could hear Junella singing below, her voice a melodic rumble. He got
out of bed and, looking out the window, tilted his head at the sunlight glinting off his
neighbor's roof. The sky was clear but still hazy as it always seemed to be in Lungast. A large
group of spots created a blotch on the sun's disk.

He took a cloak from his shelf and put it on, fighting briefly with the hooks. He stood by the
mirror wall and brushed his trunk and face, removing a few loose scales in the process. Then he
went down the ramp to join his wife.

``I thought you were going to sleep forever,'' Junella said shaking her ears playfully. They
crossed trunks and she clicked her teeth lightly. ``Let me make you some brushtik.''

Chark went into the center room and opened the windows letting in the tropical air. He turned on
the viewer to watch part of \textit{Moment by Moment}. Junella grunted with annoyance.

``Must we have that on?'' she asked. She adjusted the brushtik under the toaster, adding a touch
of genic butter to each piece. ``All they ever talk about is politics. Nothing important ever
happens in politics.''

Chark didn't bother to lay down but instead just stood and watched the program. It was a panel
discussion involving a well dressed Argenian man and someone who, due to his distinctive
markings, was obviously a Forbinite.

% TODO: Provide more description of the living space?

``You are always making a big deal over nothing just as an excuse to rattle your swords,'' the
Argenian man said. ``The Yellow Road has until recently been the only official mode of
transportation between our countries and, as such, essentially neutral ground.''

% TODO: "Rattle your swords" is an Earth idiom. Perhaps the Argenians have a different phrase.

The other panelist fluttered in obvious annoyance and replied in thickly accented Argenian. ``If
the new aeroport really is a Argenian military base, I'd hardly say that's nothing. It would be
the most provocative action taken by a First Citizen since before the Harkite War!'' He shook
with barely contained anger, growling slightly, and his \textit{grung} almost slid off his head.

``Oh come on,'' said Chark to the viewer. ``Aerosights just wants to open commercial flights to
Forbin. Jurita's Witness!''

The first panelist seemed almost pleased to have created such a reaction in his counterpart.
``You know full well that the Garnick Agreements preclude any\ldots''

``Come get your brushtik,'' Junella interrupted, ``and turn that silly thing off.''

Chark joined his wife at table, and the two of them lay down side by side sharing a common
plate. Junella sang a traditional prayer to Jixita reminding them of the virtue of patience.
Then they ate in silence for a while, enjoying the course grain brushtik and spicy genic. Chark
wrapped his tentacles around a mug of ranan juice bringing it to his lips. ``What are your next
plans?'' he asked his wife.

``The seleck is meeting Jala. I told you about it.''

``Oh yes,'' Chark said, ``another seleckia so soon.''

Junella fluttered softly. ``I know what you mean. It is quick, but we want to be complete as
soon as we can.''

``Of course.'' Chark bobbed his head slightly. ``It's just, well, we're still getting used to
Jijar and believe me when I say she takes getting used to.''

Junella shook her ears. ``I like Jijar. Anyway, you know she's a specialist in joint disorders
so when you get old and creaky you'll be glad she's around. Besides they accepted us graciously
and I am grateful for that. The least we can do is be gracious in return.''

When Chark later arrived at the base everything seemed the same as any other time. Chark
immediately went to the research center. Fanlon had gone home to sleep. Tarmala was still there.

``We should meet,'' Chark said to him.

A short time later Tarmala and Brogan were in Chark's office. Chark liked Brogan. He was also a
former Electrodyne employee and, although Chark never worked with him there, they could share
stories about life in the industrial electronics world. Despite technically being a captain,
Chark felt a greater affinity with Brogan's civilian background than with the military mindset
surrounding the project.

The three engineers gathered around Chark's table as he clipped a pad of brikken paper in front
of him, setting a short pen beside it. He raised his trunk to start the meeting. The other two
men reached out and slid their tentacles up and down his trunk to signal their readiness.
``Summarize your discovery for us,'' he said to Tarmala.

Tarmala was glad to discuss what they had found. He presented his graphs of return delays,
altitude and azimuthal slew rates, and reflection strengths. He also talked about Fanlon's
orbital calculations and showed a diagram of the orbit. Chark was impressed. Brogan studied each
paper carefully and checked a few of the calculations on his spin-dial.

``It's in an almost perfectly circular polar orbit,'' Tarmala said.

``When it's overhead, it looks like we'll be able to track it on several orbits in succession,''
Brogan added while looking at the materials.

Chark made a few notes on his pad, the clip holding the paper firmly in place as he wrote. ``We
should be ready to make more detailed observations on the next pass.''

``What about the frequency test?'' Tarmala asked. Brogan put down the graph he had been studying
and both cocked their heads and looked at Chark.

Chark scratched the side of his face just below his left eye slowly. ``I think we need to get to
the bottom of this first. The research can wait a little. If this\ldots\ object is something
we're going to pick up regularly we should understand it.''

``What do you think this thing is?'' Brogan asked.

``It's got to be a small moon,'' Tarmala said. ``They recently discovered Nermia has a moon, so
why not Rujar?''

Brogan swayed his head slightly and tilted it to look at Chark with his right eye. ``It makes
sense. A small moon could have gone unnoticed. Astronomy is a young science, after all.''

Chark tapped the short pen against the edge of the table a few times and then fluttered
slightly. ``Maybe it's something the Forbinites launched,'' he said suddenly. His comment
unleashed a flurry of objections, and he had to thump his trunk against the table to quiet them.

Tarmala was shaking his head, ears folded back. ``This thing has to be big to generate a return
like we're seeing. Launching something like that from Rujar would require huge energy. It's
beyond our technology\ldots\ and Forbin's too.''

``Are you sure?'' Chark asked.

``They'd be lording it over us if they had.''

``Maybe they will.''

Brogan fluttered briefly. ``There is no point speculating. Let's get some more data. I say we
try to set up a frequency sweep reflection measurement.''

Chark swayed his head. ``Yes, I agree. In the meantime I'll inform the general. He needs to at
least know of what we've discovered.''

\spacebreak

Chark's team worked diligently setting up the frequency sweep measurement. It was a technique
that had never been tried in an RRRF environment, and it required building some special
circuitry. The team forwent paren-sleep and took only brief farth-sleep naps at the base.
Junella was not happy.

``It's not good for your health,'' she told Chark during a phone call.

``I'll be fine, \textit{tinka}, it's just until we get this one experiment done. It's time
sensitive so we have to overload.''

Although he didn't admit it to his wife, Chark enjoyed the energy and intensity of the work. It
reminded him of his younger days at Electrodyne frantically trying to finish projects before
their competitors could. Everyone on the team was working seamlessly together, each contributing
their part toward a larger whole.

As the time of predicted reacquisition approached, the project got an unexpected visit from
General Fotkey. He came in an ordinary car without any fan fare, and with only his personal
secretary in his entourage. Chark met the two men in his office. As usual they were standing.
Chark wondered if Fotkey ever lay down.

The general was relatively short, with his back just below the level of Chark's trunk ridge, and
remarkably thin. He was wearing a casual cloak without any insignia or indication of rank,
appearing instead more like a common laborer. His secretary hung back slightly and wrote notes
on a brikken board hanging around his neck, his trunk curled back to write on the board as it
pressed against his chest.

``I'm so sorry, General,'' Chark said touching his trunk briefly to the general's left shoulder.
``I was in the lab getting ready. I had no idea\ldots''

``It's fine, Captain'' the general said. ``I wouldn't expect you to be just sitting around the
office.'' Despite his unassuming size and quiet tones, the general's voice was filled with
confidence and strength. He was a man used to giving orders that were obeyed without question.

``How was your trip?'' Chark asked politely.

``That's not important,'' Fotkey replied. ``I wanted to be here when the sweep was done.''

The general paid a visit to the lab where he got a chance to see the work being done in
preparation. He kept back, respectful of the engineer's labors, but he also asked many pointed
and probing questions. The atmosphere was tense. It would not have been a good time for the
phase locked loop to glitch or the amplifier power supply to sag.

When all was ready Brogan double checked the calculations and Tarmala moved the Ear into
position. All they had to do was wait.

``How long will it take?'' Fotkey asked.

Chark looked at Brogan but Brogan just shrugged.

``Unknown.''

The arnets ticked by. The air cooling units droned on as usual. Tarmala inspected some meters
and then adjusted the Ear slightly. They were blasting into space short bursts of power as great
as any television station and then listening for the echo. Everyone waited.

``You do have the delay scan up to maximum, right?'' Fanlon said.

``Yes, yes, of course'' said Tarmala. He rubbed his tentacles together impatiently.

Suddenly the line on the 'scope jumped. ``That's it!'' said Tarmala.

Fanlon looked at the Ear's positional controls. ``It's were it should be. The orbit hasn't
changed. Our tracking plan will be fine.''

Chark made some adjustments on the frequency scanner to account for the observed signal
strength. ``We should get four or five arnets of observing time on this pass,'' he said. ``Let's
get started.''

Fotkey stayed out of the way during the actual measurements, but was anxious for an
interpretation after even the first pass. The team gathered in a conference room down the hall
from the lab to review the data.

% TODO: Description of the conference room?

``We don't have a lot of information yet,'' Chark said. ``It might take a couple of passes
before we really know what we're dealing with.''

Tarmala was calculating spectral transforms, switching back and forth between his pad and his
spin-dial. Brogan and Fanlon were double checking his work. Fotkey sipped on a mug of water.
``Do what you have to do,'' he said. ``I want results, but I also want them to be right.''

Tarmala seemed a little uncertain. ``Huh. Are you getting this value?'' he asked Brogan,
pointing at the paper before him.

Brogan extended his trunk and slid the page closer. He studied it with first his left eye and
then his right.

``What is it?'' Chark asked.

Brogan passed the paper to Chark. On it were some shocking numbers. Chark felt his heart sink.
He tilted his head and looked over at Fotkey. The team looked at him.

``Well\ldots'' Chark began. ``This is preliminary, but it looks like this object isn't nearly as
large as we assumed.''

``What does that mean to me?'' Fotkey asked.

``With a return signal as strong as this, it means it's highly reflective\ldots\ like polished
metal maybe.''

The room fell silent.

``So you're saying it's artificial,'' Fotkey said.

``Um\ldots'' Chark stammered. ``It looks that way, yes. We will confirm this result on the next
pass.''

Fotkey leaned back. He thumped his trunk slightly against the table, thinking. ``I'll wait until
then.''

The data acquired on the next pass confirmed the preliminary result. The object was almost
certainly artificial. Chark wrote up a hasty report, and Fotkey left with it immediately. He was
taking the information directly to the First Citizen.

``This is classified at level one until you hear otherwise,'' he told the team. ``I'll be in
touch.''

After he left everyone was in shock. ``Dind,'' Tarmala cursed. ``Does this mean war?''
