% Local IspellPersDict: aspell.dunari.pws

\chapter{3}

Chark's team did their best to ignore the implications of their foreboding discovery. If
Forbinite technology had progressed to the point where Forbin could launch objects into orbit,
no place in Argenia would be safe. The enemy could choose to rain down from the skies whatever
destructive force they wished upon any city in the land. It would be far worse than the effect
of aerocars during the Harkite War.

Yet time passed and nothing more developed with the mysterious object. Chark's team was able to
observe it routinely and refined their knowledge of its orbit and size. Despite that they
received no word from Fotkey or the military office. There was little to do other than resume
their RRRF research.

``We have time lines to consider,'' Chark told them.

Chark did his best to keep his concerns out of his home and away from Junella and the other
seleckia, but she could tell he was under stress. She didn't press him for the details but
encouraged him to relax. They took long walks around Lungast finally getting to know a little
better the city they just recently started calling home. The humid air prompted frequent clouds
and rain storms that made the world seem dark and gloomy. Yet Chark found the exercise calming
and cathartic. The clouds also made the flashes less dangerous as compared to walking under the
cloudless sky to the north.

Eventually the seleckia decided to welcome Jala and held an induction ritual to bring her into
the group. As the most recently added member, Junella was obligated to organize the event. She
and Chark prepared traditional foods while Jiapita's and Jinjania's children arranged tables,
colored runners, and grass baskets. All seven homes were opened as the families roamed among
interconnected houses and hid secret gifts in the empty house Jala and her husband would soon
occupy.

The six seleckia conducted the ritual on the stubby lawn in the common space between the houses
under the open sky, which thankfully was free of rain. Jala was there too, of course, a stately
middle-aged woman with an interest in painting. The rest of the group watched from a discrete
distance, lying on temporary seats or on the grass, and listening to the harmonious drone of the
selekia singing to their newest member. At the conclusion of the ritual Jiapita raised her trunk
and proclaimed for all to hear, ``We are seven!'' There were loud bleats and yowls, and Jijar's
husband set off some crackers.

One by one all the men and children formally greeted Jala and her husband to the seleck,
brushing trunks over their faces and backs. Their only child, an adult son, was not with them
but instead was on duty at the Rangard base in the southern desert and couldn't get away. Once
the formalities were finished the group socialized for the first time as a unified whole. The
seleck had gone without complete membership for much too long.

% TODO: Say more about the social aspects of being in a seleck?

Junella and Chark had earlier prepared trays of charara leaf and toran root for the adults, and
gigat nuts for the children. Chark was in the hanaria of Jijar's home refilling a tray and
talking with two other husbands. All seemed approving of Jala, and Chark admitted that she was a
nice addition to the group.

``Don't let Jijar hear you say that,'' one man said to Chark with a shake of his ears. Chark
fluttered slightly.

``It could be worse,'' said the other man, ``at least Jijar is plenty attractive.''

``By the by,'' the first man said, ``did you hear about that new moon recently discovered?''

Chark tilted his head toward his friend.

``Yes,'' replied the other, ``I saw something about it on the viewer. It was Jar-what's-her-name
with the lovely ears.''

``Jarloz,'' the first man said.

``What's this?'' asked Chark.

``They discovered a moon going around Rujar\ldots''

``Apparently you can even see it,'' interrupted the second man, ``I saw an interview with
someone in Harlotts, on the night side of the Narlar, and he said it was like a bright star
moving steadily across the sky.''

``Probably some high altitude aerocar. Maybe it's a military experiment. You wouldn't happen to
know anything about that, would you Chark?'' he added with a shake of his ears.

% TODO: Add some descriptions of body language?

Chark fluttered nervously.

``They had an astronomer on the show. He seemed to think it was in orbit.'' The second man
said.

When Chark returned to the base some time later he intended to call a meeting of his team to
discuss public awareness of the so-called moon. Yet before doing that he needed some guidance.
He pushed aside the outer door of his office and Jeno, already there, immediately stood up.

``Call Fotkey's office,'' Chark said abruptly.

``That won't be necessary,'' Jeno replied. ``I just called you but your wife said you were on
the way. The general is waiting for you with someone named Lucastanonia.''

Chark paused. ``Waiting now?'' Jeno swayed her head toward the inner door. Chark stepped forward
both annoyed that there were people in his office without permission and also anxious about what
it might mean.

The two men stood when Chark entered. The general was dressed in a semi-formal suit looking more
like a business man than a general. He had no entourage. Chark had never met the other man
before. He was large with large ears, a long trunk, and physically precisely the opposite of the
general. Yet he had a mild mannered, easy going attitude that seemed to dispel tension whenever
he spoke. He introduced himself as ``Lucasa'' and as ``the director of the Office of Alien
Affairs.'' Chark had never heard of it before.

Chark politely greeted his guests and, after opening his windows, invited them to lay at his
meeting table. He pulled a box of tuni from a small drawer and offered it to them. Lucasa took a
sniff but Fotkey declined.

Much to Chark's surprise Lucasa took the lead. ``You might think that Parliament only wastes
money on bribes, kick backs, and free vacations for government officials,'' he said, glancing at
Fotkey briefly. ``It turns out they also waste money creating silly little departments to deal
with nonsense. The Office of Alien Affairs is such a department.''

Chark shook his ears slightly. ``I see,'' he said, ``and you work for this office?''

``I run it.''

Lucasa took another sniff of the tuni. ``Basically at the OAA we spend our time thinking about
invasions from space. It sounds crazy but, hey, now I'm a celebrity.''

Chark tilted his head to look at both men. Despite Lucasa's irreverent banter, Fotkey seemed
deadly serious. ``There is no invasion from space,'' Chark said finally.

``Oh?'' Lucasa pulled an envelope out of the pouch of his cloak and laid it on the table in
front of him.

``What do you know about bird flight, Captain Charkonaless?'' he asked.

``It's not my area.''

Lucasa opened the envelope, deftly holding it in place with one tentacle while sliding some
papers out with the others. ``Well, it turns out there is a certain gentleman in Harlotts who is
a retired pediatrician. Anyway he's very interested in birds and he likes photographing them in
flight.''

Chark tilted his head questioningly. He looked over at Fotkey again but the general's expression
hadn't changed.

``He's quite an ingenious fellow and he's built a camera that can follow a flying bird and take
pictures of it at high magnification. So he decided to try his camera on our new orbiting
visitor to see what it would reveal.'' Lucasa slid a few photographs over to Chark. ``I think
you will find them interesting.''

Chark drew the packet toward him and started to page through it.

``Jurita's Witness,'' he said more to himself than to the other men. The first picture was
little more than a fuzzy, elongated blob. The second picture was only slightly better. Yet the
third picture was amazingly clear. It showed a long, cylindrical object with several thin
extensions protruding from the ends. On one end there were clearly four extensions pointing
outward toward the corners of a square.

Chark looked up at Lucasa in disbelief. Fotkey said nothing.

``Keep looking,'' Lucasa said. ``The later pictures were taken on a different orbit and the ship
was closer to his position.''

Chark flipped to the next set of pictures. They were larger and clearer than the first. In one
all eight extensions, four on each end, were easily visible. The main body of the object also
showed some detail. It appeared to be built around a central axis. Along the side were markings
of some kind but it wasn't possible to discern them. They might have been a structure on the
side of the object, or maybe even some writing.

``Take a look at these two pictures.'' Lucasa pulled two from the pile and put them side by
side. ``They were taken a couple of arnets apart. You can see this dark spot here\ldots\ and
here\ldots\ it has moved between the shots.''

``It's rotating,'' Chark said immediately.

``My people think that's to provide a kind of artificial gravity for its occupants. Based on the
rotation rate and the size of the ship, the acceleration at the outer rim is very close to the
gravitational force here on Rujar. It seems that our visitors are from a place not unlike here,
at least in that respect.''

% TODO: Check if 1G acceleration is feasible.

Chark was stunned. ``Visitors? Do you mean Forbinites?''

Lucasa shook his ears vigorously and this time Fotkey finally spoke up. ``This thing is huge.
Our intelligence is conclusive: there is no way Forbin could lift something like that to
orbit\ldots\ and certainly not in secret.''

Lucasa reached for the tuni. ``So that's where the Office of Alien Affairs comes in.''

Chark reviewed the pictures again. ``This is incredible,'' he said at last. He felt a deep sense
of relief. Maybe there would be no war after all.

``This could be far worse than the Forbinites,'' said Fotkey. ``If these creatures are
interested in invading there may be nothing we can do to stop them. Their technology is
obviously far superior to ours.''

``Yet,'' Lucasa said, ``if they're friendly we might be on the verge of a leap in technology
such as the dunari have never seen.''

Chark pushed the pictures away and leaned back. He scratched his trunk thoughtfully with his
tentacles.

``So we want to talk to them,'' Lucasa continued, ``we want you to configure the Ear for two-way
communication. Do you think it can be done?''

``Sure, I don't see why not.''

Fotkey leaned forward and slid his trunk onto the table for emphasis. ``Understand this is a
directive from the First Citizen. Your current research is to take second priority for now
and\ldots\ you will be reporting to Lucastanonia.''

``Oh?''

Lucasa clicked his teeth. ``You see? I'm a celebrity.''

``One more thing,'' Fotkey said with a soft flutter. ``This is no longer classified. That's also
a directive from the First Citizen.''

Lucasa shook his ears and interrupted. ``It helps that our pediatrician friend has already sold
his pictures to \textit{Moment by Moment}.''

Fotkey continued, ``An official announcement will be forthcoming. The First Citizen plans to
contact the Forbin Prefect and discuss a way of presenting some kind of unified Rujar to
these\ldots\ visitors.''

As the men got up to leave Lucasa clasped Chark's trunk again. ``It will be a pleasure working
with you, Captain Charkonaless,'' he said.

``Just call me Chark.''
