% Local IspellPersDict: aspell.dunari.pws

\chapter{5}

Jernumia trotted down Following Street Darnar with the sun at her back, head up, ears out
slightly, but not excessively, and her trunk slightly curled. She was wearing a stylish business
cloak in orange and red with black leggings on her rear legs. Over her left shoulder and across
her back she wore an expensive belt that was symbolic of her professional status and yet
appropriately feminine.

With over two million people, Varnok, the capital of Argenia, was the largest city on Rujar.
Jernumia worked in the New City where the buildings were grand, modern, and beautiful. The cross
streets were wide so buildings of several stories could be constructed without shading each
other. Their sunward walls gleamed with glass windows while their arched shade walls were
covered with artistic patterns or climbing vines. Rigat trees lined many of the streets. Their
tubular leaves whistled in the ever-present Nermella and provided a natural counter-point to the
sounds of the city. In the New City many of the streets had restricted vehicle access and people
moved freely down the middle of them as if they were great promenades.

To a casual observer nothing was out of the ordinary. Despite the shocking news that the dunari
were not alone in the universe, the people of Varnok mostly went about their business as usual,
just as Jernumia was doing. She knew there had been a recent gathering on the Parliamentary
Yellow of fearful people shouting demands at their government, just as there had been similar
gatherings all across the country. Yet in the New City nobody seemed to care.

Jernumia came to the corner of Cross Street Hillart. She paused briefly to look up at the Tower
of Regents, several streets sunward, it's sharp spike penetrating the red disk of the sun, now
clear of spots. The tower always seemed symbolic to her of the dunari's rising dominance over
their world and yet at that moment it was more ironic than inspiring. A low fluttering growl
brought her attention back to her immediate surroundings as an irritated pedestrian pushed past.
Stopping too long on the busy streets wasn't advised.

She cantered across Darnar and continued along Hillart under a row of trees. The tall buildings
shielded the way from the Nermella and the air was unnaturally calm. Just ahead she saw a news
truck parked in a service road between two buildings. On the sidewalk stood a reporter
apparently interviewing passersby.

Jernumia stopped short and considered moving to the other side of Hillart to avoid them. Then
she decided to just walk by as usual. The team was busy with a young couple and they would no
doubt just ignore her. Of course she wasn't so lucky.

``Now here is a professional woman,'' the reporter said to her just as the couple moved on. The
camera aide swung around behind the reporter and pointed the camera at Jernumia.

``I don't have any time for an interview,'' Jernumia said curtly.

``This will only take a moment. Just a couple questions.''

Jernumia sighed.

``Where do you work?'' the reporter asked, extending his trunk toward her, tentacles wrapped
around a microphone.

She hesitated. ``I work at the legal office of Aratok.''

The reporter clicked his teeth and glanced back at the camera aide. ``Oh, I see. Very good. What
do you do there?''

``I'm a prosecuting adviser,'' she replied. ``I believe that was two questions, enjoy the
moment.''

Jernumia started to move on but the reporter shook his ears and stepped slightly in front of
her. ``I can see from your answers that you are, indeed, a lawyer,'' he said. ``Just one
question\ldots\ do you think the arrival of the Visitors will be good or bad for our world?''

``I certainly do not know,'' said Jernumia, ``now if you please I must get on.''

The reporter bobbed his head respectfully and stepped aside, moving on to the next person who
stumbled into his trap. Relieved, Jernumia continued down Hillart. She only went a few paces
when she heard a hoarse voice that stood out clearly against the din of the city.

``They is coming for you.''

She stopped abruptly. The voice had come from a old, broken man lying on the ground beside a
wall. He had a few nondescript trinkets spread out in front of him. He lifted his trunk and
pointed all four of his tentacles right at Jernumia.

``They is coming for you.''

He had an intensity in his eyes that sent a shiver down her back. Instinctively she held her
pouch tightly.

Someone bumped into her from behind and walked quickly around her. The old man pointed at the
newcomer as he walked away and shouted out, ``They is coming for you!'' Jernumia watched the man
momentarily and then, with a sway of her head, continued walking. She heard the old man calling
out to others, his voice fading into the sounds of the city.

The Aratok Agency was at the corner of Hillart and Following Street Varsynth. It was located in
an imposing stone building fully eight stories tall. At one time, not that long ago, it was the
tallest building in Varnok and maybe even the world. Jernumia trotted up the ramps, into the
main entrance, and walked down the interior mall, flanked by sunward windows, to the elevators.
From there she ascended four floors, looking out the elevator window down Hillart toward
Parliament and the river beyond.

She found her office and took off her pouch but she was only there a few moments before Haratol,
her court leader, burst in. ``It's about time you got here,'' he said, ``don't you have a
standard clock?''

Jernumia did not reply.

``Do you have the summary, at least?''

``Yes, yes.'' She rummaged in her pouch and pulled out a paper sleeve. From there she took out a
report, neatly typed, and gave it to Haratol. He cradled the paper in his trunk and flipped the
pages with his tentacles as he reviewed it.

``I'll read this later, but what's the finding?''

``In Runion 52 of Arnoxium 19, there was a case in Yartagard where the Bank Reform Act was
applied \textit{ax noforia}.''

``Oh?''

``The defense argued that full knowledge of the act had to be demonstrated and could not be
presumed, but because the act is a matter of public record that argument slipped.''

``Yes, I see that here,'' Haratol said as he scanned over one of the pages with his right eye.

``There's another similar case about 50 runion before\ldots''

``Good work,'' Haratol said.

Jernumia bobbed her head.

Haratol turned to walk away and she called out after him. ``Do you need me there?''

``No, I don't think so,'' he said without turning around. ``You've given me what I need.''

Jernumia returned to her table. She knelt down behind it and unpacked the rest of the papers
from her pouch, placing each carefully into its proper tray or drawer. When she was finished she
turned and looked out the window behind her. The sun hung over the city as always, sending its
ruddy light into her space. She glanced up at the sky but aside from a few wispy clouds against
the yellow-green background there was nothing to see. People say everything is going to be
different now, she thought, but nothing seemed different to her.

\spacebreak

Tuskara was not about to let a garjost stump be his undoing. With grim determination he picked
up his ax and started hacking at the roots yet again. Chips flew but somehow he never seemed to
make much progress. He pulled on the stump with his trunk. He kicked at it with his back feet.
He threw himself against it. Yet the stubborn piece of wood did not budge.

``Dind,'' he swore under his breath. ``\textit{Alnaron bi yeceit!}''

``If you had a sharp ax it might help,'' his wife called out from the open breezeway of their
greenstone home. He had no idea how long she had been watching him.

``It's sharp,'' Tusk objected. He held up the ax as if to show Jael the edge. ``I think it's
sharp,'' he added to himself. He ran the blade carefully over his exposed left front leg. It
didn't feel very sharp.

``And shouldn't you be wearing a hat?'' Jael continued. ``What if there's a flash?''

Tusk fluttered slightly and resumed hacking at the stump until Jael gave up and went back
inside. Several long moments passed and he finally started making some progress when he heard
the clang of the lifter as it stopped along the main road. He clicked his teeth and leaned the
ax up against the stump. He was panting from the exertion and grateful for the break. He held
his ears open widely to catch the cooling Nermella while he waited for his children to come up
the lane.

Mart trotted excitedly toward his father to see what he was doing. Tusk wrapped his trunk around
Mart's shoulders and stroked his face with his tentacles. Jinna just huffed and stomped into the
house with barely a word.

``What's with her?'' Tusk asked his son.

``They were picking on her 'cause of her markings,'' Mart said.

``Oh, I see.''

Tusk and Mart went inside to find Jinna arguing with her mother. ``Why do \emph{I} have to be
the one with these markings.'' She motioned with her trunk over her face and front legs. The
distinctive bands of her Forbinite ancestors were clearly, even if only faintly visible. ``I
wish I looked like you, Mom, and not like Dad.''

``Jinna,'' her mother started, ``you're a lovely girl. You look\ldots''

``I hate it. I hate it. I hate it,'' Jinna chanted. ``Mart doesn't even have these\ldots\
things.'' She left in frustration, stomping up the ramp to the upper overlook in disgust.

Jael walked to the bottom of the ramp and tilted her head to glare after her petulant daughter
with her left eye. ``Just be in a better mood when it's time to do prayers.''

``She'll be fine,'' Tusk said.

At dinner the family crowded around the small table in the hanaria overlooking the street.
Theirs was one of seven older homes in various states of disrepair that were gathered in a
circle. The common yard was small yet two families lounged on the weedy brikken grass enjoying
the moment. Jael waved her trunk up and down out the window and her two seleckia returned the
greeting.

% TODO: More social dynamics with the seleck?

Jinna had calmed down but she didn't say much and picked at her food. Mart, as usual, was the
chatterbox.

``We talked about the Visitors.''

``Oh?'' Tusk replied as he put his sarnok pod into the pod cracker.

``Yeah, we talked about why they're here.''

``What do you think?''

Mart shrugged. Then with a loud click he said, ``Maybe they can take away Jinna's markings.''

Jinna put down her tongs with a clatter and glared at him with her angry eye. ``Maybe they'll
eat your brains.''

``Now, now you two,'' said Jael. ``Jinna\ldots\ you're the oldest. Don't let him get to you.''

``Mom!''

Jael ignored her daughter and picked another pod from the center plate instead. ``It's just as
well they haven't done anything all this time,'' she said, ``whatever they want I'm sure it's
not for our benefit.''

``They're probably just trying to figure us out,'' Tusk replied.

Jael snorted slightly. ``They'd be better off staying away from us. I don't see why the First
Citizen had to meet with that Forbinite about it.''

``It's good to see Garnick being used for something positive for a change,'' Tusk said.

After meal and prayers, when the children were quiet, Jael spent some time working the family
budget.

``You need to get a job, Tusk'' she said. ``My income's not enough. Even with the seleckia
helping we'll have trouble paying our share of the rent.''

``I know \textit{tinka},'' Tusk said. ``I saw an opening at the trans-school teaching Rabik. I
was thinking of applying for it.''

Jael folded her ears close to her head. ``You always say these things but you never do them.
Ever since you lost your job with the military you've been skulking around. Apply for that
position now.'' She pushed the papers aside and put her trunk heavily on the table, leaning her
head against her tentacles.

Tusk wrapped his trunk around his wife's shoulders and she leaned against him. She shuddered
slightly as he caressed her ears. ``I will go as soon as I've had a bit of sleep,'' Tusk said.
Jael snorted.

It was several sleeps and more prodding from his wife, before Tusk finally found himself lying
in front of the school master's desk. The short, elderly man flipped through his application
with well-creased tentacles. Tusk could hear a group of children chanting vowel sounds in the
room next door using the same rhythm he learned when he was in school.

``You have an impressive work history,'' the school master said finally.

Tusk bobbed his head. ``I learned the basics of Rabik as a child while watching television, but
I studied it extensively when I was a linguistics student.''

The school master eyed Tusk carefully. Tusk shifted uncomfortably. The chant next door had
stopped.

``You might be over qualified.''

Tusk swayed his head briefly. ``I think I could give your students an exceptional course, and I
would welcome the opportunity to work with our young people.''

The school master flipped to Tusk's references. ``It says here you had a level two military
clearance. It also says you were released because of a problem with charara leaf.''

``Yes,'' Tusk replied, ``I went through a rehabilitation program. I can assure you that is no
longer a problem. There's a certificate from the program on the last page.''

The school master flipped to the back of Tusk's application and pretended to look at the final
sheet, holding the paper close to his right eye. After a moment he looked at Tusk over the edge
of the page. He returned the application. ``I'm sorry,'' he said. ``I just don't think you're
right for us.''

``Of course,'' Tusk said softly. He took the papers and returned them to his pouch. Standing up
he touched the school master's shoulder briefly. ``Thank you for your time.''

On his way out of the building he crumpled the application in his tentacles and threw it into a
wastebasket beside the school entrance. ``\textit{gar'k tic narloblia}'' he cursed softly in
Garlonick.

Tusk stood by himself on the lifter on the way home, his tentacles wrapped around the hook.
Nobody stood near him because of his markings. He was used to that. He looked out the window at
all the fine houses and businesses of Pilkity and sighed. He hadn't had a proper job in many
runion. Would they have to move to, say, Nargon in the north? That was a much cheaper area to
live. Jael would have to find a new seleck. She'd hate the idea and so would the kids.

% TODO: Do people have a problem with Tusk's "mixed marriage?"

A large sign in the window of a shop caught Tusk's eye. He cocked his head slightly to read it.
``You're not \emph{my} visitors!'' the sign cried out. He turned away and tightened his grip on
the hook.

When Tusk got off the lifter the Nermella gusted and picked up some sand in the street, blowing
it across his face. The leaflets of the trees rattled in the wind and the rigat tubes whistled
softly. Tusk slowly walked down the lane to the circle of houses he called home.

Jinna was lying on the ramp. ``How did it go?'' she asked. Tusk shrugged. ``You didn't get it,
did you?''

``No.''

``That's pathetic. How could you not get that job? Did you even apply?''

``Jinna\ldots''

``I hate this family,'' she said going inside.

\spacebreak

The Knoll was on the east side of the Marlock campus. It was a peaceful escape for the students
from the rigors of their studies and from the bustling city of Argenhost. From the top of the
Knoll, between the short larzippia trees one could see the financial district where several
tall, three story buildings rose above Cross Street Warlash.

The Nermella was whistling in the rigat leaves as Colty made his way up the well worn path. He
reached the bend where the paved walkway curved around a large rock. Instead of following it
toward the lookout at the top of the Knoll, he continued across the grass and through some
bushes. There, in an out of the way spot, was a surprisingly well groomed clearing with a small
monument to the college's founder. Jingas was there waiting for him, lying on the ground with
her pouch beside her.

``I thought you got lost,'' she said.

``Sorry I'm late,'' Colty replied with a bob. ``It took longer to submit my final paperwork than
I expected.''

Jingas patted the ground invitingly and Colty lay down next to her. She was lovely. Her dark
eyes and large, beautiful ears were mesmerizing. She had a delicate, feminine trunk, and a
slender body. She had taken off her foot pads and her toe shields were meticulously cut to
perfect points. Colty clicked his teeth and she did the same in response.

``You look very nice today,'' he said.

Jingas spread her ears. ``You flatter me too much.''

She opened her pouch and took out some plates and food. ``I brought some pressed flannen with
crocker sauce,'' she said, ``and some ligas leaves, of course.''

``Oh, ligas, sounds great,'' Colty said.

Colty took a plate from her and held it with his tentacles while she put a large, slightly
cooked leaf on it. He waited while she put a leaf on her plate as well. Then she held up the
plate and he did the same. ``We thank Jurita for this food,'' she said softly. She tilted her
head so her right eye faced the sky as she spoke the traditional prayer but Colty couldn't
resist looking at her instead. The two started to eat, holding the ligas in their tentacles and
nibbling on the tough but flavorful leaves.

``I know the department is having a thing for you, but I thought it would be nicer for us to
have our own thing.''

``Much nicer,'' Colty agreed. He stretched out his back legs to relax.

``So, Parnon, huh?'' Jingas said with a click of her teeth. She put the containers of crocker
and flannen between them. ``That's very impressive.''

``I was surprised. I assumed I had no chance with them. You know I applied on a whim. Parnon can
choose the best of the best for its faculty, and I'm just a fresh graduate with only a few minor
papers to my name.''

Jingas tossed her ears and fluttered slightly. ``You are the best of the best,'' she said, ``you
got that award for your paper\ldots\ what was it again?''

``Self-reflective mono-sequences as higher dimensional number folds.''

``You see what I mean?'' Jingas said, ``That stuff is pretty incomprehensible to regular people
like me.''

Colty took a little crocker and flannen to spread on his leaf. ``It's not, really. It's just a
cute curiosity. Nothing more.''

``You'll do fine at Parnon,'' Jingas clicked. ``A place like this,'' she added with a nod toward
campus, ``would bore you before long.''

``Maybe. I only hope I can live up to the expectations. I'll have a lot of work to do before I
can get full status. More research, more papers, more conferences, not to mention teaching.''

``You love it.''

Colty looked at Jingas as she chewed her leaf. He almost said, ``I love you,'' but held himself
back. He wasn't sure why.

They finished their leaves and lay back on the grass comfortably to rest. It can take some time
to digest ligas, even with the flannen, and it was best not to jump up too soon after eating.
The sky was beautifully clear and green without a cloud visible. The Nermella, slowed to a
gentle breeze by the forest and warmed by the bright sun, played over their bodies and brought
fragrant smells to their nostrils.

``It's hard to believe we are in the middle of Argenhost,'' Jingas said finally. She paused for
a moment. The sounds of traffic could be heard faintly on the wind. ``It's like the city is a
million karnons away.''

Colty stared into the sky pensively with his left eye.

Jingas reached out her trunk and touched his leg with one tentacle. ``What are you
thinking about?''

Colty paused uncertainly. ``The Visitors.''

``Them again?'' She pulled her trunk back and folded her ears.

``They circle around up there looking down at us, never doing anything. They're studying us like
bugs under glass.''

Jingas turned her head to look upward as if she expected to see the alien ship passing over
Argenhost at that very moment. ``At least they are letting us get on with our lives.''

Colty just watched the sky and Jingas slid over to lean against him. The two lay quietly
together for a time and then Jingas sighed. She sat up on her front knees with her legs folded
under her and regarded Colty closely.

``Yes?'' he asked.

She bobbed her head a couple times. ``Colty, I want you to know that I'm not going to ask you to
marry me.''

Colty's ears folded back and he turned toward her. ``I\ldots''

``I know, I know, you weren't expecting that.'' She bobbed her head again. ``I think you are
though.''

``Um\ldots''

% TODO: Remove the noise word?

``I'm not going to because, well, because right now you love your work too much. I don't want to
distract you.''

Colty felt his heart sink. ``You're not a distraction.''

``I would be. I would have to be. I can't be the wife of a man like you. I would take too much
from you.''

``Jingas\ldots''

``No, let me finish.'' She took a breath. ``I think you're very special, but you work very hard,
you travel a lot, and you are totally focused on your math. There is nothing wrong with any of
that. It's good. It's right. What I need to do now, before you go to Parnon, is set you free.''

Colty got up on his front knees as well. ``Jingas, I really like you. I'm sure we could work
it out.''

She shook her head. ``We've had some good times, Colty, and I'm really, really happy I got to
know you. But when you leave for Parnon I think that should be the end for us. I'm sorry.''

Colty turned his head away and curled his trunk. ``I'm sorry too,'' he said.

``You're a good guy. I know you will meet the right girl eventually.''
