% Local IspellPersDict: aspell.dunari.pws

\chapter{7}

Modifying the Ear to send modulated audio messages instead of RRRF pulses had been very easy.
The team decided to use a simple amplitude modulation scheme on the idea that it was so basic
that it could certainly be decoded by the Visitors without prior arrangement. The Office of
Alien Affairs had already worked through some scenarios and the team quickly agreed to use the
hyperfine hydrogen emission frequency as the carrier.

The tricky part had been figuring out what to say. After much debate between Chark's team and
Lucasa's office it was finally agreed to first send a beacon in an effort to get the Visitors'
attention. The beacon would consist of sets of pulses with a prime number of pulses in each set:
two, then three, then five, and so forth. Such a signal, it was felt, would be unusual enough to
stand out against the background din of dunari radio traffic, and also be unmistakably intended
for the Visitors. Once the Visitors replied, if they did, they would take it from there.

Most of the time RRRF research continued at the Ear as usual. However, each time when the
Visitors' orbit passed over Lungast the focus shifted to establishing communication with them.
At first Lucasa was always there, but as time passed without any response from above, Lucasa's
busy schedule started taking priority and he left the regular sessions entirely to Chark's team.

Brogan argued that they should try some other approaches: different frequencies, different
modulation schemes, even a different message. ``Maybe the Visitors are monitoring aerocar
control channels,'' he suggested, ``and will never listen to what we're broadcasting now no
matter how long we do it.''

Yet Chark felt consistency was important. ``If we dance around the spectrum we'll be more likely
to miss each other. If I was them I'd want to make a complete survey of all frequencies. They'll
find our signal sooner or later.''

``Maybe they just don't want to talk to us,'' Fanlon said.

One particular time, much like any other, the team gathered together to listen. They had
dedicated a small corner of the lab for this purpose, having moved a few comfortable seats in
from the lounge. Brogan and Tarmala tried to guess the odds that they'd hear a response. Chark
and Fanlon just relaxed, taking the moment as a break in their RRRF work.

Tarmala configured the Ear and quickly picked up the Visitor ship. He was practiced at it so it
was easy. Brogan changed a few plugs and threw a few switches. The familiar sounds of their
transmission filled the room. After the final pulse was sent the circuitry automatically
switched to listening mode and the soft sound of static washed over them.

Everyone waited quietly, listening to the hypnotic hiss. Brogan leaned over to Chark. ``When are
you and Junella going to take that vacation?''

Suddenly the static disappeared and was replaced with shocking silence.

``Is something wrong with the listener?'' Chark asked. Tarmala was already twisted around and
looking at the equipment over his head.

``It's a carrier!''

Suddenly a strange, alien voice came from the speaker. It spoke slowly, deliberately, and with a
eerie, hissing screech.

``\alien{Weee requessst meeeeting\ldots\ at Garnick\ldots\ hour 75\ldots\ thisss run-ion,
  Argen-ian massster tie-em.}''

``Jurita's Witness\ldots'' Brogan said, his voice a whisper.

Chark sat up and anxiously pointed toward the tape recorder. ``Are we recording this?''

``Yes, yes,'' said Brogan quickly glancing at the device, ``definitely.''

Fanlon checked the Ear's slew path and made sure the plotter was working as well. Everyone
crowded around the speaker.

``\alien{Weee requessst meeeeting\ldots\ at Garnick\ldots\ hour 75\ldots\ thisss run-ion,
  Argen-ian massster tie-em.}''

``They're repeating the message,'' said Chark. ``They want to be sure we get it. Can you
configure the Ear for transmission again?''

Brogan objected, ``we might miss something.''

``We have to ACK.''

``\alien{Weee requessst meeeeting\ldots\ at Garnick\ldots\ hour 75\ldots\ thisss run-ion,
  Argen-ian massster tie-em.}''

Tarmala quickly went to the transmitter panel and switched a few plugs. The alien screech went
silent. He threw some switches and plugged in a microphone.

``They'll be below the horizon soon,'' Fanlon said.

Tarmala passed the mic to Chark. ``You're on the air Captain Charkonaless.''

Chark held the mic to his lips, wrapping his tentacles tightly around it. He paused uncertainly
and then spoke. ``We hear your message and understand it.'' The team looked at him intently.
``We hear your message and understand it.'' Chark paused again and then with a shrug repeated
his message one more time.

``That's it,'' said Fanlon. ``They're below the horizon.''

Everyone started breathing again. ``Okay,'' Chark said. ``Okay.'' He rubbed his trunk ridge with
his tentacles. ``Get Lucasa on the phone.''

\spacebreak

The grass was lush and the gloranga flowers were in bloom. Tusk was lounging in the back yard
flipping through the help wanted section of the community docket. He wondered if he should apply
for the waiter position at the Wrap-and-Go down the street. It was a humiliating thought.

He looked up at the clouds, rubinum-red, surrounded by a deep orange-green sky. The world seemed
to glow in a golden light filtered through the dusty winds high above. The Nermella was little
more than a gentle breeze of summer air, and despite Tusk's troubled thoughts all was at peace.
After a few moments he got up and took a watering can over to the garden. Mart came out of the
house to help, sent there by his mother. Tusk was glad to have him.

When they finished Tusk put the tools away, picked up the docket, and went back inside. Jinna
and Jael were talking and when Jinna saw him she ran over to him.

``Dad,'' she said, ``my friends and I want to start a seleck. Is that okay?''

``A seleck?'' Tusk asked. He glanced over at Jael but she just shrugged in that helpless way
parents do when their children just won't let something go.

``I think you are a little young for that.''

``Lots of girls are doing it now. And it would be really good for me. It could help me get
better grades.''

Tusk shook his ears. ``Well, that sounds good.''

``So I can?''

``Um\ldots\ well, I think your mother and I should talk about it.''

The phone chimed and Jael stepped over to answer it.

``It would just be a starter seleck,'' Jinna persisted. ``To help me between now and went I join
a real seleck later.''

``A starter seleck,'' Tusk said clicking his teeth.

``Tusk?'' It was Jael. She held the phone in her trunk with a concerned look on her face.
``There's a government person on the phone for you. It's someone named Lucastanonia.''

\spacebreak

Chark waited in the central room with his bag beside the door while Junella and Jiapita bustled
in the hanaria. He hadn't traveled without his wife since the military had flown him down to
Lungast to inspect possible locations for the Ear. He didn't like traveling without her now. She
didn't like it either.

``Please be careful,'' she said standing in the arched opening of the hanaria. ``Garnick is so
far and we don't know what these Visitors are going to do.''

``Don't worry,'' Jiapita's voice came from behind, ``he'll be surrounded by the military.''

``Why do you have to be there, anyway?''

``Lucasa trusts me,'' Chark said.

Junella fluttered slightly and tried to distract herself by helping Jiapita. Finally Chark heard
a car pull up in the lane. It had the distinctive rattle and deep moaning engine of a military
vehicle. Junella went to the door.

``I'm here for Captain Charkonaless,'' a young voice said.

``Yes, yes,'' Chark said getting up. Junella reached out for her husband as he gathered his bag.
They crossed trunks, and then again. She shook slightly. Chark tapped her and then Jiapita as
well.

``Why Garnick?'' Junella asked.

``I don't know, \textit{tinka}. Maybe they like the heat.''

The young driver picked up Chark's bag and slung it over his back. Then the two of them walked
down the stone path toward the waiting vehicle. As they drove off, Chark saw his wife standing
in the doorway waving her trunk with Jiapita behind her. He put his tentacles against the window
and wondered what the world would be like when he returned.

The first part of Chark's journey was an aerocar flight from Lungast to Doggenbrash. Chark had
been on an aerocar only once before, and he hadn't enjoyed the experience very much. That
aerocar had been cramped and noisy, and the ride was rough and bumpy. It had made him sick. So
it was with trepidation that he walked out onto the landing pavement at the Lungast base. Yet
the aerocar before him was not like the little bi-wing that he had been in before. Instead it
was new, modern, and big. It had two fixed, swept back wings and a tall tail.

``Captain Charkonaless, sir,'' came the greeting from the pilot as Chark stepped on board. He
saw not one cramped little seat as he was expecting, but rather four spacious seats, two on each
side of the compartment.

``I'm impressed. Who else is joining us?''

``You must be a VIP,'' the pilot replied. ``It's just you today, sir.''

Chark settled down in one of the seats and looked out the window while the pilot did his
pre-flight check. The sun slanted into his right eye, its ruddy light cutting across his face
and splashing onto the opposite side of the compartment. In the distance, past the runways and
past the fence around the base, Chark could see the roof tops of Lungast. He saw the trees
blowing in the Nermella, and he saw a bird flying from one branch to the next, probably for
nuts. Chark sighed. Where he was going there were no trees, no birds, and no nuts.

``All set,'' the pilot said. ``Get ready, sir.'' Chark sat up straight and adjusted his
restraining straps while the aerocar rolled out onto the runway, facing the Nermella to get some
air speed for free.

The take off was not nearly the ordeal Chark had feared. The new aerocar was obviously more
powerful and could get into the air with less struggle than the one Chark had been on before. He
watched with fascination as the streets and buildings of Lungast rushed under him. They were
climbing steadily and as they did Chark could see more and more.

``How you doing, sir?'' the pilot called out loudly to be heard over the roar of the engines.

``I'm fine,'' Chark replied in kind. It was noisy but the ride was smooth and he didn't feel
queasy at all. ``This is a nice bird.''

Chark settled down for the flight. At first he spent most of his time looking out the window.
It's not everyday when one sees Rujar from the air. The land seemed so lush. There was yellow
everywhere you looked. The towns and roads cris-crossed an endless plain of forests and fields.
They flew over the Zagenhos River. He could see the twisting banks rimmed with tall trees,
looking like little sticks. Far away, against distant clouds, he imagined that he could just see
the wall of night. The Narlar, the boundary between everlasting day and everlasting darkness,
was not far from Lungast.

``Sir\ldots\ the Argenian Sea.''

Chark looked down and there he could see the shore of the great sea of Argenia. Floating
carelessly were some boats and beside them a beach. Yet soon they passed over all that and were
flying across the open water. It was not entirely featureless. The dark yellow-red color varied
from place to place depending, Chark supposed, on the concentration of microscopic organisms
living there.

They flew for many karnons over the sea. The distant shore was barely visible as a murky line
against the horizon when finally the opposite shore passed under them. Chark breathed a sigh of
relief. He felt more comfortable knowing there was dry land under his feet, even if it was far
below.

Eventually the aerocar turned more directly west out over dryer regions on its way toward
Doggenbrash. With the sun in Chark's face and the Narlar at his back the feeling that he was
heading for an historic moment was strong. He read for a while, and then distracted himself with
a puzzle. For a time he dozed in farth-sleep and even some paren-sleep. When he awoke he could
see they were over the Northern Plains. The ground was flat and speckled with hargbush and riven
weed. Chark was glad to see it. For one thing it meant they were getting close to their
destination. Yet also this was the land where he grew up.

The pilot was on the radio chattering with flight control in Doggenbrash. ``You'd best buckle up
again, sir,'' he said. He cocked his head toward the slowly looming city in front of them.

When Chark finally got off the aerocar he breathed deeply. The air was dry and fresh. He really
did prefer it to the dank, moist air around Lungast. He had time to wander around a little
before starting the next leg of his journey. He couldn't resist getting some mosen cake. It had
been so long since he had any.

% I wonder if Chark could do something useful, in the sense of forwarding the plot, during his
% stop at Doggenbrash.

When he returned to the base he was greeted by a military officer. Beside him was a man of
obvious Forbin descent. Chark nervously wondered what it was about.

% TODO: Describe the officer better?

``I'd like to introduce you to Tuskara,'' the officer said. Chark politely, even if reluctantly,
clasped trunks with the Forbinite. ``This is Captain Charkonaless,'' the officer continued his
introductions.

``We are well met Captain,'' Tusk said, tilting his head slightly.

``What's going on?'' Chark asked.

``I'm a language guy,'' Tusk said. ``Apparently they think that might be useful.''

``Are they expecting the Visitors to be speaking Forbin?''

Tusk snorted.

The aerocar to Garnick was specially modify to endure the heat of the deep day. The engine
design had to compensate for the high temperature air intake. The top surface of the wings and
fuselage were coated with reflective paint. It was also much less comfortable than the aerocar
Chark took to Doggenbrash. There was barely space for the pilot along with all the supplies that
were packed into the cabin. The military was not about to waste a flight to Garnick by just
carrying people. So much for being a VIP, Chark thought.

The two men pushed their bags into a cranny and squeezed into the seats directly behind the
pilot. Chark was not one to complain but he didn't like being in such close quarters with the
Forbinite. Fortunately his seat was a little forward so he could talk to the pilot more easily.

The take off was terrifically bumpy; the complete opposite of Chark's experience at Lungast. He
started to wonder if the aerocar was breaking up just as it finally lifted off the runway. The
pilot leaned back slightly and clicked his teeth. ``The rubber in the tires is designed for the
landing pavements of Garnick.''

The flight was uneventful but uncomfortable. Chark avoided touching Tusk. Instead he made small
talk with the pilot. Tusk spent most of the time looking intently out the window, and not paying
attention to anyone.

The flat plains changed to rolling hills covered with brikken grass, and they flew over karnons
and karnons of open grassland. It was the ancestral home of the dunari people. They had evolved
on those grasslands, in the far distance past, and small, trunked quadruped animals still grazed
it. Yet as they flew, the brikken grass became progressively patchy, struggling more and more to
find a living in the increasingly unforgiving desert.

Gradually the hills became larger and more rugged. Serrated crags reached toward the sky. The
pilot had to climb high to stay a safe distance above the rising mountains. Chark glanced over
at his traveling companion but Tusk appeared to be in farth-sleep at least.

``We'll catch the Nermellum on the other side of these mountains,'' the pilot explained. ``It
makes the trip a little faster.''

%%%%%
% INFODUMP
% The following paragraphs change point of view. Can this material be better integrated?
% Can the comments about Garnick can be moved to Chapter 2 and the Light Sights program?

The frigid Nermellum wind, blowing from the deep night, passed over Rujar's poles into the day.
As the Nermellum made its way across the face of Rujar, it would provide an extra boost to their
aerocar and cool them. On the ground the Nermellum moderated the heat enough for the brikken
grass to grow. Yet at Garnick there would be no relief and no grass.

Garnick was located at the northern edge of the great Barlok Expanse. Directly beneath the
perpetual sun, Barlok was a searing wilderness long thought to be devoid of life. It wasn't
until aerocars made exploration of Barlok feasible when strange life forms were found there,
living in miniature ecosystems entirely contained in the shade of an overhanging rock, isolated
from each other by the blistering sands.

Garnick was built many runion ago by Argenia, ostensibly as a military base for training cadets
in deep day survival. Everyone knew that it was really a way for Argenia to assert its
dominance, some say arrogance, by creating an outpost so far forward that it was nearly on the
doorstep of Forbin. Neither side ever had any designs on the wasteland that was Barlok, but the
existence of an Argenian base there was a long standing thorn in the Forbin hide. Although the
strategic importance of Garnick was questionable, the symbolic significance of it loomed large.
During the Harkite War Garnick changed sides many times as it was taken and retaken, ultimately
ending up under Argenian control again as per the treaty of Visnik.

% END
%%%%%

Chark pondered why the Visitors would choose such a spot for a first meeting. Was it as he
suggested to Junella? Were they more accustomed to the scorching heat of the Rujaran deep day?
He felt a certain anxiety. How strange would the Visitors be? It was all very well to talk
abstractly about meeting advanced aliens but when it came down to actually doing it, Chark was
having second thoughts.

``The Road to Hell,'' the pilot said suddenly. Chark looked up and saw him pointing toward the
ground. Tusk also stirred and looked out the window. Scrawling across the barren rocks they
could see the faint remains of a road. Here and there it was covered by wind blown sands but in
other places it was clear.

Garnick was built before aerocars existed. Incredibly a road was cut across the harsh desert for
thousands of karnons to connect the outpost with the rest of Argenian civilization. Chark tried
to imagine what it must have been like to be stationed at Garnick at that time, knowing your
only way home was that thin, narrow road scratched into the shifting sands.

``Are you married, Captain?'' Tusk suddenly asked.

``Huh? Ah, yes,'' Chark replied.

``Kids?''

Chark shifted a bit in his seat. ``No we don't have any children. We had a son, but he died from
Puget's Disease.''

``Oh, I'm sorry.''

Chark wasn't sure why he told the Forbinite that. He didn't normally discuss Chartik with
strangers. Maybe it was something about being so far from everything, packed into a small, hot
aerocar cabin together.

Tusk pulled a wallet from his pouch. He took out a couple of pictures. ``My kids,'' he said.
``This is Mart\ldots\ and here is Jinna.''

Chark looked at the pictures with some interest. They seemed like ordinary children and the boy
didn't have a trace of Forbinite markings.

``This is my wife, Jael,'' Tusk said. ``We live in a little town called Pilkity, not far from
the Sea of Organite. Jinna likes to go swimming there.''

``Swimming?'' Chark said with surprise. ``She's an adventurous girl.''

Tusk seemed pleased.

Finally, after what felt like an eternity in the air, they came to Garnick. Chark tried to
stretch his legs a little in anticipation of finally getting out of his seat but there was no
room.

As they descended the pilot had to circle around to approach the proper runway, giving them a
good view of the base. It didn't look like much. There was one main building of single story
brick construction. There were a couple of smaller out buildings, several hangers, and a
building attached to a large tower bristling with radio antenna.

``Buckle up,'' the pilot said as he started his final approach. ``We'll be on the ground
shortly.''

The landing wasn't as shocking as the take off had been. The heat of the deep day had softened
the tires just enough for them to absorb the worst of the bumps. Yet the runway was short and
Chark felt nervous that they wouldn't be able to stop in time. Somehow they managed it.

After touching down the pilot taxied toward one of the hanger areas. There were some people
standing on the landing pavement waiting for their arrival. Both Tusk and Chark stirred
impatiently. The pavement might have been hot enough to boil water, but anything seemed better
than being stuck in that aerocar for even another moment. When they finally parked and the side
door swung open, Chark felt a blast of baking heat fill the cabin. Standing there wearing
reflective clothing was Lucasa.

``Welcome to Garnick!''

