% Local IspellPersDict: aspell.dunari.pws

\chapter{11}

Chark was briefed by Lucasa in a small room at the corner of the main building. The windows were
covered with insulated shades and the room was lite entirely with artificial lights. At Garnick
you did what you could to keep the sun out. Also there was no conference table nor any seats.
Such comforts took up valuable space for minimal return. The room was utterly barren.

``You should know,'' Lucasa said, ``that the Forbinites have been invited to the party. We are
expecting a delegation to arrive within the hour.''

``Forbinites! What about that man who flew in with me?''

``You mean Tusk? His father was a Forbinite defector during the Harkite War. He's worked for
Fotkey before. He's been called the most brilliant linguist in Argenia---maybe even in the
world.''

``So he's one of us?''

Lucasa swayed his head. ``You provide the channel over which we can communicate with these
creatures. I want Tusk to help us figure out how to actually talk to them.''

Chark was distracted. ``How did the Forbinites find out about this meeting?''

``The Visitors contacted them as well. After that they are here on the good graces of the First
Citizen.''

As part of the Argenian delegation Chark was on hand when the Forbinite aerocar arrived. Lucasa,
Tusk, and General Fotkey, as well as some military police guards, were also present. As was
traditional they met the arrivals on the landing pavement even under the broiling sun. The
reflective clothing helped, but it seemed to Chark that every breath burned his lung.

The Forbinites were larger than the Argenians, even larger than Lucasa, and had bold markings on
their faces, trunks, and legs. The precise pattern was unique for each individual, but always
symmetric from side to side. Their cloaks were fitted closely to their bodies covering even
their undersides. Each wore a traditional \textit{grung} on their head, a small cloth cap that
clipped to their ears. Their clothing was mostly all yellow gold and decorated with the
serpentine motif that symbolized their country.

Chark felt intimidated by their size and the atmosphere was tense as they approached. The guards
stood at ready, trained to be alert in the presence of the enemy. One of the Forbinites stepped
forward approaching Lucasa. Chark could see that it was a woman.

``I am Kenellio,'' she said in Argenian with a heavy accent, ``the special High Council
appointee to the alien situation.''

``I am Lucastanonia, the Director of the Office of Alien Affairs. We are well met.''

Lucasa extended his trunk but the Forbinite did not return the gesture. Instead she introduced
the other members of her delegation: Field Marshal Noxan of the Forbin military, and Sar.
Tulurkian from the astronomy department at Farlozia University. Lucasa introduced the Argenian
delegation as well.

Thankfully from Chark's point of view the three Forbinites all spoke reasonable Argenian. Yet
Noxan spoke briefly to Tusk in Forbin. Tusk swayed his head slightly in response. ``I trust your
father is well,'' Noxan added.

``He died several runion ago.''

``My condolences.''

Tulurkian greeted Chark with enthusiasm as one scientist to another. ``It's a shame we can't sit
and talk,'' he said as he clasped Chark's trunk, ``maybe in better times.''

The Forbinites were shown into the building and Chark was grateful to return to the cooled
space. They removed their reflective outer cloaks and then Lucasa invited the guests to review
their quarters.

``We should like to discuss the aliens,'' Kenellio replied. ``You must tell us what you know.''

``Perhaps you would like to rest a little first after your long flight,'' Lucasa said, ``maybe
have something to eat.''

``We do not need rest such as you.''

Fotkey led the way to the small conference room that Lucasa had used for his briefings. The
Argenian military police brought up the rear and waited outside the door while the seven dunari
crowded into the space.

% Fotkey, Lucasa, Tusk, Chark, Kenellio, Noxan, Tulurkian

``Are we to be escorted by armed guards the entire time we are here?'' Noxan asked.

``You are guests of Argenia,'' Fotkey replied, ``but understand this is a military base.''

``I expected nothing less from the `hospitality' of Argenia,'' Kenellio said.

``You're here, are you not?'' Chark blurted out.

Kenellio glared at him. ``We are here because we choose to be. We do not need to be here.''

``We don't need you here either,'' Chark said coolly.

Fotkey raised his trunk. ``The Forbin delegation is here at the invitation of the First Citizen.
I suggest we focus on the matter before us.''

Kenellio started intently at Chark until he turned his head aside. ``Quite right, General,'' she
said.

Lucasa then asked Chark to summarize his efforts to communicate with the Visitors. Lucasa played
a recording of the initial contact. Kenellio and Tulurkian listened to the alien hissing with
interest. ``That's identical to the message we received,'' Tulurkian said. ``I think it's the
same word for word except our message was in Forbin, of course.''

``So the aliens are actively trying to engage both our nations,'' Kenellio said.

Tusk snorted slightly.

``Do you have a comment?'' Noxan asked with obvious disdain.

``If it takes a voice from space to bring our peoples together, I'm all for it.''

``These events do not change Forbin policy toward your nation,'' Kenellio said.

``Of course not,'' Tusk replied.

The meeting went on much longer than Chark expected. It was mostly without incident but tempers
flared now and then. Chark found it exhausting and was glad when the Forbinites finally agreed
to inspect their quarters. There was still some time before the Visitors arrived and Chark
wanted to just relax as best he could before then.

Tusk touched his shoulder as the Forbinites headed down the hall with their escort. ``Are you
hungry? We should eat a little.''

Chark wasn't entirely comfortable eating with a person he had just met but he needed the food
and at Garnick one couldn't be particular about social customs. Anyway, compared to the
Forbinite delegation, Tusk seemed like an old friend.

The food dispensary was at the other end of the base. Like everything at Garnick it was small
and cramped. There were a few tables but no privacy partitions. Eating times were normally
staggered so only one other table was in use when Tusk and Chark arrived. The food was nothing
special. Mostly it was dried or canned or pre-cooked in some way.

Chark took a plate and picked out an edible, even if uninspiring, collection of roots and dried
fruit. Tusk was more interested in the crackers and nizar spread although he took a few roots as
well.

``It's not like Jael makes,'' Tusk said as he lay down across from Chark, ``but it will have to
do.''

They ate in silence for a time, and avoided looking at each other out of respect for their
mutual privacy.

``I don't like the Forbinites,'' Chark began. He stopped abruptly.

``They're an arrogant bunch,'' Tusk said.

``Why do they have to be that way?''

``It's called \textit{huzaro}\ldots\ a way of being. In their case it's a kind of bluntness.
They see it as being honest. They tend to think all our polite talk means we are trying to hide
something.''

``I'm an engineer, not a diplomat,'' Chark replied with a sigh.

``We are all diplomats to an alien race. Who knows what \textit{huzaro} they have.''

As the meeting time approached all seven dunari gathered in an observation room overlooking the
runway where they donned reflective clothing and got ready to meet whatever came on the landing
pavement of Garnick. The Forbinites objected to the procedure saying that it was unnecessary to
actually stand outside in the oppressive heat.

``It's a quaint Argenian custom that makes no sense in this situation.'' Kenellio said. After
much discussion it was agreed that they would wait in the observation room until the Visitors
appeared and then go outside as they were landing.

Fotkey had a portable radio strapped to his front leg that he used to communicate with the
control room. This allowed the delegation to keep track of events as they unfolded and Fotkey
spent much time talking on the device.

Just as Chark was hooking his clothing he heard the roar of aerocar engines. Through the large
observation room widows he could see a fighter speeding down the runway to take off. Shortly
after another fighter followed. ``Are we expecting a battle?'' he asked Fotkey discretely.

``I'm not expecting anything. Just taking precautions.''

Once all was ready everyone waited in silence. It was an eerie sensation. Seven dunari, from
both nations, wearing special cloaks and standing together in the middle of the wilderness
waited to hear a voice from the heavens.

Chark whispered quietly to Tusk, ``We are the seleck of Garnick, I think.''

Lucasa raised his foot and looked at his watch. ``If they are punctual it should be soon.''

Outside, the barren landscape was ablaze with the bright orange light of the Rujaran sun. The
lighting at Garnick was direct, coming down as it did from almost overhead, and very different
than the muted light Chark had become used to at Lungast. The shadows were almost non-existent
and the rocky hills about the base seemed to glow with the heat.

Fotkey scanned the sky with a small trunk telescope.

``Look!'' said Tulurkian. He pointed above a ridge. In the distance there was a streak. It was
very far and very high, yet it moved with unnatural speed. Chark felt a shiver down his back.
They were coming.

The streak turned toward the base, approaching and descending quickly. The dunari shifted
uneasily. Fotkey talked on his radio. He turned up the volume on his set. A crackle came from
the speaker.

``This is Garnick base,'' said a dunari voice, ``to incoming craft: please identify.'' The reply
was quick and to the point.

``\alien{Weee requessst land.}''

Chark's heart pounded. Although everything was as expected, it all felt very surreal. He looked
over at Tusk but Tusk was studying the approaching streak intently.

Fotkey clicked the button on his radio. ``Go.''

The speaker crackled again. ``Your request is granted. Please proceed to runway three as marked
by the flashing orange lights.''

The Visitors did not reply but their aerocar, or whatever it was, raced toward the base with
incredible speed. There was no chance of the Argenian fighters keeping up with such a craft.
Chark suddenly felt very vulnerable. Dunari technology was obviously no match for that of the
Visitors.

Fotkey pointed toward the door and the group made their way outside into the furnace-like heat
of the deep day. They trotted toward the end of runway three, realizing that they had less time
than expected.

``Gods of \textit{Laronia}'' Noxan said. ``It's fast.''

All was strangely quiet. There was no roar of engines or chopping of propeller blades. The craft
approached in silence and without slowing down at all. Suddenly there was a bright green flash
and the craft abruptly, almost instantly, slowed to a more normal speed. Chark had no idea how
such incredible deceleration could be accomplished or how the occupants of the craft could
possibly survive it.

Only moments later a deep booming explosion washed over the base. It seemed to fill the air and
Chark could feel it in his bones. The ground shook. Everyone started, unsure what to do, and
Fotkey barked orders into his radio. Yet nothing more happened. The Visitor craft flew smoothly
toward the runway, going much more slowly as any aerocar might.

The craft landed easily, using only about half the runway to do so, and came to a halt not far
from the dunari delegation. It was a surprisingly small vehicle. It had short, swept back wings
and a stubby tail. It had a smooth body as if it was made in one piece, and it rested on landing
gear of some kind. The craft gleamed in the hot sun and sparkled as if it was coated with some
unusual material. Yet overall it looked amazingly normal. Kenellio was busy taking pictures.

The group waited nervously but nothing happened. Finally after a few moments Fotkey's radio
crackled again.

``\alien{Weee have communication devissse. Requessst tooo deliver.}''

Fotkey turned to Lucasa. Lucasa looked at Chark.

``It's some kind of radio,'' Kenellio said. Chark shrugged.

``It sounds like they can't come out of there,'' said Tulurkian.

Fotkey held the mic to his lips. ``Go ahead.''

A door of some kind opened underneath the small craft, and two trunk-like structures descended.
They were holding a large, flat rectangular object. Chark fought the urge to back away. Kenellio
took more pictures and Fotkey observed the process with his telescope. ``They're some kind of
mechanical trunks,'' he said.

The trunks retracted and the door closed, leaving the rectangular object behind, standing on the
ground in an upright position. The alien voice once again crackled from the speaker.

``\alien{Weee requessst depart.}''

Permission was granted and the small craft was soon racing down the runway. It lifted off
smoothly and turned more tightly than any dunari aerocar ever could. It receded into the
distance and then, suddenly, there was another green flash followed by a powerful explosion. The
craft sliced upward at extreme speed, vanishing against the glare of the sky in only moments.

Everyone turned to look at the runway where the rectangular object stood alone.

``That's it?'' Chark said.

Fotkey talked some more into his radio. ``Bring that thing into hanger three and start the
cooling units in there.'' He turned to the group. ``It looks like we have a present.''

When the delegation arrived at hanger three the ``present'' was already there. Several armed
guards were there as well. Before them was a large, thin slab standing upright on a narrow base.
The slab was taller than even the Forbinites and maybe twice as long as a typical dunari. It
appeared to be nothing more than a solid sheet of dark, unreflective material, like a wall.

Everyone was reluctant to approach the slab but Tusk had less fear. He studied the object and
the various symbols that were faintly printed around the edge, but the meaning of them escaped
him. Chark and Tulurkian stepped forward also to inspect the device. Chark reached out to touch
the object. The surface was terrifically smooth, almost like glass, and yet felt cool to the
touch. There was no reflection at all, and there were no controls to be seen.

``I don't think it's a radio,'' said Chark. ``I think it's a viewer. I'm almost sure this smooth
side is some kind of screen.''

``You'd think they'd tell us how to turn it on,'' Tusk said as he walked around behind the
device. ``Maybe we're just too stupid to understand it.''

Tusk's timing couldn't have been better for the device chose that moment to act. There was a
collective gasp from the group and everyone instinctively stepped back. All gazed at the dark
sheet as it began to come to life. In the center some strange, alien symbols briefly appeared.
Tusk tried to study them but he wasn't able to get a proper look. The guards stood ready with
their trunks on their weapons as more symbols appeared in the upper left corner of the screen
and vanished. Then, suddenly, the gray dissolved into a view\ldots\ a view of a room.

It was a room like any other, remarkable only in how ordinary it was. The room was small, with
no windows, and had an attractively tiled floor. The walls were plain with a kind of spiral
pattern along their base. The room looked like the sort of room Chark had seen a thousand times.
It vaguely resembled the conference rooms at Electrodyne.

The device almost seemed to have vanished. It was as if they were looking through a large, open
door or perhaps a window without any panes. The thin, black frame that once surrounded the dull
sheet was all that marked the edge of their world from the other. Again Chark reached out his
trunk.

``Captain, be careful!''

As he touched the device he could feel the surface. It was still there. It was still smooth and
cool. Tentatively he looked around the frame of the device the way one might look around the
edge of a door jam. Just as with a real doorway he could see past that edge to parts of the
other room at either side.

``Amazing,'' Chark said.

At that moment they heard sounds coming from the room. The source was out of view but obviously
close. Yet as mundane as the room appeared, the sounds were strange and alien. They were a kind
of chittering mixed with short clicks and long hisses. Everyone stepped back. The guards did as
well.

Then, suddenly, \emph{it} stepped into view. It came from the side and stood in the middle of
the screen, motionless.

``Jurita\ldots'' someone said and then there was silence. All stood and stared. The creature was
taller than a dunari, yet it was also thin and stood on only two legs. Its front legs were
atrophied and hung at its sides. Its head was perched on top of its body, high above the rest of
its torso on a short stalk. Its eyes were small and forward facing. It did not have a trunk but
instead there was a small stub in the middle of its face where a trunk would normally be. Its
ears were tiny flaps at the side of its head. Its skin, if that's what it could be called,
appeared to be thin and without scales, and had a pale color that looked utterly alien.

Roughly the creature resembled a trunkless chortak, a tree climbing animal that lived in the
Gunon forests of Forbin. Yet unlike a chortak the creature was obviously wearing clothing of
some kind. It was covered with a light colored fabric that wrapped around its torso and back
legs.

It studied them as well. It moved its head without moving its body just as a chortak might, and
looked squarely at each of them.

``It's a two-way system,'' Chark said suddenly. In retrospect it wasn't the most inspired thing
to say as the first words to an alien civilization. Yet Chark couldn't help blurt them out when
he realized the true nature of the device.

``\alien{Yesss},'' the creature spoke slowly. ``\alien{Communication syssstem}.''

Nobody said a word.

``\alien{It let usss\ldots\ talk},'' the creature continued, ``\alien{without risssk of\ldots
  sicknesss}.''

There was a pause and then with a deep breath Lucasa stepped forward. ``I am Lucastanonia,'' he
said, and the creature turned to look at him. ``I am the director of the Argenian Office of
Alien Affairs. I bid you welcome.''

The creature took a small device from the folds of its clothing and using one of its front legs
raised it up to look at it. Its toes were remarkably long and flexible and it wrapped them
around the device to hold it the way a dunari would use tentacles. ``\alien{Your language\ldots\
  difficult. We learning. I sssorry for\ldots\ misss-takes}.'' It lowered the device again and
said, ``\alien{We are well met}.''

Lucasa went around the room and introduced everyone, Argenian and Forbinite alike. After each
Argenian introduction the creature replied, ``we are well met.'' In the case of the Forbinites
the creature replied in Forbin: ``\alien{Ick narla it forgonnen}.''

When the introductions were finished the delegation looked at the creature. It seemed to
understand what was expected. ``\alien{I am Michael},'' it said, ``\alien{You can call me
  Mike}.''
