% Local IspellPersDict: aspell.dunari.pws

\chapter{13}

Jael was scrubbing the floor of the hanaria when she heard a vehicle pull into the lane in front
of the house. She went to the window and gasped slightly, clicking her teeth with delight.
``Your father's here!'' she called out. She trotted to the door and slid it aside just as Tusk
was dragging his bag from the back of the military truck. When he turned and saw his wife he
spread his ears widely and shook his head in greeting.

``You're beautiful, \textit{tinka},'' he said coming up the walk. The two of them embraced
tightly, wrapping their trunks around each other. Jinna and Mart came to the doorway and Tusk
embraced them as well, affectionately stroking his children's shoulders and faces.

``You're home early,'' Jael said.

``I ended up flying out on a waste transport. It wasn't exactly luxury accommodations but I was
tired of Garnick.''

``What are the Visitors like?'' Jinna asked.

Tusk shook his ears. ``Can I at least go inside first?''

With Mart's help Tusk dragged his bag into the entryway while Jael went to call the seleckia
with news of his arrival. He stood for a while in the center room just breathing in the air and
smelling the familiar scents of home. Finally Jinna could be patient no longer. ``When you
called you told Mom they are like chortak.''

Tusk lay down on his favorite seat and his children lay down on the floor beside him. ``They
don't have a trunk but instead they use their toes for gripping, almost like tentacles. They
have very strange skin with no scales, and small, round eyes on the front of their faces.''

Mart shivered. ``Did you talk to them?''

``Yes I did. I even learned a little of their language.''

Tusk looked up and saw his wife standing in the doorway of the hanaria with her head tilted
slightly. ``They're friendly,'' he said. ``They're curious about us, and they're looking for a
place to build a telescope.''

``A telescope?'' Jinna asked.

``Yes. That's why they're here.''

Jael started to shake and Tusk got up and went to her, wrapping his trunk around her. She
reached up and held him tenderly in her tentacles. ``It's going to be fine,'' Tusk said softly.
``There's nothing to fear.''

Jinna rolled onto her knees. ``What's going to happen now?''

Tusk clicked his teeth slightly. ``I almost forgot!'' Reaching into his cloak he took out a
money slip and handed it to Jael. ``A little something for my efforts.''

She looked at the amount and flapped her ears. ``Oh Tusk. This is wonderful.''

``I am now the head of linguistics at the Office of Alien Affairs,'' Tusk announced.
``Unfortunately my new job is going to require a lot of travel. I have to leave for Varnok soon.
I'll be testifying before Parliament.''

Jael swayed her head. ``My, my\ldots''

``The whole world is going to know about this, \textit{tinka}. We're going to shout it from the
rooftops.''

Tusk only stayed in Pilkity for a short time. All too quickly he had to pack his bag and say
goodbye to his family again. He took a local train to Tappentos on the eastern edge of the Sea
of Organite. From there he transfered to the Vinden line that went to Varnok. Across the top of
his ticket was a bold warning about the Vinden Tunnel. The city had grown large and finding
space for new tracks on the surface had become difficult. The proposal to lay rails underground
had been controversial and occupied the news rags for many runion. Yet, like so many things,
after the deed was done everyone lost interest. Many avoided the Vinden line because of the
tunnel, but Tusk had always been attracted by the adventure and was looking forward to trying
it. He was in an adventurous mood.

He had been to Varnok before, of course, but not often and not lately. As his train raced into
the center of the city he marveled at the tall buildings and at the bustling energy of the
streets. Finally a gentle, recorded voice warned the passengers of the upcoming tunnel. As the
train sped under the Luskin Insurance Building, Tusk braced himself. Artificial lights came on
automatically in the cab moments before the train was enveloped in darkness. It was a small
piece of night in the heart of Argenia's jewel.

The woman in the seat next to Tusk's specifically avoided looking out the window but Tusk
couldn't help but watch the wall of blackness. There was no sense of motion aside from the noise
and rocking of the train. There was only a great, empty void. The window was the edge of the
world beyond which there was nothing. Yet in moments the emptiness ended as the train
decelerated into the open, well lite expanse of Varnok Station. Tusk could hear the woman next
to him breath a sigh of relief. They had made it through the void.

Tusk wrestled his bag from the luggage drawer and made his way out of the train onto the
platform. Varnok Station was mostly underground with the light of the sun angling down from high
windows above the the first floor roadside shops. Tusk looked for signs to lead his way to the
street. The path was well marked, but the flow of people was daunting. It was nothing like
Pilkity. He didn't have time to really study the signs so he made his way forward as best he
could. Briefly he wondered if there were places like this on Earth.

When he reached the street level he was in the main terminal of Varnok Station. There on one
side was a long row of ticket counters. There were public phones, some places to eat, and
tourist information. One wall was all windows looking out onto the street.

He bought a map and then lay down on a seat for a moment to get oriented. The station was only a
few blocks from Parliament and from the Hoskin Hotel where he was staying. The Hoskin was one of
the finest hotels in the city and, due in part to its proximity to the Parliament buildings,
often hosted people from all corners of Argenia.

He decided to walk and, slinging his bag over his back, took to the streets. Again he had to
dodge many people but at least it didn't seem so oppressively crowded in the open air. The
street was in the shade and chilly but the ruddy light of the sun glanced off the third story
windows of the building across the way.

He walked by shops and stores of all kinds. A chew-shop caught his eye. He paused, but only
briefly. In previous times he would have felt compelled to buy a little something for later, but
the idea didn't enter his mind.

At the corner with the next following street, a trinket table was spread just behind a wall
protecting it from the Nermella. On the table were several stone and cloth figurines of space
aliens. Tusk paused to inspect them. He picked up one that looked like a dunari man with
exceptionally long legs, two trunks, an over sized head, and thin slit-like eyes.

``I can do better on those prices,'' a short, grizzled man said. He was lying in a seat behind
the table.

Tusk put the item back. ``Not interested.''

The Hoskin was an impressive building of stone and quartz-glass with broad steps leading up to
an elegantly adorned entry. At either side were imposing, intricately carved statues of Rujar
and Nergium, the two figures of ancient myth representing light and darkness.

The lobby was long and large, like a great hall with a polished marble floor and stone pillars
supporting the tall ceiling. On the floor were several finely woven harsna rugs, and here and
there were some large tables topped with inlaid rosenwood for the guests to use. A smartly
dressed man lay at one of the tables writing in a booklet.

Tusk approached the guest counter and was greeted warmly by a young woman with a beautiful trunk
sash carefully bound with a finely woven trunk cord.

``How may I help you?'' she asked.

``I'm Tuskara. I believe I have a reservation.''

The woman flipped through the pages of her book. ``Yes, indeed,'' she said. ``We have you right
here, sir. It's room 308. It has a beautiful view of the city and of the Parliamentary Yellow.
\textit{Ick narla it forgonnen}.''

``Thank you,'' Tusk replied with a bob of his head. ``\textit{Ick narla nat elan}.''

Tusk knew he would have ordinarily had trouble getting a room at a high class hotel because of
his Forbinite markings. Yet in Varnok that was not a problem. The Hoskin no doubt saw many
Forbinite dignitaries. He was impressed they even hired staff who knew the language, at least a
little.

Normally he was quick to clarify he was as much an Argenian citizen as the rest of them. Yet on
that day it didn't seem necessary. He felt proud of his heritage, and did nothing to disabuse
the woman of her assumptions.

She had Tusk sign some papers. ``It looks like the room has been paid for by the OAA. I believe
you're all set. Enjoy your stay.'' He was about to leave when she stopped him. ``Sir Tuskara? I
almost forgot. There is a message here for you as well.'' He took the envelop from her, noticing
the emblem of the OAA in the corner.

The room, of course, was immaculate. Tusk opened the curtains and looked out. The view from the
third floor was impressive. Only a block or two away, sprawled out before him, were the
Parliament buildings. They were arranged in a semi-circle with a large expanse of grass between
them. Walkways lined with shrubs and flowers criss-crossed the Yellow, and at the center was a
fountain. He peered around the curtain at the sun, hanging above the city. There was a haze in
the air, typical of tropical climates, and he saw some clouds on the horizon. It was so
different than Garnick, he could hardly believe he was on the same planet.

The message was from Lucasa. ``I hope you're hungry,'' it said. ``Can you meet Chark and me for
dinner at Frankens? Hour 69:100, Garnick time.'' He looked for a clock and found one in a
drawer. He had a little time to kill.

He turned on the viewer and flipped through the channels, stopping when he came to \textit{News
  in Depth}. There, much to his surprise, was Lucasa being interviewed by Jarloz. Tusk turned up
the sound and stood watching the program.

``So these\ldots\ \textit{humans} as they call themselves,'' Jarloz was saying, ``want to build
what, exactly?''

``They want to build an astronomical research facility on Nermia,'' Lucasa replied. ``Actually
in orbit around Nermia.''

``Why Nermia?'' Jarloz asked. ``Surely Rujar is much more hospitable.''

Lucasa shifted a bit the way he did when he was about to launch into a long explanation. ``They
said it's too expensive for them to transport building materials from their home 5000 light
runion away\ldots''

``I can imagine,'' said Jarloz.

Tusk snorted slightly in disgust. He wondered if Mike was also watching this.

``So they mine the raw materials and fabricate what they need on site. Apparently Nermia has the
necessary resources in an easily accessible place.''

``I'm sure it's easy for them,'' Jarloz said. She flopped her perfectly manicured ears and then
said, ``We understand they spoke to you in Argenian. How\ldots''

There was a knock at Tusk's door, and he turned off the viewer. It was the greeter with a tray
of fresh gingin. ``Compliments of the hotel, Sir Tuskara.''

Frankens turned out to be six blocks south of the Hoskin at the edge of the Varsynthia River.
Tusk walked again, happy to get more exercise. When he arrived, Lucasa and Chark were already
there. He was lead to their partition and when they saw him they bleated softly to greet him.

Tusk touched the shoulders of both men. During the time at Garnick they had become close. The
world was changing, and they knew they were at the focus of it. They lay down again as the
waiter gave Tusk a menu.

``I saw you on television with Jarloz,'' Tusk said to Lucasa. ``Is she as annoying in real
life?''

Lucasa shook his ears slightly. ``She's actually very smart, but she has to jazz it up for the
show. News is mostly about entertainment these days.''

``It's just a little embarrassing to hear her inane comments knowing the humans are probably
watching.''

Lucasa snorted. ``If inane comments worry you, you'll have a lot more to worry about when you
testify in front of Parliament.''

``About that,'' said Chark, ``exactly what is going to happen?''

Being the director of a controversial government office, Lucasa was no stranger to Parliamentary
politics. He had gone before various committees and subcommittees numerous times to justify the
work of his office and to account for his minuscule budget. Now they were going to testify
before the whole Parliament to brief them, in detail, on just who the Visitors are and what they
want.

``They've heard the basics already,'' Lucasa explained, ``but they want to hear it from us and
they want to ask questions.''

The men ate their dinner and talked for a long time. They talked about Parliament's likely
reaction to their testimony. They talked about the people's likely reaction to the humans. They
talked of a bright future where humans and dunari and Argenians and Forbinites all worked
together to make the world a better place. The food was excellent, the chew was delightful, and
the mood was light and optimistic.

``I understand you live around here,'' Chark said to Lucasa.

``I live in the West City,'' Lucasa said. ``My wife teaches at the University.''

``At Parnon?'' Tusk asked, ``isn't that a long commute?''

``There's an express train. It only takes a few arnets to get out there.''

When they finally left Frankens the sky was cloudy and the Nermella was blowing strongly. The
men instinctively huddled together to protect themselves from the sand and grit. It was a
response evolved into their bodies millions of runion ago when their distant ancestors roamed
the open plains.

``I have a car right over there,'' Lucasa said pointing his trunk. ``Why don't I drive you two
back to the hotel. Otherwise you might get rained on.''
