% Local IspellPersDict: aspell.dunari.pws

\chapter{17}

At the front of the Parliamentary chamber a tall rostrum rose as a dominating edifice built out
from the wall. On top of that rostrum the Speaker of the People normally stood beside a carved
stone table where he presided over the session. Below and to the left was a seat for the First
Citizen and to the right was a seat for the Second Citizen. Both seats were usually empty since
Parliament conducted its business independently of the executive government.

In front of the rostrum was a polished tonstone floor spreading out in all directions. Its dark,
patterned surface evocative of the shifting sands of the Argenian deserts. Slanting upward and
away was row upon row of seats for the Delegates of the People. The light of the Rujaran sun
filtered through windows placed high along the back and spilled onto the floor where history was
routinely made.

Lucasa, Fotkey, Chark, and Tusk sat behind a long table set out not far from the First Citizen's
seat, microphones ready, waiting for the session to begin. Delegates gradually filled the
chamber, standing in the aisles talking with each other, or lying at their desks reviewing
notes.

When Tusk was a child he once went on a tour of Parliament with his parents. At that age he
didn't appreciate the ornate carvings on the pillars or the tapestries on the walls. It was the
same room where Zendengiat demanded the dismantling of the Regents over 10 arnoxes ago. It was
the same room where Nilokiliox inspired a despairing nation to rebuild after the Endless War.

Tusk read the inscription carved into the top molding, just below the high ceiling above the
rostrum. It was written in Elocum, an ancient language that was reserved for the most formal
proceedings. Loosely translated the inscription read, ``Here speaks the Voice of Argenia, to the
benefit of all.'' Tusk snorted slightly. All unless you're a Forbinite, he thought to himself.

After a short time the Speaker of the People ascended the rostrum and took his position beside
the table. He carefully placed on its stone surface a heavy gavel. It had been many arnoxes
since it had last been used, yet its presence there beside the Speaker was deeply symbolic. He
stood tall and tilted his head slightly as the delegates took their places by their seats and
stood. Lucasa and Fotkey also rose. Chark and Tusk followed suit. There was no microphone on the
rostrum. The acoustics of the chamber were good, and the voice of Argenia could be loud.

``I convene this special session, on this date of Hour 70:050, Runion 90, Arnoxium 22 since
confederation, Argenian Master Time. I invite the First Citizen of Argenia to attend. Do you
agree?''

The delegates stomped their feet in a single blow.

``We are in agreement,'' the Speaker said.

At that moment a small door at the side of the chamber opened. The First Citizen walked
confidently across the floor and took his seat. He was wearing the cloak of office, yellow
trimmed with gold. As he lay down everyone else in the chamber, except the Speaker, did so as
well.

% TODO: Describe the First Citizen more (what he looks like, not just what he's wearing).

``We will now hear of matters concerning the \textit{human} situation. The testimony will begin
with that of Lucastanonia, Director of the Office of Alien Affairs. Do you have a statement?''

``Yes, Sir Speaker,'' Lucasa said.

Lucasa then recounted in detail all the specifics of the human situation. Tusk knew most of it.
He talked about the first pictures taken of the human ship. He talked about Chark's attempt to
communicate with the Visitors, and about the eventual response. He also talked about how the
First Citizen was contacted by the Forbin Prefect. That's when the delegates started to ask
questions.

``How did the Forbinites know of our communication?''

``The Forbinites were contacted directly by the humans.'' Lucasa replied. ``It was not us who
told them of the meeting at Garnick.''

``And so the Forbinites invited themselves to our secure military base?''

``No, they were invited by the First Citizen on my recommendation. The humans were unaware of
the true nature of Garnick. They suggested it because they thought it was neutral ground.''

Lucasa continued by describing the details of the meeting itself. He focused on how the humans
were interested in getting to know the Argenian and Forbin languages better. This raised some
concerns as well.

``Why would they want to know both languages?'' one delegate asked. Tusk fluttered softly to
himself. Fortunately the microphone in front of him was off.

``The humans understand we are two nations,'' Lucasa said. ``They're trying to avoid taking
sides in our conflicts and aim to interact with both Argenia and Forbin equally.''

``So you're saying,'' the delegate continued, ``that any technology they give to us they will
also give to Forbin.''

Lucasa lowered his head briefly and then turned back to the microphone. ``There was no talk of
technology transfer. The focus was on learning how to communicate better and on their explaining
to us why they're here.''

Lucasa then went on to talk about the observatory orbiting Nermia the humans wanted to build.
``Their ship is called \textit{Summer Breeze}. We were told it is a scientific vessel with a
crew of only 21. Its purpose was to follow up on an earlier automated survey looking for a
suitable location for their observatory.''

``So they've visited Rujar before?'' asked a delegate.

``Only an automated survey from a distance. The humans were clear that they didn't realize we
were here until \textit{Summer Breeze} arrived.''

A delegate in the back of the hall stood up suddenly. Such locations were reserved for the
juniors, typically just elected by their districts. ``I request right of first question,'' he
called out loudly.

The Speaker turned toward Lucasa. ``Do you acknowledge right of first question?''

Lucasa bobbed his head.

``I'm\ldots\ I'm a bit flabbergasted by the questions so far; no disrespect intended but they
seem so provincial. Lucastanonia, did you ask how many other civilizations are out there? Are we
about to become part of some kind of interstellar federation?''

``We asked, yes. It turns out there is no 'interstellar federation' as you said. In fact, we are
the first civilization the humans have encountered. Furthermore they described Rujar as very
remote for them.''

The young delegate seemed satisfied so Lucasa continued. ``The humans are interested in
establishing diplomatic relations with both Argenia and Forbin. They're sending another ship,
the \textit{Golden Light}, with two ambassadors on board, one to each of our countries.'' Lucasa
consulted his notes. ``The ambassador to Argenia is named\ldots\ let me see if I can get this
right\ldots\ \alien{Tomar'all}.'' Lucasa turned to Tusk, ``Is that right?''

Tusk turned on his microphone and leaned forward. ``\alien{Thomas Marshall}''

``When they get here they want to talk, officially, about the observatory,'' Lucasa said. Then
he leaned forward and spoke directly and deliberately into the microphone. His voice louder and
more distinct than usual. ``They have concerns about the suitability of our solar system for
their project. They don't want to invade our space or embroil themselves in our political
turmoil.''

``What turmoil?'' one delegate shouted out indignantly. ``We've been at peace with Forbin for
many runion now.'' Other delegates voiced similar feelings and the chamber was astir with
activity.

``Maybe we don't want them here,'' another delegate cried out. ``I, for one, am not comfortable
with the idea of an alien race, with who-knows-what motivation and who-knows-what abilities
creating some sort of machine on another planet in our solar system. It's a location that just
happens to be out of our reach and out of our sight. Who knows what its \emph{real} purpose
might be?''

Many other delegates shouted out their support with cries of ``Yea be yea.''

``What are you talking about?'' said another delegate loudly. ``Think of what we could stand to
gain from these humans.'' He pointed his trunk upward to emphasize his point. ``Their technology
is obviously considerable. With their help we could potentially cure diseases, improve our
transportation and communication systems, and solve our energy problems. This is not a time to
be fearful. Where is the Argenian boldness we so love to speak of in this chamber?''

``Yea be yea!''

Several other delegates shouted their objections and then, suddenly, the Speaker stomped his
foot and let out a long bellowing bleat. It was a disconcertingly primitive sound in the great
hall of Argenian government.

``Gentlemen,'' he said finally. ``We have time to debate this matter later. Right now, let me
remind you, we are here for the testimony of these experts and to ask \emph{clarifying}
questions only.''

Tusk watched the First Citizen carefully. As was traditional he did not say a word since,
technically, he was outside the scope of his jurisdiction and only present by invitation. Yet it
was clear that he was being very attentive and listening closely to all that was said.

After the clamor abated Lucasa continued. ``Another point the humans made clear was that they
regard this entire system as our space and they won't build anything on Nermia without
permission from both Argenia \emph{and} Forbin. They understand this is not something we've had
to think about before, but when the ambassadors arrive they will be looking to negotiate with us
about this.''

``We have no jurisdiction over Nermia,'' said one delegate, ``and certainly neither does
Forbin. What they ask is absurd.''

``What if we refuse to answer?'' said a delegate from the back.

``I think,'' replied Lucasa, ``they would simply build their observatory elsewhere. What impact
that might have on future relations with them is not clear.''

``Maybe we should assert jurisdiction over Nermia,'' another delegate said. ``We could talk
with Forbin and sketch out a rough understanding.''

``The Forbinites would never participate in such talks. The idea is ludicrous.''

The Speaker again bellowed. ``Clarifying questions \emph{only}, gentlemen.'' Once again after
the delegates settled down he indicated to Lucasa that he could continue.

``That's all I have, Sir Speaker,''

Next it was Fotkey's turn. The Speaker introduced him as ``General Fotnarkey, Supreme Commander
of the Argenian Armed Forces.''

``Do you have a statement?''

``Yes, Sir Speaker, I do,'' Fotkey replied.

As expected, Fotkey gave his military analysis of the humans. Tusk found it distasteful that
they would be evaluating the humans in terms of threats and countermeasures but he supposed
that, at some level at least, it was necessary. During his stay at Garnick he had spent a lot of
time, more then anyone else, talking with Mike. They mostly talked about nouns, verb tenses, and
sentence phrase structure. Still, after working with Mike so much Tusk felt a certain affinity
for him.

``Although we only saw evidence of a single aerocar-like device,'' Fotkey was saying, ``it was
clear that their technology is far superior to ours. That said, my perception is that the ship
currently in orbit around Rujar is too small to be much of a threat by itself. However, they
admitted to sending another ship and it is unknown how many others might be on the way.''

``What do you know about this other ship?'' one delegate asked.

``The humans claim it is bringing equipment for their construction project, but we have no
intelligence to substantiate that claim.''

Fotkey paused and then continued. ``I would advise preparedness. We have no independent
confirmation of their intentions, and so it would be wise to expect the worst. Our defenses
might not be sufficient, but then again we don't really know their capabilities.''

Tusk checked his watch and looked over at Chark. He was fidgeting with some papers in front of
him. Tusk thought that he probably should have at least written down some notes about what he
was going to say.

``It's important to note,'' Fotkey continued, ``that so far the humans have done nothing
aggressive. I regard that as a hopeful sign. Nevertheless I think it would be unsafe to assume
their non-aggressive behavior will continue indefinitely.''

Fotkey then gave a summary of the non-classified standing of the Argenian armed forces. His
conclusion was that they were ready for battle should that prove necessary.

Next it was Chark's turn to testify. He mostly focused on the technical aspects of their
communication. He briefly described the device the humans used at Garnick.

``Why are the humans afraid of meeting face to face?'' asked a delegate. ``Is our air
unbreathable by them?''

Chark explained that the humans were concerned about biological contamination. ``They said that
Rujaran life and Earth life might be `incompatible' making us immune to each other's diseases.
However, they would need to examine some samples of Rujaran life before making that
determination.''

``What kind of `samples?' ''

``Bacteria, fungus, that sort of thing.''

Chark went on to explain that the battery in the communications device used at Garnick had run
low, and the humans retrieved the device to recharge it. ``We agreed to communicate via
conventional radio in the meantime, and I've already applied to the CB for an official frequency
allocation.''

``Are you saying,'' asked a delegate, ``that we could get a message to the humans at any time?''

``My people are setting up the necessary equipment as we speak. In fact, we intend to establish
a permanent communication station at Lungast.''

Chark then explained that the humans wanted to deliver another communications device by the time
the ambassadors arrived. ``We gave them the specifications of our standard power outlets. They
intend to build a converter circuit of some kind. We will literally be able to plug their
equipment into our outlets.''

This created some clamor in the chamber. ``Was that wise, Captain Charkonaless?'' one delegate
asked. ``I don't approve of you revealing information about our systems without authorization.''

Chark leaned back and glanced briefly at Lucasa.

``Let me reassure you,'' Chark said, ``there is no chance of them being able to attack our power
system through a standard outlet. Our systems are heavily fused and filtered.''

``Who knows what they can do?''

There was more shouting from the delegates and several of them tried talking at once. Again the
Speaker bellowed.

``Sir Speaker,'' said one delegate finally. ``I submit that without specific knowledge it is
easy for our imaginations to run wild. Yet we can't let ourselves become afraid of shadows. I
for one accept our expert's testimony. Please continue Captain Charkonaless.''

Chark finished up with a few general comments about human electronics technology as he saw it.
Then he closed his testimony and it was Tusk's turn. Tusk took a deep breath and leaned forward
toward his microphone. He heart was pounding.

The Speaker introduced him. ``Now we will hear from Tuskara, Head Linguist at the Office of
Alien Affairs. Do you have a statement?''

``Um\ldots'' Tusk said. ``Yes, I do.''

Tusk began, tentatively, to describe his experience talking with the humans. He felt uncertain
at first, but as he talked he grew more confident. ``Their language is complex. It is also very
difficult for us to speak, but that is mostly because they use consonant sounds that are hard
for us to make with our vocal equipment. Yet it seems to work both ways; they have difficulty
speaking our language as well.''

Nobody seemed interested in the linguistic details of English versus Argenian so Tusk skipped to
the more pragmatic aspects of their communication. ``Their technology makes it easier for us
both. They have an electronic speech synthesis system that can translate speech on the fly. I
spent a lot of my time working with one human in particular to help him tune the device to
Argenian and Forbin.''

``You assisted the humans with Forbinite communication?'' one delegate asked. He was obviously
quite shocked and Tusk felt suddenly nervous. He didn't want to land in prison as a traitor due
to his testimony before Parliament.

``Ah\ldots'' Tusk hesitated. ``The humans are intent on learning both languages. If it wasn't me
it would have been one of the Forbinite representatives. I happen to know both languages and
the humans were comfortable working with me.''

Nobody seemed to object to Tusk's explanation so he continued. ``I was also able to learn a
little of their language. They did nothing to prevent that and, in fact, encouraged it.''

``Can you understand them well enough to following conversations between them?'' asked a
delegate.

``Only partially,'' Tusk said. ``I hope to continue learning. I'm planning to return to Lungast
with Chark\ldots\ ah\ldots\ Captain Charkonaless and continue working with the humans over the
communications link he's establishing there.''

The delegates asked a few additional questions about how communication with the humans would
work, but soon Tusk's testimony was over. He switched off his microphone and leaned back onto
his seat with great relief.

The Speaker then stomped his foot twice. ``Now that testimony is complete, the leaders of the
three associations will each make a preliminary statement.'' He tilted his head slightly to look
at an elder statesman lying in the front row. ``Hap?''

Haptair was the most senior member of Parliament. It was said by some that he carried more
influence than even the Second Citizen. With slow, painstaking movements he rose from his seat
and shuffled out onto the floor, turning to face the hall.

``It is\ldots'' he coughed and cleared his throat and then tried again with a voice that was
surprisingly full and vibrant. ``It is the inclination of the Narosta Association to be
optimistic in the face of the unknown. We agree that this is a time of great opportunity;
opportunity that we must not squander out of fear. Yet it is also natural to be cautious in a
situation such as this. Thus we will advocate a carefully measured approach that will move
Argenia forward boldly and yet also safely.''

After Haptair lay back down, the Speaker nodded to another delegate just two seats away. The
delegate, considerably younger, stood up and walked confidently onto the floor.

``The wisdom of my esteemed colleague can not be doubted. Yet his words offer little in the way
of concrete counsel. It is likely that we are at the mercy of these humans, and that all we say
or do here will be for naught. Yet if they do have any sense of decency and are willing to
negotiate with us honestly, then it is the position of the Polictaria Association that we
encourage them to build their\ldots\ whatever it is\ldots\ somewhere else. We can't debate
territory on another planet. What is its ultimate value to us? What reasonable payment could we
demand? If the humans renege, to what court do we appeal for restitution?''

Finally the Speaker nodded to Zarlonloti, impeccably dressed as always, lying at the end of the
row near the wall. He rose and walked onto the floor with a casual, unhurried pace.

``The Limzar Association finds it strange that we are discussing the humans as if they are
dunari. These creatures are aliens from beyond the world. They have never known the beauty of
Rujar or any of the Seven's other great works. What makes us think they share our concerns? What
makes us think we can negotiate with them at all? We are insects to them. We are something to be
ignored when we aren't in their way and brushed aside when we are. The Limzar supports the
position of the Polictaria Association; the humans should not be allowed in our midst. I pray to
Jurita that we even have that choice.''

With that the Speaker closed the session, bellowing the dismissal, taking the gavel, and
descending from the rostrum. The First Citizen exited through the same door from which he
entered as the delegates immediately started to mull about, talking again among themselves.

Tusk stood up and stretched. He put his trunk over Chark's shoulder. ``It's going to be a longer
road than I thought.''

\spacebreak

The First Citizen's offices were located in a relatively unassuming two story building across
the street from Parliament. The grounds were beautifully groomed and covered with thick, yellow
grass, studded with red flowers. The building itself was one of the first government buildings
in Varnok, constructed in simpler times before viewers, aerocars, and humans.

General Fotkey trotted up the steps dressed in full uniform and showed his identification to the
security guard on duty. The guard knew Fotkey, of course, but checked the paperwork carefully
anyway. The guard then slid aside the broad door and Fotkey stepped into the main lobby of the
building. On the floor was a mosaic of the seal of Argenia, permanently illuminated by the
sunlight shining through tall side windows. Fotkey walked across the pool of light and was
warmly greeted by a receptionist at the counter on the far wall.

``We are well met General,'' he said. ``Go on up. He's expecting you.''

Fotkey climbed the curved ramp to an upper floor and then down a wide hall to another security
guard. As he approached the guard turned and knocked on a dark narwood door. He opened it a
crack and stuck his head just inside for a moment. Then the guard stood back and let Fotkey
enter.

``General!'' The First Citizen was standing beside his table. He stepped forward, meeting
Fotkey, and the two men clasped trunks briefly. Fotkey touched the First Citizen on the
shoulder, and then he heard the door click behind him.

% TODO: More description of the First Citizen?

The First Citizen motioned to Fotkey to lay down and they both positioned themselves on adjacent
seats in front of the First Table. ``There is quite a lot to discuss,'' the First Citizen said.
``To start, give me an update on the special weapons program.''

Fotkey shifted uneasily and glanced briefly toward the closed door where he had come in. ``We
have eight devices complete, and four more are in various stages of production.''

``How is security?''

``The Forbinites seem to be buying our cover story that we're using Rangard for deep day tank
experiments.''

The First Citizen slowly stroked the side of his face with one tentacle. ``And now we have a new
consideration.''

``The humans.''

``It's important they don't find out about the special weapons.''

``I agree,'' Fotkey said, ``they may be our only defense against them.''

The First Citizen tilted his head and fluttered softly. ``It is my fervent hope that we never
have to use them in any case. Frankly, I'm more concerned about the humans telling the
Forbinites.''

Fotkey bobbed his head briefly.

``Since we know very little about the surveillance capabilities of the humans, I want to put
production on hold for a time. Finish the devices in progress but don't start constructing any
new devices.''

Fotkey hesitated. ``Are you sure that's wise, sir, considering what we know about the Forbinite
program?''

``It's a different world now, General,'' the First Citizen interrupted. ``We need to understand
this new world before we lunge ahead with old plans.''

``Yes sir,'' Fotkey said reluctantly.

The First Citizen smoothed his cloak with his trunk the way he did when he was about to change
the subject. ``What is your personal assessment of the humans? Are they sincere about this
telescope project they're talking about?''

Fotkey took a deep breath. The question was almost impossible to answer. He looked down and
thought for a moment. Finally he looked up. ``I think they are sincere.''

The First Citizen swayed his head. ``That's my impression also. I'm going to grant them
permission to build on Nermia and invite the Forbin Prefect to do the same.''

``Yes sir.''

``The Prefect won't be able to refuse. He'll be much too worried about us making friends with
these humans.''

``Yes,'' Fotkey said, ``a technology transfer from them would give us a strategic advantage.''

The First Citizen clicked his teeth slightly. ``You are a military man, General, and you are
good at what you do, but sometimes you don't think big enough.''
