% Local IspellPersDict: aspell.dunari.pws

\chapter{19}

When Jernumia arrived at the office everyone was strangely subdued. She ignored it and instead
unpacked her pouch as usual, putting each piece of paper in its appropriate place. Lying down
behind her desk she opened her case book and went over the checklist. There was still the matter
of precedent finding for her current case. After reviewing the details again she packed up a few
papers and made her way toward the door.

``I'll be in the law library,'' she told the receptionist.

``Not now,'' Haratol said overhearing her, ``the First Citizen is about to speak on
the viewer.''

``Yes, yes,'' Jernumia replied, ``I'm sure that's very important but I have \emph{your} case to
prepare.''

``You don't need to work so fanatically,'' Haratol said, ``you have nothing to prove here.''

``I only want to prove the defendant guilty,'' she said curtly, ``if he is,'' she added.

She pushed by Haratol, her pouch looped over her back and headed for the door. As she went into
the hallway she passed by the small huddle room where people often relaxed and socialized. She
almost never went inside, but today there was a group of her colleagues gathered around the
viewer. On the screen was a picture of the Parliamentary Yellow, now covered by a large crowd.

Jernumia hesitated.

``Take a moment to listen,'' Haratol said as he came up behind her, ``his speeches are never
long.''

Jernumia glanced at the wall clock and decided she could afford a few moments. It was, after
all, an historic event. The others, all standing in the room, made space for her as she entered.
Jarloz was talking on the viewer.

``This will be the first public appearance of the First Citizen since the arrival of the humans.
Kopta, what do you think he will be covering today?'' The camera changed to a shot of the studio
where Jarloz had a guest Jernumia didn't recognize.

``I think the people of Argenia are hungry for leadership and for reassurance. We don't know how
to act or how to feel. People want the First Citizen to tell them.''

``This First Citizen has been criticized, especially by the Limzar Association, for having a
lack of vision,'' Jarloz continued. ``Do you think that criticism will hurt his credibility
now?''

``I don't think so. In exceptional times people often forget\ldots''

Jarloz raised her trunk slightly and appeared to be listening intently to the ear piece she had
hooked over her head.

``I'm sorry Kopta, but the First Citizen is about to start. We will come back at the end of the
speech.''

The scene changed again. This time it focused on a podium erected at the sunlit side of the
Yellow. There, with banners of Argenia at either side, a single man stood ready. He raised his
trunk and then spoke into the microphone.

``To all gathered here,'' he began, ``the First Citizen of Argenia!''

The man stepped aside and the First Citizen walked lightly up to the podium. There were loud
bleats across the Yellow and the First Citizen flopped his ears and waved his trunk.

``Always a politician,'' Haratol said at Jernumia's side.

Jernumia said nothing but instead watched as the First Citizen finally lowered his trunk and
moved to speak.

``Now we know that we are not alone in the universe,'' he began. ``This discovery has caused us
concern and, yes, even fear. That's natural. We fear what we do not understand and, lacking
understanding, assume the worst possible outcome. Yet these Visitors, these humans, appear to be
much like us. They are curious explorers seeking to deepen their knowledge of the universe. They
come here not to overthrow our society but instead to build an astronomical observatory.

``Is it true? How can we be sure? There is no way to answer rationally. Instead we must listen
to our feelings. I believe any civilization advanced enough to travel between the stars will
also be advanced enough to treat us with respect and consideration despite our backward nature.
The evidence is exactly this. While we fret about their intentions what the humans have actually
done is plain enough: they've worked at communicating with us so that they can \emph{negotiate}
with us to use our space. If they wanted to destroy us, I very much doubt we'd be having this
discussion.''

He leaned into the microphone slightly. ``I therefor implore you, the people of Argenia, and
your Parliament, as well as the Prefect of Forbin to grant the humans the permission they
seek\ldots''

Suddenly a shot rang out and the First Citizen stumbled back from the podium. Everyone in the
room gasped and Jernumia instinctively stepped forward as if to help. A second shot was fired
and the First Citizen fell to the ground. In moments the podium was swarming with IIB agents but
the First Citizen could not be seen.

``Someone has fired on the First Citizen,'' Jarloz said. The camera swung to the left where it
scanned over the side of the Hoskin Hotel. ``Someone has fired on the First Citizen of
Argenia!''

\spacebreak

Colty waited impatiently in the Fargon conference room. Fargon, the home of the mathematics
department, was on the Old Campus and one of the original Parnon University buildings. It was a
historic structure with a graceful elegance that the newer buildings on the South Campus lacked.
Even a simple conference room was decorated with intricate moldings and a beautifully polished
floor. Colty had always thought it fitting that the department of mathematics, representing one
of the oldest intellectual fields, should be in such a venerable building.

Dargonick and Lucastanonia came into the room interrupting Colty's reflections. Sar. Dargonick
was the leading professor of the mathematics department and well known to Colty, but Lucasa, the
Argenian ambassador to Earth, was someone he had only seen on the viewer.

Colty stood up while Dargon did the introductions. ``We are well met, Sir Lucastanonia,'' he
said as he touched the ambassador's shoulder respectfully.

The three men lay down around the table and Lucasa took some papers out of his pouch, clipping a
few to the available table clips. Dargon looked uneasy. ``We're expecting a fourth party,'' he
said, ``Sar. Rocallanon from the physics department should be here any moment.''

Lucasa fluttered slightly and leaned back but said nothing. Dargon tried to fill the time with
small talk. He asked Colty how he liked Parnon.

``I'm really enjoying it,'' Colty said politely. Then after a pause he added, ``What is this
meeting about, anyway?''

Lucasa leaned forward again, grateful for having a reason to speak. ``As you know, it has always
been the intention for some human scientists and engineers to come to Rujar to educate our
people.'' He unclipped one of the papers sitting on the table in front of him and slid it over
to Colty. ``It seems like that's going to happen sooner than we thought. Apparently one of their
scientists is en-route to Rujar even as we speak. This individual is an astrophysicist and would
like to observe the construction of their telescope.''

At that moment Rocallanon arrived. As the leading professor in the physics department he wore
the formal cloak of enlightenment, and on his chest the insignia of the Argenian Order of
Physical Scientists. ``I was delayed,'' he said as he entered the room.

Dargon did the introductions again and Rocalla clasped trunks briefly with Lucasa and Colty.
``Where are we?''

``I was just explaining to Sar. Coltinarly about the human coming to Parnon,'' Lucasa said.

``Oh, good,'' Rocalla said as he lay down. ``Please go on.''

``It's coming to Parnon?'' Colty asked. ``You mean to the human compound at the base?''

``More than that,'' Lucasa said. ``She teaches at the human equivalent of a university. She has
expressed interest in being a visiting professor here at Parnon while she waits for the
observatory to be completed. We are actually very lucky to have her. She is apparently one of
their most respected scientists.''

Colty was shocked. ``Are you saying it's going to be wandering the halls and teaching classes?''

``She.''

``What?''

``\emph{She} will be wandering the halls and teaching classes. Yes, that's the idea.''

``Like a farg Forbinite,'' Rocalla said softly, ``with their females doing science.''

Colty struggled to absorb the unsettling idea of a human on the Parnon campus. It was one thing
for them to have a small compound on the military base just outside town. The proximity of
Parnon to Varnok made such a presence there understandable even if unpleasant. Yet one never saw
any humans in the city of Parnon itself. Even the special vans that occasionally transported
them to Varnok were a rare sight.

``What does this have to do with me?'' Colty asked.

It was Dargon who answered. ``Of course you know it's traditional for a visiting professor to
have a host who can help him find his way around and generally get oriented to the campus. This
human has specifically requested a host from the mathematics department.''

Colty folded his ears back against his head. ``Wait a moment\ldots\ are you suggesting that
\emph{I} be its host?''

``It doesn't make sense,'' Rocalla said. ``It\ldots\ she\ldots\ is a physicist so it only seems
natural that someone from the physics department be her host.''

Dargon looked at Lucasa.

``She was clear about her desires,'' Lucasa said. ``We received a message directly from her
about it.''

``Perhaps the humans don't understand how our academic disciplines are organized,'' Rocalla
said. He turned to Colty. ``I don't doubt your work is important in your field, but you're a
pure mathematician and your research is just not relevant to physics. In contrast, I have many
publications in the area of stellar dynamics. I'm sure the human and I would have much more to
talk about.''

Dargon looked uncomfortable. He curled his trunk slightly. ``Of course the role of host is not
intended to be primarily academic.''

``I don't like the idea either,'' Colty said. ``I'm a new instructor. I need time to develop my
full status presentation. I also don't know the university that well myself.''

Lucasa slowly twirled a pencil in his tentacles. He looked over at Dargon.

``I actually recommended you, Colty,'' Dargon said. ``I thought that perhaps since you are
younger you might be more open to the adventure. Think of what it might do for your career.''

``I don't see how it will help him,'' Rocalla said. ``He's in the wrong field.''

Lucasa fluttered. ``There is also the matter of her office. We'll need to do some renovations to
make a suitable space for her. Renovating space in Fargon is quite impossible.''

``It's going to have an office too?'' Colty asked.

Dargon disregarded Colty's rhetorical question. ``What's wrong with 105? That's a free office.''

Lucasa swayed his head. ``The problem is that we have to add an air cooling unit and toilet
facilities. Fargon is a small, historic building so making those kinds of changes is very
difficult. Cogart is much easier to change. We were thinking of using the mini-conference room
there, 227.''

``That's just down the hall from my office,'' Rocalla said with satisfaction.

``Wait,'' Colty said, ``what's this about toilet facilities?''

``The humans aren't that different from us,'' Lucasa explained. ``They need such things just as
we do.''

``What about the compound?''

Lucasa shrugged. ``If she spends significant time on campus, as expected, she'll want something
a little closer. That part will actually be easy for us. The humans will install it. They seem
intent on recycling all organics. We don't even need to bother with waste ducts.''

``I submit,'' Rocalla interrupted, ``that the humans are unfamiliar with the way we do things
and that once the situation is explained to them they will agree that someone from the physics
department would make a better host.''

Lucasa twirled his pencil. He glanced at Dargon and then Colty. ``I can get a message to her,''
he said at last, ``and maybe ask for clarification.''

``I think you should do that,'' said Rocalla.

``Sar. Coltinarly?''

``Yes, certainly.''

``Okay,'' said Lucasa. ``It will take some time. She's about 3000 light runion away right now.
In the meantime it makes sense to go forward with the renovations of Cogart 227 in any case.''

Satisfied with the outcome of the meeting, Rocalla moved to get up. However, Lucasa did not
budge and Rocalla hesitated.

``One last thing,'' Lucasa said. ``Her name is \alien{Rachel}.''

\spacebreak

After the meeting Colty returned to his office to continue his work. As a junior faculty without
full status his space was little more than a closet. His one window faced north and received no
direct sun. On his small table, amid a collection of books and papers, sat an artificial light
that he used on the gloomy, rainy times that frequented Parnon. He had an awkward, uncomfortable
seat for the occasional student, his wasn't much better, and a bookcase along the wall that made
his narrow space seem even narrower.

He lay down at his table and flipped through a paper on lars-extended co-sequences. He had read
it several times before but he kept coming back to the proof of Theorem Five. The technique used
was standard yet Colty's instincts told him there was something wrong with it. He pondered the
reasoning again but he had trouble concentrating. The meeting with Dargon and Lucasa had
rattled him.

With a fluttering growl he skipped to the end of the paper and scanned down the reference list.
The author seemed to rely heavily on one reference in particular that was in a Forbin journal.
Colty wondered if he could get a copy of it. He decided a walk would do him good so he packed
the paper into his pouch, locked his office, and headed out onto the Yellow.

Fargon was just one building that surrounded a broad, beautifully tended expanse of thick grass,
criss-crossed by walkways, dotted with gardens, and lined by large rigat trees. The sound of
their tubular leaves singing softly soothed Colty's agitated mind.

He walked briskly along the wide, brick path that lead from the spire of Tartuk Hall on one end
of the Yellow to the library on the other. The library was an impressive stone building with a
wide, arched entrance below tall sunward windows. Just outside the door several students stood
huddled around a table offering a petition. Tied to the legs of the table, and flapping in the
Nermella, was a sign with a quote from the Limzar Association's position statement: ``The humans
haven't done any harm, nor have they done any good.''

Colty ignored the provocative sign and slid the library door aside. He walked across the
polished tonstone floor in the large inner court beneath the high, domed ceiling. All was quiet
and serene; an oasis of thought in a troubled world. He came to the librarian's counter where an
attractive young woman was busily checking in books.

``Good singing, Joleia,'' Colty said.

``Sar. Coltinarly,'' she replied with a click of her teeth. ``What can I do for you today?''

He took the paper out of his pouch. ``I'm hoping to get an Argenian translation of this
article.''

She took the paper and looked at the citation closely with her right eye. ``This won't even be a
challenge,'' she said. ``The National Library probably has the journal and we can get one of our
staff translators to make the translation for you.''

``That's wonderful, thank you.''

``If you can wait, I'll call Varnok right now and let you know.''

She walked back to a table behind the counter and picked up the phone. Colty could hear
indistinct sounds from outside. There was some shouting and bleating. After a few moments Joleia
returned. ``It's on its way. I'll send you a note in campus mail when the translation is
ready.''

``Thank you so much.'' He paused and then motioned toward the main entrance with his trunk.
``What's going on out there anyway?''

Joleia fluttered disapprovingly. ``Freedom of speech,'' she said. ``Is there anything else that
you need?''

He hesitated. ``Actually I do have one other question. There's going to be a performance of the
Yarlantan thrumbatas at the concert hall at the end of the runion. I\ldots\ I was wondering if
you would be free to come with me to hear them.''

Joleia spread her ears slightly and clicked her teeth several times. ``Sar. Coltinarly, are you
asking me out?''

Colty clicked his teeth as well. ``It appears that I am.''

``In that case, the answer is yes!''

\spacebreak

The rest of the runion passed uneventfully. Colty divided his time between preparing for
classes, grading assignments, and trying to get some work done on his research. Yet progress was
painfully slow. His topic was stellating co-sequences, a natural extension of his dictatorium
work on mono-sequences. There was, however, scant literature on the subject. The most recent
paper was almost 100 runion old and largely treated the material as closed. Yet to Colty's eye
it seemed like there was something more.

He was glad for the distraction when the time for the concert finally arrived. In his apartment
he first changed into a professor's cloak but then thought better of it. Joleia was a
sophisticated woman, but not at all the snobby type. He didn't want to overplay it, so he
switched to casual garb and felt more comfortable in the process.

He waited for Joleia on the stone patio at the entrance to the concert hall. A cloth banner for
the Yarlantan thrumbatas was tied to a large pole in the center of the patio, flapping in the
Nermella. It was warm and he watched the people stroll by. Students, faculty, staff, and
citizens of Parnon all came to see the performance. He looked out over the street toward Old
Campus where he could see the spire of Tartuk Hall rising above the trees. Beyond that, outside
of town, was the military base and the enigmatic human compound.

``Lost in thought, Sar. Coltinarly?''

He turned to see Joleia standing near him. She was wearing a beautiful rubinum cloak trimmed
with orange and gold. On her head she wore a matching cap. Behind her but standing with her were
two other women about her age.

He clicked his teeth at her. ``Please\ldots\ `Colty' is fine.''

Joleia introduced the two women as seleckia and Colty gently touched each woman's shoulder as he
bobbed his head politely. They both seemed as pleasant as Joleia which was in Colty's mind both
encouraging and unsurprising.

The four of them entered the hall and were directed toward their seats on the west side. As was
traditional Joleia's seleckia lay beside her and opposite Colty. As they settled he looked
around. He had never been in the Parnon concert hall before. It was, of course, open to the sky
with walls to direct the Nermella away from the audience. The barriers were carefully designed
to let in the sun while at the same time reducing wind noise and trapping the sound from the
performers.

The seats were arranged in a circle around a central stage except for the east side. Seats there
would have been facing the sun and not very useful so instead the entrance ramp for the stage
was in that direction. On the stage were several thrumbatas waiting for the start of the
performance.

``You look very nice,'' Colty said.

``You do too. Thanks for inviting me.''

The two chatted casually for a time as the hall gradually filled with people. Colty found it
easy to talk to Joleia. Her quiet, relaxed manner was refreshing. As she spoke it was obvious
that she was observant and intelligent.

% TODO: What are her seleckia doing?

``Do you think the humans have music and concerts like this?'' Joleia asked.

``I have no idea,'' Colty said. He hesitated and then added, ``you know, I've been asked to host
a human visiting professor.''

Joleia tilted her head with surprise. ``Really? A human professor is coming to Parnon?''

``Yes. It will be here in a few runion.''

``That's wonderful! The department must think highly of you to give you the honor of hosting
it.''

Colty shifted uncomfortably. ``I haven't agreed to do it. I'm not sure I'll have time since I'm
working on full status.''

``I understand. Of course.'' Joleia looked pensive and Colty wondered what she was thinking.
``At least the humans are still interested in us despite the Junar attempting to assassinate the
First Citizen.''

``You think it's good for them to be interested?''

``Of course! Imagine the library they must have.''

Colty shook his ears, but Joleia continued more seriously. ``What must they think of us when we
have these religious fanatics trying to gun down our own leaders?''

``I guess that doesn't make a good impression.''

Joleia fluttered softly to herself. ``If I was them I would stay far away from the likes of us
until we've grown up a little more.''

After several arnets the two performers walked down the ramp and onto the stage. They turned
back to back and raised their trunks while the audience stomped their feet. The performers
waited a few moments for the applause to stop. Finally when the hall was completely quiet they
walked, without looking back at each other, to their instruments.

Lying on specially designed seats with one leg wrapped around the backs of their thrumbatas,
they raised their trunks to the strings. Glancing at each other they paused, nodded slightly,
and started to play.

The music flowed from the stage and filled the hall, mixing with the breeze and rising into the
open sky. Colty cocked his head slightly to let the complex melody wash over him without
distraction. The interplay of the notes formed a pattern in his mind. It had a beautiful
structure that constantly changed and flowed, like an elegant proof.

After a time he turned and looked at Joleia, but she was listening intently as well and appeared
to be lost in her own world. The performers were renowned players and justifiably so. When the
first kos ended, they paused briefly and immediately started the second kos. Yarlantan's music
never seemed to stop but always transitioned smoothly into a fresh kos, changing in style and
mood as it did. When played by the masters it was more than music. It was a story in sound.

When the concert was over the hall exploded with enthusiastic stomping. Joleia flapped her ears
with pleasure.

``That was lovely,'' Colty said loudly to be heard over the din.

The performers knelt as bleats of approval sounded from the crowd. When they rose they waved
their trunks and then exited the stage.

After the concert Colty returned to his apartment to relax and sleep a little. Things had gone
well with Joleia and he hoped they would get together again. After cleaning up he went through
his checklist to ensure that his space was all in order. Finally, he pulled down the mesh
screens on the windows to block the harsh light of any flashes, pausing for a moment to notice a
few fresh spots on the disk of the sun. He lay on his bed, his mind filled with thoughts of
co-sequences and thrumbatas, of music and mathematics, of Jingas and Joleia. Farth-sleep dulled
his reason until paren-sleep stopped his mind.

\spacebreak

Colty was in his office reviewing student papers when Lucasa kicked at the door frame. He was
surprised to see the ambassador.

``Lucastanonia,'' he said. ``I had no idea you were coming.''

``This is an unofficial visit. May I come in?''

Colty took the books off his spare seat so Lucasa could lay down and stacked them up on his
table instead. ``I'm sorry I don't have much space here,'' he explained. Despite his large size
Lucasa didn't seem concerned.

``Have you given any more thought to the idea of hosting our human guest?''

``I've tried not to,'' Colty admitted. ``I thought it was decided that Sar. Rocallanon would be
a more appropriate host.''

``It was decided that I would contact her, and I wanted to show you her reply.'' He took a
folder from his pouch, removed a single sheet of paper with the letterhead of the OAA, and
passed it to Colty. ``These are her words translated to Argenian, of course.''

Colty hesitated yet he had little choice but to take the paper and read it. The note was short.

\begin{quotation}
Ambassador Lucastanonia,

Thank you for describing the situation at Parnon University in more detail. It's true that I do
not know your culture nor anything about the university's policies and traditions. I suggested a
mathematician only because I consider myself to be a mathematician as well. I do not wish to
cause problems, and I apologize for any confusion I have created. I will leave it up to you and
the university's faculty to decide who should be my host.

Rachel
\end{quotation}

``She writes graciously, don't you think?'' Lucasa said.

Colty was struck by how ordinary the note seemed. It could have been written by a dunari
professor from any one of a hundred colleges or universities sprinkled over Argenia. He had to
remind himself that it was actually written by a member of an alien race who was currently
thousands of light runion away.

``She says she's a mathematician,'' Colty said. ``Didn't you say she was a physicist?''

Lucasa shrugged. ``I don't think they distinguish those fields the way we do.''

``So in other words, it doesn't really matter to her from which department her host comes.''

``It might not matter to her,'' Lucasa said, ``but it does to me. I'm hoping that you're
willing to do it.''

Colty looked over the note again, lingering on the phrase, ``\textit{I do not wish to cause
  problems}''

Lucasa clicked his teeth softly. ``You don't get into a position like mine without a certain
ability to read people. I like what I read from you, Sar. Coltinarly, a lot better than what I
read from Rocalla.''

Colty bobbed his head slightly. ``Okay, I'll\ldots\ I'll do it.''

``Good!''

``I just hope I don't regret it.''

``Don't worry,'' Lucasa said. ``You'll have the full support of my office. In fact, I want you
to meet my linguist Tuskara. He can brief you on what we know about the humans, including things
the press does not report.'' Lucasa shook his ears slightly. ``You'll like him. He has an
amazing way of making the humans seem like dunari.''
