% Local IspellPersDict: aspell.dunari.pws

\chapter{23}

Colty stood on the landing pavement of the Parnon base under a temporary tent. He was surprised
by how few people were there. Lucasa, Tusk, the university chancellor, and Rocalla were the only
dunari representatives, besides himself. The human ambassador was also present, standing off to
the side talking to Lucasa. Colty had never seen a human in real life before, and it was eerie
to watch it moving its otherworldly body as it talked. He felt like he should go and join the
conversation but he couldn't bring himself to do it.

The press was there as well but only a few local reporters, one of whom was from the student
newspaper. The humans were old news, and the arrival of one individual no longer captured
national attention. One of the reporters interviewed Colty briefly but Colty gave carefully
prepared answers designed to sound good in print. He kept his real feelings to himself.

Parked to the side was one of the modified vans commissioned by the Office of Alien Affairs. It
would be used to transport the human to the compound at the edge of the base. Operated by an OAA
driver the van would also serve as her personal taxi service, taking her from the compound to
the university and back.

The compound itself had been a storage building. The OAA, together with the military, renovated
and secured the building for the benefit of the human visitors. The humans themselves did some
additional renovations inside, the exact nature of which was not known. No dunari had ever been
in the compound since the humans took residency, and Colty did not like to think about what
manner of activities took place there.

Colty stood with Rocalla and Tusk mostly just listening to their conversation. He didn't feel
like he had anything to add. Instead he worried about pronouncing the human's name properly. He
didn't want to start their relationship off by making a fool of himself.

``There!''

In the distance Colty could see a fast moving streak. Even though it was obviously still karnons
away, it seemed like it would be there in moments. The group assembled in preparation. Lucasa
and the human ambassador were first, followed by the chancellor, Rocalla, Tusk, and finally
Colty, all in a curved arc.

Despite the great distance, Colty could see the distinctive green glow surrounding the craft. It
was much easier to pick out than he expected. Suddenly there was a bright, green flash and
almost instantly the craft began to move more like a dunari aerocar. A few moments later, a deep
noise rumbled over them like distant thunder on the desert.

``That's the transition,'' Tusk explained. ``It's really quite something when it happens right
over your head.''

Once the craft landed it taxied over to the vicinity of the tent and stopped. Colty was
surprised at how small it was. It looked hardly any larger than the OAA van. A hatch opened and
a single bipedal figure climbed out. It started walking toward the assembly, swinging its front
legs in time with its back legs despite their not touching the ground. The dunari greeters made
last minute adjustments to their clothing, their ears, and their trunk cords.

As the human drew closer she slowed down and approached the head of the line. Although taller
than a dunari, she was not nearly as tall as Colty had expected. She was wearing plain, light
colored clothing that covered her body and back legs but her face and front legs were bare. She
had a kind of mane of black fur growing from her head, tied in the back, and small, forward
facing eyes. Yet it was her human skin that seemed the most alien. Instead of the creased scales
common among animals of Rujar, her skin was smooth and lightly colored. It was unlike any skin
Colty had ever seen.

The human ambassador stepped forward and the two humans grasped each other's front legs almost
they way dunari clasp trunks. Colty wondered if it was a gesture done for their benefit. He
could hear the human ambassador speaking in Argenian. His voice was smooth and well practiced.

``Welcome, Sar. Rachel. It's an honor to have you here.''

``Thank\ldots\ you,'' she said haltingly. She also spoke directly in Argenian without the aid of
a machine translator. Her voice was strange to Colty's ears. It was high pitched but soft and
light, almost like an airy whisper. It hissed and clicked slightly and she spoke slowly but her
diction was good and her speech was easily understood. ``I\ldots\ glad to-be here. We are well
met.''

``We are well met.''

Lucasa then stepped forward. 

``We welcome you, Sar. Ra'hel, to Rujar, the Barnal province of Argenia, and to the city of
Parnon. I am Lucastanonia, Argenia's ambassador to Earth. Greetings.''

``Thank\ldots\ you, Luc\ldots\ Lucas'anona,'' she said. ``I\ldots\ apreiate that. We are well
met.'' A couple of reporters snapped some pictures. Despite tradition, Lucasa did not touch the
human.

Lucasa then turned to his right and introduced the chancellor of the university. The human
looked directly at him with her small, round eyes. ``We are well met,'' she replied slowly after
his introduction. Her soft voice was in stark contrast to the deep, rumbling, raspy sound of
dunari speech.

``We are so pleased to have you with us. Let me know if there is anything you need.''

Rocalla was introduced as a professor of physics. ``It's an honor to meet you,'' he said. ``I
look forward to talking with you.''

``We are well met,'' she repeated. Colty was surprised at how polite she was. Although she was
no doubt following a protocol that had been described to her, she seemed oddly sincere.

When Tusk was introduced he bobbed his head slightly and then spoke to her in her own language.
It sounded like little more than noise to Colty's ears but she seemed pleased. She replied to
Tusk in her language as well. It was an utterly alien sound coming from her, and it made Colty's
skin crawl.

Lucasa then introduced Colty as ``Sar. Coltinarly'' and as her host. As he had been told to
expect, she looked directly at him while Lucasa talked. Her resemblance to a chortak was
remarkable. Colty was fascinated by her circular pupils. He wondered what impact that shape had
on her vision.

``We are well met, Sar\ldots Ra'el,'' he said after Lucasa finished. He knew that he had
destroyed her name after all. ``I look forward to getting to know you.''

``Thank\ldots\ you,'' she replied slowly. ``I\ldots\ apreiate your hossspitality. We are well
met.''

After the introductions she stepped back slightly and nodded toward the reporters as a group.
They took several pictures of her, first by herself and then standing with Lucasa and the human
ambassador. Colty studied her carefully but her expression appeared emotionless and impossible
to read. Finally she got into the van that would take her to the compound. The introductions had
been short and to the point. She didn't make a speech nor did she answer any questions from the
reporters.

``They often seem rather cold,'' Tusk explained after the van drove off. ``I sometimes get the
feeling they aren't quite sure what to say to us.''

Colty was relieved it was over. He returned to his office where he spent some time working. He
found it comforting to be thinking about math again in his dark, cramped space. The encounter at
the base seemed almost unreal by comparison. He knew he had to be at Cogart in time for the
presentations but meanwhile he was grateful for a few moments of normal life. A couple of
students stopped by unexpectedly to ask about the latest assignment and that kept him nicely
distracted for a time. After they left he decided to visit the library. He felt a need to talk
with Joleia.

She was in her office when Colty arrived but as soon as she saw him she came out to speak with
him. ``How did it go?'' she asked.

``Nerve wracking,'' Colty said, ``but fine, I think.''

``It must be exciting to meet a human face to face.''

``They are strange creatures. It seemed so distant. It spoke Argenian but was very formal.''

``It's probably strange for her too.''

``I suppose so,'' Colty replied.

The two went outside into the library courtyard and lay for a while under a small rigat tree to
talk. Colty found it a great relief to speak with Joleia and soon he felt much more relaxed.

``Have you ever been to Dolarta's?'' she asked.

``Is that the improv club on Following Street Tanis?''

``Yes. I like to go there sometimes to listen to music. It's not the Yarlantan thrumbatas but
it's still fun. Would you like to go with me sometime?''

Colty clicked his teeth. ``That would be wonderful.''

The time with Joleia flew by, and all too soon he needed to leave for Cogart. He thanked her for
a nice chat, took a deep breath, and stood up. ``Good luck,'' she said as he headed down the
path.

When he got to Cogart he saw a small group standing near the side entrance beside an access
road. He recognized Lucasa, Tusk, and Rocalla but there were a few other people as well. Dargon
was not there. The university chancellor was not there. Nobody from the press was there. It
looked like nothing more than an impromptu gathering like so many such gatherings that take
place all the time around campus.

Colty joined the group and was greeted by Tusk. Only a few moments later, the human's van pulled
up the service road and came to a halt not far from them. Everyone quieted and stood nervously
waiting while Lucasa walked over to the van. The driver got out and opened the back door,
sliding a ramp down to the ground. Moments later the alien stepped out.

She walked down the ramp easily. It amazed Colty that she could keep her balance so well on only
two legs. He was also struck by her clothing. She had evidently changed since her arrival.
Unlike the humans he'd seen on the viewer she was dressed rather colorfully. She wore a kind of
tunic that covered the upper part of her body and her front legs. It was light yellow with tan
trim and red highlights. On her back legs were leggings in a similar style. On the left side of
her chest, small yet distinctive, was some kind of symbol in black. Colty wondered what it
meant. Her fur was tied in a bundle and hung down the back of her neck to her shoulders.

``It's good to see you again,'' Lucasa said.

``Yesss, thank you,'' she replied quietly.

She greeted Tusk, Rocalla, and finally Colty. As before she looked at Colty directly when she
spoke to him.

``This way,'' Lucasa said. He indicated the side door into Cogart with his trunk.

``We thought you might enjoy a short tour of our labs first,'' Rocalla said.

``Yesss,'' she replied, ``that fine.''

They entered the first floor of Cogart with Rocalla in the lead, side stepping, Lucasa and Tusk
on either side of the human, and Colty just behind Tusk. As they walked down the hall the
human's head nearly reached the ceiling despite the modern, spacious construction. Colty
doubted if she would even fit in the halls of Fargon.

Rocalla talked more or less constantly describing the labs and giving an overview of the
research done in each. He spoke with pride about the equipment they had, emphasizing its
uniqueness. ``Parnon has one of the most respected physical science research facilities in all
of Argenia. I look forward to talking with you about the equipment you use.''

The human looked around each lab with obvious curiosity but she didn't say much. She listened
attentively to Rocalla which only seemed to encourage him. She was particularly interested in
the signs. She tried to pronounce the words printed above each door. ``Mag-net-ics\ldots\
Struct-ural Phys-ics\ldots\ Crys-tall-ography.'' Rocalla tried to help her but she seemed to
rely more on Tusk when it came to linguistic matters. Tusk spoke to her now and then in her own
language but that seemed to annoy her.

``Mussst prac-tice,'' she said to him.

Colty was surprised at how like a dunari she seemed. She certainly didn't come across as a
creature with an all powerful intellect. She made many mistakes trying to speak Argenian and
struggled at correcting them just like anyone learning the language might. He found himself
thinking about what it must be like for her being the only human surrounded by dunari. Colty
wondered if she was afraid.

After the labs the group passed by the lecture hall in the center of the building. Rocalla
opened the doors and invited her to go inside.

``Thisss where teach?'' she asked.

``Yes,'' Rocalla replied. ``This is the space we use for the introductory classes. The advanced
classes are taught in smaller rooms on the second level.''

The room was strangely empty of students, no doubt cleared out for the tour, but the board was
covered, as usual, with many exercises and notes. The human looked at the board curiously.
Rocalla demonstrated it by writing a small formula in the corner. She watched him closely and
then picked up a piece of chalk with her amazingly dexterous front toes. She drew a curved line
on the board which seemed to please her.

``What made from?'' she asked holding up the chalk.

``It's chalk,'' Rocalla said.

% Modern chalk on Earth is made from calcium sulfate. It used to be made from actual chalk rock,
% calcium carbonate. Perhaps the dunari are still using the older material. Note that calcium
% carbonate is laid down (on Earth, anyway) in a marine environment. Presumably Rujar wouldn't
% have as much of it as Earth does.

``It's made from calcium sulfate,'' Tusk said.

``Cal-cium\ldots\ sul-fate,'' she repeated slowly.

The group then went to the second floor where she was shown her office. The air cooling unit
installed by the humans was running and Colty was shocked at how cold it was. She didn't seem to
mind. In fact she walked up to the unit and briefly raised her front feet so that the cold air
would blow on them.

``Thank much very,'' she said to Rocalla. ``I\ldots\ appreiate what you are done for me.''

``You are most welcome,'' Rocalla replied.

She closed the door to her office and Rocalla locked it and then gave her the key. He nervously
held it out in his tentacles and let it dangle from a short chain while she wrapped her toes
around it.

``There is a large conference room at the end of the hall where we will have the
presentations,''

When they reached the conference room there were already a number of people gathered. Colty
recognized a few of the faculty from physical sciences, but most of the people were unknown to
him. He felt like an outsider, in some ways even more alien than the human.

As she entered, the people nearest to the door stepped back. Colty couldn't tell if they were
just trying to give her space or if they were reluctant to be close to her. She was conspicuous
not only in her humanness, but also in her height, standing well above even Lucasa.

Rocalla directed her toward the front where he invited her to lay down on a seat there. She
paused and then lowered her body by bending her back legs. Instead of lying down the way a
dunari might, she folded her legs underneath her, keeping her torso upright. This made her seem
taller than ever. It was a strange posture and looked decidedly uncomfortable but she didn't
seem to mind.

Everyone else also took their seats. Tusk and Colty lay down on either side of the human and
Lucasa next to Tusk. Rocalla stepped up to the podium and waited for everyone to settle. When
all was quiet he spoke. No microphone was needed in the small space.

``We are here to welcome Sar. Ra'hel, a visiting professor from\ldots\ Earth to the physics
department of Parnon University.''

The group stomped their feet. The human jumped slightly and looked around curiously. Colty
reluctantly leaned toward her. ``It's traditional to stand up at this point.''

She raised herself back up and turned toward the room. When the stomping ended she lowered
herself to her seat as before. ``Thank,'' she said softly to Colty. He was again taken by how
ordinary she seemed.

``We have a few presentations,'' Rocalla continued, ``to help our guest get familiar with the
research being done here. We promise to keep it short.'' Several people shook their ears
slightly. The human looked around uncertainly. ``Without further ado let me introduce Nartoga, a
dictatorium student in the crystallography group.''

The student came up to the podium from the back of the room. He avoided looking at the human and
spoke too quickly. He talked about the work being done in his group using highly technical
language, drawing many diagrams of crystals on the blackboard that Colty did not understand.

The human watched the presentation intently, but Colty found her impossible to read. She looked
bored, but for all he knew she was fascinated. At the end of the presentation there were some
questions from the audience. The human looked around as they were asked, but she did not have
any questions herself.

As the next student came up to give his talk, the human reached into a small pouch in her
clothing and took out a slightly flexible cup-shaped object. It had a clear tube dangling from
it that led back into her pouch. She put the cup over her mouth and stubby nose and took several
deep breaths. She then quickly returned the device attempting, it seemed, to be as discrete as
possible.

They sat through several other presentations all of which Colty did not understand. He wondered
again what he was doing there. However it did give him a chance to study the human. He looked at
her discretely with his left eye. Her body was strange, but not hideous. Her resemblance to a
chortak, an ordinary Rujaran animal, made her seem more familiar than he expected. She wasn't
identical of course. Her front toes were far more flexible, her head was higher and larger, and
her skin was altogether alien. Yet her basic shape and even her demeanor was not as strange as
Colty feared it would be. He wondered if there were any animals on Earth that resembled the
dunari and, if so, what they might be like.

The last presentation was by Rocalla himself. He spoke in a loud voice and with a confidence
that the other presenters, mostly students, had lacked. He talked about his work on stellar
evolution with an emphasis on the outstanding problem of how stars derive their energy. He
showed that no model involving the burning of fuel in the ordinary sense could account for the
longevity of stars, even after considering the primordial heat of compression, and he speculated
that some unknown source of energy must be involved. Unlike the other presenters he frequently
looked at the human. As before she watched the presentation carefully but otherwise made no
sign.

When Rocalla was finished there were a few questions from the audience and then everyone
politely stomped their feet. The human did the same with one of her front feet, something she
had learned while watching everyone else. Rocalla returned to the podium. ``I would now like to
invite our guest to say a few words, if she would like, about her work as well.'' He spoke
directly to the human. ``It's traditional for a visiting professor to do a short presentation,
but under these unusual and special circumstances I think it is fine if you'd rather not.''

``I underssstand,'' she said. ``I say some words.''

The human stood up and walked to the front of the room. She turned and paused. Colty could see
her looking over the audience. ``I mathemat'ian,'' she began. ``I search for\ldots\
in-consissstencies in laws of physssics as we understand them.'' Colty sat up a bit. ``There is
problem.''

She turned and using her dexterous front toes picked up a piece of chalk just as a dunari might
do with his tentacles. She looked at it briefly and then stretched out her leg and started to
draw on the board. She drew an elongated oval, at a slight angle, and roughly filled it in with
chalk. From each end of the oval she drew a large, curving arc that wrapped three quarters the
way around and spiraled outward slightly.

``This our galaxy,'' she said. ``It has bar of starsss in center 280,000 light run-ionsss
long,'' She touched the chalk to the oval and spoke slowly and softly, ``two major branchesss,
and some minor branchesss I not show.''

Rocalla was riveted and several students scribbled down some notes. As far as Colty knew the
exact shape of the galaxy was not previously known.

``There is clustersss. Very tight.'' She curled the toes on one of her front legs into a ball.
``You call them ssspherical clustersss. Many of them go around galaxy.'' She made a waving
motion above her drawing. ``But\ldots\ some go right around bar.'' She moved one of her legs as
if to show an object rising up out of the board, around the oval, and then back into the board.
``Very strange orbitsss. Do not make sense.''

She put down the chalk and looked at her toes. She seemed disturbed by the chalk dust on them
but she tried to ignore it. ``I think our theory of ssspace-time not quite right. Maybe
structure of space hasss special symmetriesss. Maybe explain strange orbitsss.''

She walked over to the podium but she was much taller than it was. ``From Earth can't see far
side of galaxy much. Earth in center of disk. Too much starsses in way. Ru-jar on top of disk
above clouds of starsses. I come here to make measurementsss of far side spherical clustersss.
Get better numbersss\ldots\ check my theory.''

She paused. ``That all I have.''

It had been a short talk but in that time she had already revealed things never before known.
There was a palpable energy in the room.

``What did you mean when you said `structure of space?' '' Rocalla asked.

The human seemed at a loss. She glanced toward Tusk who cocked his head slightly in readiness.
``Space\ldots'' the human began, ``hasss physical\ldots\ realness. Even empty it isss
filled\ldots'' She glanced out the window and ran one of her front feet through the fur on her
head. She spoke in her own language briefly and softly, but more to herself than to the
audience. ``Um\ldots\ space has math structure. Can study it. I\ldots\ I sorry. I do not know
wordsss.''

The room fell silent.

``Maybe we can talk more about it another time,'' Rocalla said. The human nodded her head in a
surprisingly dunari way.

``What is strange about the orbits?'' someone from the back asked. Colty couldn't see who it
was.

``They is not\ldots'' she struggled again. This time she spoke to Tusk in her language.

``Random?'' Tusk said. ``It means arbitrary.''

``No,'' she said. She made a kind of stirring motion with her front leg. ``More mixed up.'' She
spoke some more in her language.

``Chaotic,'' Tusk said. ``I think you mean chaotic. It means like you said: complex and
unpredictable.''

``I so sssorry. I need to learn so much before I right talk.''

``You're doing fine,'' said Tusk.

She took a breath and then raised her foot toward the back of the room. ``The orbitsss are
not\ldots\ chaotic. Them\ldots\ locked.''

``Your people have interstellar space flight,'' another student said, ``can't you just send a
ship to the center of the galaxy to get the measurements you need?''

She nodded her head slightly. Colty wasn't sure what it meant. ``It too far.'' She picked up the
chalk again and drew a mark on her diagram about two thirds the way from the center and between
the main spiral branches. ``This Earth.'' She then drew another mark right next to the first
one. The two marks were touching slightly. ``This Rujar.'' She put the chalk back down. ``You
see how close we is? Yet this as far we go.''

Again the room was profoundly silent.

She pointed one of her flexible toes at the mark for Earth and then moved it slowly toward the
center of the oval. ``This very long way. We have probe going to center galaxy. Take many
run-ionsss.''

She turned back toward the audience. ``Much easier look from here. You have nice view.''
Colty clicked his teeth softly. Was that a joke?

``Are you going to teach any classes?'' someone asked.

She hesitated and then glanced toward Rocalla. He got up and walked to the podium. ``We will
give our guest a chance to get acclimated. After that, we'll see.''

``I would like teach,'' she said.

There were no other questions and the human went back to her seat. Colty noticed a film of water
on her skin that was not there before. It made her seem completely alien and Colty fought a
strong revulsion.

Rocalla said a few final words to wrap up the event. ``Again we welcome you, Sar. Ra'hel. If you
have any questions or need anything I'm sure any one in the department would be happy to help.
Of course you can also ask your host, Sar. Coltinarly.''

``Thank,'' she replied.

The event ended and people got up. As everyone was leaving Rocalla came over to her. He talked
about her presentation but Colty didn't pay attention. He had reached his limit. He had nothing
to say to the alien, to Rocalla, or for that matter to anyone else. Discretely he slipped out
the door and made his way down the empty halls of Cogart.
