% Local IspellPersDict: aspell.dunari.pws

\chapter{29}

Colty was in his office puzzling over a paper on co-sequence correlation. He was having trouble
with one of the proofs. He took out his notebook and flipped back several pages. Partfolio's
work on mono-sequence rotation seemed inconsistent with the co-sequence correlation theorem.
Colty wrote a few formula on a fresh page of his notebook and stared at them with his right eye.
His concentration was broken by the chiming of his phone.

``Yes,'' he said brusquely as he answered.

It was Joleia. ``Colty?''

``Oh, hi! Yes, it's me.''

``I don't mean to disturb you, but I just wanted to let you know that the human is here at the
library. It\ldots\ she was asking about math books.''

``Math books?''

``Yes, but I guess I shouldn't have bothered you. I'm sorry.''

``It's fine,'' Colty said. ``I was thinking of taking a break anyway. I suppose I should go over
there and do my hostly duty.''

It was a beautiful moment. The Nermella was blowing lightly and the sky was filled with brightly
colored patches of clouds. Yet Colty didn't notice the breeze, the sky, or much of anything else
as he walked rapidly down the path toward the library. It was exactly an interruption such as he
feared would happen.

When Colty got to the librarian's counter Joleia was there. She pointed toward the ramp. He
nodded and took a deep breath. He knew the way very well, and bounded up the ramp to the second
floor. At the top of the ramp was a small lounge for study or reading but it was empty. He
walked down the wide aisle past engineering, past chemistry, and past physics and finally to
mathematics at the end. Colty slowed and looked down the stacks at either side of the aisle but
he saw nobody. The smaller lounge at the end of the aisle was also empty. It seemed as if the
human had already left.

Yet behind the lounge in the corner of the building was a section of elementary texts. There he
found her on the floor with her rear legs folded beneath her and her body upright the way she
had been during the presentations. She had two books on the floor beside her and she held
another in her flexible front toes.

As Colty approached she looked up. ``Um\ldots\ hello,'' she said tentatively.

``Good singing,'' Colty said.

She looked at him closely. ``Sar. Coltinarly?''

Colty wasn't surprised that she had trouble recognizing him. He hadn't made much of an
impression at their first meeting. He swayed his head but she just watched him uncertainly.
``Yes,'' he added finally.

``Good\ldots sssing-ingg,'' she said. She parted her lips slightly and bared her teeth. Colty
instinctively folded his ears against his head and stepped back, bumping into the bookshelf in
the process. He had no idea what the strange creature was about to do, and he was suddenly aware
of how alone they were.

The human abruptly covered her teeth again. ``I sorry,'' she said, ``I am happy you isss here.''

``It's fine,'' Colty replied nervously. ``We just need to get used to each other.''

``I find good readsss here.'' She held up a book on elementary algebra. Colty felt braver and
stepped closer to her so they were properly next to each other. Even with her legs folded under
her she was almost as tall as Colty was standing. Yet without her towering over him she did seem
a bit less intimidating.

``That's very basic material,'' Colty said as he looked at the cover of the book.

``I needsss learn your math writing. It hard for me.''

Colty noticed what appeared to be a piece of glass leaning up against the bookshelf, about the
size of a sheet of paper. Yet it was covered with symbols in an alien script. The human saw him
looking at it and she picked it up.

``My machine,'' she said. She touched the surface with her toes and then, much to Colty's
amazement, folded it in half and then in quarters. She put it into a small pouch that she had
strapped around her waist. ``I should take notesss dunari way. How you do it?''

``I just use standard research notebooks. You can get them at the university supply store.''

``Can you show?'' she asked. ``Am-basss-ador Lucas'anona gave me money.'' She took a small bag
out of her pouch and showed Colty a few bills.

``That's more than enough. In fact, you probably shouldn't carry that much around.''

``Oh.''

``Are you sure you want to go to the store? I could get you the notebooks you need.'' Colty
wasn't thrilled by the prospect of taking yet more time away from his work. Yet it seemed like
the appropriate thing to offer.

``It safe? I will like to see.''

Colty agreed to escort her and she put the books back on the shelf, carefully returning each to
its rightful spot. ``You can check those out at the counter and take them with you,'' Colty
said.

``I can?'' She seemed surprised. ``How othersss read?''

``That's how it works here.'' He was struck by the creature's unassuming attitude. ``Any of
these.'' He waved his trunk as if to indicate the entire shelf.

She picked a couple of the books off the shelf again and looked at them. She seemed undecided
for a moment but then put one back. ``I take thisss one,'' she said, holding the other up
slightly.

Colty stepped aside and pointed the way with his trunk. ``It's this way back to the counter,
Sar. Ra'el.''

She raised herself up on her hind legs, a process that both fascinated and disconcerted him.
Then they walked toward the library entrance side by side. She said, ``You can just call me
\alien{Rachel} if you like.''

When they reached the main lobby of the library there was a short line at the counter. The
people waiting saw them coming and shifted uncomfortably. The human seemed fine waiting off to
the side some distance away. Colty looked for Joleia, but didn't see her.

When everyone in the line had been served they approached the counter where a woman who looked
as old as the library itself waited for them with obvious disdain in her eyes. ``What can I do
for you, Sar. Coltinarly?''

``She wants to check out this book,'' Colty explained. He gestured toward the book the human was
holding. She put it on the counter, retracting her front leg quickly.

The librarian looked at her balefully. ``Do you have a library card or student ID?'' she said
curtly.

To Colty's surprise the human opened her pouch and took out a brand new visiting professor card,
complete with her picture. The librarian studied the card suspiciously as if she thought it
might be some kind of trick.

``She's a visiting,'' Colty said as if the librarian didn't know.

The librarian fluttered and recorded the book with the human's name and ID number, which she
copied from the card, into her borrower's log. The human watched the process with interest.

``How long me keep?'' she asked.

``Standard check out is one runion,'' the librarian said without looking at her.

The human took the book and the two of them made their way toward the entrance. She walked
slowly and Colty felt frustrated by the delay. When they reached the inner court the human
paused and looked up at the great, domed ceiling. She raised one of her front legs and extended
a long toe to point at the inscription carved beneath the high windows. ``What do words mean?''
she asked.

``I don't know.''

``Maybe Tusk know.''

``Probably. The entrance is right over here.''

Colty watched the reaction of people as they made their way across campus. Some simply stared
and others avoided looking at them at all. Most kept their distance by using other paths or
walking on the grass to go around them. The human looked around warily, clutching her book close
to her body with her front legs.

They came to a garden full of tulara flowers in bloom. The human stopped to look at them. Behind
the flowers was some nara bush, its broad yellow-gold leaves curled slightly in the sun. She
lowered herself to get a closer look by bending her rear legs. On the ground beneath the nara
leaves was a thin vine dotted with tiny red flowers.

She asked lots of questions about the plants. What were they called? Did they need to be
watered? Did they grow naturally around Parnon? At first Colty kept his answers brief but her
curiosity was infectious, and soon he started wondering about the answers himself. He guessed
there were no flowers on Earth. The human acted as if she'd never seen one before.

In the middle of the Yellow they came to a seat. The human stopped to ask what it was.

``It's a place to lie down and rest. Students sometimes come here to study or to just enjoy the
sun. You'll find them all around campus.''

She lowered herself onto the seat in her usual way with her legs folded underneath her. She
pointed to the top of the spire on Tartuk Hall and asked about the ring.

``It's Parnon's symbol,'' Colty explained. ``The top represents thought, ideas. The bottom
represents things, actions. Ideas to actions and actions to ideas,'' he said repeating the
university's motto.

``Oh.''

Colty lay down beside her. For some reason he didn't feel as hurried as he had earlier. 

``How old Parnon?'' she asked.

``I really don't know. It's old. It was one of the first universities chartered during the age
of the Regents. That was before confederation. One of the first Regents, I forget who exactly,
decided that he wanted his reign to be known as a time of enlightenment. So he created two
universities to be centers of learning. Parnon was one and Parlola was the other. Parlola
doesn't exist anymore\ldots''

Colty stopped abruptly. He became aware that the human had turned to look at him. Her forward
facing eyes with the strange, round pupils regarded him closely. He shifted uncomfortably. ``I'm
sorry,'' he said. ``I was rambling.''

``It interesssting,'' she said.

``I don't really know much about it. I'm new at Parnon myself. I'm not a very useful host.''

``You good. You not fear me.'' She looked out across the grass toward the steps of Tartuk Hall.
``I apreiate that you take time.''

They lay quietly for a moment. ``How did you pronounce your name again?'' he asked.

``\alien{Rachel}''

He took a breath. ``Ra\ldots\ hel''

``Yesss!'' she smiled briefly but quickly suppressed it.

``I'm Colty.''

Rachel reached into a pocket and took out the same clear, flexible cup she had during the
presentations. As before she put the cup over her face and breathed deeply several times. She
quickly returned the cup to her pocket and then noticed Colty watching her.

``What is that?'' he asked.

``It just ox-i-gen,'' she replied. ``Your air has less ox-i-gen than Earth air. It helpsss me to
have a little more.''

``So you need that to survive here?''

The human shrugged in a very dunari-like way. ``My body get used to air here in future.''

They lay together for several moments before Rachel finally suggested they continue on their
way. A gust of wind kicked up some dust on the path as they walked. Rachel's hair blew in the
breeze.

``It alwaysss windy,'' she said, ``ssso they tell me.''

``It's the Nermella,'' Colty explained. ``It blows, more or less, all the time. Sometimes it
brings dust from the desert but here at Parnon it mostly brings clouds from the sea.''

They came to the opposite side of the Yellow and the path took them between two buildings.
Rachel seemed fascinated with the landscaping. The shadows from the buildings formed the
outlines of gardens and walkways. Inside the permanently shaded areas were plants adapted to
such life while the sun loving flowers were planted in the open.

She compared the shapes of the shadows with the shapes of the tops of the buildings, looking
back and forth between the two. ``It amazing you do thisss,'' she said.

``What do you mean?'' Colty asked.

She pointed along the walk where it had been lined up against the shadow of Minar Hall. ``You
plan on shadow being here alwaysss.''

``Of course. Where else would it be?''

``My sun brighter and hotter. It yellow, not red. Earth far away from it. Earth not\ldots\
locked.''She raised up one of her front legs and held it out to the sky. ``As Earth turn sun
movesss across sky.'' Then she looked at the ground and gestured over it. ``Shadowsss move
across ground.''

It was hard to imagine. ``It must get very hot with the sun high in the sky.''

``It get warm, not hot. Only short time. Then night.''

``Night?''

``Yesss.'' She pointed at the ground and moved her leg in a slow arc again. ``Earth turn. Sun on
other side for a time. Then rise. On Earth you don't go to night. The night come to you.''

Colty was fascinated and horrified at the same time. ``How do the plants survive it? Do you even
have plants?''

``Plantsss used to it. Like that for billionsss of run-ions. It strange here for me. Day
alwaysss. Sun never move. Confusesss my body. We sleep at night. At compound we switch light and
dark so more natural for usss.''

It was hard for Colty to think of anything more unnatural than purposely alternating between day
and night. How could their society operate like that? he wondered. What kind of creatures could
live on such a world? He realized with a start that he was walking next to one of those
creatures.

They came to Following Street Tanis, a road that formed the boundary between the university's
Old Campus and the town of Parnon itself. There was significant traffic on the road and many
pedestrians on the sidewalk. Rachel seemed nervous. ``How far?'' she asked.

``It's just up there,'' Colty said pointing down the other side of the street a short distance.

She hesitated. ``That fine.''

As they walked along the sidewalk Colty became very aware that they were being stared at. He
heard whispered voices behind him and people moved out of the way as they approached. His heart
pounded. He wasn't used to being a spectacle. He glanced over at Rachel. She had her head down,
looking at the sidewalk in front of her. She didn't seem to notice the reactions of others.

They came to a crosswalk, stopping a short distance back from the road behind a small group of
dunari gathered there. ``We need to wait here,'' Colty said, ``until the sign says we can go.''
He indicated the crossing signal on the other side of the street.

``That fine,'' she said softly. She didn't look up.

A military officer stepped up to the crosswalk and the other dunari seemed to relax a little. He
looked steadily at the human but said nothing. When the light changed the group moved across the
street. Colty waited slightly so they would follow behind. Rachel held her body close to
herself.

Once they got inside the store Rachel seemed to relax a little. Very few people were shopping
and the aisles protected them from the staring eyes of onlookers. Rachel also got interested in
all the items for sale and started asking many questions about them. What were they for? How
much did they cost? Why were there so many different kinds of paper clips? They were questions
from a stranger in a strange land. Colty had to remind himself that this was the same creature
who only an hour before had told the physics department of Parnon University that space itself
had structure.

With Colty's help she picked out a small notebook and a pen for writing. He was amazed at how
easily she held the pen between her long front toes. He never would have guessed that any kind
of bony appendage could have the necessarily dexterity to accurate manipulate a writing
implement.

``What these?'' she asked regarding a rack of photo sheets. Colty explained they were for
decoration.

``A collection of nice things to look at.''

She was fascinated. One sheet featured images of Morgolan Park. ``It beautiful,'' she said
looking at a picture of a snow capped mountain. ``I think I need dec-oration for office.''

``Human!'' The two looked up to find three young men, probably students, standing at the end of
the aisle. The one who spoke stepped forward with his friends just behind. ``What do you want
here?'' he said with a demanding tone.

``I\ldots\ I want dec-oration.'' She glanced at Colty.

``We don't want you here,'' the student said in a loud voice. ``Go back where you came from!''
He stepped forward menacingly.

Rachel stepped backward and ran into the rack sending several photo sheets clattering to the
floor. Without thinking Colty put himself between the human and the approaching anger. ``Look,''
he said calmly. ``We're just here to get a few things. This is not the time or place for this.''

``Is there a problem?'' The manager of the store was standing at the end of the aisle looking
generally displeased. His ears were back, his legs were locked straight, and he fluttered
softly.

The student looked back and then again at Colty. He pointed his trunk at Rachel. ``You watch
yourself.'' He slapped the backs of his friends. ``Let's go.''

After they left Rachel started to pick up the photo sheets that had fallen on the floor. The
manager motioned to Colty. ``Get it out of here,'' he said in a low voice.

Colty returned to help Rachel. ``I so sssorry,'' she said. ``Made chaotic messs.''

``It's okay. Let's just pay for what we have and head back to campus.''

Rachel's leg shook as she put the photo sheets back on the rack. Colty understood that she was
distraught. She replaced the photo sheet of Morgolan Park but as she turned to go, Colty grabbed
it again.

A small crowd had gathered as a result of the altercation but Rachel did not look at anyone. She
kept her head down and went to the cashier with Colty. She put down her notepad and pen. Colty
added the photo sheet. She fumbled with her pouch. Her amazing front toes didn't seem to work as
well as usual. ``I've got it,'' Colty said handing the cashier a bill. ``You can pay me later.''

``Thanks for coming,'' the cashier said as she gave Rachel a bag. Rachel didn't reply but
instead took the bag and headed for the door.

The two made their way back to campus in silence. Colty wasn't sure what to say. ``I'm sorry
about what happened,'' he said finally. ``It's my fault. I should have gotten what you needed.''

``It fine,'' she said, ``This how we learn about each other.''
