% Local IspellPersDict: aspell.dunari.pws

\chapter{31}

Colty walked up the side ramp of the Cogart lecture hall to the back row. The students looked at
him curiously. He was sure none of them recognized him or understood why he was there. He lay on
a seat in the corner and watched the rest of the class wander in. At Parnon the opening of every
course was scheduled and all students where expected to be there at the same time.

From his vantage point almost a full story above the lecture floor he could easily see the
entire room. The walls were plain, the seats were functional, and all was in a modern style
designed more to save money than for aesthetic appeal. Light from the meshed windows illuminated
the chalkboard without creating too much glare for the lecturer to see. White sun panels hung
from the ceiling to reflect some of that light throughout the rest of the space.

The opening time drew near and yet no instructor appeared. Some students lay quietly, notebooks
ready. Some chatted awkwardly with their neighbors. Colty eyed the large line clock on the wall
above the chalkboard. The class grew restless. It wouldn't do if the class opened late. If
Parnon expected the students to be there, the instructor had to be there as well.

Suddenly the door opened and \emph{she} walked in, carrying the course book in her arms. She
didn't look at the students but instead carefully climbed the steps to the platform. She put the
course book on the podium and faced the hall. She was wearing thin, white clothing around her
torso but her arms and most of her legs were bare, exposing her strange pale skin openly. The
black hair on her head was pulled back as usual and hung below her shoulders.

Silence.

``Hello,'' she said tentatively. Her alien voice was soft but carried surprisingly well. Colty
could hear her fine even at the back. ``My name is Ra'hel and this is Introductory Physics. I
will be your instructor.''

``Dind,'' someone swore rather loudly.

Colty heard a noise coming from his right. Two students had just left through the back doors.
Rachel seemed not to notice. She walked to the blackboard and, picking up a piece of chalk in
her fingers, wrote some notes. ``My office is in 227,'' she said pointing in the general
direction. ``I will be there when I can, but I have a pad beside the door so you can leave a
message if I'm not around.'' She turned back to the class. ``Of course I will leave a pad here
as well.''

``I'm having trouble understanding your accent,'' one student called out suddenly.

She turned toward the source of the sound but the student did not identify himself. ``I
apologize,'' she said. She stepped forward to the edge of the platform in front of the podium.
It was something a dunari professor would never do. Colty wondered if he should be making notes.
``I've been practicing, but I know I need more practice. If you don't understand my speech
please say. It will help me learn.''

% TODO: Think about the kind of grammatical mistakes Rachel makes at this point and apply them
% consistently to her dialog. She should be speaking better than in the last chapter but still
% not perfectly.

A student in the middle of the hall called out, ``Are you a girl?'' There was some embarrassed
shifting, and Rachel paused uncertainly.

``Our genders are a bit different than yours, but essentially yes, I'm a girl.''

Another student at the far side raised his trunk abruptly, his first tentacle pointing straight
up insistently. ``How can you teach us?'' he asked without waiting to be acknowledged, ``What do
you know about how we learn?'' There was a murmur of voices in the hall.

Rachel paced across the platform, still in front of the podium, and ran her fingers through her
hair. ``I understand\ldots'' she raised her voice. ``I understand your concerns.''

The class settled down and Rachel continued. ``I have the course book so I know what material to
cover.'' She stopped at the end of the platform nearest the student who had asked the question.
``Last term I attended one of Sar. Rocallanon's classes so I got a good idea of style of
teaching used here. With your help I think I do a good job for you.''

Colty overhead one of the students right in front of him whisper to his neighbor, ``At least
it's not a Forbinite.''

Rachel walked back to the center of the platform. She paused and looked over the class. Nobody
said anything. She finally stepped behind the podium and opened the course book.

``Shall we begin?''

She carefully transcribed the notes on vectors onto the board, pausing occasionally to highlight
some point. Her style was very methodical and she followed the course book closely. Colty
clicked his teeth softly with approval.

Relieved by the normalcy of her presentation most students dutifully copied the notes into their
notebooks. Nobody asked any questions, and although Rachel paused and looked at the class often
they just waited for her to continue.

Finally she wrote the first exercises on the board: several simple problems in vector
manipulation. The students started to work the problems but Colty watched Rachel. She seemed
grateful for the break and leaned against the wall to rest. She took a few breaths from her
oxygen canister. Then she pulled a small device from another pocket in her clothing, unfolded
it, and used it to fan air over her face as she gazed out the window. As was traditional she
didn't acknowledge Colty at all. They would meet afterward to discuss any issues Colty
identified.

After a few moments she realized several students were watching her, and she hastily put the fan
away. Standing up straight again she spoke to the class. ``Does everyone understand how to work
these problems?'' Hearing no response she started to write the answers down next to each
question. ``You should be sure you can derive these\ldots''

``Do you teach on Earth?''

Rachel stopped and turned back toward the class still holding the chalk up slightly.
``Yes\ldots\ ah, yes I do,'' she said.

``Like this?''

Rachel put the chalk down and tried to brush the dust off her fingers.

``Let it finish putting up the answers,'' another student blurted out.

``Would you be quiet?'' yet another student said, ``As Jurita's Witness it's just a question.''

The class shifted uneasily. Colty tensed wondering if he should do something.

``I really don't care. I just want the answers to the problems.''

Rachel held her hands up. ``Please, pleassse\ldots\ no upssset. I\ldots\ will finish writing out
the answers in just a moment. In the meantime is natural to be curious.''

She paused for a moment and the class settled down again. ``My classses on Earth are different.
We use our technology to organize them, and we present the material in a different way.''

Nobody said anything more so Rachel finished the answers and, after consulting the course book,
started on the next topic. The rest of the presentation was uneventful. Rachel moved through the
material at the carefully prescribed pace and by the end of the class she had introduced the
basic equations of motion and had derived the idealized trajectory of a projectile.

Finally a soft but clear beeping started and Rachel touched a small device she was apparently
wearing on her waist to silence it.

``That is the end of my presentation for now,'' she announced. ``I will leave additional
materials on the board. Please come to my office sometime to introduce yourself. I will let you
know when we have our next sssynchronized meeting.''

Rachel stood well away from the students as they filed out. She looked exhausted. Her skin was
damp and she fanned herself often. Colty hung back and carefully approached her after everyone
else had left.

``How did I do?'' she asked.

``It was fine. I think it was fine.''

``Good,'' she said rubbing her forehead.

``Don't go in front of the podium. It's considered inappropriate to be that familiar with the
students.''

``I sorry. I must return to my office to cool.''

\spacebreak

Colty went to Fargon, grateful to be back in more familiar territory. He returned to his office
and spent several moments reviewing his notebooks and organizing his thoughts. He wasn't used to
being unprepared and the feeling unsettled him. Being host to the human hadn't taken as much
time as he expected, for she mostly kept to herself either in her office or in the library. Yet
because of her class he hadn't had the opportunity to properly meditate on his work and he found
himself having trouble focusing.

Finally the appointed time arrived. He rubbed his trunk ridge with this tentacles and then
returned his materials to their prescribed places in the trays of his table or on his shelves.
He carefully closed his most recent notebook and tucked it into the wide pocket of his cloak.
Then, locking his office behind him, headed down the long leeward hall to the Overseer room at
the end.

Colty took a deep breath and pulled aside the heavy wooden door. The room behind was small and
overflowed with history and tradition. It was in this room where Potarian first presented his
masterpiece \textit{Components of Numbers}. It was in this room where Rustuk proved the Finite
Folding Theorem by, as the story goes, scratching folding notations into the wall while being
grilled by an unforgiving Overseer.

Slowly, almost reverently, Colty walked across the ornately tiled floor. It was inlaid with
colored wood forming an image of the emblem of Menola, the ancient symbol of mathematical
purity. At the front, on a raised platform covered with garn stones, were three heavy, wooden
seats facing toward the center. The platform extended along the adjacent wall under wide windows
of lightly colored rosen glass. Lying on the seats were Colty's Overseers, awaiting him. They
were the three most established faculty members in the department who would ultimately decide if
he was to be granted full status.

Sar. Dargonick greeted him. ``Thanks for coming, Colty,'' he said good naturedly. Sar. Molanalor
and Sar. Tarnockalia said nothing. Tarnock seemed to be intently studying a document on the
small desk attached to his seat.

``As you know this is just routine,'' Dargon said. ``It's the start of your fourth term and we
need to check in with you.''

``Of course.''

``I hope you've been settling in well,'' Dargon continued.

``Yes, I think so,'' Colty replied. ``I have a small apartment\ldots''

``Tell us about your research,'' Tarnock interrupted.

Colty hesitated but his Overseers just waited. He wondered if he should take out his notebook
but in the end he didn't. ``Um\ldots\ I've been looking at co-sequences,'' he started. ``The
topic is a natural extension of my dictatorium on monosequences and my literature searches show
that it hasn't been extensively researched.''

Molan pushed aside a booklet that Colty recognized as his dictatorium. ``Hasn't Sar. Nulari and
his group at the Walash Institute already covered co-sequences?''

``I'm familiar with his publications of course,'' Colty said. ``His work is very general. I'm
actually looking specifically at co-sequences with stellating bases.''

``Why?'' asked Tarnock. He regarded Colty coolly as if waiting for an answer he thought Colty
could not provide.

Colty shifted on his feet slightly. He had often asked himself that same question. ``Well\ldots\
um\ldots\ it seems to me that there is a relationship between stellating co-sequences and prime
numbers.''

``It seems to you?'' Tarnock said with a flutter. He flicked his trunk to the side dismissively.
``What kind of relationship?''

Colty adjusted his cloak slightly. He took a deep breath to calm his nerves. ``I'm\ldots\ I'm
not sure\ldots\ yet.''

``Colty,'' Dargon jumped in, ``you're interest in co-sequences is understandable given your
background, but we're concerned that it's not a rich enough area to satisfy Parnon's
expectations of its professors.''

``At best there might be a minor paper in it,'' Tarnock said.

Molan turned toward Tarnock. ``Yet if stellating co-sequences are in some way primal, that's an
interesting result, yes?''

``If it's true,'' Tarnock said, ``but then what? The field is a dead end. It's probably not true
anyway.''

Colty raised his voice slightly to break back into the conversation, ``I feel like there is a
rich structure there\ldots\ although,'' he added more softly, ``I'm having trouble discovering
it.''

``There is no rich structure in co-sequence theory,'' Tarnock said. ``Nulari has already said
everything worth saying in that area.''

Dargon fluttered softly. ``Colty, we have big expectations for you, and we just think you can
aim a little higher. Given that it is your fourth term already and you haven't made progress on
co-sequences, perhaps you should give some thought to another topic.''

Colty held his ears close to his head. ``You\ldots\ you're right. You make a good point. I've
also been doing some reading about graph extension fields.''

``Excellent,'' said Tarnock. ``That's an important area, and nobody else in the department is
working on it.''

``That topic is also well aligned with your specialty,'' Dargon said.

``Yes,'' Colty replied softly.

Dargon made some notes on a pad in front of him while Colty waited patiently. Finally he looked
up and Colty stood ready to be released. ``One more thing,'' Dargon said. ``How are things going
with our guest?''

``I just came from her opening session,'' Colty replied.

``Do you talk with\ldots\ her\ldots\ it's a \emph{her}, yes?'' said Molan.

``Yes,'' Colty said. ``She comes to me regularly but she asks about little things. The last time
we spoke she asked if every building on campus had skylights.''

``Do you ever talk about math?'' Dargon asked.

``Sometimes she has questions about things she's read, but she learns quickly, and doesn't need
a lot of explanation. She\ldots\ well, to be honest she's very ordinary.''

After Colty left his overseers, he spent the rest of the runion trying to forget about
co-sequence theory and, for that matter, about his research entirely. He prepared for classes.
He spent time with Joleia. He distracted himself by helping Rachel. Her class seemed to go
smoothly, and she made good progress through the course book. He even found the time to explain
the rules of conifi to her, a game with a rich and fascinating mathematical structure. He was
pleased that Rachel seemed to appreciate it.

It was late in the runion when Colty finally spent nearly an entire hour in the library
gathering resources for his new research direction. Returning to his office he spread three
papers on graph extension fields across his desk. Each was a classic that introduced a distinct
branch of the topic. He read the abstracts and flipped through the pages. Graph extension fields
was an active area with mathematicians from all over Argenia and Forbin working on it. The
extent of the literature was daunting.

Colty took a deep breath and leaned back. He needed to find a niche where he could carve out a
useful and significant contribution. He picked up one of the papers and skimmed it. The figures
illustrating the primitive graph operations were certainly pretty. Yet Colty did not feel
inspired. I'm a number theorist at heart, he thought, not really a combinatorialist, and
certainly not an algebraist. Had he chosen the wrong specialty?

He tossed the paper back onto his table and, making some space, opened his notebook to a blank
page. Down the side he wrote in big letters, ``GEF.'' He let his mind wander a bit, reflecting
on various aspects of the subject. Do Hubort graph operations commute? he asked himself. He drew
a couple of figures on the page. ``Of course not,'' he said out loud. It was an entirely trivial
observation.

Colty sighed. After a moment he closed his notebook. He felt a need to take a walk. Locking his
office he headed outside with the intention of just walking once or twice around the Yellow. Yet
he found himself heading for Cogart instead. For some reason he felt an urge to talk to Rachel.

The atmosphere in the physical sciences building was very different than in Fargon. Everyone
seemed to move a little more quickly, and the sound of feet against the hard floors clicked a
little more loudly. ``Is Ra'hel here?'' Colty asked at the main desk.

Jarnol tilted her head toward the human's office. ``Last I knew,'' she said.

Colty walked down the hall. Rocalla's office door was open and Colty could hear him talking to
someone. As he approached he recognized Rachel's soft voice.

``\ldots\ it's frustrating for me,'' Rocalla was saying, ``to be working on problems for which
you already know the solutions. You say you are here to teach us. Why not give me some direction
at least?''

``I understand\ldots'' Rachel began. At that moment Colty lightly kicked the base of the door
frame. Rachel stopped and turned. ``Oh, Colty,'' she said with pleasure.

``I just came over to check in with you,'' Colty said, ``but if you're busy I'll come back
another time.''

``It's fine,'' said Rachel. She turned toward Rocalla who looked distinctly annoyed by the
interruption. ``We can talk about this more later. Okay?''

``Sure.''

``I was about to go to the library,'' Rachel said to Colty. ``Will you walk with me?''

Rachel led the way back toward the entrance of the building. She seemed anxious to get out of
Cogart and, without speaking, walked more quickly than normal. When they finally got outside she
paused. The sun lite the clouds with bright rubinum light and cast a tinge over the buildings
and grounds of Parnon. Three spot complexes were obvious on the sun's bright disk. The Nermella
gusted strongly and Rachel smiled. It was an expression Colty had come to know and understand.
It pleased him to see it.

``It is so nice to be outside,'' she said. She reached up and gathered her hair into a bunch and
tied it with some kind of clip.

Colty watched with fascination. ``Is that uncomfortable?'' he asked.

``Oh no,'' she said. ``and tying it back keeps the Nermella from blowing it in my face.''

They started toward the library. When they got to the main walkway Rachel turned toward Colty.
``You came at the perfect time,'' she said in a loud voice to be heard over the blowing wind.
``I needed an excuse to be somewhere else.''

``Why was that?''

``Sar. Rocallanon wants me to tell him where stars get their energy, but I can't.''

``You don't know?'' Colty asked.

``Oh, I know,'' Rachel replied. ``It's just\ldots\ I'm not supposed to talk about certain
things.'' She hesitated. ``You see there are some on Earth who feel it is too early for me to be
here. My government only allowed it because I agreed to\ldots\ to be discrete about certain
topics.''

``Your people want to maintain the advantage,'' Colty said.

``No, oh no,'' She stopped and faced Colty directly. ``I'm a scientist and a teacher. I'd love
to tell you everything I know. It's just\ldots\ knowledge can be dangerous. For your own
protection we don't want to teach too much too fast.'' She ran her hand through her hair. ``I
understand this,'' she continued, ``but I don't like it any more than Rocalla does.''

Colty reflected on the odd idea that knowledge could be dangerous to the knower, as the two of
them continued to walk. ``Have you explained this to Rocalla?'' he asked finally.

``Yes,'' Rachel said. ``I've tried. He thinks I'm being\ldots\ um\ldots\ e-va-sive.''

``Evasive.''

They came to one of the gardens along the path and Rachel stopped to look at it. ``Oh look,''
she said as she stooped down. There growing close to the ground where the same tiny red flowers
she had asked about the first hour they met. ``These are called topolia flowers. They bloom
along the vine in sequence from the root toward the end.''

Colty flapped his ears with surprise. ``How do you know that?''

``I read about it. What do you think I do in the library all that time?'' she added with a
smile.

Colty clicked his teeth loudly and Rachel watched him closely.

``That means you are amused, right?'' she asked. ``When you click your teeth together like
that?''

``Well, uh, it can mean that, yes,'' he replied. ``It can also just mean we're pleased with
something or happy to see someone.''

``So if I do it,'' Rachel clicked her teeth together a few times, ``how does that come across?''

``That's a bit eerie, honestly,'' Colty said. ``A little shocking even. I don't think you should
do that.''

The two of them lay on seats next to the garden for a short time. Colty felt surprisingly at
ease. ``Do you have flowers on Earth?'' he asked.

``Oh yes,'' Rachel said. ``Many beautiful flowers. They are not like here though. Plants on
Earth are green but the blooms are different colors.''

``Green plants,'' Colty said softly, trying to imagine it.

The Nermella gusted strongly and Rachel looked up at the rigat branches over their head. She
closed her eyes and appeared to listen to the whistling of the air as it rushed through the
leaves. Colty watched her silently.

``I've decided to give up on stellating co-sequences,'' he said.

She turned to look at him.

``My overseers convinced me that the work is probably a dead end.''

``I thought you told me they had fascinating structure.''

``It seemed like they had fascinating structure. I can't prove any of my conjectures so I guess
I don't know.''

He looked out over the Yellow as he spoke, and then the two of them lay silently for a while.
When Colty turned toward Rachel he found her looking at him.

``What will you study instead?'' she asked.

``Graph extension fields. It's an important topic.'' Rachel didn't reply but seemed to reflect.
He watched her for a moment. ``I don't suppose you know anything about stellating
co-sequences.'' As soon as the words left his mouth he regretted saying them. He didn't want to
sound like Rocalla, and he certainly didn't want to just be told the answer.

``I don't,'' Rachel replied. ``Honestly, I've never heard of anything like them.''

``That's more evidence that they probably aren't worth pursuing.''

Rachel looked at Colty closely, studying his expression. After a moment she asked, ``how do you
feel about this new direction?''

``It's fine.''

``Yes?''

``Sure.''

She made a low humming sound that seemed to come from deep inside her body. It was a sound Colty
never heard her make before. He turned to look at her.

``You don't like doing what everyone else is doing,'' she said.

``What makes you say that?''

``I'm getting to know you.''

Suddenly there was a soft, beeping sound almost like the chirping of a cican bug. Rachel took
from her pouch the device she simply called her ``machine'' and unfolded it. Colty could see its
surface covered with exotic human symbols that she seemed to study briefly. Then she quickly
folded her machine up again.

``A flash is coming.''

Colty tilted his head to look up at the large disk of the sun, shading his eye slightly with his
trunk. There was a rich collection of spots but the glare of light made it difficult to see much
else.

``We should get inside,'' Colty said, ``let's get over to the library.''

The two of them got up and started walking again, more quickly than before. Everyone else went
about their business as usual without human technology to alert them. Just as they reached the
far side of the Yellow the sun started to brighten and the color of the sky started to change.
The hue first went from a rubinum-red to a sandy orange and then to an eerie, otherworldly
green. Many people ran to the shade cast by a tree or a building. Yet others didn't care and
stood quite happily beneath the changing sky.

Colty and Rachel trotted up the steps of the library but Rachel stopped before going in.
Standing beside a buttress to protect them, she turned to look across the Yellow to witness a
Rujaran flash.

Ripples of light flowed across the scene: pink, yellow, and golden orange. The clouds flashed
and flickered with riotous color. Yet the intensity grew and the light became almost purely
white, a harsh glaring brightness that was yet completely silent.

All about the Yellow the bulbs of the rigat trees opened, spreading their triangular petals to
expose the luxuriant red of their pulp to the nourishing light. Thousands of blooms transformed
the non-descript trees into towers of color. Rachel gasped and spoke excitedly in her own
language, more to herself than to Colty.

``It is best not to look too long at the flash,'' Colty said holding his trunk in front of his
right eye.

After several long moments the intensity began to wane. The white light gave way to ripples of
yellow and orange and then even that stopped. Rachel moved away from the wall of the library and
looked up at the sky. Colty joined her at her side. Although the sun was returning to normal,
long streamers of color stretched out across the sky as shimmering, flickering sheets of green,
red, and gold. It would be many arnets before the after effects of the flash were completely
gone. Rachel watched the spectacle transfixed and then finally turned to Colty.

``That was incredible!'' she exclaimed.

``You've never seen a flash?''

``There are no flashes on Earth. The sun always looks the same.''

``You are lucky then. Honestly they just cause trouble. There are always a lot of accidents
during a flash and they interfere with radio transmissions.''

Rachel wasn't paying attention. Instead she turned back to the sky.

\spacebreak

Early the next runion, Colty lay at the desk in his office with student papers stacked neatly in
two piles before him. On his left were the papers he had not yet graded and on his right were
the papers he had finished grading. He scowled slightly at the paper in front of him. The
student obviously hadn't been paying attention in class. How else could he have missed even the
most basic points? Colty held the pen in his tentacles and started writing notes neatly on the
page. As he wrote he heard a knock from the general direction of the door.

``Yes?'' he said without looking up. There was no reply, and Colty finished the comment he had
been writing. ``Can I help you?'' he turned to see Rachel standing patiently in the doorway.
``Oh,'' Colty said, ``come in, please.''

``I don't mean to disturb your work.''

``It's fine,'' Colty replied. ``Any excuse to take a break from grading papers is welcome.''

Rachel knelt down on the seat beside Colty's desk. She was quite close to him but he had gotten
used to her so he didn't mind.

``I came here to give you something,'' she said. She opened the pouch tied around her waist and
took out what looked like a small book. She put the book on Colty's desk where he could see the
title printed in Argenian, \textit{The Mathematics of Logic}.

``What's this?'' Colty asked. The title was intriguing but he didn't recognize it. Also the
author's name was unusual.

``It's a gift. For you.''

``That's not necessary.''

``Please, I would like you to have it.''

Colty clicked his teeth and flapped his ears slowly. He slid the book toward him a little.
``Well then, thank you,'' he said and Rachel smiled. He flipped open the book and started paging
through it. It was filled with symbols he had never seen. ``Where did you get this?''

``From the library on \textit{Golden Light}. It's a classic. I studied from this book myself
when I was a student.''

Like a rush the full impact of what she was saying swept over him. ``Wait. This\ldots''

``Yes,'' Rachel continued, ``it's one of our books.''

``But the text is in Argenian.''

``I had it translated. We have machines that can do most of it. This is technical material,
though, so I did help a little. I'm afraid the translation isn't perfect. It's a shame because
the original is so beautifully written.''

Colty studied one page carefully. The mixture of Argenian words and alien symbols was both
surreal and exciting. ``This is wonderful, Ra'hel. I am very grateful.''

``I should warn you that the material gets fairly challenging by the end. However, I think you
will like that.''

``I do like that,'' Colty said with a click of his teeth. ``What's the point of reading things
that are obvious?''

Rachel laughed lightly.

``I'm surprised you're sharing this, after what you told me,'' Colty added.

``It's fine. Thankfully my government feels abstract mathematics is safe because it is so far
from the applications.''

``Yes, quite useless.'' Colty clicked his teeth again as Rachel chuckled.

``Anyway, they can't complain about what they don't know.'' She added slyly.

After Rachel left Colty settled down again with the student papers. He looked over the paper he
had been commenting to remind himself about the work. He wrote a few more notes on the page and
then circled a line of reasoning that looked approximately correct. He reflected for a moment
and wrote a grade at the top of the paper. Then he neatly added it to the stack of finished
work.

He leaned back and fluttered slightly. There were still quite a few papers to review. He glanced
at the small book Rachel had given him. Again he flipped open the cover and looked at the title
page. He turned to the first chapter and started to read. In his mind he heard words written by
an alien intelligence who lived thousands of light runion away. Yet he understood those words.

\textit{When engaging in a precise discussion about the nature of reasoning itself, it is first
  of utmost importance to distinguish between the logic and the meta-logic\ldots}

