% Local IspellPersDict: aspell.dunari.pws

\chapter{37}

Contrary to his expectations Colty found himself enjoying Rachel's physics class. He had taken a
similar class when he was a student, of course, but it was good to get a review. He even
recognized some of the problems on rotating reference frames and angular momentum that had
appealed to his mathematical mind. Monitoring Rachel's class made him feel nostalgic about his
time at Marlock.

The students, being students, and first term students at that, didn't take to the course as
enthusiastically. When Colty stopped in between scheduled lectures, there were usually only a
few students in the back, apparently chattering amongst themselves instead of working the
problems Rachel left on the board. The pad was usually empty, and as far as Colty knew not one
student had come to Rachel's office. It was understandable, but not entirely acceptable. Colty
felt like he should say something, but he wasn't sure just what.

During Rachel's forth scheduled lecture Colty lay in his usual seat while she went through yet
another electrostatics problem. She carefully transcribed the steps from the course book, but
the class was restless. Some students copied the example into their notebooks but others just
sat and stared more or less blankly. Colty even saw a couple of them in farth-sleep. Several
students spoke to each other in hushed voices, obviously not paying attention at all.

Rachel finished putting the example on the board and then turned to the class. She consulted the
course book again, reading the notes there briefly before she spoke. ``In this example,'' she
began, ``the first electrosource is fixed while the second is attached to this spring-loaded
rocker. As the rocker moves the distance between the two sources change, and so you must employ
both the spring force formula and the electroforce formula in a single balancing equation.''

She looked over the class. ``Does that make sense?''

The students simply waited.

Rachel walked through the math. She set up the balancing equation by opposing the rocker torque
produced by the spring against the torque produced by the electroforce. Using some basic algebra
she then solved the equation for the rocker angle in terms of the electrosource strength and the
spring flexure constant.

``Are there any questions?''

Nobody said a word.

She flipped to the next page of the course book and paused while she reviewed it. She twisted
her head in a very non-dunari way and looked up at the line clock above the board. ``I'm almost
out of time. There are some problems here for you to do. Let me copy them to the board.''

Most of the students dutifully opened their notes and got ready to record the problems. Rachel
wheeled the podium around so that she could easily consult the course book while she wrote on
the board. All was silent in the hall except for the soft scratching sound of chalk against hard
slate. When she finished she turned toward the class again. ``Please work these problems and
submit your answers to the slot for this course as usual.'' She paused. ``That is all for now.''

Aside from monitoring Rachel's class Colty didn't see her very often. He was busy with his own
classes and struggled to make progress on graph extension fields. Yet his mind constantly
returned to the problem of stellating co-sequences. Like a well worn stone in his pouch, he
couldn't help but turn it over and over.

``Isn't it appropriate as a host to invite a visiting professor to share food?'' Joleia asked
him later that runion. The two of them were lying on seats in the library courtyard taking a
short break.

``What are you saying?'' Colty replied.

``I'm just saying you should be a better host.''

Colty looked at her incredulously. ``It's appropriate, if the host and guest are getting along.
However, it isn't expected, and anyway this is a different situation.''

``You and Ra'hel get along. We could invite her to \textit{Juju} this runion. You wouldn't even
have to talk to her much. She could lay with my seleck. She is a woman after all.''

``What would Jijana think about that?''

``She'll love it. Everyone else will probably see it as an adventure.''

With Joleia's prodding Colty soon found himself kicking the door frame of the human's office in
Cogart. The door opened a crack at first but when Rachel saw who it was she invited him in
happily.

Colty was shocked by how cold the space was, and Rachel hastily adjusted the air cooling unit.
``I sorry. I'm more comfortable with the cool air.''

Her office was very sparsely decorated. It was hard to believe she had been at the University
for almost four runion; it looked like she had just moved in. She had a simple table with a
phone and a few empty shelves. Above the table was the photo sheet she bought at the store in
town shortly after her arrival, but otherwise the walls were blank. She had a large window but
with the mesh more closed than Colty liked. The course book for her class was lying on the
bottom tray of her table but there were no other books, no papers, and no notes to be seen.

% TODO: Rachel has her machine out and running...

Rachel knelt on the seat in front of her table, bending her legs under her the way she did, and
gestured toward another seat to the side. Colty just stood.

``I\ldots\ um\ldots\ Joleia and I want to invite you to \textit{Juju}. It's a runional
celebration that some selecks do. We thought you might enjoy socializing a little.''

She looked at him for a moment with an expression he couldn't read. ``I've heard of it. Will
there be many people there?''

``Just Joleia's seleck and few guys like me, companions of the seleckia. No children.''

Rachel looked down and thought for a moment and then, as if abruptly coming to a decision,
looked back at Colty and smiled. ``I'd be happy to come. Thank you.''

The rest of the runion passed uneventfully. Colty and Rachel hardly crossed paths outside of her
class and even Joleia was busy with cataloging new acquisitions. Colty was apprehensive about
Juju with Rachel there. Not everyone accepted her, and he worried that it had been a mistake to
invite her into a social situation with so many who didn't know her.

When Juju came and he arrived at the hall, most of the seleckia were already preparing food in
the hanaria and chattering among themselves. Rocktar was arranging seats around two large,
round tables at diagonal corners of the hall. Jujaran streamers were tied to wall clips around
the space, their bright yellow-red colors creating a festive atmosphere. The ceiling windows
were slid aside letting fresh air and light into the building. The Nermella gusted strongly with
hot, moist air but most of the wind flowed smoothly over the rounded roof and only a gentle
breeze filled the hall to softly rustle the streamers.

``We are well met,'' Colty said in mock formality.

Rocktar clicked his teeth. ``I'm glad to see another male here. I don't think the other two are
going to make it this time.''

Joleia's seleck was new and composed entirely of young women. Even Jijana, the oldest, was only
20 runion Joleia's senior. Only one seleckia was married and three were without companions.
Rocktar was about Colty's age, and by profession a banker in Varnok. Colty helped him finish
arranging the seats.

``We're getting an electronic tabulator at the bank,'' Rocktar begin. He always tried to engage
Colty in matters mathematical. ``It's something Electrodyne produces. it's supposed to be faster
and more accurate than the models from Calctronics.''

``How are these things different than punch calculators?'' Colty asked.

``They're much different. For one thing they can be configured to do different kinds of
calculations\ldots''

``Colty!'' Standing in the doorway to the hanaria Joleia beckoned to him with her trunk.

Leaving Rocktar to finish the seating, he followed Joleia into the hanaria. There platters of
toran root and flannen where spread over the counter. The smell of roasted barak filled the air.
Down the long, narrow space the seleckia chopped, scrapped, rolled, and mashed the various
components of their Juju feast. When Colty entered Jijana paused and faced him.

``She's bringing her own food, yes?''

``What is this creature like?'' came a voice from the back.

``She's very soft spoken and polite,'' Joleia called in return.

``One of Jurita's wonders,'' came another voice.

``That may be,'' said Rocktar. He was standing directly behind Colty, ``but does that mean we
have to share food with it.''

Preparations continued and before long Colty found himself lying at one of the tables with
Rocktar and Polis. The new arrival was Jimlara's husband, and a technician at the Parnon medical
center. Colty never knew quite what to say to Polis but it mattered little as Polis seemed happy
to talk even to stone. Rocktar tried to interest Polis in tabulation machines while Colty
watched the clouds, brightly lite with rubinum, race past the ceiling windows. They seemed to be
moving from the north, an unusual direction and a sign that a worstora was on its way.

``She is here!'' Joleia called out. The men and several seleckia turned to see the tall, slender
figure of the human standing tentatively next to the door frame of the main entrance. Joleia
went to greet her and Rachel, recognizing her, came forward. She was wearing light tan clothing
trimmed with gold that left her arms exposed but that covered her legs. Around her neck she wore
a kind of silver chain that glinted even in the subdued light. Her hair was tied back, as usual,
and she carried a small translucent container made of some unknown and perhaps unknowable
material.

``I wasn't sure if this was the right place,'' she said in her soft voice.

``Oh yes,'' replied Joleia.

Everyone gathered around as introductions were made. Rachel looked very intently at each person.
``I don't think I'll be able to remember all the new names,'' she said.

``We should make name cards,'' Jimlara said.

Joleia turned to Rachel, touching her arm briefly with her tentacles. ``How do you pronounce
your name again?''

``Most people call me Ra'hel.''

``Ah,'' Jijana said, ``but this is Juju. Perhaps we should call you Jahelle.''

Rachel tilted her head slightly in a remarkably dunari gesture and then, in a uniquely human
way, she smiled, careful to keep her lips together so as to avoid alarming her hosts. ``I would
like that.''

Food was brought forth and laid out upon the tables. Joleia got a plate for Rachel. Opening the
container she had brought, and much to the interest of all, Rachel took out her food and
arranged it carefully on the plate. She had several large, red berries and something like
garconal covered with a thick spread. She also had what looked like nuts and a small cake of
some kind. Her meal seemed at the same time both exotic and mundane.

``Those nuts look good,'' Jernum said.

``Everything looks good to you,'' Jimlara said clicking her teeth.

``You mustn't eat them,'' Rachel said seriously. ``I was told they contain substances that are
poisonous to you.''

Rachel's food was taken to the table and set before her place, along with everyone else's plate.
Finally, standing in the center of the room beneath the largest ceiling window, Jijana called
the feast. The group assembled around her, lying on the ground. Rachel knelt beside Joleia
approximately to Jijana's right. Colty, Rocktar, and Polis lay in the back, largely out of the
way but still part of the group.

Moments before the room had been filled with the sounds of talking and the bustle of activity.
Yet as the group took their places around Jijana they grew solemn. Rachel clasped her hands on
her lap and waited.

Jijana raised her trunk toward the cloudy sky above, tentacles open. She then began to sing. Her
voice started soft and low, like a distance rumble, but grew in intensity and pitch until it
became a strong undulating hum. The other selekia joined her, one at a time, their well
practiced voices both in opposition to and harmonizing with Jijana's lead.

Colty watched Rachel as she sat quietly taking in the experience. Her eyes moved to each singer
as they joined. Finally Rachel closed her eyes as the combined song of all seven women washed
over them and through them to vibrate into their very bones.

Abruptly the sound stopped and, except for the Nermella rushing over the rooftop, silence filled
the room. Nobody moved. Jijana called out, ``We proclaim the moment of Jujar. May all of us
share in her peace. May we bring her peace to all whom we touch.''

``Our love is my love,'' Jijana said loudly.

``Our strength is my strength,'' the seleckia replied.

``Our wisdom is my wisdom.''

``Our spirit is my spirit.''

``Our home is my home.''

``Our patience is my patience.''

``Our truth is my truth,'' Jijana finished.

Jijana lowered her trunk and again all was silent. One moment passed, and then another. Finally
Jijana said, ``Let's eat!''

The seleckia made their way toward one table while the men made their way toward the other.
Joleia again touched Rachel's arm and lead her toward the table with the women. Rachel looked
over her shoulder briefly.

``They eat separately?''

``Yes,'' said Joleia, ``don't worry. It's fine.''

``Once the seleck is more mature,'' Jijana explained, ``there will be more men and things will
be better balanced.''

``Who is Jujar?'' Rachel asked.

``She is Fifth of the Seven,'' Joleia said, ``she created community.''

``Before Jujar,'' Jijana continued, ``we were nothing more than animals.''

They took their places at the table. Rachel knelt between Joleia on her left and Jernum on her
right. The plate of alien food she had brought was carefully set in front of her. Nobody moved
to eat and Rachel again waited patiently.

``We bless the food first,'' Joleia quietly explained. ``Don't worry, it will only take a
moment.''

When all were ready Jijana raised her plate before her. Everyone else, even Rachel, did the
same. Then, as if on cue, the seleckia began to chant.

``\textit{Asam nolan}.''

Each person put their plate down in front of the person on their right.

``\textit{Nolan fornor}.''

Everyone then then picked up the new plate that had appeared before them.

``\textit{Fornor itok}.''

Everyone again passed their plate to the right.

``\textit{Itok jinok}.''

The pattern repeated with each phrase of the blessing flowing from the phrase before. Rachel did
not know the words but she participated in the movement, caught up by the energy that seemed to
swirl around the table. The growling rumble of the dunari voices and the fluid, choreographed
motion had an almost hypnotic quality, and Rachel found herself naturally staring upward through
the ceiling window toward the ruddy clouds.

Finally after eight passes everyone's plate was once again returned to them. The chanting
stopped as abruptly as it had started and Jijana said softly in Argenian, ``We thank you,
Jurita, for this food.'' Then, suddenly, the eating began and the group burst forth with social
chatter.

``How's that telescope coming?'' Jattaka asked from across the table.

``Actually, it's ahead of schedule.''

``You \emph{are} an advanced race,'' said Jernum and everyone in the vicinity shook their ears.
Rachel smiled, not worrying so much about her teeth.

``We've even decided on its name,'' Rachel said.

``It didn't have one before?''

``It had a project designation but not a real name. The tradition is to name telescopes only
after construction has begun so a name can be found that suits the flavor of the place where it
will be located.''

Jijana bobbed her head slightly. ``So what name have you chosen?''

``It will be called \textit{Nitooli},'' Rachel said.

The seleckia looked at her questioningly.

``It's actually a name Tusk\ldots\ Tuskara suggested. A nitooli is a holy man in a certain tribe
of nomads from before the time of the Regents,'' Rachel explained. ``When taken by urges from
beyond, the nitooli would climb the highest mountain in the vicinity and stare out across the
desert for many long moments until he received a vision from the gods.''

``And so also your telescope will stare into the sky,'' Jijana said, ``are you expecting any
visions?'' The seleckia shook their ears.

``I don't know what we'll see,'' Rachel answered seriously, ``that's what we're here.''

Rachel continued with answering questions about the telescope and about her work as an
astronomer. Jijana eventually raised her trunk showed the seleckia a small ornately decorated
ball with a flattened side so that it would rest steadily on the table. Several of the women
bleated and there was some scattered stomping.

``It's the Juju stone,'' Joleia explained. ``Our truth is my truth.''

Jijana held the stone in her tentacles and pressed it firmly onto the table next to her plate.
She cocked her head slightly. ``I see\ldots\ a tall rigat tree covered with mosen flowers\ldots\
growing in the Parnon trash dump.'' She paused and let go of the stone.

``\textit{Har-za!}'' the seleckia cried out loudly.

``You share with the group,'' Joleia continued softly to Rachel, ``some wondrous work of Jurita.
It could be anything, but it has to be something real; something you've actually seen. It
reminds us there are miracles around us every day.''

The stone then moved to Jimlara on Jijana's right. She tapped the stone lightly a few times.
Then with a click of her teeth she said, ``I see Polis finally finishing the new walk to our
house.''

``Har-za?'' Joleia said in her normal voice and the seleckia shook their ears.

Jimlara gripped the stone more firmly and pushed it against the table. She thought for a moment.
``I see children rolling in the grass, unconcerned about their future.''

``\textit{Har-za!}''

As the stone was passed around the table the group continued to eat and talk. Some of the
visions were serious and deep. Some were silly and funny. All were seen as works of the creator
of the universe. After each vision everyone, even Rachel, shouted ``Har-za'' with increasing
power and clarity.

When the stone reached Joleia she held it in her tentacles for only a moment before pressing it
to the table.

``I see the light shining through the library windows and over the books, tables, and people. I
see quiet, warmth, and safety.''

``\textit{Har-za!}''

Rachel held out her hand and Joleia passed the stone to her. She took a deep breath and pressed
it to the table.

``I see a waterfall in a lush forest spilling into a deep, clear pool.''

``\textit{Har-za!}''

After everyone finished eating, Jijana called the closing blessing and the seleckia then busied
themselves cleaning up the remains of the meal. Rachel helped by moving plates and other dishes
into the hanaria and by packing her scraps back into the translucent container she had brought.
She spoke briefly with Colty but was soon swept back into the conversation with the other women.

Once things were in order, Jattaka took out her tinar and started playing a soft, bubbly melody.
People gathered around in smaller groups, some lying on seats, some standing, talking together.
Rachel found herself with Joleia, Colty, and Jimlara. She knelt on a small pad that had been
found in a closet while the dunari stood so that their faces were all at about the same level.

``That was a lovely vision,'' Joleia said to Rachel, ``when you had the stone. Is that a place
near Parnon?''

Rachel smiled slightly. ``Actually it's a place on Earth not far from the home my parents had in
the mountains when I was a child.''

``Then Earth must be beautiful.''

``Parts are.''

``Ra'hel, I hear you are a teacher,'' Jimlara said.

``Jimlara teaches also,'' Joleia explained, ``at the underschool here in Parnon.''

``How is it going?'' Jimlara asked.

``It's fine, I guess.'' Rachel glanced at Colty. ``Actually I don't feel like it's going that
well.''

``You're doing fine,'' Colty said. ``You are very precise and that's good for them to see.''

``What's not going well?'' asked Jimlara.

``The students don't seem very engaged. Also the style of teaching feels unnatural to me. I'm
just not used to the way you do it here.''

``Why not teach the way you do it?''

Rachel hesitated and Colty replied instead. ``The university has certain traditions.''

``Maybe her approach is better,'' Jimlara said. ``It's an incredible opportunity for your
students to have Ra'hel as a teacher. I don't understand why the university doesn't encourage
her to do whatever she wants.''

Colty flapped his ears slightly, ``To be honest, I don't think it was ever discussed.''

Joleia put her trunk over Colty's back. ``He is such an academic sometimes.''

The conversations continued and shortly gargon nuts were passed around. Joleia offered some to
Rachel but she declined. ``I can't risk eating any of your food without it being analyzed
first.''

As time passed the sky grew gradually cloudier and darker. The Nermella increased in strength
still blowing from the north. The gathered dunari became uneasy, looking frequently out the
ceiling windows at the mottled, rubinum clouds rushing past.

``We should close the windows,'' Jijana called out, ``a worstora is coming.''

Rocktar took the hook from the wall by the entrance and started to slid the windows to their
closed position. Standing on the tips of her feet Rachel was able to just reach the window
handles without the help of a hook and assisted with closing the many windows around the room.

``Can we get some artificial light, please?'' a voice called out.

``I don't think this building has any,'' Jimlara said. We'll have to go outside, under the
canopy.

At that moment a low rumble enveloped the hall and seemed to vibrate the air around them.

``My trunk sash!'' Jattaka called out, ``I think I left it in the hanaria.'' She took a step
toward the hanaria door but hesitated.

``You can pick it up after the storm clears,'' said Jijana.

Jattaka took another tentative step but then wrapped her trunk around her shoulders and
fluttered softly.

Rachel was standing near her. ``What's wrong?''

``I don't want to leave it, but it's too dark in there now.''

``I'll get it for you.'' With that Rachel sprang forward and disappeared into the gloom beyond
the hanaria door. She was only gone a moment before returning to the grateful Jattaka with her
sash.

Following Jijana the group went out onto the walkway in front of the building. There, under a
rigid canopy the Nermella gusted strongly yet the overcast light of the Rujaran sun was brighter
than the light inside. Rachel was wedged between Jattaka and Joleia, huddled against the
building. Along the street other people were similarly gathered under other canopies, short
roofs, or entrance arches.

Just after everyone assembled, a bright, searing flash split the sky followed by a crack and a
powerful explosion of thunder. The seleckia shifted uneasily, and Jattaka muttered a brief
prayer just as a second flash filled the landscape with its white light.

``A worstora,'' Joleia said leaning toward Rachel. She spoke loudly to be heard above the
rushing wind. ``Do you have anything like this on Earth?''

``Actually, yes,'' Rachel replied, smiling slightly.

In only a moment or so the rain started. It came heavily, roaring down from the sky in great
sheets of water. Thunder shook the air and lightening flashed again and again. Soon the road in
front of them was a flowing stream. The seleckia chanted in their low voices, rumbling almost
like the distant thunder.

Finally the rain abated and the clouds thinned. The sky brightened and the dunari grew more at
ease. All the people returned inside the buildings and the walkways cleared. Pools of water
spotted the street.

Rachel paused at the doorway to speak with Colty. ``I need to get back to the compound. This is
really the middle of my usual sleeping time.''

``Of course.''

``I didn't get a chance to talk with you much.''

``I should have told you what Joleia had in mind,'' Colty replied.

``It was fine, wonderful really. But I did want to give you this.'' She took a small, loosely
wrapped package from her pouch and held it out to Colty.

Deftly he cradled the package in his trunk and used his tentacles to pull open the covering
paper. Inside was another small book. The title glimmered in the brightening sun, \textit{A
  Theory of Numbers.}

``Joleia told me you were almost finished with the first one.''

``Thank you, Ra'hel,'' Colty said.

Rachel put one arm on Colty's back and patted him softly with her hand. ``I hope you enjoy it.''
