% Local IspellPersDict: aspell.dunari.pws

\chapter{41}

Marvia Base was a few karnons outside Varnok on the north side of the Varsynthia valley. Unlike
most bases, Marvia was primarily administrative, holding the military's main offices. It had
basic defenses, including a section of elite fighters intended to protect the base and also
Varnok from Forbin air attack. Situated deeply in Argenian territory Marvia had never been
reached by Forbin forces in Argenian history but in such matters caution was always indicated.

A military car pulled up to the gate at Marvia and was greeted by the guards on duty. After a
few hasty words it was let through. The car drove between the well maintained two story
buildings and around a neatly landscaped rotary at the far end of the base. From there it went a
short distance beyond the main campus to the athletic building where it parked next to a half
karnon oval running track.

A solder wearing the uniform of the Messenger Service climbed out of the car and trotted over to
General Fotkey's personal secretary who was standing beside the track. On the far side of the
oval, wearing a tight running jacket and moving at a gallop was the general himself. His
secretary held a stopwatch in his tentacles.

``Urgent word for the general, sir,'' the messenger said.

``The general hates to be interrupted during his run,'' the secretary replied. He didn't look at
the messenger but instead watched Fotkey intently as he rounded the end of the oval.

``This can't wait, sir,'' the messenger said. With that he immediately began to gallop toward
Fotkey.

``Oh, you are going to regret that!'' the secretary called out after him.

The messenger intercepted Fotkey just as the oval started to turn back toward the gate. They
were too far away for the secretary to hear what they were saying, but he could see Fotkey
waving his trunk in annoyance at first. Then he stopped. The two talked for a moment and then
Fotkey turned and started galloping directly toward where the secretary was standing. When he
got into shouting distance he called out.

``Make my aerocar ready,'' he said.

``Yes, sir,'' the secretary replied.

``I'm going to change,'' Fotkey said in a more normal voice as he reached the secretary. ``Pull
my car around back and get me to Parnon at once.''

``Where are we going?''

``Shunalia.''

In the southern part of Argenia, not far from the Sumar Sea the broad valley of the Shunal
stretched north and south for many karnons. On the west side of the valley rose the Tukonook
Mountains, the second highest range on Rujar known. Their rocky ramparts blocked the Nermella
and enclosed the moist air from the sea making a region of unusual lushness considering its
location in the middle day. Clouds often formed in the Shunal valley and the unnaturally still
air was thick with humidity.

Fotkey's aerocar landed at a makeshift aeroport in Shunalia, the largest town in the valley. He
was greeted by a small entourage of military personnel including Yutar, an investigator in the
Security Division.

The general had only two words for Yutar: ``Show me.''

Fotkey was taken by military car to a small base outside the city. On the way he was briefed.
``They left the checkpoint here at Shunalia at hour 62:065,'' Yutar said. ``The checkpoint at
Hark's Corner radioed here at 63:010. They were twenty arnets overdue then.''

``Why did Hark's Corner wait a full twenty arnets?'' Fotkey growled. He fluttered loudly.

``The roads are sometimes slow. This is fregon migration season so it's not unusual to have to
wait for a herd of them to pass,'' Yutar explained. ``After that call, both Shunalia and
Hark's Corner sent out trucks. That's when they found them.''

They arrived at the clinic and Fotkey quickly climbed out of the car and trotted to the doorway.
He was greeted by the physician on duty.

``This way, sir,'' the physician said.

In the back room, lying on surgical tables were the bodies of two, young Argenian solders. They
were covered with a long sheet, displaying the symbol of their country. Fotkey stepped up to the
first table and pulled back the sheet. The body had been laid out neatly but a large hole had
been blasted into the side of the head. The sheet was stained with orange blood still seeping
from the wound.

Fotkey fluttered. He pulled the sheet back further to find more wounds in the solder's chest and
back. His clothes were stained with blood and matted to his body.

``We've recovered a couple of the bullets,'' Yutar said. They are standard issue Forbin
autorifle bullets, thirty-five caliber.

``Thirty-five?''

``Yes, strange,'' Yutar replied, ``Harkite War vintage.''

Fotkey covered the solder with the sheet again. ``Clean this man up.'' He walked to the next
table and pulled back the second sheet. As with the first solder, the second had a large hole in
the side of his head. He also had smaller wounds on his shoulder and front leg. Fotkey fluttered
again.

``The wounds on his legs don't look serious.''

``No,'' the physician said. ``They probably hurt. He would have fallen down but they didn't kill
him. It was the shot to the head that did that.''

``Execution style,'' Fotkey said softly. ``Show me the truck.''

Yutar led Fotkey out of the clinic and then several paces to the garage a short distance away.
Inside was a rather non-descript truck of the kind you might see anywhere. It had Transportation
Agency markings and looked like a road construction vehicle. Fotkey went to the back.

``Open it.''

They swung the doors wide and Fotkey looked in. There wasn't much to see. In the middle of the
cargo area was a low frame designed to hold the device but it had obviously been cut.

``They had torches with them,'' Yutar explained. ``They knew what to expect.''

``So we have a leak,'' Fotkey said softly.

``No doubt.''

``Seal the Rangard base immediately,'' Fotkey said. ``Nobody gets in or out until we resolve
this.''

``Already done,'' said Yutar.

Fotkey climbed into the truck to get a closer look. The frame had been cut with obvious
efficiency. Only the minimum number of braces that needed to be sliced where touched.

``We found the button on the floor here,'' Yutar said.

The men climbed back out and Yutar led the general to a small table with several items laid out.
Yutar opened an envelope and took out a simple button with some thread still attached.

``It's Forbinite,'' Yutar said, ``but the strange thing is that they don't use this style on
their uniforms any more.''

Fotkey studied the button closely. He seemed disgusted.

``It may be that we are dealing with a rouge faction.'' Yutar said. ``They are using old
ammunition and old clothing. This might not be the work of the normal Forbin military.''

Fotkey did not reply at first. He turned the button over in his tentacles studying it with his
right eye. ``If I were doing an operation like this,'' he said at last, ``I wouldn't wear my
uniform at all.''

``You think it is a set-up of some kind?''

Fotkey passed the button back to Yutar. ``You tell me.''

After examining the truck Fotkey insisted on going out to the scene. ``There isn't much there
now,'' Yutar explained. ``We've moved everything back here for detailed study.''

``Nevertheless I want to see it,'' Fotkey replied.

The truck had been found at the end of a short service road and behind a small rise that blocked
it from the highway. Yutar explained that it looked like they had been stopped on the main road
and killed there. The attackers probably then drove the truck out of the way so they could
remove the device.

``That's bold of them,'' Fotkey said, ``to attack like that on the open highway.''

``There isn't a lot of traffic to Hark's Corner,'' Yutar explained.

``The attack must have been unexpected,'' Fotkey said looking down the road to get a feeling for
the terrain. ``The drivers are trained solders of the Argenian military. They wouldn't have just
allowed uniformed Forbinites to approach freely.''

``Perhaps the attackers were disguised in some way.''

``Then what of the button?''

Yutar showed Fotkey where the truck had been parked, slightly off the road, and pointed out
additional tire tracks that were clearly from another vehicle. ``It looks like another truck,''
he said, ``probably smaller than the one we used, a sand carrier perhaps or a naraza transport.
There are a lot of naraza ranches in this part of the Shunal so a vehicle like that would be
inconspicuous.''

Fotkey scanned over the mountains to the west, in the direction of Forbin. The rocky peaks of
the Tukonook rose up through the hazy air and towered over them. Their flanks covered with
yellow forests touched with red.

``Is there a way over the mountains here?'' Fotkey asked.

``No, sir,'' Yutar said. ``The nearest pass is Palarbia at least 100 karnons to the north and
that's a very rough road. I doubt you could get a naraza transport, or anything similar, over
it.''

Fotkey studied the mountains with a grim expression. ``What about an aerocar? Could the
Forbinites have one that could lift something that heavy over the ridge?''

``We'll have to ask the Intelligence Service. I'm skeptical, but it's our best guess right
now.'' Yutar paused and then added. ``I'm sure the humans could do it.''

Fotkey fluttered loudly. ``And why would the humans be interested in delivering a secret
Argenian weapon to Forbin? As far as we can tell they don't even know about it.''

``The Forbinites aren't supposed to have known about it either,'' Yutar reminded him.

Fotkey turned and studied the long, lonely highway again. As he watched a single car sped down
the road. It was an older, private vehicle, probably on its way from Hark's Corner with business
in Shunalia. It was the first vehicle they had seen since arriving. Fotkey followed the car with
his eyes as it receded into the distance and then he looked to the east, in the direction of
the Argenian heartland.

``Something about this doesn't feel right to me,'' he said at last. ``This is a fine spot to
mount an ambush but a terrible spot to get the spoils back to Forbin.''

``You suspect it is not the Forbinites?''

``Old ammunition\ldots\ old clothing\ldots\ it's as if they rummaged around in Forbinite cast
offs looking for something to throw us off track,'' Fotkey replied. ``How far is the Sumar Sea
from here?''

``Not far,'' Yutar said. ``It's only 50 karnons from Shunalia to the port of Ikatty.''

``And from there by boat to Hazenfat or Nuk,'' Fotkey said.

``Yes, sir.''

Fotkey thought for a moment. ``I will, of course, need a full report from you as well as regular
updates.''

``Of course,'' Yutar said. ``We've been searching the area for evidence of an aerocar and we've
been talking with people in Hark's Corner and Shunalia but, unfortunately, we haven't come up
with much yet.''

``Continue your efforts,'' Fotkey said. ``In the meantime get me back to my plane. I have work
to do.''

They drove back to the base in relative silence. Fotkey pondered all the evidence he had been
presented. When they arrived a young communications officer was waiting for them.

``General, Inspector,'' he began. ``We just received word that an officer at Rangard has
committed suicide.''

``Who?'' Fotkey demanded.

``Captain Dolappit.''

``Dolap!'' Fotkey repeated.

``There's more,'' the officer said. ``The investigators there found encrypted Forbin messages in
his personal materials. They're working on decoding them now.''

Fotkey swayed his head slowly. ``Thank you,'' he said. ``You are dismissed.''

After the officer left Fotkey and Yutar walked back toward Fotkey's aerocar. ``Isn't it
convenient,'' Yutar said wryly, ``for the source of the leak to identify himself like that,
complete with evidence incriminating Forbin?''

``Too convenient,'' Fotkey replied.

\spacebreak

Colty found his usual seat and waited patiently for the scheduled lecture to start. The
students, as they always did, talked quietly or got their notes on electroforce ready. Yet when
Rachel finally appeared she was not carrying the course book. Instead she held a walkway brick
from the resurfacing project outside Tartuk Hall.

She stepped lightly up the steps to the platform and carefully put the brick on the floor out of
the way. The students found their seats and grew quiet as Rachel, without looking at the class,
wheeled the podium off to the side, pushing it up against the wall by the end of the board. She
then grabbed hold of the metal table, wrapping her hands around the legs just under the top and,
leaning back, dragged the table to the middle of the platform. The students watched, puzzled by
her strange behavior.

Finally, when everything was arranged to her satisfaction, Rachel stood in the center of the
platform immediately behind the table.

``Shall we begin?'' she asked.

She raised one arm into the air above her, fingers straight, waiting. The room grew silent.
Nobody dared to speak or move and all eyes were on her. After a moment, and without a word, she
curled her fingers into a tight ball and then, bending her knees slightly, brought her fist down
against the table with a rapid stroke. An explosive, metallic bang echoed throughout the hall.
Everyone was startled. Many cried out from the unexpected suddenness of it.

``Why didn't my hand go through the table?'' Rachel asked.

The students shifted, uncertain if they should answer. Rachel walked to one end of the platform
and spoke directly to the class. ``It seems like a simple question. Surely you must have some
answer.'' She waited.

Suddenly a voice called out from the other side of the hall: ``Because you didn't hit it hard
enough!'' There was a flurry of shaking ears and Rachel laughed lightly. She walked across the
platform and looked into the class.

``Who said that?''

With reluctance a student about three quarters the way back, along the wall raised his trunk.

``What's your name?''

``I'm Hiptarik.''

Rachel nodded her head slightly. ``A good answer Hiptarik, but not the one I'm looking for.''
She turned to pick up the brick, and put it on the center of the table. Then she knelt on the
platform and looked carefully at the brick by peeking just over the edge of the table. She
reached up and touched the top of the table with one hand.

``It's made of atoms, yes? And so is the brick. We know atoms are mostly empty space.'' She
stood up again and, putting both palms on the brick, leaned forward and pressed down on it with
almost her full weight. ``Why doesn't the brick just pass through the table, its atoms slipping
between the those of the metal, and then after that, through the floor and eventually all the
way to the center of Rujar?''

She raised the brick above her head as if to fling it against the table top. ``Maybe it just
needs a little help.'' Yet instead she hesitated and then set the brick back down. ``Well, you
get the idea.''

Nobody said a word.

``Jilarnia---you have a lovely name, by the way---what do you think?''

A young woman a few rows back in the center started. ``Me?''

``Yes, you.''

``I\ldots'' she looked around anxiously, ``I don't know.''

``Imagine yourself as a very tiny mite crawling on the table top. You are so small you can
actually see the atoms.'' Rachel knelt down again and pointed at the where the brick sat
unmoving. ``What do you see?''

``Little balls, I guess, kind of pressed against each other.''

Rachel stood up and scanned over the class. ``Ticgon are you here?''

``Um, yeah.'' A trunk was raised on Rachel's left.

``What are these balls Jilarnia spoke of?''

``Atoms.''

``Yes, and atoms are\ldots''

A voice called out from the right. ``It's the chemical bonds. That's why the brick doesn't go
through the table. The bonds hold the atoms together.''

``Thank you, Hiptarik,'' Rachel said, ``but please don't interrupt the thoughts of others.''

``Sorry.''

``The bonds give materials their strength,'' Rachel continued, ``but what are these bonds,
exactly? Ticgon?''

``Ah\ldots\ well they're little wooden sticks in chemistry class.'' Again the class shook their
ears.

Suddenly an annoyed voice came from the back, ``What does this have to do with electroforce?
Isn't that the topic of today's lecture?'' The student was only a few seats down from Colty.

Rachel tilted her head. ``You think my question has nothing to do with electroforce?''

``It's the atomic particles,'' came another voice, this time from near the front. ``The black
electrosources orbiting the atomic cores repel and prevent the atoms from passing through each
other.''

``That's crazy,'' said the student from the back. ``That's a tiny force. It would be
negligible.''

``Yeah but there are many, many atomic particles per square area. The total repulsion would be
a lot. Besides, what else could it be? It's not magic.''

Rachel leaned back against the table and crossed her ankles without saying a word.

``But then,'' the student in the back continued, ``if you pushed hard enough the brick would go
through the table, yet it never does no matter how hard you push.''

``Why not let Ra'hel explain it?''

The class grew silent and all eyes turned to Rachel. She smiled slightly. ``You are doing a fine
job without me. Yet perhaps the time has come for us to explore the plausibility of this idea
with some calculations.''

She went to the board and began to write down some figures and formulas. She did not consult the
course book because there was no course book to consult. Instead starting from first principles
she simply derived the equations she needed, pausing now and then to ask the students how to do
certain steps. Finally she took a spin dial from her pouch and did some quick computations,
manipulating the dial without its stand by holding it in one hand while she used her other hand
to operate the pointer leg.

As the class continued Colty noticed that Rachel asked far more questions than she answered. It
was as if she pulled knowledge from the students that, in some sense, they already had. Her
questions guided them toward the right conclusions, but ultimately they were conclusions the
students derived themselves.

Rachel finally summarized the lecture standing in the middle of the platform out in the open.
There was no podium to stand behind. She spoke with an energy and passion that Colty had never
heard her use.

``We have seen that the electroforce is what gives materials their rigidity and strength. We
have also seen that it is the electroforce that provides the energy of chemical reactions\ldots\
reactions that drive your vehicles and power our bodies. If the electroforce did not exist or
behaved differently than it does, the universe would be a very different place.''

Rachel leaned forward and held one hand out to focus the class. ``Yet the real lesson here today
is about the importance of intuition. The equations and calculations are important too, but to
understand the universe you must feel it, you must live it, as a tiny mite or as a giant sun.
You must try to know the universe as Jurita does, and see it as She can see it.''

She paused and the class waited attentively. ``That is all for now.''

``Are there problems?''

``Oh, ah, finish reading the section in the text about the electroforce. I'll put some problems
on the board before I leave.''

The students filed out and Rachel sat down on the top step of the platform. She took a few
breaths from her oxygen canister. Colty approached her just as the last students were leaving.

``It was different,'' he said.

``I just hope I don't get in trouble for it. I guess you'll be telling the department.''

Colty clicked his teeth. ``Only if they ask.''
