% Local IspellPersDict: aspell.dunari.pws

\chapter{43}

Fotkey slid open the door to the conference room on the second floor of the Argenian Executive
Offices. The First and Second citizens were there, lying side by side at the end of a large,
polished black rosenwood table. The table was lite by reflecting panels mounted on the ceiling,
catching the sun.

Fotkey stood at attention. ``I'm sorry I'm late.''

``You're not late,'' the First Citizen said, ``but you know that.'' He gestured toward a seat at
the opposite end of the table. Fotkey put his pouch on the smooth surface before lying down.
When Fotkey got positioned the First Citizen spread his tentacles in a gesture of beckoning.

``The investigation is continuing,'' Fotkey began. ``The encrypted messages found in Captain
Dolappit's belongings have been decoded. They used an older Forbin code that the MIS has already
mastered.''

Fotkey paused but nobody said anything. ``The messages are in Forbin. Most are from an operative
providing instructions for gathering information about the work of Sar. Parnez, especially, but
also of the other scientists working on the weapon.''

% TODO: Maybe reference Sar. Parnez earlier in the story in some way?

The Second Citizen rubbed his face with his trunk and tilted his head slightly. ``Would Dolap
have access to those notes?''

``Yes. He was a logistics aide on the project. He would have been close to all three of the Sars
on many occasions. The messages also reference the schedule for moving the devices. It seems
Dolap sent that information to his contact but we have none of those messages.''

``What do you know about this operative?'' the First Citizen asked.

``Nothing,'' replied Fotkey, ``we have no intelligence about him at all. Of course the
Forbinites have been cracking down; several of our agents have been recently jailed\ldots\ or
worse.''

``What I want to know is how Dolap got through our security. He had high level clearances.''

``My special investigator is looking into that. We've been trying to trace Dolap's activities,
find out who he's been talking with, but it takes time. So far he appears to have been a model
citizen. He had a wife in a reputable seleck, a family, a long, distinguished career in the
military, and was active with the Limzar Association.''

``You must continue your investigation, of course,'' the First Citizen said, ``but right now we
need to worry about the future, not the past. General, put that military mind of yours to work
and tell me what the Forbinites are trying to do.''

Fotkey thought for a long moment. Finally he tilted his head slightly to glance at the Second
Citizen and then he looked at the First Citizen with his right eye. ``I believe the Forbinites
want to study the device. The last we heard their program was well behind ours. They want to
catch up and, if they can, using Parnez's notes, jump ahead.''

``So you don't think they plan to use the weapon?'' the First Citizen said.

``I don't think so. At least not now. They want to replicate the technology and extend it. To do
that they'll need the device in one piece.''

``Do you think there's any chance they'll be able to complete Parnez's work?''

The room fell silent.

``I think it's doubtful,'' Fotkey said finally. ``None of their scientists have the equal of
Parnez's genius. They wouldn't even have a program if it hadn't been for that leak\ldots''

``We shouldn't underestimate them,'' the Second Citizen interrupted. ``They are ruthless. If by
some chance they manage to build second generation devices before we do, I fear they won't
hesitate to use them at once.''

The First Citizen pondered. He leaned back slightly and studied both of the other men. ``Should
we be considering a first strike?''

``That would mean certain war, of course,'' Fotkey said.

``And let's not forget they have one of our devices now,'' added the Second Citizen.

``They'll likely have more than one if we give them much time,'' the First Citizen replied.

Fotkey swayed his head briefly. ``Sirs, I must add that I have concerns about the evidence we've
gathered so far. It doesn't make sense. An operation like this would be at the highest levels of
Forbin intelligence. Yet they used obsolete codes in their messages. Why didn't Dolap destroy
the messages he received? Any sensible protocol would require that. It's as if we were intended
to read them.''

``What are you saying, General?'' the First Citizen asked.

``I'm saying that maybe the Forbinites didn't steal the device at all, and whoever did is trying
to frame Forbin.''

``That's crazy,'' said the Second Citizen, ``who else could possibly have stolen it? The
humans?''

``I doubt they did it.''

``Then who?''

The First Citizen leaned forward. ``General, if you're right then executing a first strike would
be the worst mistake we could make.''

``And if I'm wrong\ldots'' Fotkey continued.

\spacebreak

Colty closed the door of his office to shut out the bustle in the halls of Fargon. He wanted to
concentrate and for that he needed some quiet without the threat of immediate interruption. He
returned to his table and clicked on the artificial light. Spread out before him where the
unbound pages of \textit{A Theory of Numbers}.

Colty studied the text carefully, reviewing each proof closely. The style was incredibly formal
with very specific rules describing even the presentation. Yet the creativity and imagination of
the original mathematicians still shone through, mathematicians with strange alien names like
\textit{Euler}, \textit{Hilbert}, and \textit{Ranjana}. Colty felt a certain kinship with them.
He recognized many of the problems they had solved and admired their attention to detail and
desire for elegance. Mathematics was more than science. It was art. He could see the humans
understood that too.

Ordinarily Colty found flaws when he read the mathematical texts of the dunari. Minor omissions
or invalid steps, usually easily patched, peppered the published works. Yet in the human text
there were none. The proofs were perfect.

``They're checked by machine,'' Rachel had explained.

Colty looked at the exercises at the end of the chapter. The first several were trivial. He
scanned to the bottom of the list looking for the ones marked with stars. Opening his notebook
to a fresh page, he worked through several of the one-star exercises. They weren't difficult but
he enjoyed the chance to practice human mathematical technique.

On the next page a three-star exercise caught his eye. He read it twice and then put his pen
down. Leaning back he reflected for a moment. He clicked his teeth slightly. ``Interesting,'' he
said aloud. He wrote some ideas into his notebook and then stopped. Looking at the problem again
he flapped his ears slowly and fluttered.

Suddenly he had an idea. Twisting in his seat as best he could, he scanned over the books behind
him. Reaching his trunk out and using two tentacles, he slid one from the shelf. He opened it on
the table and flipped through the well worn pages until he found the section on Garacken
Reduction. He scratched his shoulder slightly as he re-read it.

Colty set the book aside and started on a clean page of his notebook. It seemed as if the
three-star exercise could be solved as a G-Reduction instance. He wrote some steps using human
notation, but then crossed them out. The notational systems were incompatible but the concepts
looked applicable. He tried again, this time casting the exercise into dunari form. The
converted exercise reduced readily, producing the desired result.

Colty put his pen down and leaned back. Had he solved it? To convert the exercise he had abused
the notation. The lovely elegance of the human presentation was lost. Could he prove that the
converted exercise was still the same?

The sound of someone knocking at the kick panel of his door jarred Colty out of his reflections.
``Yes? Come in.''

The door slid aside and Dargon stepped in. Colty straightened himself and started to anxiously
gather together the loose pages of the human book.

``I don't mean to interrupt your work,'' Dargon said. ``I just wanted to see how you were
doing.''

``I'm\ldots\ I'm fine.'' Colty slid the book, roughly assembled, back into one of his trays. He
closed his notebook. ``Please, have a seat.''

Dargon lay down next to Colty's table and looked around. ``It's a cozy office,'' he said
finally.

``I don't mind it. Not a lot of distractions.''

``I think it's great that you've decided to look at graph extension fields. That's such an
important area so it's good to have someone on our faculty specializing in it.''

``Yes.'' Colty put his trunk over his notebook and tapped the cover of it slightly. ``I'm\ldots\
I'm still getting my bearings in the literature.''

``Of course. I imagine there are many interesting lines of research possible in that field. I
understand it would be hard to choose.''

Dargon continued chatting about life around Parnon in casual tones. ``I hear a rumor that you
are companion to one of our librarians.''

Colty clicked his teeth. ``It's true. Joleia and I have been together for several runion now.''

``That's wonderful. It's good not to be too consumed by your work. You have to give time to
other things---things that really matter.''

Dargon stood up and made his way through the cramped space toward the door. Just before leaving
he turned to Colty. ``I'm looking forward to hearing more about your graph extension work at
your next overseers meeting.''

Colty bobbed his head respectfully.

After Dargon left Colty opened his notebook again and studied the symbols scratched there. Could
some sort of formal mapping, a kind of isomorphism, be created between the human and dunari
mathematical systems? he wondered.

Tentatively he wrote a few more formulae beneath his solution to the exercise, yet he found it
hard to concentrate. He stood up and paced back and forth in his narrow office trying to clear
his mind. He understood all too well Dargon's message. Setting aside one notebook, he took out a
second for his graph extension work and flipped through the pages impatiently. Suddenly he
stopped and rubbed his face with his trunk.

``I need a walk,'' he said aloud.

The Parnon student center was the hub of student services and activities on campus. It was a two
story building in an elliptical shape oriented parallel to the Nermella. In the middle of the
building was a large courtyard with an impressive rigat tree growing at the center. At the east
and west ends of the courtyard gaps in the building let in light, air, and people.

As Colty wandered into the courtyard he saw a small group collecting in front of a temporary
stage erected on the south wall. At the top of the stage a podium stood waiting, microphones
ready. Several people worked behind the stage setting up cables, while two others handed out
pamphlets to the onlookers. A news crew with a portable video camera interviewed a pair of
students.

Colty remembered seeing something about a Limzar speaker being on campus but he hadn't paid much
attention. Perhaps this was just the distraction he needed. He lay quietly on a seat under the
rigat tree watching people gather as final preparations were made.

The crowd grew steadily and soon there was a sizable herd of dunari filling the courtyard. Some
lay on the ground but others just stood. After a few moments a middle aged woman stepped up to
the podium. She leaned forward into the microphones and said a few words but they were
inaudible. She turned and said something to the people behind the stage and then tried again.

``Are we on?'' her voice rolled out of the speakers and filled the courtyard. ``Yes? Good.''

She paused a bit while the crowed settled down. ``It's a great moment here in Parnon,'' she
continued, ``it's good to be back!'' There was some stomping of feet and cries from the crowd.
``I know you didn't come here to listen to me so without further delay, I would like to present
Chancellor Zarlonloti of the Limzar Association.''

There was more stomping of feet and some bleats of support. The woman withdrew and was replaced
by a stately gentleman. He was thin with dark scales and exceptionally well dressed. He walked
calmly forward, feeling no particular rush, letting the applause continue for a long time. When
he reached the podium he waited patiently for the crowd to once again fall silent.

``Change is upon us,'' he began simply, ``but is it change for good or ill?'' He paused and
scanned over the crowd as if to be sure everyone was paying attention.

``The First Citizen believes his plan of industrialization is change for good,'' Zarlon
continued. ``He believes poisoning our land, sickening our rivers, and choking our air with the
smog of our machines is the change our country needs. I say it isn't so. Such change is change
for ill. Do you agree?''

A loud ``Yea!'' rose up from the crowd.

``Immediately prior to coming here I visited the wind generators at Karackia. Dozens of mills
taller than the tallest buildings in Parnon, driven by the Nermella, produce enough electricity
to power a small city. It is a marvel of engineering. Yet what the government pamphlets don't
tell you is the metalworks where the mills are made routinely dump huge quantities of free
copper, zinc, and lead into the Notalios River, turning it an unnatural green.''

Zarlon paused again and unhooked one of the mics, clipping it to his cloak. ``Nothing lives in
the water. Nothing lives along its banks. For dozens of karnons the people depending on the
Notalios are broken and despairing, yet nothing is done.''

He walked across the stage slowly, talking directly to the people in the crowd. ``Marvel of
engineering? This is not the Argenia I know. This is not the Argenia I want. Do you agree?''

An even louder ``Yea be Yea!'' filled the air.

``The Limzar offer a different plan, a plan for a different kind of change. We embrace the
values that made this country great: working hard, helping each other, respecting the world that
Jurita made for us. Our plan is to use these values to guide our path into the future. I believe
that in this way Argenia can be made clean again. The Argenia of our mothers will be the Argenia
of our daughters. \emph{Do\ldots\ you\ldots\ agree?}''

``\emph{Yea be Yea!}''

Zarlon returned to the podium and appeared to look at some notes. Colty shifted in his seat and
snorted slightly.

``Technology has its place,'' Zarlon continued, ``but it is up to us to control it rather than
letting it control us. We are not machines, nor will we be slaves to machines. We only need look
at the humans to see the consequences of rampant technology. They hide in their compounds, away
from the sun. They talk to us through their machines and look at us with their machines. They
are not building a telescope on Nermia; their machines are doing it. They are unfeeling, remote,
and little more than machines themselves. I, for one, do not want Rujar to ever become the way
Earth no doubt is: controlled, scheduled, and contained.''

Zarlon started walking across the stage again as he continued speaking. He spoke of limiting
technology and ways the government should intervene in its expansion. He spoke of returning to
the basic values that inspired the founders of Argenia. He endorsed the Limzar candidate for
First Citizen and spoke of how his election would usher in a new era.

``We must never forget,'' Zarlon said as his speech drew to a close, ``that it is Jurita who is
the source of all authority. Her ways have not led us astray, and they will not. We must look to
Her for both inspiration and guidance. We deserve a First Citizen who understands that. We
\emph{demand} a First Citizen who understands that. Thank you for your attention.''

The crowd burst forth with loud stomping, bleating, and cheers. Colty stomped his foot as well.
Joleia had always spoken highly of the Limzar and out of respect for her he contributed to the
applause.
