% Local IspellPersDict: aspell.dunari.pws

\chapter{47}

Jernumia stood on the planks beside the prosecutor table and consulted her notes.

``Eminence,'' she began, ``it may be true that the accused brought sufficient cash to the
transaction, but without bank receipts we have no way to know where that cash came from\ldots''

``This is insane,'' objected the defense adviser. He pressed his trunk against the defense table
defiantly. The Eminence gave him a baleful look causing him to clench his teeth. Then, turning
toward Jernumia, the Eminence nodded slightly. The three judges, sitting as they always did in
the box above the planks watched the proceedings with detached eyes.

``The Real Estate Accountability Act is very clear,'' Jernumia continued. ``In section 8.3,
paragraph six it stipulates that all cash used for transactions of this magnitude must be fully
documented.'' She stepped toward the stand and handed a page to the Eminence and a copy to the
defense adviser. ``Several documentation methods are enumerated in the act; none were used in
this case.''

The Eminence glanced over the paper. He turned toward the defense adviser. ``I don't think you
did your homework.''

``It might be that the letter of the law was not followed exactly in this case, but the
intent\ldots''

``How do we know the intent?'' Jernumia interrupted. ``The Real Estate Accountability Act was
passed precisely to avoid, for example, money laundering schemes and other criminal
activities.'' She raised her trunk slightly to ward off the objection. ``I'm not accusing your
client of money laundering; I only mean to point out that the law is written the way it is for a
reason.''

The defense adviser took a deep breath. ``I request \textit{hutia noro}.''

``I respectfully request \textit{hutia noro} be denied in this case,'' Jernumia said. ``The
defense should have known the details of the Act.''

The Eminence looked closely at both advisers. He glanced up at the clock mounted high on the
back wall of the court. ``Section 8.3 is a technicality of the act,'' he said at last. ``I'm
sure this is not the first time it has been overlooked. I'm going to grant \textit{hutia noro}.
We'll reconvene at hour 71:050, Legal Time.''

Jernumia bobbed her head.

``Thank you, your Eminence,'' the defense adviser replied.

As the court emptied Jernumia gathered her materials and slipped them back into her pouch. The
defense adviser talked in hushed tones to his client. ``Don't worry, we'll work it out,'' she
heard him say.

``What woman would be a lawyer anyway?'' the client asked.

Haratol came up to Jernumia and clicked his teeth slightly. ``Very nice. Let's get out of
here.'' She slung her pouch over her back, smoothed her cloak slightly, and walked with Haratol
toward the prosecutor's door.

``It is a technicality,'' she said when they were well out of earshot of the defense.

``And you're so good at finding them,'' Haratol replied.

Jernumia stopped. ``Then why I am I still trying these silly real estate cases?''

``These are hardly silly cases. Fraterhof is a major corporation. Aratok doesn't take silly
cases.''

``Come on,'' Jernumia said as she flapped her ears, ``we are talking about the questionable acts
of some minor, ignorant Fraterhof manager. Any new attorney could handle this case, you know
that.''

Haratol slid his trunk over her shoulder and encouraged her to walk again. ``Look, the guy in
the big office wants you on these cases because you make them easy. Do you really think a new
adviser would have found that stuff in the Accountability Act? That was brilliant.''

Jernumia snorted slightly.

``Are you hungry? Maybe we could get some food together.''

``What? No, I don't think so.''

Jernumia ended up walking alone on the streets of Varnok, letting the Nermella blow her
frustration away. When she got the position at Aratok she was excited to be the first woman in
their team of court advisers. It was a prestigious job and a fitting reward for her dedication
to the profession and her exacting attention to detail. Yet two arnoxes later and she had only
just started to see the inside of a courtroom. The goal she wanted always seemed to be one step
beyond. Perhaps, she thought, she expected too much.

The smell of fresh mosen cakes wafted from a eatery at the corner, yet she wasn't hungry.
Instead she turned down Following Street Polnet toward the river and the museum district. She
had 50 arnets before court convened again and while she knew she should be reviewing
counter-attacks to possible defenses, she didn't feel interested in that.

\spacebreak

The lifter stopped in front of the train station, and Colty climbed out scanning the side
parking lot. There, in a spot reserved for taxis, Rachel's van was parked. He quickly walked
toward it and Rachel came out the back, staying close to the wall to be less conspicuous.

``I hope I haven't kept you waiting long,'' he said when he reached her.

She spoke to the driver briefly and then snapped her pouch about her waist. Facing Colty, she
took a deep breath. ``I'm ready,'' she said.

The two of them made their way into the station. As usual Rachel looked around with interest.
The people of Parnon had grown accustomed to the human presence and although many avoided her or
watched her cautiously, most went about their business. There was a ticket counter to one side,
a waiting area across from it, and from there a ramp that went down to the trains. The station
in Parnon was nothing special; Colty wondered what she was going to think of the one in Varnok.

Colty paid for the tickets, complements of the university, and the two walked down the ramp to
wait for the train. They stood close to the wall. Rachel squatted and leaned her back against
the bricks, in part to make it easier to talk with Colty, in part to make her less tall.

``I would like to ask you,'' Colty said at one point, ``if your people have anything similar to
our high level arithmetic.''

``We do, yes, very similar. We call it abstract algebra.''

``I\ldots\ I would be interested in learning more.''

Rachel smiled slightly. ``Would you like a book about it?''

``I know I shouldn't be asking.''

``I'd be happy to get a book for you.''

When the express train finally arrived they hung back and let the others board first. The train
itself wasn't particularly long and was organized for commuters. There were two seats on either
side of a central aisle. About half of the seats where occupied. Everyone stared at them as they
made their way down the aisle looking for a pair of free seats. Rachel kept her head down to
avoid hitting the roof of the train.

They found a suitable place behind a mother and daughter, and Colty offered Rachel the seat by
the window. The woman in front of them seemed distinctly uncomfortable but she didn't move.
After a moment the young daughter turned around and stared at Rachel. ``Hello,'' Rachel said.
Colty could tell she was suppressing a smile. The child's mother spoke to her daughter in a
hushed voice and reluctantly the child faced forward.

Colty could hear some other voices behind them muttering something about the human but he
couldn't make out what they were saying.

``How are you doing?'' he asked Rachel.

``I'm fine. I'm okay.''

It wasn't long before the train pulled out and accelerated down the track. Rachel looked out the
window intently watching the town rush by. ``There's Tartuk Hall,'' she said pointing. In the
distance Colty could see the distinctive spire rising above the trees and buildings of Parnon.

The train followed the river. Its dark waters glinted orange-red in the sun's light. Rachel
watched as they passed fields of nanik and florenstalk. The ground was covered in a blanket of
red and yellow. Large rigat trees lined the river bank in some places but in other places the
view was clear. In one spot a boat floated lazily along the current.

% TODO: More specific descriptions?

The track crossed roads now and then, often with cars parked waiting for the train to pass.
There was also the occasional house, farm, or business. In the distance there were some red
clouds against a greenish sky.

Colty looked at Rachel. She had her head turned to face out the window so he could mostly only
see her thick, black hair hanging down to her shoulders. As usual she had it gathered together
in a neat bundle, this time with a bright yellow tie. Her clothing was a lighter shade of yellow
with some tan and a bit of green. Despite the strangeness of her body her clothing seemed rather
ordinary. The colors and even the style were not unlike something an Argenian woman might wear.

As they approached Varnok the view became progressively more urban. There were fewer fields and
more buildings. They had veered away from the river and were crossing a kind of industrial park
on the outskirts of the city. Rachel tilted her head as she tried to look forward to see what
was coming.

The train passed by a large, brown area. No plants grew on the ground. Instead, some distance
from the tracks, were several tall piles of orange and red dirt. Two small buildings were squat
on the ground with large pipes connecting them. From each building a pipe emanated and after
going a short distance made a right angle turn and burrowed into the ground.

Rachel studied the scene. ``What was that?''

``To be honest I'm not sure.'' Colty felt vaguely embarrassed. It wasn't Varnok's nicest
introduction.

Soon the train began to slow and the city rose up around them. There were no more trees, only
streets, buildings, and traffic. The buildings themselves grew taller and soon they were
towering over the train, three, four, or even five stories high. The train ran parallel to
Cross Street Hark, moving in and out of the shadows of the buildings.

Colty looked ahead. ``Okay, here we go,'' he said.

At that moment a voice came over the intercom, ``We are approaching the Vinden Tunnel,'' it
said. ``We'll be arriving in one arnet. Welcome to Varnok.'' The train dipped below the street
level and suddenly was engulfed in the darkness of the tunnel. Lights, glowing rubinum-red
flipped on automatically. The girl in front of them gasped and her mother wrapped her trunk
around the child to comfort her

``Oh wow,'' Rachel said.

``Varnok Station is underground,'' Colty explained. Rachel turned and looked out the window into
the absolute darkness, staring intently. What she could see Colty had no idea.

In moments the train slid into the open, underground platform of Varnok Station. Colty was
thankful that it was well lite both with natural light from high windows and supplementary
artificial lights. The train eased to a stop and then, after a moment, the doors opened. People
began to stand up and gather their things. The child in front of them turned and stared at
Rachel again. Rachel nodded slightly, and the child spun around without a word. When the mother
stood up she glanced at the human but quickly looked away and hustled her daughter forward.

The platform was, as usual, crowded with people. Many backed away from them with alarm, bumping
into each other to keep their distance. But others actually seemed very interested in Rachel and
bobbed their heads as they passed. Rachel nodded politely but she kept her arms tight to her
body and stayed close to Colty. ``There are so many people,'' she said.

Colty wrapped his tentacles around Rachel's arm and led her forward, moving toward the ramp for
the city trains. When they got there the ticket master was reluctant to let Rachel on. ``It will
have to stand in the back,'' he said.

``That fine,'' said Rachel softly to Colty.

``\emph{She} will be happy to do that,'' Colty said as he paid the fare. They got on the train
and made their way to the rear where they held on to poles for balance. Rachel wrapped her hand
around the pole, gripping it tightly. On the seat next to her was a woman dressed in a business
cloak.

``Jurita preserve me,'' she said softly with her head bowed. She clutched the rail of her seat
tightly in her trunk. Rachel looked at her and then at Colty but didn't say a word.

The ride on the city train was short since the museum district was just in another part of
downtown. The train had a stop in the center of Museum Park overlooking the river. As it pulled
into the stop the train jostled suddenly and Rachel put out her hand against the seat in front
of her to catch herself from falling. The woman next to her gasped and muttered something to
herself, huddling away from Rachel as best she could.

``I'm sorry,'' Rachel said to her. She looked over at Colty with an expression he couldn't read.

When they got off the train everything was more open and less crowded. ``I'm so glad you're
here, Colty,'' Rachel said. ``I wouldn't have been able to do this without you.''

They walked a short distance from the train stop into the park. The grass was neatly trimmed and
the park was beautifully landscaped. There were seats here and there, many in the shade of kilar
trees. Rachel, as always, was interested in the flowers. They talked for a time and enjoyed the
air. In the shade and away from the busy train she seemed more relaxed and more like herself.

Surrounding the park were three stately buildings. Colty pointed to each in turn. ``That's the
art museum,'' he said, ``and the science museum, and finally the natural history museum.'' He
turned to Rachel. ``Where would you like to go?''

``Oh, the art museum, definitely.''

The Argenian National Museum of Art, as it was officially called, was itself a historic
architectural landmark. Wide steps lead up to three broad doors over which was spread a tile
banner. The tiles were different colored stones set into the rock that, taken together, depicted
Roskala as he for the first time looked over the Varsynthia valley where Varnok would eventually
rise.

Colty paid the admissions fee, again compliments of the university, and they entered the museum.
The main hall was fully three stories high and decorated with long tapestries of the Mingala
Era. Stone columns rose from the marbled floor to hold up the vault of the ceiling. Along the
ceiling, windows where placed precisely to allow the light from the sun to shine against a wall
covered with another tile image. It was a depiction of the ancient Hall of Yartolas before the
Bizat War, one of the lost wonders of the world. The shadows in the image where aligned exactly
with the incoming sunlight so that it almost appeared as if the sun was shining into the image
itself.

``That's beautiful,'' Rachel said.

Colty was surprised at how few people were in the museum, but then again it was a moment when
many people were working. He was grateful it was quiet. Rachel was more at ease that way. There
were some people of course, and even a number of children, but they were respectful. They mostly
kept their distance but they did not seem afraid.

Colty flipped through the guide book given to them when they entered. ``Do you have anything in
mind you'd like to see?'' He asked.

``All of it,'' Rachel said. She was still looking at the tile image. She laughed slightly and
then turned to Colty. ``No, not really. We can just wander.''

At that moment Colty noticed a man approaching. He was dressed formally and seemed pleased to
see them. ``I heard you were visiting,'' the man said as he drew near. ``I'm the museum
director. It's a great honor to have you here. I'd be very pleased to give you a personal
tour.''

With the director's guidance they spent many moments walking through the various galleries and
admiring works of many styles and from many eras. Rachel seemed interested in everything but she
seemed to particularly enjoy the romantic realists. She spent the most time studying the works
of Mumulia, Arkitalos, and especially Zugola. Colty could see her eyes scanning the paintings
closely. She asked many questions about the symbols and history in each work. They were
questions the director was more than happy to answer.

Many people greeted her and spoke with her, often asking questions about her world as well.
Rachel chatted comfortably with the small groups and became more animated and talkative.

``That way is one of Zugola's most famous works,'' the director said, pointing down a short
hall. ``I think you might like it.''

Separate from the other galleries, \textit{The Creation} was kept in its own room under
continuous guard. Colty, Rachel, and the director walked down the short hallway to the special
gallery. They entered the room to find one wall almost filled with Zugola's masterpiece. It was
taller even than Rachel and at least as wide as four dunari standing lengthwise together. The
painting was cordoned off with a guard at one end. The lighting, hanging down from the ceiling,
was carefully crafted to mimic the natural light of the sun so the colors were accurate. At the
back of the room were some steps. Colty assumed they were used, perhaps during tours, so people
could get a good view.

A young couple was in the room, standing on the far end of the ropes. They looked at Rachel with
surprise but they didn't move. Rachel didn't seem to notice them. She was transfixed by the
image before her.

On the left, standing on the side of Jenas Mountain was Varsynth, one of Jurita's prophets. He
had his trunk outstretched pointing into the distance over the Erebus valley. The sun shone full
on him, his golden cloak tousled by the wind. On the right, above the distant Argenian Sea, a
great light shined down from heaven. Riding on that light was the Seleck of Jurita: the seven
forces of the world.

Zugola's mastery of color and lighting was exquisite. The darkness of the distant clouds, the
rubinum of the valley floor, and the bright, living yellow of Jurita's light all combined to
give the image a vibrant intensity that few artists had ever achieved.

Rachel looked at the painting silently for some time. ``What does it mean?'' she asked quietly.

``It's about the moment Jurita breathed life onto Rujar,'' the director replied in a hushed
voice.

``How do you like it?'' called out the young man standing across the room. He and his companion
were looking at Rachel. ``It is said Zugola worked on it for over 200 runion.''

``It's an incredible piece. Truly beautiful.''

Rachel's answer seemed to please them.

``Do your people have anything like this?'' the young woman asked.

Rachel ran her fingers through her hair. She looked momentarily lost in thought. ``We have a
couple of pieces that might be similar.''

The young man clicked his teeth. ``Sometime, I'd love to see them.''

\spacebreak

Jernumia lay on the seat in front of \textit{Women Working}. It was a painting from the reign of
the Third Regent depicting a group of women grinding nuts and baking larka for a Juju feast.
They used simple trunk-operated nut grinders and roasted the meat over an open fire. Yet it was
the expressions that fascinated Jernumia the most. It was as if each woman's story was written
on her face, stories that were sometimes joyful, sometimes sad, and sometimes deeply mysterious.

There was a commotion in the hall outside the gallery and Jernumia caught a glimpse of what
appeared to be a human walking past with several dunari about it. She went to the doorway in
time to see the creature, it's tall thin body standing high above the two men on either side.

``What's happening?'' she asked a museum guard.

``The human is going to speak in the Morgolan Room just down the hall.''

Jernumia checked the clock in the gallery. It was a good time to leave. The last thing she
wanted was to be caught in a crowd around one of the mysterious aliens, and she had
counter-attacks to prepare in any case.

``Which way is the quickest exit?'' she asked the guard.

In the Morgolan Room there was already a substantial gathering. Rachel lay on the floor by
folding her legs directly under her. Colty lay down beside her and the director stood in the
center of the room waiting.

Around the walls forty or fifty people gathered and several leaned against the wooden posts that
punctuated the space. A number of children where there also, mostly lying on the floor where
they wouldn't get in the way of those behind. At the last minute two reporters arrived,
obviously rushed to get there, and forced themselves into the back.

``I realize this is very spontaneous,'' the director started, ``but we are honored to have
with us today the visiting professor from Parnon. Um\ldots\ how did you pronounce your name
again?''

``Just call me Ra'hel.''

Her soft voice was in striking contrast to the rumble of the director.

``Well, Ra'hel,'' the director continued, ``let me get out of the way and let you speak.'' He
lay down at the edge of the circle of open space in front of Rachel and the room drew silent.

``Sar. Coltinarly brought me here to see the museum,'' Rachel said. ``I wasn't expecting to
speak so I'm not sure how to start.''

``Why do you look like a chortak?'' a voice called out. It came from one of the children in the
front row. There was some shaking of ears.

``I'm sorry,'' said the child's father, ``he doesn't mean any disrespect.''

``It's fine,'' said Rachel. ``In fact, it's a very good question.'' She addressed the child
directly as she continued. ``It's because of parallel evolution. Do you know what that means?''

The child swayed his head.

``Creatures that evolve in similar environments develop similar traits. Millions of runion ago
my ancestors on Earth lived in trees just as the chortak do now. I have gripping hands and
forward facing eyes that allow me to sense the distance of something by just looking at it.''
She held out her hands, partly open. ``If you are jumping from one branch and reaching for the
next\ldots\ well\ldots\ you don't want to miss.''

``What is Earth like?'' came a voice from the back.

``In some ways it's like here. Many worlds have similarities and life on those worlds are
similar too. It's parallel evolution again. Yet there are differences. Our sun is hotter than
yours and shines with a different color. My eyes are adapted to the light from our sun just as
your eyes are adapted to the light from yours. I can't see rubinum. To me that color is just
black. But I can see ultra-green. In fact the sky of Earth is ultra-green and so it looks very
different than here to my eyes.''

``It must be very hot on Earth if your sun is hotter.''

``Earth is far away from our sun. In fact, I'm actually used to a cooler climate than here. I'm
an endotherm. My body generates its own heat. It's a useful thing when the temperature varies a
lot, but here where the sun never sets it's more of a disadvantage, I think.''

One of the reporters in the back spoke up. ``Is it true that Earth is covered by your technology
as Zarlon of the Limzar Association has claimed?''

``Zarlon has never been to Earth so what would he know?''

Everyone in the room shook their ears.

``We have great cities, like you do, and large works of engineering. Yet there are also millions
of square karnons of wilderness on Earth. Even in our cities, as in yours, we have trees and
plants\ldots\ and flowers.'' She glanced briefly at Colty.

``It's true that we caused damage to our world with our technology, just as you are doing now to
yours. We are trying to undo the damage we caused, but it will take time. I'm sure you will do
the same eventually.''

``Are you saying that we are like you?'' a woman called out.

Rachel turned to look at her. ``In many ways you absolutely are. We estimate you are only two
thousand runion behind us technologically. Compared to the age of the universe that is but a
fleeting moment. Even without any help from us you will be building star ships soon enough. If
things had only been slightly different our first meeting would have been from you visiting
Earth rather than the other way around.''

``But we are not as smart as you,'' another child called out.

``I don't know about that,'' Rachel said, ``what is `smart?'\,'' She smiled slightly.
``Anyway, I think I can safely say that one of the smartest people I've ever met is dunari.''

Colty wondered who she meant.
