% Local IspellPersDict: aspell.dunari.pws

\chapter{53}

% Zarlon (leader of the Limzar party and jilfora of the seleck of Junarit)
% Numtala (vascular surgeon at a teaching hospital)
% Gaalets (member of Parliament)
% Qunart (sub-general of the Argenian armed forces)
% Mepel (judge in the provincial court of Polaria)
% Ratfornin (professor of history at Parnon)
% Hearbunbia (Lieutenant president of Toxalla Media)
%
% Junarit is in the Seleck of Jurita. She wrote several "controversial" documents called the
% "Songs of Truth." Their authenticity is disputed by the more traditional religious community.
% One of the things mentioned in the "Songs of Truth" is the idea that men should also form
% selecks.

Zarlon turned off the main highway amid a grove of larzippa trees and onto a rough, poorly
maintained dirt road. Many karnons from the nearest town, the remote location was wild and
primal. A low cloud of dust rose behind his car only to be dispersed by the forest breeze. The
road slanted upward slightly and after a short distance came upon an open field of harsna grass,
finally reaching a small parking area at the base of a rise.

Six other men were gathered there, talking among themselves and waiting for Zarlon's arrival. As
he got out of his car they came forward and greeted him by momentarily touching their trunks to
his left shoulder.

``Greetings this moment my fellow seleckia,'' Zarlon said, ``a time of greatness is upon us.''

``Dangerous greatness,'' one of the men replied. ``I fear my aides are suspicious of my
unscheduled absence.''

Zarlon swayed his head and touched the man's back with his trunk. ``Fear not. Soon secrecy will
no longer be needed. Once the Last War has wreaked its purpose our order will lead the way
toward rejuvenation.''

``Maybe,'' the man said, ``but until then let's get on with it. I only hope none of us were
followed.''

Zarlon swayed his head and with solemn purpose the men put on ceremonial cloaks usually reserved
for female seleckia. Each was individually adorned with a symbol of one of the seven forces. As
the leader Zarlon wore the cloak of Jilfora, decorated with the binding force of love.

Silently the group walked up a narrow path to the top of the small rise. There the ground
leveled forming a wide and surprisingly well-tended lawn. The knoll was the only feature in an
enormous plain of harsna grass. The tall, lush yellow plants surrounded them like a sea
surrounds a boat, stretching into the vast distance as far as even dunari eyes could see. The
sky above was clear and green with a touch of golden sand driven from the distant desert by the
Nermella wind, the breath of Jurita Herself.

At the center of the lawn was a low stone pedestal made from rocks that were carefully
positioned and mortared. A flat rock, like the top of a table, rested on the pedestal, awash
with the deep, orange light of the sun.

The six men took their positions around the pedestal. They stood facing inward about a trunk's
length apart. Zarlon knelt inside the circle facing the pedestal with the sun on his face.
Bowing his head slightly he laid his trunk lightly on the table. The others put their right
front knees on the ground.

``\textit{Huminia fa legori},'' Zarlon said softly and with a practiced confidence.

``\textit{el itara!}'' the others replied in unison.

``\textit{Huminia fa tarola}''

``\textit{el tarara!}''

``Let Jurita guide our thoughts.''

They all lay comfortably on the grass.

``I still have concerns,'' Numtala began after a few moments, ``about the human reaction.''

``It doesn't matter what they do,'' Zarlon said. ``Even if they do nothing the First Citizen
will attack.''

``Or perhaps it is the humans that will make the Last War manifest,'' Ratfornin continued.

``Yes, pouring down Jurita's wrath from the skies,'' said Numtala.

Zarlon flapped his ears. ``The humans are cowardly. They will hide behind their technology and
do nothing, all in the name of being gracious guests.''

``What of the super weapon?'' Qunart said.

% TODO: Can I do this scene without mentioning the names? All the new names are overwhelming and
% they are only used once.

Zarlon turned to face him. ``We will use it when the time is right. As you know, the target has
already been chosen. We need to even the sides in this impending conflict, for Argenia is too
strong. Without our help Forbin will be overrun and there will be no Last War, only Argenian
dominance.''

Zarlon swayed his head slightly to his left and then to his right, taking in the group of them
with both his eyes. ``We have debated this at length. The time for discussion has past. Now is
the time for action.''

The perverse seleck of men all rose as Zarlon draw a small dagger from his leg pouch and laid it
on the stone table. The others sat back on their rear legs and Zarlon took a small piece of
cloth paper from the folds of his cloak. He held it high into the air.

``Here is the order to kill the unholy creature that by its very presence defiles the city of
Parnon, the country of Argenia, and the world. Yet its death will ignite the hot, purging flame
of Jurita's anger, and by so doing fulfill Junarit's prophecy.''

``\textit{So it was, is, and will be,}'' replied the group.

Zarlon put the paper in the center of the table and weighed it down against the Nermella with
four polished lamstones. The seleck of men lifted their front legs into the air, with their left
foot raised slightly higher than their right, trunks outstretched. Zarlon leaned back and turned
his face toward the open sky, his trunk reaching upward as well.

% Some mega-foreshadowing here!

``We call upon you, Mother of the Universe, to witness the actions of your humble servants. May
your judgment be swift and true.''

``\textit{el tarara!}'' everyone called out in unison.

Zarlon stepped forward and took the dagger. With a quick motion he plunged the tip into his left
front thigh between two scales. Without any expression of pain he dug the dagger through his
thick, leathery skin until orange blood started to ooze down the blade. ``\textit{el itara},''
he called out loudly.

``\textit{el itara},'' replied the seleckia.

The seleckia lowered their front legs back to the ground and stood up normally, still remaining
in a circle around Zarlon and the stone table.

Zarlon held up the blade. ``The will of Jurita acts through the seleck of Junarit as foretold in
the \textit{Songs of Truth.}'' Touching the point of the dagger to the paper, Zarlon stained the
corners with several drops of blood. Putting the dagger aside he bowed his head and began to
chant. At first his voice was little more than a low, rumbling growl. Yet it was taken up by the
seleckia and soon the sound rose like the wind, vibrating, it seemed, the air around them and
the ground beneath them. It became not the voice of one man or another but instead the revealing
voice of the false Goddess, raining down from the sky above and enveloping them.

\spacebreak

Colty paced back and forth in front of the classroom. His notebook was open on the table next to
a collection of co-sequence papers and the three books Rachel had given him. The blackboard was
covered with several Nulari sequence diagrams and a mixture of dunari and human mathematical
notation.

He paused in front of a critical formula. It was Nulari's theorem of co-sequence divisibility.
He stared at the formula as he had done a hundred times before. It wasn't the whole story.
Something deep inside him knew there was more. He walked over to the other side of the board
where he had reformulated the basic axioms of co-sequence theory in terms of human sets. It was
an elegant isomorphism. He made an adjustment to one of the steps of his proof showing the
logical equivalence of the two forms.

The human mathematics was extraordinarily precise and it revealed several areas where the
original theory was lacking. Colty walked back to the dunari side of the board and looked at
Nulari's second lemma. He compared it to the longer, but more accurate version he had derived on
the human side using his isomorphic axioms. The dunari math was fuzzy, unclear, and slippery.
The human math was unforgiving and exact. The human form of Nulari's second lemma didn't seem to
say the same thing.

% TODO: Then it isn't an isomorphism, is it?

Colty concentrated. Ignoring the notation he tried to discern the essential difference between
the two forms. The logic of Nulari's development seemed inescapable, yet somehow it arrived at a
slightly different result than Colty's derivation using human notation. He walked over to the
table and flipped through Nulari's seminal paper. Was there some extra step he had missed? He
knew the paper word for word. How could he have missed anything?

He walked over to the human side again and stared for a while at the alternate version of the
lemma. Something about it looked familiar to him. He picked up one of the books Rachel had given
him and flipped through it. There it was on page 215: an example of hyper induction. The
starting formula in the example had a similar structure to the one he had written on the board.

Colty looked at the version he derived, then back to Nulari's version, and then to his version
again. He reflected on hyper induction, an advanced human mathematical technique. In his mind
the concepts drifted about, touching and bumping against each other like pieces in some
fantastically intricate puzzle. Reaching out with his trunk he tentatively scratched a few
symbols on the board in an effort to put two of those pieces together. A third fit with the two,
followed by a forth. With a rush everything clicked into place.

He knew.

Nulari's derivation made a hidden assumption. It was something so subtle that it had gone
unnoticed by dunari mathematicians. That assumption forced the reasoning into a specialized
setting. Colty realized that if he did not make the assumption he would be in a more general
world. That was why the form he derived using human notation looked different. Colty also
realized that now he could use hyper induction on his version of Nulari's lemma.

He started writing quickly, his trunk found it difficult to keep up with his mind. The math
poured out of him onto the middle of the board, and he freely mixed the notations of two
species, passing back and forth between them fluidly, as he sketched the outline of his proof.
It was easy.

When he finished Colty let the chalk drop into the tray, and then stepped back to review what he
had created. Quietly he scanned over the board.

``Obvious,'' he said aloud to himself, and then clicked his teeth.

He flipped open his notebook to a fresh page and carefully copied down his proof sketch for
later study. There was still a lot of work to do to fill in the details, but he knew that now it
would just be a routine matter. He had done it. He had the proof that would grant him full
status.

After he finished copying the details into his notebook he erased the board, pausing slightly
just before erasing the final steps. He then walked down the hall to his office, notebook tucked
safely in his pouch. He had to tell someone about what he just discovered, and he knew who that
person needed to be.

He dialed Rachel's office number. The line beeped several times and he wondered if she was at
the library. Eventually she picked up.

``Yes?'' her light voice said on the other end.

Without any introduction Colty simply said: ``All stellating co-sequences have prime bases.''

There was a pause.

``And you have a proof,'' Rachel said in reply. It wasn't a question.

``I do,'' Colty said, clicking his teeth loudly.

``That's wonderful! What a lovely result.''

``I still have to clean it up, but I'd love to show you. I had to use hyper induction.''

``You used hyper induction on co-sequence math?'' Rachel asked.

``Yes, after realizing Nulari made a hidden assumption that was blocking the depth of his
result\ldots''

``Wait,'' said Rachel, ``how did you justify that?''

``I created an isomorphism between human and dunari mathematical systems, so I was able to map
your techniques onto dunari concepts.''

There was another pause.

``That's amazing, Colty. I'm very interested in talking with you about this. I have to teach
right now, but after class I'll come right over, if that's okay.''

``That would be just fine,'' Colty replied.

\spacebreak

Karn tapped one tentacle impatiently against the steering stick as he peered out the window of
the truck. Gat followed his gaze along the cross street to where a small walkway opened onto the
road by a patch of harsna grass.

``How good is our intelligence on this?'' Yarlon asked. He leaned his trunk on the ledge of the
small window in the barrier separating the front seats from the back area, pulling himself up to
see a little better.

% TODO: The name Yarlon is too similar to the MIB agent helping Fotkey.

``It is what it is,'' Karn replied, practicing his Forbin accent. He looked at the clock on the
truck's main panel.

``There!'' Gat pointed down the street to an alien, bipedal figure striding toward them. Its
front legs swinging ridiculously without even touching the ground.

Karn sank back slightly in his seat. ``Very punctual.'' The creature turned into the walkway and
quickly disappeared behind the harsna. Karn engaged the transmission and moved the truck gently
forward, stopping again just where the human had turned. ``Yarlon,'' he said simply. He gestured
toward the back and then slid open his door.

Gat moved over to the driver seat while the two other men trotted down the path. Karn quickly
surveyed the area. As expected the way was empty. They soon rounded a bend and saw the human
ahead, apparently oblivious to them. Karn slowed to a rapid walk and Gat followed him.

``Can we have a word?'' Karn called out as they caught up to Rachel. She turned around and
regarded them uncertainly. ``We would like for you to come and speak with the Forbin
ambassador,'' Karn continued.

Rachel tilted her head slightly. ``What? I don't understand.''

``You should come right now.''

``Now?''

Karn reached into his cloak and held up a trunk pistol more just to show Rachel he had it than
to directly threaten her with it. ``Yes,'' he said, ``now.''

Yarlon wrapped his trunk around Rachel's waist and she started at the intrusion. Karn turned so
the two men flanked her on either side and Yarlon butted against her with his trunk ridge.

``What\ldots\ what's going on?'' Rachel said anxiously.

``Just a meeting, that's all,'' Karn said.

Rachel let herself be led back toward the main road. Yet before they cleared the harsna grass
another man approached them going the other way. Karn leaned over and spoke quietly to Rachel.
``If you say anything we will kill him.''

She tensed but kept walking. The man passed trying not to take notice of the human as he went.
By the time they reached the road Gat had backed the truck onto the walkway. Karn quickly opened
the back and motioned for the human to climb in. Rachel hesitated.

``Go on. We have a short drive to get there.''

``Where are we going?''

Karn fluttered. He grabbed Rachel's arm and pulled her forward powerfully. ``The sooner you get
in, the sooner you get back.''

Rachel scrambled into the truck and Yarlon followed. Karn turned briefly and threw a cloak hook
on the ground before he joined them. Once inside he latched and locked the truck doors. There
were no proper seats and the windows were narrow and mounted high on the sides. A large skylight
let the indirect light of the sun into the space.

Yarlon opened a long box mounted half way up one wall and took out two autorifles, passing one
to Karn. Rachel shrank into the corner. Karn checked the magazine. He looked over at the human
and clicked his teeth. ``Just a little insurance,'' he said.

Karn swayed his head and Yarlon started to reach into the pouches in Rachel's clothing. Rachel
pulled away and made a sound like a short squeal unlike any sound a dunari might make. Karn
leaned his weight against her body and held her firmly while Yarlon dumped her belongings onto
the floor of the truck. There was her machine, a smaller rectangular device, and her oxygen
canister with its tubes and clips.

Karn picked up her machine and tentatively unfolded it. ``What is this?''

Rachel panted slightly but didn't reply. Karn tossed the device to Yarlon. ``Destroy it.''

Yarlon repeatedly folded the device in half but that seemed to have little effect on it.
Standing on one edge of the clear sheet he then tried to rip it. After some struggle he managed
to distort the shape of the device, stretching it and deforming it, but not actually tearing it.

``Good enough,'' Karn said. ``Now this one.''

The second device was smaller and thicker than the first. Yarlon put it on the ground and using
the butt of his autorifle hammered on the device repeatedly. After only a few blows it gave a
satisfying crack as the front surface collapsed into the body. Rachel looked on but said
nothing.

Karn thumped his trunk against the barrier behind the driver and then slid open the small
window. ``Stop at the canal,'' he called out and then closed the window again. He then picked up
the canister.

``I need that,'' Rachel said softly.

``What is it?''

``Oxygen. I need it for my breathing.''

``We are aware.'' Karn inspected the canister carefully. After a moment he slid it back toward
Rachel.

She picked it up, placed the cup over her mouth and took several rapid breaths.

``Now take off your clothes.''

Rachel stared at him.

Karn tilted his head. ``That wasn't a request.''

Rachel didn't move.

Karn reached into his cloak and took out his trunk pistol. He pointed it at Rachel. ``We need
you alive, but I'm guessing if I shoot you in the leg you'll live. Is that right?''

Still Rachel didn't move, and Karn slid back the latch on the pistol. With shaking fingers
Rachel started to undo the hooks of her clothing. There were two main parts, one covering the
top of her body and the other over her back legs. Beneath her outer clothing was another,
thinner layer. Karn insisted she remove that as well.

``Disgusting,'' Yarlon said softly at the sight of Rachel's body. She huddled in the corner with
her arms wrapped around her.

Karn passed her clothing to Yarlon who then started inspecting every fold. With a small knife he
tore through the seams and ran his tentacles along the material.

``Just making sure you don't have any concealed weapons or tracking devices,'' Karn said
casually.

When Yarlon finished Karn threw Rachel's clothes at her and she quickly slid them back on
despite the damage.

They rode in silence for a time. Finally the truck came to a stop and Rachel looked up. The
window slid open. ``We're at the canal,'' Gat called out. Karn took the two damaged devices and
put them through the window.

``Dispose of these.''

The door opened as Gat got out. Moments later he returned and started the engine again.

``Where are we going?'' Rachel asked.

``A Forbin military base,'' Karn replied.

``There are no Forbin bases around here.''

``Of course not. We're going to cross the western desert.''

Rachel's eyes widened. ``That will take too long! I can't be away from the compound that
long.''

``We won't tell.''

``You don't understand. I need water\ldots\ much more than you. I can't go without water.''

Karn leaned back against the side of the truck. ``There will be water at the base.''
