% Local IspellPersDict: aspell.dunari.pws

\chapter{59}

Fotkey stood on the observation platform overlooking the parade grounds. Behind and about him
rose the imposing ramparts of Fort Yarta, and beyond those walls the modern city of Yartagard.
Yet Fotkey concentrated on the grounds where herd after herd of highly trained solders marched
to the sound of their callers. Each herd stopped before Fotkey, turned in unison, and, leaning
forward slightly presented their weapons to him at ready. He scanned each group carefully before
swaying his head slightly. An aide at his side barked a command and the callers directed the
herd down the grounds to a large, arched receiving doorway.

It was a standard drill well known to the solders. Their charge had always been one of
preparedness. Yartagard was over four thousand karnons north and west of Varnok and was the last
major city in Argenia before the great desert. If ever Forbin should invade it would be the
troops stationed there that would first respond.

When the presentations were over, Fotkey turned and walked through a door behind the platform
into the controller space at the center of the fort. At one side he was flanked by sub-general
Torman, the highest ranking resident officer at Yarta. At his other side was his trusted aide,
two captains, and some assorted, minor officers.

``If it's war Forbin wants, we're ready,'' Torman said. Yet Fotkey did not reply.

``Will it come to that over this human?'' one of the captains said.

Fotkey paused and turned toward the captain. ``The First Citizen will act, as always, according
to the best interest of Argenia.''

``Of course, sir.''

They continued walking toward the Strategy Room where Fotkey intended to review deployment and
other issues of operational status. Just as they rounded the corner a communications officer
trotted toward the group.

``General! We have word that the First Citizen wants to talk with you on the secure phone in
just six arnets.''

Fotkey fluttered slightly. ``Fine. I'll go to my private office. Torman, I'll met you in the
Strategy Room right after the call.''

With only his aide a short distance behind him, Fotkey turned down a different hall eventually
reaching a small and unassuming office. The aide stood beside the outer door while Fotkey
carefully opened it and locked it behind him. Alone in a small, cramped foyer with a skylight
far above him, he opened the inner door to his private space.

The decor was sparse and functional. A few cabinets lined one wall and a small table stood in
the center with a secure phone at one corner. The windows were deep and placed high above him so
that even someone standing on the fort's ramparts outside could not reach him with a bullet.

Fotkey settled onto the seat and turned on the secure phone. He took a small book from a
mini-pouch slung around his shoulders against his skin under his cloak. He opened the book and
carefully adjusted the twelve dials on the phone according to the numbers on the first page,
together with a special number known only to him. He then ripped out the page and placed it in a
metal tray beside the phone where he burned it using a small lighter from one of the table's
trays.

He didn't have to wait long. At exactly the expected time the light on the phone began to flash.
There was no chime. Fotkey answered it.

``Hello, general,'' came the First Citizen's voice. There was a slight echo and some soft
squealing in the background as an artifact of the scrambling circuits. ``I trust all is well.''

``Yes, we're fine here. You?''

``We're fine here also.''

Fotkey relaxed a little. This simple, scripted exchange meant that neither side was under
duress. Had the First Citizen used the singular rather than the plural greeting a very different
sequence of events would have unfolded.

``I know you are in the middle of things,'' the First Citizen continued. ``However, I'll be
meeting with the ambassador earlier than expected and I wanted to talk with you first.''

``Yes, sir.''

``The human has been seen in Clickora, traveling west in the company of three men, at least one
of whom has a Forbin accent.''

``Clickora,'' Fotkey repeated, ``that's in aerocar range of Yartagard.''

``I want you to send an elite group to look for them.''

``Surely the humans will be looking as well,'' Fotkey said.

``It doesn't matter. If we find them first, rescue the human if you can.''

``Yes, sir.''

``Keep it out of Forbin at all costs.'' The First Citizen paused. ``General, you are authorized
to kill it if necessary to do that. Just\ldots\ make it appear to be an accident.''

Fotkey suppressed a flutter. ``Yes, sir. I understand.''

``You know as well as I do that with our weapon and that creature's knowledge we'll have a major
mess. It must not be allowed to happen.''

``The human will likely not cooporate with the Forbinites,'' Fotkey said.

``Maybe not, but we can't count on its ability to resist being broken---or tricked.''

``What of the IIB? This is more their purview.'' Fotkey asked.

``Yes, yes\ldots\ the IIB is having trouble tracking them. In this case I think we need to bend
the rules a little.''

\spacebreak

Karn rounded an outcropping of rock. A short way ahead was a small, dilapidated, two story
building. It was about the size of a house with a wooden sand wall, quartz windows, and a
tonstone step. Around the front and side was a covered porch. The roof was flat, typical for
desert construction, to cut down on the surface exposed to the abrasive wind.

Behind the house was a small garage, just as dilapidated. A large fuel tank, partly buried in
the sand, was pressed against one of the outside walls. The area was desolate and remote. They
had passed no other buildings or signs of civilization for many karnons. The Rujaran sun beat
down on the sandy yard, angling in from high above a rocky ridge behind the house.

Karn pulled the truck off the road and halted in front of the house. He banged on the wall
between the front seat and the back area. ``We're here,'' he said in a loud voice. Karn and Gat
climbed out and opened the back. Yarlon was lying on the floorboards beside the motionless body
of the human.

``Glad to hear it,'' Yarlon said as he started to drag himself out of the truck.

The men looked at Rachel and Gat poked her with the butt of his autorifle. She didn't move. Karn
wrapped his trunk around her legs and, leaning back, dragged the human to the edge of the
gateway. Rachel moaned softly. Her body was covered with a film of water and her clothing stuck
to her skin. It was a repulsive sight.

Karn took the knife from his thigh belt and cut Rachel's bonds. He then dragged her limp body
over the side of the gateway and she stumbled onto the ground. ``Get up,'' he said simply.
Rachel moaned again. Karn kicked her lightly. ``Get up now.''

She struggled. Her hair was matted against her forehead and neck. Her face and arms were covered
with wet grim. Gat wrapped his trunk around her arm and unceremoniously pulled her to her feet.
She leaned wearily against the side of the truck and looked at her captors.

``I\ldots\ need water.''

``Inside first.'' Karn pointed toward the house with his autorifle.

Rachel grabbed her oxygen canister and tried to put the cup over her mouth but Karn took it away
and pushed her forward. She plodded across the sand toward the building but paused after a few
steps. ``It's not an army base,'' she said.

``Inside.''

With an armed dunari at either side and another walking behind, Rachel made her way to the step
and into the building itself. The main room was filled with clutter. Several wooden storage
cabinets were strewn across the floor. Boards, buckets, used motor parts, oil covered rags, and
other junk was scattered about. Along one wall was a dirty cook stove, and in the middle of the
room was an open ramp that led through a hole in the ceiling to the upper level.

The men directed Rachel to the left through a wide door into a smaller room that was nearly
empty. In the middle of the room was a stained, torn seat. On the wall beneath the wide quartz
window, was a small table.

``Down.'' Gat pointed at the floor with his autorifle and Rachel got down and leaned her back
against the wall. Karn lay on the seat a short distance from where Rachel sat, her knees drawn
up against her body. Gat and Yarlon stood at either side of Karn.

Karn looked at Rachel for a time, watching her watching him over her knees.

``This is not an army base,'' she said finally.

``No.''

There was no trace of Forbin accent in Karn's voice. Rachel did not reply. Her eyes moved to Gat
and then Yarlon before returning to Karn. She hugged herself tightly.

``Why are we here?'' she asked.

Karn slid his tentacles into his cloak and took out his pistol. ``To kill you,'' he said simply.

Rachel shifted uneasily. Her breathing was ragged and her eyes glistened, but she said nothing.
Karn watched her. Then he raised the pistol and aimed at the center of her chest. Rachel stared
at Karn but did not move. Several moments passed and her body shook.

Karn lowered the pistol. ``You're braver than you look,'' he said. He nodded toward the other
two men. ``Find out what color its blood is.''

Gat and Yarlon stepped forward. Yarlon drew a knife from his thigh belt. Rachel tried to move
away but Gat grabbed her arm with his trunk and pressed one of his front feet down against her
hip. Rachel cried out and struggled but Karn stood up and raised the pistol again. Stepping
forward he held it close to her chest. ``Stop or it ends now,'' he said in a commanding voice.

Rachel panted. Water streamed down the side of her face and she glared at Karn. ``If you're
going to kill me anyway, why should I care?''

Karn leaned forward, holding the pistol to her head. ``Because in the end you're nothing more
than an animal, and all animals have an instinct to survive.''

Resigned Rachel stopped fighting against Gat's grip. Karn nodded toward Yarlon and he pressed
his foot against Rachel's other wrist. With a firm pressure he drew the knife blade down
Rachel's arm. A thick, red liquid flowed from the wound. Rachel grimaced but she didn't move or
make a sound. She panted rapidly, eyes open and staring at the ceiling, as Yarlon finished.

Karn watched her closely. He nodded. ``Very good.''

Yarlon poked at Rachel's finger tips with the knife. ``Maybe we should try cutting one or two of
these things off.''

``What are you?'' Karn said, ``A barbarian? We need a different kind of test.'' He looked
around, his eye falling on the door to the storage closet at the back of the room. ``Put it in
the dark.''

Rachel looked toward the closed door and then back at Karn. ``No,'' she pleaded. ``Please, I beg
you.''

Gat and Yarlon grabbed her arms tightly in their trunks and dragged her toward the door as Karn
unlocked and opened it. Inside was a rough closet with a few wooden cabinets and a pile of sand
jackets on the floor. There was no window and no artificial light. Inside, with the door closed,
the darkness would be absolute.

``Please, no!'' Rachel screamed as they dragged her into the closet. It took both dunari to hold
her inside as she kicked and thrashed against them. Finally Karn raised his pistol to silence
her. ``Learn to deal with your fear,'' he said as Gat closed and locked the door.

Rachel howled and started to pound on the inside of the door. ``Please! Please! Please!'' she
begged. ``Let me out!'' Suddenly there was a crash. For a moment all was silent but soon she
started again. She sobbed and moaned and there was more banging, this time it sounded like
against the wall. Karn stood quietly listening, his trunk against his forehead. Gat and Yarlon
waited.

There was more silence and then a groan and another crash. After that the silence was complete.
Karn waited but all was quiet. ``Have you mastered your fear?'' he said loudly. There was no
reply. He motioned toward Gat who opened the door.

She was gone.

Astonished Karn stepped forward. In the back corner of the closet, behind a wooden cabinet, one
of the vertical planks had been pulled out creating a narrow gap that lead into the space
between the main wall of the building and the outer sand wall. Karn moved to take a closer look
but the the gap was filled with the blackness of night. Instinctively he recoiled.

``Light!'' he said. ``I need light!''

Gat rushed into the main room and pulled a trunk-held artificial light from one of the cabinets.
Karn flicked it on and, taking a deep breath, kicked the broken vertical plank aside and leaned
into the gap to see.

It was much too narrow a space for a dunari but he could see how the human might have been able
to squeeze into it. Yet the creature was nowhere to be seen. In the red glow of the artificial
light he could see the rough cut boards of the outer wall on one side and the siding of the
original building on the other. Some distance away the gap made a right angle turn where it
wrapped around the side of the house and beyond that point he could see nothing.

``Dind,'' he muttered.

He pulled back and once again the hole filled with darkness. He snapped at Yarlon. ``Stand
here,'' Karn said. ``If anything comes through that hole, shoot it.''

Karn motioned to Gat and the two men trotted into the main room. Karn raised his autorifle and
started shooting at the wall. The powerful gun easily sent bullets through the wooden planks,
sending a shower of splinters into the air. Gat followed suit, starting at the other end of the
wall and worked toward Karn. The noise was deafening as the two autorifles fired round after
round through the dry boards.

``Down,'' Karn motioned. ``Aim low in case it's lying on the ground.''

More splinters flew through the air. Finally Karn went back into the storage closet while Gat
stayed in the main room. Grabbing the light he again looked into the gap. There was no sign of
her. ``More!'' he called out loudly. Gat fired his weapon, raking the wall with more holes. Karn
could see the bullets breaking into the gap as he watched. Finally he nodded.

``Enough!''

All three dunari gathered in the main room. The air was filled with the burnt smell of firing
mixture and the wall of the room was littered with bullet holes. Karn walked along the wall
kicking it just above the ground. About half way down the far side, the hollow sound gave way to
a heavy thump. Karn clicked his teeth.

``So that's it,'' Yarlon said.

``I want to see its body,'' Karn said. He grabbed the wall plank tightly in his tentacles and
pulled, but it held fast. ``Let's dig from outside.''

All went out into the open Rujaran sun and around to the side of the house where the body of
the human lay. Gat got some shovels from the garage and the three of them started to move the
sand away from the outside of the sand wall. The hot Nermella blew across their backs.

% Rachel emerges from the wall, goes upstairs, and takes refuge on the roof.

After a short time they came to the bottom of the sand wall and began to dig under it.

``It's not here,'' Yarlon said as his shovel found open space on the other side of the wall.

``Not here either,'' Gat said as he broke through a short distance away.

The men kept clearing away the sand finding more and more open space. There was no sign of a
body where they expected it. Karn had a sinking feeling. He threw his shovel on the ground and
motioned for the other two to follow him back inside. Once in the main room he reloaded his
autorifle and started shooting a stream of bullets into the wall, this time cutting a line a
short distance above the floor. Gat did the same and soon they had hacked off the lower part of
the wall all the way around.

Karn kicked the loose boards free. The ``body'' they had located before was nothing more than a
pile of desert sand that the Nermella had blown between the cracks in the outer wall.

``Light!'' Karn commanded.

He lay on the floor and slid under the broken remains of the wall to look around. The gap was
empty but supporting the sand wall there were several horizontal boards nailed to the both sides
of the gap. They were thin and frail, yet a tree climbing chortak would have been able to
scramble up them like a series of steps. Karn slid back out into the main room and looked at the
wall. At the top, near the level of the highest board there were no bullet holes. He slammed his
trunk against the floor in frustration.

``It must have escaped while we were outside,'' Gat said.

Karn stood back up, his ears tightly folded against his head, fluttering loudly. He faced
Yarlon. ``I told you to stand by that hole and shoot anything that came out.''

``It was dead!'' Yarlon said.

Karn curled his trunk and pressed it to his forehead. ``Okay,'' he said finally. ``Yarlon, go
outside and trot around the house. See if there are any fresh tracks leading away.'' Yarlon
swayed his head. ``Gat, she's got to be here. You and I will rip this place apart if we have
to.''

Yarlon turned to leave but Karn stopped him briefly. ``One more thing,'' Karn said. ``Don't
underestimate her again.''

Karn and Gat started opening any box or cabinet that looked large enough for a human to squeeze
into. At one point Karn glanced out the window to the truck parked beside the road. He felt in
his pouch and pulled out the key. It was the only key. He nodded to himself in satisfaction.

``Leave everything open,'' Karn said. ``If she comes back in here we don't want her hiding
somewhere we've already searched.''

While Gat worked on the first floor, Karn carefully ascended the ramp to the upper floor. Going
slowly he surveyed the space before taking the final steps. The second floor was in just as much
disarray as the first. There were boxes strewn about, several mining tools along one wall, and a
metal cabinet filled with old radio equipment. Karn moved about the room cautiously, checking
every potential hiding place. Against his back he felt the warm Nermella. He turned to find one
of the windows open. Yet on the floor beneath that window there was no accumulation of wind
blown sand.

Karn approached the window, with his autorifle raised. Staying well inside the room he looked
through the window first in one direction and then the other. It opened onto the narrow roof of
the porch, covered with broken hack tiles. Cautiously Karn stepped out on the roof. He looked
over the side at the ground but there was no sign the human had jumped. It was a full story to
the sand below. If she had been desperate enough to jump she would have surely broken a leg.

% Rachel is actually on the roof but since the dunari are poor climbers Karn does not even
% consider that possibility. In other words, he underestimates her again. :)

Karn climbed back into the room and returned to the first floor. ``Well?'' he asked Gat.

Gat just shrugged. ``Maybe it's in the garage,'' he suggested.

The two men went outside and trotted to the garage. There were many tracks around the house so
it was difficult to tell if the human had gone that way or not. The garage itself didn't have a
vehicle inside but was being used more as a storage area for junk. Along one wall was a drum of
gasoline and collection of tools. A pile of broken auto parts was in the corner along with some
oily sheets.

Leaning against another wall were several planks of fabri-wood. The space underneath them was
small and dark, but large enough for a human to hide. Karn motioned to Gat to move around to the
far side as he approached one end. Yet the only thing under the planks was a box of empty
bottles.

``There is nothing here and no place to hide,'' Gat said.

Karn kicked the wall in several places but it was obviously just a single layer of cheap
construction. It was amazing it was still standing at all.

% Back at the house Rachel has opened the propane valve on the cook stove and the valve on her
% oxygen canister and gone outside. She's hiding along the wall of the house opposite the garage
% and Yarlon.

``It can't have just vanished into hot air,'' Karn said.

Karn and Gat went back outside and Karn motioned to Yarlon who was standing some distance away.
Yarlon came trotting over to the other men. ``Well?'' Karn asked.

``Nothing,'' Yarlon said. ``It's got to still be here.''

``Then where in \textit{fran} is she?'' Karn fluttered in frustration. ``Let's take another look
in the house.''

Karn felt a need to again inspect the gap where Rachel had disappeared. Maybe she was still in
there, hiding in some nasty corner crawling with desert mites. The three men tore off more of
the inner wall and used two artificial lights to illuminate every part of that troublesome
space. Rachel was not to be found.

% Rachel has run over to the garage where she is making her Molotov Cocktail.

Gat and Yarlon checked upstairs again while Karn kicked the wall between the two rooms. When
they returned with nothing Karn stomped on the floor. ``What's under here?'' he asked.

``Sand?'' said Yarlon

Karn stomped again. ``Does that sound hollow to you?''

Gat was distracted. ``Didn't you put that gas bottle in the tray by the door?''

``What?''

``Why is it on the table?'' Gat picked up Rachel's oxygen canister and shook it slightly. ``It's
empty.''

Yarlon raised his trunk. ``I hear something.''

Everyone stood silently. From the other side of the room there was a faint hiss. Gat walked
toward it slowly at first and then more quickly. ``It's gas from the cook stove!'' he said. He
held up the line where it had been disconnected from the tank.

``Dind,'' Karn muttered. He sprang into action. Rushing to the nearest window he quickly
unlatched and opened it. ``Yarlon,'' he said turning back to the others, ``open that window near
you right away.''

At that moment a bottle filled with gasoline came flying through the window Karn had just
opened. Tied around the end was a flaming rag. Karn spun around just in time to see the bottle
smash against the edge of a metal shovel and burst into flame.

Years of training took hold and Karn flung himself to the ground. Almost instantly the entire
room was engulfed in a ball of fire. The very air he breathed was filled with scorching heat,
and the house shook with a thundering explosion. Fire roared over his back, engulfing him. In an
act of sheer desperation, Karn lept forward and out the window he had opened, still clutching
his autorifle tightly.

He landed heavily on the ground, rolling to absorb the blow. He got up quickly and backed away
from the house, now fully ablaze. He could hear cracking wood as the feeble building, weakened
by the explosion, started to collapse. He glanced briefly to either side, but there was no sign
of the human.

``Gat! Yarlon!'' Karn rushed around to the front of the house in time to see Gat stagger out the
doorway, his body burning. Gat felt to the ground and Karn quickly piled sand on top of him to
put out the flames. The odor of burnt flesh filled the air.

Karn turned toward the house and ran up the step. ``Yarlon!'' he cried out. He moved forward to
enter the house but heat radiated from the door as if from a furnace in a way that not even his
dunari body could withstand. ``Yarlon! Yarlon!'' There was no reply.

Karn ran back to Gat and pulled him away from the house. Flames roared over the building,
consuming the dry wood like kindling. A towering plume of smoke rose over the desert and into
the Rujaran sky, driven into the air by the Nermella.

Gat was badly burned. Every part of his body was blackened and blistered. He moaned softly.
``You're fine,'' Karn said. Gat opened his eyes and looked at Karn but said nothing. He was
incapable of speech. ``The pain will be gone soon,'' Karn said, ``and then you will be with
Jurita.'' Gat closed his eyes and never opened them again.

Karn looked around nervously. He did not like being in the open. He felt a need to hide. He
needed to regroup, rethink. He needed to plan. He reminded himself that he still had the
autorifle. His adversary was unarmed. He picked up his weapon with grim determination. He had to
hunt the creature down.

He ran some distance from the house and started to circle it in an ever widening spiral. He
stopped every dozen paces or so to study the scene. He paid close attention to the sand near the
house to see if it gathered human footprints each time he went around. Yet there were none.

All was strangely quiet. The house burned steadily, the wood crackling as it was consumed. There
was no sign of life. Karn knew that somewhere nearby she was waiting. Perhaps she was watching
him that very moment. He looked toward the garage. She had to be there.

He ran toward the small building, galloping as quickly as he could move. He reached the door in
moments and grabbed it tightly in his trunk. With a quick motion he slid the door aside and then
deftly swung the autorifle up. There was a noise toward the back. Immediately he aimed and
fired, sending a stream of bullets in the general direction of his target. The back door swung
closed, and Karn fired into the back wall for several moments. Then, stopping, he ran to the
outside of the garage.

He hugged the wall inching forward to the corner. The smoke from the fire cast ghostly shadows
on the sand about him. Folding his trunk back he started to peek around the corner of the
building. Suddenly, with a ferociousness he did not expect the human lunged at him. She swung a
heavy metal pipe, hitting him squarely across his trunk ridge and sending a stabbing jolt of
pain over his trunk and down his back.

Karn jerked back, reeling. Rachel lept around the corner and came at him again. Karn raised his
autorifle but Rachel grabbed the barrel in one hand and prevented him from pointing it properly.
Nevertheless Karn fired a few rounds, sending a blast of bullets into the wall of the garage.
Rachel swung the pipe with her other arm, hitting him again on top of his head.

Karn dug his feet into the ground and wrapped his trunk tightly around the autorifle pulling on
it powerfully. He dragged Rachel toward him. Yet her grip, far stronger than Karn imagined
possible, held fast. Rachel hit him again, and then again. Karn fell to his knees.

He fired the autorifle in an increasingly desperate attempt to dislodge the human, hoping that
perhaps the heat of the barrel would loosen her hold. Yet she did not let go. Instead she
pummeled him with the pipe. Each blow by itself wasn't enough to cause much damage, but the
effects accumulated. Karn was in agony and he started getting woozy. The world began to fade and
finally he let go of the weapon and collapsed on the ground.

Rachel jumped back, autorifle in hand. She threw the pipe behind her and pointed the weapon at
her one time captor. The autorifle wasn't designed for human hands and she held it awkwardly.
Her short arms couldn't properly wrap around the holds and her fingers barely reached the
trigger.

Karn panted heavily. The ground seemed to spin and he fought to stay conscious. Rachel took a
few steps back but otherwise didn't move, keeping the heavy autorifle trained on him. Finally
Karn looked up.

``It won't fire that way,'' he lied. ``You have it upside down.''

Rachel hesitated but didn't move.

Karn rubbed his head. Orange blood was running down the side of his face. Then, using his trunk
to support him, he slowly stood up. Rachel shifted nervously and took another step back. Karn
looked at her and sighed. ``What are you going to do?'' he asked.

``What do you think?'' she hissed. ``I'm an animal with an instinct to survive.''

She fired.
