% Local IspellPersDict: aspell.dunari.pws

\chapter{61}

Muldon drove the mini-truck along a rough path through the short brikken grass. A cloud of dust
rose behind him only to be blown quickly away by the desert wind. He crested a small rise and
stretching out before him and below him was a broad valley of low ridges covered with dark,
yellow grass. The rocky crags of the Handark mountains formed an imposing wall rising above the
far side of the valley floor many karnons away. Sculpted by eons of blasting sand the steep
flanks of the mountains were smoothed to impossibly rounded forms.

With a soft grunt Muldon stopped the mini-truck and took a compass from his pack.  Kneeling on
the seat he carefully read a bearing to two prominent knobs along the distance ridge. Satisfied,
he turned the mini-truck off the trail and drove, more slowly than before, over the sandy
ground. He raised himself up on his knees as he went, looking carefully ahead.  After a half
karnon or so, he stopped again and took some more bearings.  He swayed his head and turned off
the mini-truck.

Climbing off the small machine, Muldon threw an empty pack over his shoulder and then checked
his gun. He hooked it to the belt of his canvas cloak and started walking down the incline,
wading through the grass and leaving a trail of disheveled brikken stalks behind him. He paused
now and then and stood up straight to see as far ahead as he could. His dunari eyes were adapted
to see small objects hidden in the grass, and it wasn't long before he spotted them. It was a
field of branhask.

Muldon came to the first plant and knelt down to inspect it. ``Ah Jan,'' he fluttered. It was
immature. The skin had a yellow hue and had not yet darkened to the ripe orange color he wanted
to see. Despite the recent flashes he was still too early. He walked from plant to plant until
he finally found one that was ready. With a deft motion he cut the thick, succulent stalk from
the base and put it in his empty pouch.

Muldon found only a few other ripe stalks as he methodically searched the field. Periodically he
paused to scan the horizon in all directions. It was an instinctive response but it made sense
as well. The dunari were not the only ones attracted to branhask fields and certain other
creatures were not nice to meet. Muldon felt his gun to make sure it was in easy reach.

City people could buy their branhask in the food markets, already cut and peeled. However, in
Tunar Provence, far from the branhask farms, it was expensive to buy. Muldon had been taught
that it was stupid to waste money on something you could just drive out into the brikken grass
and cut yourself. Besides, Jillziah thought it tasted better wild.

After his fifth stalk, one that was still rather too yellow, Muldon paused again to look
around. Far to the west, just beyond the mountains, the sky had an ominous brown color. He
watched it for a few moments and then pulled a portable radio from his pouch. With a flick of
his trunk he extended the antenna fully.

``Maka?'' he said into the device. ``You there?''

There was a burst of static as the squelch control found its level.

``Ya, Paps?'' came the voice of his son.

``Looks like a dayenstat's coming.''

``I see it too, Paps.''

``We should get back.''

Muldon returned to his mini, and put the pack with the branhask into the basket attached to the
dragboard. Starting up the mini's engine he knelt on the seat and headed back the way he had
come. It was easy to retrace his path and soon he was racing along the rough trail over the rise
and down the hillside to where the family flat truck was parked. In the distance he could see
another cloud of dust approaching the same location.

He arrived at the flat truck a little before his son and had time to drive the mini up on the
bed and strap it into place. He was tightening the bands when Maka pulled up.

``Find anything?'' Muldon asked.

``Not much. It's all too young.''

``We'll come back after the next flash,'' Muldon said.

They strapped the other mini onto the bed beside the first, and then put their disappointing
collection of branhask into a single basket also strapped to the bed. ``It looks like more if
it's in one basket,'' Muldon said with a click of his teeth.

The two of them climbed into the flat truck's cab and started home. Muldon eyed the darkening
sky. A thick, brown cloud was crossing the valley blotting out the view of the distant
mountains. ``It's coming fast,'' he said. Maka looked at the storm but said nothing. Already the
the sand was starting to block the sun. Muldon felt chilled.

It was several karnons back to the main road but they were able to go a little faster once they
passed the turn to the copper mine. The road to the mine was well worn and well maintained. Many
generations of Muldon's family had been minors, spending most of their working lives in dark
tunnels lit with primitive gas fired artificial lights. It wasn't the kind of life most would
understand, but it worked for Muldon's father, for him, and for his son. Muldon glanced at Maka
with pride.

Sand started to blow. It slithered across the road and clouded the air. Muldon turned on the
artificial lights. Most dunari vehicles didn't have them but when you live in Tunar they were a
necessity.

By the time they reached the main road the dayenstat was almost in full force. A howling
wind dragged sand over the truck scraping away what little was left of the finish. Muldon turned
on the sand wipers and stared into the brown, dusty fog with his right eye.

``Jurita's witness,'' Maka said. ``Where did all this come from?''

``Quick to come, quick to go,'' Muldon said quoting desert lore. The saying was usually applied
to love affairs but the Tunarian residents knew what it was really about.

Maka adjusted the radio mounted into the frontboard of the cab. ``Ma? You there Ma?''

There was a few bursts of static and then a woman's voice came out of the speaker. ``Where in
seven are you?'' she said.

``We're on the main road, Ma. Should be home soon.''

``Don't get yourself buried out there, 'cause I ain't digging you out!''

Maka turned off the radio and clicked his teeth.

Muldon leaned forward peering more intently than usual into the blowing sand. ``You see that?''
he asked Maka. Ahead of them, beside the road, was a indistinct rectangular shape barely darker
than the air. Muldon slowed as they approached.

``Looks like a truck, or maybe a van.''

As they got closer it was unmistakably a vehicle parked partly on the road and partly in the
sand. Muldon pulled up behind it and stopped. Without any discussion the two men buttoned up
their sand jackets and then, taking a deep breath, opened the doors of the truck and plunged
into the storm.

The wind bit their faces harshly but their thick dunari skin was resistant to the damage. They
walked over to the van, leaning forward slightly to let some of the abrasive air pass over them
at least. The cab was empty. ``You think they tried to walk?'' Maka said in a loud voice.

``Let's hope not,'' Muldon replied. He went to the back of the van and tried the handle. It was
unlocked. Maka joined him as he opened the door. The van was mostly empty but curled up in a
ball in the far corner was the body of some kind of creature covered in a dirty, sandy cloth.

Muldon climbed into the van and cautiously approached the body. He wondered if he should get his
gun first, but the creature looked dead. Muldon reached out with his trunk to pull down the
cloth but he discovered, to his surprise, that it was actually some kind of clothing. The
creature moaned and moved. Muldon jumped back. Its front leg fell free and landed on the floor
of the van. It wasn't a leg like any Muldon had ever seen.

``Jurita\ldots'' Muldon said softly. He climbed quickly out of the truck. Maka stared at the
leg. Muldon closed the van door.

``What is it?'' Maka asked.

``Don't know,'' Muldon replied. He paused and then opened the door again. The creature hadn't
moved. Muldon climbed back into the van and Maka followed him. ``Be careful,'' he warned. Muldon
took as close a look as he dared, but didn't touch. The exposed leg was thin and frail with
strange, light colored skin. Its toes were long and thin and slightly curled. The creature's
head, at least he assumed it was its head, was covered with some kind of fibers. He couldn't see
its face.

``I think it's a human,'' Maka said.

``That's crazy.''

``I'm serious, Paps,'' Maka replied. ``They have legs like that.''

Muldon leaned over the creature. He could see that it was breathing. It was alive but obviously
sick. The two scrambled back out of the van again and Muldon closed the door. He went around to
the front and opened the driver's door. The key to the van was still in the ignition.  Climbing
in, he tried to start the vehicle. The engine responded but didn't engage.

``No fuel,'' he said.

Maka didn't need to be asked. Making his way through the blowing sand he returned to the flat
truck and opened one of the lower storage hatches. He pulled out one of the spare canisters of
gasoline along with a sand sleeve. With Muldon's help they emptied the canister into the van,
using the sleeve to keep the sand from getting into the van's gas tank.

``I'll drive this thing back to the house,'' Muldon said loudly to Maka. ``You take the truck.''

``No problem, Paps.''

The van started fine and Muldon made his way down the road carefully. The van didn't have any
artificial lights so he had to rely on the lights on the flat truck behind him. Fortunately the
dayenstat seemed to already be slackening.

Their house was a simple structure just off a bend in the road. It had an attached garage, an
important feature when storms raged, and a metal shop in back where Muldon did various repairs
for extra income. After entering the driveway he came out of the van and, with Maka's help,
opened the garage door. He pulled the van inside and closed the door tightly behind them. By the
time they had parked, Jillziah was standing impatiently in the doorway to the main house.

``First, what took you so long? Second, who's van is that?''

``You won't believe what we found,'' Maka replied.

Muldon was opening the van doors again and Jillziah came down the step to see. The three of them
stood at the end of the van looking at the helpless body lying there.

``It's sick,'' Muldon said.

``It's a human,'' Maka said.

Jillziah just stared. ``It can't be a human.''

Muldon climbed back into the van and leaned over the creature.

A small voice came from the doorway to the house ``Momma?''

``Go back inside, Jinna,'' Jillziah said firmly.

``What is it?'' the girl asked.

``Inside. Now.'' Jillziah went up the steps and herded her young daughter back into the house.

Maka got into the van again as well, and both dunari just watched the human for a time. Finally
it moaned and rolled onto its back, one leg flailed a bit and then its eyes fluttered open.
Muldon and Maka jumped back.

The human chittered some and looked around vacantly. Then it noticed the dunari and seemed to
focus on them. It backed away in fright for a moment and then stopped.

``Who\ldots\ who are you?'' it said softly in Argenian.

``Dind,'' Maka said to himself.

``I'm\ldots\ I'm\ldots\ I need water,'' the human said. ``Water and cold.''

``Maka, go bring some water out here.''

``Yes, Paps,'' he said.

``And bring some of Jinna's ice.''

The human tried to sit up but only grimaced in obvious pain. Muldon moved forward, uncertain of
what to do. ``Can I help?'' he asked. One of its front legs looked cut and was stained with a
dark reddish blood. There was a black mark on her other front leg that didn't look normal to
him.

The human struggled to sit up again and Muldon reached his trunk behind her to support her as
she did so. Eventually she managed to position herself so she was leaning against the side of
the van.

``Thank you,'' she said with a soft wheeze.

The door to the main house opened and Muldon could hear his wife's voice. ``What's this about
Jinna's ice?'' she asked. ``Don't you know how expensive that stuff is?''

Jillziah came around to the back of the van and stopped dead.

Rachel looked at her silently for a moment. ``I\ldots\ I have some money\ldots''

Jillziah briefly raised her trunk to her forehead. ``I'll get water.''

``Ice too,'' Muldon called after her.

When Jillziah returned she had a pan of water and a small insulated bucket with a few ice bricks
in it. Muldon passed the water to Rachel who struggled to hold it. Muldon held one end with his
tentacles while Rachel sipped water from the other. She grimaced again.

``Are you in pain?'' Muldon said.

She touched the black mark on her arm. ``I hurt myself.''

``What are you doing here?'' Jillziah asked.

``I was\ldots\ kidnapped. I escaped.''

Muldon glanced briefly at his wife while Rachel sipped some more water and then reached for the
ice. Jillziah took some tongs from her pouch and tried to pass them to Rachel but the human
ignored them. Muldon moved the bucket closer and to their amazement Rachel put her hand into the
bucket and grabbed a small ice brick without tongs. She held the ice tightly in her fist, and
leaned back in pleasure as the coolness flowed into her body. The ice melted quickly and she
raised her cold, wet hands to her forehead and then ran them down her arms. The water rinsed off
somewhat the dirt on her face and arms but when she got to her bloody injury she hissed with
pain.

``I'll get something to bandage that,'' Jillziah said.

Rachel slowly regained some strength. Muldon called the provincial police on the radio and Maka
got an old sleeping pad from their storage area. Muldon and his son helped Rachel get positioned
more comfortably on the pad. It was dusty and dirty but she seemed appreciative. Jillziah
returned with sand cloths to wipe up some of the grim and a few medicated pads. Rachel insisted
on using some water to clean her wound but then let Jillziah gently wrap a pad around her arm.

Rachel drank several cups of icy water, slowly at first but with greater gusto as she improved.
Muldon had never seen anyone or anything drink so much. Yet her breathing wheezed and she was
obviously very weak. 

Eventually the deputy arrived, and Muldon ushered him into the garage. He was wearing his
official uniform cloak of the Tunarian police and carried a small trunk pistol in his front leg
strap. When he came to the open end of the van he froze when he saw the human.

``By the seven sands, you weren't kidding!''

``We are well met,'' Rachel said weakly.

``It even talks,'' the deputy muttered.

``It\ldots\ she was kidnapped,'' Muldon said.

The deputy looked at Rachel and she swayed her head slightly.

``Who kidnapped you?''

``They pretended to be Forbinites but I don't think they were.''

``Forbinites! \textit{cartak!},'' the deputy stepped back. ``This is way over my head. I'll have
to call it in.''

``Can you get a message to ambassador Marshall in Varnok?'' Rachel asked. ``They will be looking
for me.''

% TODO: The reader might not remember who Marshall is. At this point in the story he has had a
% small part.

The deputy went back to his vehicle to use the radio there, and Rachel closed her eyes briefly.
When she opened them again Jinna was standing quietly in the doorway of the van, and staring at
the human. Muldon had his trunk over her shoulder as she leaned against his leg.

Rachel bobbed her head to the girl. ``Thank you for the ice.''

``I didn't need it,'' Jinna said. She looked up at Muldon.

``It's for her medicine,'' Muldon explained, ``but those injections have already been used.''

The deputy returned and told Rachel that he sent a message to the provincial police main office.
From there Varnok would be contacted. ``We just need to wait now.''

With Muldon's and Maka's help Rachel managed to get out of the van and up the ramp to the inner
doorway. They led her into their home and helped her get comfortable in a seat in their living
area. Jinna, with braces on one of her front and rear legs, limped behind them.

At first Rachel spent some time giving information to the deputy but it wasn't long before the
deputy was talking with Muldon about the state of the mines, the timing of the next flash, and
if his daughter should try to form a seleck, ``like those city people do.'' Maka tended to the
flat truck while Jillziah busied herself peeling branhask and keeping Jinna out of the way.
Rachel dozed.

When Rachel awoke many moments later Muldon and the deputy were nowhere to be seen. Instead,
standing in the doorway to the hanaria was Jinna watching Rachel intently. When she saw the
human was awake she limped over to her seat.

``Those are pretty ribbons,'' Rachel said remarking on the decorations on Jinna's braces.

``Seven colors,'' Jinna replied. She took each ribbon one at a time between her tentacles.
``Um\ldots\ light orange is for Jinjana, dark orange is for Ji\ldots\ Jixita\ldots'' Slowly and
carefully she recited a name for each of the colors while Rachel listened quietly.

``Ma says they watch over me,'' the girl explained when she finished.

Rachel bobbed her head slightly. The tangy smell of baking branhask faintly filled the room.

``Do the Seven come for you too,'' Jinna said, ``when you die?''

Rachel looked at her for a moment. ``I don't really know, but I think so, yes.''

``Jinna!'' Her mother stood in the doorway. ``Don't be a pest. Leave our guest alone.'' 

At that moment the front door opened and Muldon came bustling in. ``Your people are coming for
you,'' he said to Rachel, ``they're gonna land one of those\ldots\ those aerocar things on the
road right here in front of the house.''

``On the road?'' Jillziah said.

Rachel moved to get up. ``I should wait outside.''

``Jinna, call Maka,'' Muldon said and his daughter scampered out of the room as best she could.
Muldon held out his trunk and let Rachel use it to steady herself as she stood. With Muldon's
and Maka's help she made her way down the steps and out onto the sandy yard toward the road. The
sky had cleared and was a beautiful golden green. The mountains to the west were clearly visible
and the wind from the north and was strangely cool.

``The Nermellum,'' Maka said.

When they reached the road Rachel knelt between Muldon and Maka to rest. Yet they didn't have to
wait long. In a few arnets a fast moving streak, greener than the green sky, could be seen
coming from the southwest. In only moments, it seemed, it was upon them. The craft made the
transition a safe distance away and then circled around the house twice. Finally, with awesome
maneuverability it swung into its final approach, descending to the road as it rushed toward
them.

Gripping Muldon's trunk Rachel stood up again, resting her arms against the backs of Muldon and
his son. They waited until the human craft came to a stop some distance down the road and then,
with their help, Rachel made her way toward it.

A small platform descended from the bottom of the aerocar and several figures began to trot
toward them, each on just two legs. Following behind was a kind of long, narrow, stretcher that
moved with the help of four mechanical legs attached to each corner. As they approached it was
clear the humans were wearing some kind of protective suits. Even their heads and faces were
covered.

Neither Muldon nor Maka moved. They remained at Rachel's side supporting her. When the humans
got to them, they chittered and hissed in their strange language. Rachel responded in kind,
talking to them with her alien voice. They took her from the two dunari and helped her lay on
the stretcher. They examined her eyes, inspected the bandage Jillziah had prepared, and attached
some kind of exotic device to Rachel's arm.

Just as they were about to put a breathing cup over her face she turned toward Muldon and Maka.

``Thank you for everything,'' Rachel said. ``May Jurita be with you.''

The humans then whisked Rachel back to the aerocar. Muldon and Maka stood together at the side
of the road as the craft lifted off, spun around, and streaked away, leaving behind only the
golden green sky and the cool Nermellum wind.
