% Local IspellPersDict: aspell.dunari.pws

\chapter{67}

Colty walked up the steps along the side of the Cogart lecture hall and found his usual seat in
the back. The students chattered among themselves much as they always did yet there was a
palpable energy in the room. Two of the students directly in front of Colty turned in their
seats. ``Sar. Coltinarly, she is alive?'' one asked.

Colty swayed his head. ``Yes, thankfully.''

``We heard the Forbinites got her.''

``I don't know,'' Colty replied, ``I've hardly spoken to her.''

The clock moved forward, and the students grew more agitated. Nobody expected her early, of
course, and the delay just increased the excitement in the air. Some students waited with their
notebooks ready, as usual, but most talked to their neighbors in passionate tones.

Just then the door swung open and Rachel walked in. Right behind her a man dressed casually took
a post beside the door as unobtrusively as possible for an IIB bodyguard. Rachel wore tan
clothing with red trim. A bandage ran up the length of one arm but she otherwise appeared
perfectly healthy.

She walked briskly across the floor and up the steps to the podium but before she reached the
platform someone in the back of the room started to stomp. Soon the tentative sound was joined
by another, and then another, from all corners of the room. Rachel reached the platform and
turned uncertainly as the stomping spread across the hall like fire on a field of brikken grass.
Colty gripped the table in front of him and the floor vibrated from the thundering ovation.

Rachel brought her hands to her face. She shuddered, blinking her eyes, yet the stomping went on
and on. After several long moments she raised her hands, shaking them slightly to discourage the
sound. ``Thank you,'' she said. ``Please\ldots\ please\ldots\ thank you.''

The stomping stopped and a profound silence filled the hall. Rachel closed her eyes for a
moment. Finally she addressed the class.

``I\ldots\ apologize for missing some of our sessions,'' she begin. She took another breath and
then walked to the other side of the platform. The class watched her intently, listening to her
every word. ``I'll have to take a few things out of the schedule to catch up, but I think we can
still cover all the important topics despite my absence.''

Casually she walked to the center of the platform and put one hand on the desk, her fingers
gently touching the top surface. She turned and looked at the class.

``Shall we begin?''

\spacebreak

When Narlock arrived at the IIB facility, he found ambassador Marshall waiting for him in one of
the interview rooms.

``He just got here,'' Narlock's secretary explained. ``He's very intent on talking to you.''

Narlock fluttered. Despite the human, he stopped at his office first. He wasn't about to be
rushed. After organizing his table and reviewing his schedule he took some writing materials and
went to find out what the ambassador wanted.

The two greeted each other politely and then got down to business. Marshall leaned his front
legs on the top of the table the way humans tended to do and spoke. ``I'm anxious to know
what you've learned about the note Rachel took from the body of her captor.''

Narlock didn't answer immediately. Legally there was no reason to give the human any information
and maybe even some good reasons to avoid it. Although normally an excellent judge of character,
Narlock found the strange creatures almost impossible to read. Despite directives from the IIB's
governing board he felt no particular inclination to trust the ambassador. Instead he made a
show of carefully clipping his pad to the table and unsheathing his pen before he replied.
Marshall's impatience was obvious even to him. He didn't care.

``It does appear to be a Junar assassination order.''

``Naming Rachel?''

``Well, yes,'' Narlock fluttered slightly. He paused and tilted his head a bit.

``What is it?'' Marshall asked. Narlock might have had difficulty reading the human, but
Marshall's experience with the dunari made it easy for him to read Narlock.

``Understand that the primary investigation is being handled by the Military Investigative
Service, not the IIB. The presumption is that this was a Forbin act; my involvement is only
a\ldots\ formality.''

``I see.''

Marshall tapped this front toes together for a few moments. ``Rachel's impression was that they
were Junar pretending to be Forbinites.''

``And the MIS believes they were Forbinites trying to frame the Junar.''

``Wouldn't it be better to find the truth rather than presume it?''

Narlock fluttered loudly and tossed his head slightly. ``We have our way of doing things, and it
works. I'd appreciate it if you let us do our job. Don't worry. We will find the truth.''

``Yes, but now my people are at risk. I hope you can understand my desire to\ldots\ help in the
investigation.''

Narlock swayed his head and turned to look at Marshall with his right eye. Marshall took a small
photograph from his pouch. It was printed on a kind of shiny, slippery paper.

``This is a picture of the note,'' Marshall began. Narlock turned the photograph so he could see
it.

``That orange splotch is dunari blood,'' Marshall continued.

``Yes, of course it is. That is standard for Junar orders. They are legitimatized, as it were,
by the blood of the leader issuing the order. The assassin who carries out the deed must have a
so-called 'legitimate' order on his person when he does it.''

``So if we can identify who's blood that is, we would know who was responsible for the order
against Rachel's life?''

Narlock snorted slightly. ``In theory, yes. But one quarter of all dunari have the same blood
factors as are in that sample. We already checked them. A match would prove nothing.''

Marshall nodded. ``Using our technology we mapped the DNA of this blood sample exactly. Since
every dunari's DNA is unique finding a DNA match would definitely identify the perpetrator.''

Narlock thought for a moment and then wrote a few notes on his pad. ``When did you do this?''

``When?''

``Yes, when?'' Narlock fluttered.

``I'm not sure, but we do have that information recorded. Why?''

Narlock put down his pen. ``We have something called the Rules of Yotal that limit the kind of
evidence that can be gathered before an investigation is started. Your analysis might be ruled
out of order in court.''

Marshall leaned back slightly. ``We could do the analysis again if that is necessary.''

``Jurita's witness,'' Narlock said softly, ``the whole thing is a blumber. The defense would
make a game of it, what with violations of chain of evidence, use of untested technology, and so
forth. Wouldn't you also need a sample of the suspect's blood to make the match?''

``Yes, or any biological material. A scale or even a toe shield would be fine.''

Narlock rumbled. ``If it could work legally, it would be helpful. I appreciate what you're
trying to do. I'll look into it.''

When Narlock returned to his office he organized his notes carefully. He considered contacting
the MIS office at Marvia but he decided to wait. He didn't want to lose what little initiative
he had. Instead he called his secretary. ``We need a lawyer well versed in evidence
procedures,'' he told her. ``Who can we get?''

\spacebreak

The mood in Parliament was tense. Even before the hearing began several of the delegates were
speaking loudly to each other in voices that could easily be heard across the hall. On the floor
Fotkey, and Lucasa waited patiently behind a table to give their testimony. To Lucasa it seemed
like only a few short moments since he was last in that hall talking about the humans. He
glanced over to Fotkey but both men remained silent. Fotkey looked grim.

The Speaker of the People called Parliament to order. Yet one delegate couldn't help but cry out
excitedly, ``There will be war before we are done here!''

The Speaker raised his voice to gain control of the assembly. ``Gentlemen, our purpose is to
hear discussion regarding the Forbin interference with our affairs. Withhold any motions until
after the initial statements are complete.''

The Speaker asked Fotkey to give his statement and the hall grew quiet. Fotkey stood up and
walked to the center of the floor, directly beneath the rostrum, disregarding the microphone on
the table. He spoke with authority and power as any general might. His voice seemed strangely
large coming from his small body.

``The facts of this case are confusing,'' he began. ``Some physical evidence does implicate the
Forbin government\ldots''

Loud bleats erupted from the delegates, particularly the younger ones seated in the back.
Fotkey continued, raising his voice a bit more.

``Yet the hostage recovered a Junar execution note from the body of one of the kidnappers naming
the human as the target. The holder of the note was none other than Karnolia, a notorious Junar
assassin.''

``Everyone knows Karn is a mercenary,'' cried one delegate, ``who will work for whoever pays the
most. He knows the Junar ways. If the Forbinite dinders wanted to confuse us, who better to do
the job?''

``Besides,'' called out another delegate, ``what would the Junar want with the human? There is
no sense to that.''

``But what would the Forbinites want with the creature?'' called out another voice. ``There are
already humans in Forbin anyway.''

Fotkey raised his trunk and stretched out his tentacles to draw the attention of the hall.
``Forbin has, in the past, conducted operations deep inside Argenian territory just to goad us
and to demonstrate their imagined superiority.''

Zarlon stood by his seat and raised his trunk high, signaling his desire to speak. Unlike the
rabble around him, Zarlon continued to honor the Protocols of the People. The Speaker stomped
his foot heavily to gain the attention of the hall and then said, ``We recognize Zarlonloti of
the Limzar Association.''

``Thank you, Sir Speaker. I submit that the real reason Forbin singled out the human is for the
creature's knowledge. I have read the reports published by your office, general, and it is clear
the Forbin government is working on some kind of weapon of enormous power.''

``Our intelligence,'' Fotkey interrupted, ``is that their program is dead. The weapon they seek
is nothing more than a fantasy, and they know it. You would also know that if you truly read
those reports.''

``Ah, but would it be a fantasy if they had access to human knowledge? Perhaps your intelligence
is faulty. Perhaps they are developing weapons even now, with or without human help, that
threaten our security!''

Fotkey fluttered and tossed his head. ``If I didn't know better, I'd say the Limzar Association
was trying to incite a war.'' He raised his voice to head off the clamor. ``This investigation
has been under the auspices of the Military Investigative Service. I'm not at liberty to discuss
all aspects of this case publicly but there are other lines of evidence that are distinctly
incompatible with the idea of Forbin responsibility.''

``It's all very well,'' a delegate in the back called out, ``for you to invoke secret military
proceedings, but how does that help us? We can only base our actions on information brought
before us.''

``The humans are the cause of this,'' Zarlon continued. ``Just having that creature freely
walking the streets is a lightening rod for trouble. It puts our citizens in danger simply by
existing.''

``Yea be yea!''

Zarlon took this as license to continue. ``The knowledge the alien is presumed to have makes it
a tempting target for enemies of all kinds. We're just lucky that in this instance we have not
incurred the wrath of these creatures. Jurita only knows what destructive power they might be
able to wield.''

The noise in the hall increased and cries of ``yea be yea'' filled the air.

``It is the humans who will incite a war, if not with their own actions, then with their
existence in our midst. How long will it be before the Forbinites try again? We must dis-invite
them from our world and send them back from whence they came.''

The Speaker was about to bellow to gain control of the assembly when Lucasa leaned forward and
spoke loudly into his microphone. ``Sir Speaker, may I address the concerns of the esteemed
delegate?''

Lucasa adjusted his papers. The hall quieted. Finally he spoke. ``I am, frankly, disgusted with
these proceedings\ldots''

The hall erupted into loud objections. Voices overlapped. Delegates called out. The Speaker
bellowed several times times before the room was brought under control.

``Lucastanonia,'' the Speaker warned, ``you are not to disrespect this assembly.''

Lucasa leaned forward into the microphone again. ``I apologize, Sir Speaker. Yet to send the
humans away would be a crime against the dunari people for which our progeny would never forgive
us.''

He started to speak more loudly. ``But more than that it would also be acquiescing to the
demands of a terrorist organization. Surely that is the real agenda of the Junar\ldots\ to have
us reject the humans and set us against our adversaries out of fears grown in our dark past. Is
that the action of a brave and powerful nation? Is that the action of a people who reject
terrorism in all its forms? Or is that the action of a nation of cowards?''

Loud cries were raised, but Lucasa ignored them and continued forcefully. ``There is much talk
about the bravery of Argenia in this hall. Yet if you want to see actual bravery you apparently
have to go to Parnon.'' He pointed his trunk in the general direction of the city. ``There she
who almost died in the desert is teaching her classes \emph{as if nothing has happened}.''

Lucasa gathered his materials together. ``I must apologize again Sir Speaker, I can not abide to
remain here.'' He stood up, and amid a storm of cries and bleats and bellows, walked out.

\spacebreak

% TODO: Make a transition here without a \spacebreak?

Lucasa went quickly down the path leading from Parliament and across the Yellow. The flowers
beside the walkway caught his eye, and he slowed to look at them. Finally he stopped in front of
a lovely bed of garlangias, their large red blooms and thick yellow leaves danced in the
Nermella. Lucasa clicked his teeth. A short distance away a small fountain surrounded by green
harnolick flowers bubbled playfully.

He took a deep breath and looked back at the Parliament buildings. They were old, imposing, and
impressive. Flying from the top of the main building the Argenian flag flapped proudly in the
wind. On the opposite side of the Yellow from Parliament rose the city of Varnok. Lucasa could
hear the sounds of traffic drifting on the breeze. Finally he looked into the green sky,
punctuated with a few rubinum clouds. He knew that somewhere far above \textit{Golden Light} was
flying silently over their troubled world.

Lucasa continued down the Yellow, more slowly, taking in the flowers as he walked. When he got
to the street he took a bus to the Office of Alien Affairs and walked to his space on the second
floor. Jinlo, his secretary, was surprised to see him.

``The hearing ended early,'' Lucasa said. He went into his private office. Just before closing
the door he turned and said, ``Oh, and I think I lost my job.''

Lucasa lay down beside his table. It was a very nice table. It had lots of space and nice trays
and drawers. He opened the drawers one by one. He had a lot of junk he'd have to throw out when
he left. He spun around and looked out the window at the Nizon Bank across the street. He
wondered how much thought the people there gave to the humans. Did they care at all? Did they
wonder about the future of Rujar or of the dunari race? Or were they content to focus on only
getting by day to day? Lucasa knew that could be his focus now as well. Maybe he could teach.
Maybe he could write a book.

He pressed the intercom button on his phone. ``Jinlo, could you come in here?''

She appeared at the door in moments.

``Would you, by chance, want something to eat? I have a little rigat fruit in here.'' He opened
one of his drawers and took out a small box.

She clicked her teeth. He opened the box and offered it to her. She took one of the fruits and
lay down on the opposite side of the table to eat it. Lucasa chewed his fruit absent mindedly.

``What happened?'' Jinlo asked.

``I couldn't listen to fear any longer.''

Jinlo nodded slightly. ``You admire her, don't you?''

They ate their rigat fruits in silence for a while. Lucasa offered her another but she shook her
head.

``I think your job is safe,'' he told her. ``They'll find someone else for you to work under.''

``I don't want to work under someone else.''

At that moment the phone chimed. Almost instinctively Jinlo answered it. She spoke for a moment
and then passed the phone to Lucasa. ``It's the Secretary of Parliament,'' she said. Lucasa took
the phone and Jinlo left to give him some privacy.

``I just wanted to let you know,'' the voice on the phone said, ``Parliament voted against
declaring war on Forbin\ldots\ and to keep the humans here as our guests.''

``That's wonderful.''

``Also, some people are saying you should run for First Citizen.''

Lucasa shook his ears. ``Oh? Well, that's the last job on Rujar I'd want.''

After he hung up the phone he clicked his teeth slightly. He pressed the intercom button again.
``Don't pack your bags just yet, Jinlo.''
