% Local IspellPersDict: aspell.dunari.pws

\chapter{71}

Narlock disliked the dungeon. It smelled funny.

The dungeon was the name given to the lowest floor of the IIB building. It was below the level
of the street but still had windows in the back that looked out over the Varsynthia. In other
parts of the city such a view might command a high value, but the IIB offices were in a
renovated manufacturing plant that once used the power of the river to drive saws and lathes.
Rusted chutes and metal frames still sprouted from the water and anchored to the foundation of
the building, too expensive to properly remove. Instead of being surrounded by parks and gardens
as in the commercial district, the river there was lined by decaying brick work and a crumbling
access road.

Narlock sat in the office of a young lawyer on the IIB legal advisory team, waiting with rising
impatentice while the lawyer collected his notes. The east facing window at his back had no
direct sun and the room seemed dim and gloomy, an effect enhanced by the dreary view outside.
Taped just below the top edge of the window was a small film drawing of the sort a young child
might make.

``This certainly is an interesting situation,'' the lawyer said looking up at Narlock at last.
``I don't think there's ever been a case quite like this.''

``You don't think?'' Narlock said with an edge of sarcasm. The lawyer either didn't care or
didn't notice. The later case Narlock found more likely.

``I've reviewed your notes and the problem as I see it is twofold. First the aliens are not
under our jurisdiction and thus not subject to our purjury laws. There is no treaty in force
between Argenia and, ah, the humans. And second, the alien technology is unknown to us and won't
be acceptable to the Standard Citizen.''

Narlock squirmed a little. The seat was not very comfortable. ``The ambassador from Earth has
stated humans on Argenian soil will be subject to Argenian law.''

``Yes, but it doesn't really mean anything until a treaty act is passed and signed by both
governments.''

``What's the problem with the technology?'' Narlock asked.

``It isn't reasonable to ask the Standard Citizen to accept technology that isn't understandable
by us. Precedent requires the nature and limitations of any technology used to acquire evidence
be comprehensible so the quality of the evidence can be properly evaluated.''

Narlock suppressed a flutter and shifted in his seat again. ``I spoke with a biology professor
at Parnon and he is convinced that what the humans claim to have done is possible in principle.
He even agrees that a DNA match would identify the owner of the blood with 100 percent
certainty.''

The lawyer swayed his head. ``The problem is the Standard Citizen test. You can try to use the
professor as an expert witness, but it isn't his technology he'd be testifying about. The
Standard Citizen can't rely on his word about it. One way or another he would have to rely on
the word of the humans and legally they aren't trustworth.''

This time Narlock did flutter. ``Fine. I get it. So this means I can't use DNA evidence, yes?''

``It would be a very difficult case to make.''

``Right.''

``Maybe one day when we have such technology ourselves it will be different.''

``I can't wait that long,'' Narlock said. He stood up and gave a brief nod to the lawyer, turned
and left. At least he didn't have to spend any more time in that smelly basement.

\spacebreak

Colty heard a light tapping at the door frame of his office. At that moment he was focused on
his paper about the primality of co-sequences, fussing with the wording of a particularly tricky
passage. ``Yes?'' he said finally looking up. It was Rachel. ``Come in, come in,'' Colty said
excitedly curling his trunk toward him.

``I'm hoping we can walk, maybe down by the river?'' she said.

Colty swayed his head and hastily slid the papers on his desk back into his notebook. Then they
made their way out of Fargon and along the campus pathways.

``I heard it wasn't Forbinites that\ldots\ well, that took you,'' Colty said.

``They were Junar.''

``Junar?''

``Honestly, I don't want to talk about it.''

The two walked silently for a time. Rachel's IIB bodyguard kept a few paces behind but otherwise
stayed out of the way. Colty discretely eyed the bandage on Rachel's arm but said nothing.

Some students walked by and greeted Rachel warmly. She thanked them graciously, as always. Colty
clicked his teeth as he watched. When they left, Rachel turned toward him and said, ``There's
something I want to show you?''

``Yes, of course,'' he replied.

She pointed toward the narrow path that ran along the river. ``Let's go down there.''

As they turned off the main path the bodyguard followed more closely behind. Colty felt nervous
having him there. ``Everyone insisted,'' Rachel said glancing over her shoulder. ``I even agreed
to get a tracking implant. They told me I couldn't come back unless I did.''

Finally they came to one of the seats along the path. ``Let's lie down,'' she said. The
bodyguard stopped a few paces behind and stood beside the path leaning against the trunk of a
tree.

The weather was beautiful with a clear, green sky. The embankment beside the path fell away in
front of them down a grassy slope to the edge of the river. The brush that normally crowded
along the water had been partially cleared leaving a small space where the grass touched the
swirling flow. Large rigat trees arched overhead, their leaves gently humming in the ever
present breeze.

Rachel took a folded piece of paper from her pouch and gave it to Colty. On it were several
colorful bands printed horizontally across the page. The bands were many thin vertical lines
densely packed together, with each line printed in its own color producing an overall effect
like that of colored confetti. Here and there were thicker black lines spaced in irregular
intervals that partitioned the other lines into random sized blocks.

``It's a fragment of the dunari genome,'' Rachel explained. ``The lines represent individual
nucleotides.''

``Where did you get this?'' Colty asked as he looked at the image with interest.

``It's a long story but it was taken from a sample of dunari blood. The biology lab on
\textit{Golden Light} mapped the full genetic structure of whoever's blood it is.''

Colty studied the image.

``You see,'' Rachel continued, ``this fragment is special. We call it \textit{Kessler-17}. This
very same fragment is in the genome of all vertebrate animals on Earth. It's in my genome as
well.''

Colty reflected on the significance of Rachel's words. It seemed almost too incredible to
believe. He cocked his head slightly toward her. ``Does that mean we're related somehow?''

``I don't see how. Our biology is incompatible. We use different nucleotides. It's just the
pattern that's the same.'' She leaned over a bit to look at the image herself. ``Our lead
biologist on \textit{Golden Light}, his name is Brendon, explained it to me. Not all genes code
for proteins. The ones that don't are called non-coding sequences. Yet they mostly do have other
purposes, or at least are understandable as ancient viral genes or evolutionary cruft. Some
non-coding sequences are passed faithfully from generation to generation over long time scales.
They are said to be ultra-conserved.''

Rachel shifted slightly and looked over the grass toward the river. ``The Kessler sequence is
the only ultra-conserved non-coding sequence we can't explain. Now we find it on Rujar in dunari
blood. It's hard to understand.''

Colty briefly turned over the page as if to search for some explaination on the back, but it was
blank. ``You said all vertebrates on Earth have this?''

``Yes,'' Rachel said, ``it must have arisen on Earth at least three billion runion ago. Brendon
wants to find out if it exists in other Rujaran life forms. He looked at the bacterial samples
we used to check for biological compatiblity, but the Kessler sequence doesn't show up there.''

Colty handed the paper back to Rachel but she waved him off. ``That's for you,'' she said. ``I
thought you'd be interested.''

They sat quietly for a time while Colty digested the startling revelation. They watched the
swirling water of the Varsynthia move slowly past.

``It pleases me to think,'' he said finally, ``that we might be cousins somehow.''

\spacebreak

Fotkey's office at Marvia Base was large and plush. His table was polished to a fine sheen and
on the floor before it was a wide, circular harnskin rug. On the south wall, and almost the
length of a dunari, was a large map of Argenia. On the north wall was a similar map of Forbin.
Elsewhere the room was decorated with pictures of Fotkey with various first citizens, the fort
at Yartagard, and even Fotkey's yellowed and tattered certificate of achievement for completing
his initial training as a third class herdsman. Behind his table was a broad window that looked
out onto the Marvia campus, and in one corner a heavy, locked door made of steel separated
Fotkey's office from the document room where the secrets of Argenia were stored.

Fotkey lay at his desk scanning over routine reports when a sharp kick sounded at his office
door.

``Enter.''

The door slid open and his guard leaned in. ``A courier, orange class, here for you, Sir. His
papers check.''

``Yes, yes.''

A small, lean man came into the room wearing the uniform of the Courier Service. He had an
orange band down the left side of his cloak running from his shoulder to just above his left
front knee. He waited discretely until the door behind him slid closed again.

``I have reports from the realm,'' he said formally.

Fotkey didn't bother with the formal reply. ``Show me.''

The courier unclipped a flap in his cloak and deftly untied several knots to loosen a thick
pouch. He approached Fotkey's table and lay the pouch on it more or less directly in front of
the general. With a practiced motion, he unlocked it and slid a diverse collection of papers
onto the table and then held up the empty pouch, open for inspection.

% TODO: Opening the pouch should be a bit more complicated. Locks?

Fotkey waved him off with a gesture. He began to flip through the reports, setting them to one
side or another to organize them. He made sure everything was in order before he released the
courier.

After bending his left knee slightly in respectful salute, the courier turned and left without
hesitation.

Fotkey reviewed the reports carefully. They spoke of troop movements and supply reenforcements.
Several herds were moving into position along the Forbin frontier. Aerocars and tanks were being
made ready for service. Reservists were being notified of the possibility of being called to
duty. Everything was happening discretely, like a kind of great dance, designed to avoid alarm
by giving the appearence of ordinary troop rotations.

The general put down the last report and thought for a moment. Then he fluttered softly.
Gathering the reports together into a tray he got up and approached the steel door to the
document room. Putting the tray on the ground briefly, he spun the combinations and used his key
to open the door and store the shocking truth away from the eyes of Argenia.

Returning to his table, he flipped on the intercom to his support staff. ``Is Yutar here yet?''

``No, Sir.''

``Why not?''

There was a confused pause. ``I don't know, Sir.''

Fotkey disconnected and fluttered. He paced a bit behind his table and then stood staring out
the second story window over the rolling, yellow hills surrounding the base. Some distance away
he could see two fighters parked on the landing pavement, fueled and ready to fly at a moment's
notice. It was as it always was yet that day the tension in the air seemed high.

A movement caught his eye and he turned to see a vehicle coming down the road toward the office
building. Fotkey recognized the markings of the Investigative Service painted on the side. With
a growl he turned and quickly left his office, heading down the hall briskly while his guard
struggled to catch up.

He intercepted Yutar just as he was coming into the building. Fotkey dismissed his guard and
gestured for Yutar to join him. The two men left the building and followed the walkway beside
the road for some distance. The Nermella was strangely calm, and the air was thick with humidity
from the distant seas.

``I trust your trip was uneventful,'' Fotkey said casually.

``Yes, exactly,'' Yutar responded saying nothing more than an answer to Fotkey's question.

After a time Fotkey raised his trunk and pointed off the path and the two men made their way
through the low harsna grass and down an embankment. There, below and behind the office
building, ran a small stream, its waters flowing lazily toward the distant Varsynthia. Tall
reeds lined its banks and in many places the surface of the water was covered with red mats of
algae criss-crossed by insect tracks.

Before reaching the stream itself they came to a rough path that wound alongside the water,
following it to a broad, flat, marshy space at the edge of the Marvia campus. They walked along
the path at a comfortable pace that was neither rushed nor as stagnant as the stream.

``Tell me what you know,'' Fotkey said finally.

``Intelligence is hard to come by,'' Yutar replied. ``Forbin is clamping down. One of our
operatives has been jailed and I fear for his life. Another has gone silent. I hope he is only
keeping quiet to avoid detection.''

``What of their weapons program?''

``We believe it is still alive. There is an unsettling rumor of a breakthrough in the refinement
process but we can't verify that.''

``They have no devices yet?''

``Not as far as we know.''

They walked while Fotkey reflected. ``Tell me about our missing device,'' he said finally.

Yutar tossed his head and swatted away a fly with his trunk. ``We have found nothing\ldots''

``Nothing?''

``I'm sorry, Sir, but it's as if it has just vanished off the face of Rujar.''

Fotkey fluttered. ``That's not good enough. It \emph{must} be found. Does Forbin have it?''

``We\ldots\ I'm sorry, Sir, we can't tell for sure. But their increased security says to me that
they do.''

Fotkey stopped abruptly and Yutar turned to face him. ``Maybe the Forbinites are just
responding to what we're doing,'' Fotkey said softly.

The two men stood silently for a time. Fotkey stared off over the stream and into the distant
west as if, somehow, he could see all the way to Forbin. ``Maybe the Junar have the weapon,'' he
said at last.

``The Junar? Impossible,'' objected Yutar.

``Imagine it. The Junar steals the weapon and then assassinates the human creature making it
look like Forbin kidnapped it. What do we do? We stampede toward a war neither side wants,
destroying each other and leaving the Junar the victor. They don't even need to use the weapon.
It's a strategic masterpiece.''

``No, Impossible,'' Yutar repeated. ``The Junar is nothing more than a two bit group of
religious fanatics with delusions of grandeur. They don't have the resources---or the
intelligence---to do something like that. Leave the Junar to the IIB, that's what I say.''

Fotkey turned to look at Yutar again, putting his trunk on the man's shoulder. ``You do
understand that before we can fight a battle we need to know who is the enemy.''

``Of course, Sir.''

``Just find the weapon.''
