% Local IspellPersDict: aspell.dunari.pws

\chapter{73}

Peno lay at a small table beside an open window sipping ranan juice. He enjoyed a short tray of
cakes while he read a few pages from the latest mystery best seller. Jameia bustled in the
hanaria and then came to join him, lying on the opposite side of the small table. She took a bit
of cake in her tentacles and tilted her head to look out the window. Across the street a small
clothing boutique looked back at her. Yellow vines climbed the walls on either side of its
doorway with green flowers turned toward the sun. Next to the boutique was another apartment
similar to the one where Jameia and Peno lived; the home of another seleckia.

Cars and mini-trucks drove down the following street as people made their way to work in
Yartagard's center. Children gathered at the lifter stop in front of the boutique chattering
freely, teeth clicking and ears shaking.

``Must you go this moment?'' Jameia asked. ``It's going to be so crowded.''

Peno sighed. With two tentacles he put a nut from the cake between the pages of the book to mark
his place. ``I want to get that paperwork done, and you know how hard it is to get an
appointment with him.''

``Let me go with you, at least.''

``You never like going into town, \textit{tinka}. Anyway, I'll be home before you know it.''

He got up and found his regular bookmark. Slipping the nut between his teeth he cracked its hard
shell and then carefully placed his book in its usual spot on the hallway shelf. His wife stared
out the window absent mindedly.

Peno glanced at the clock. ``I should be going. The lifter will be here soon.'' He walked down
the short ramp to his study where he slid several files of papers into his pouch. He grabbed his
purse and his keys and slid them into his pouch as well, making sure all was carefully stowed
away. When he returned, his wife, still sipping on ranan juice, was watching the people along
the street outside. He touched her shoulder with his trunk and she turned toward him.

``When I get home we can clean the attic finally,'' he said.

``That would be wonderful.''

They crossed trunks. Peno stepped out the front door of their modest home and down the ramp
toward the walkway. His wife stood in the doorway watching as he paused in the alcove
momentarily. The Nermella blew fresh air over him, laden with the scent of harsna from the
grasslands beyond the city. He turned to face the wind and for the briefest of moments, the
world seemed to pause with everything and everyone as it should be.

Suddenly the brick wall beside the boutique flashed with a sharp, white brillance greater than
the sun. It stung Peno's eyes and he instinctively flenched and twisted his body away from the
onslaught of light. He could feel a blast of heat envelop him. The people in the street cried
out in shocked surprise and raised their trunks to shield their eyes from the unnatural glare.
Yet almost as quickly as it had come the light faded to a shimmering yellow glow. Everyone
stopped, even the cars, and stared toward the city exclaiming in horror.

Peno stepped out from behind the wall of the alcove to witness a terrifying sight that defied
description. There, in the direction of the city a huge, boiling cloud was rising. It glowed
with yellow-white heat and billowed upward and outward with overwhelming speed.

``Jameia!'' he called out to his wife, but she was transfixed, standing in the doorway looking
at the cloud with an indecisive stare. Peno ran back up the ramp.

``What\ldots\ what is happening?'' she asked, her voice an anxious flutter. She reached out to
him. Their trunks touched.

Then it came. At first it was a dull rumbling like the sound of distant thunder. Yet soon the
ground started to shake with rising violence. The people in the street began to bellow,
uncertain of which way to run, uncertain of what to do. ``Jurita's witness!'' Peno yelled
as he pushed his wife back through the door. The rumbling became a roar. The roar became a
howling gale of burning death that blew back even the Nermella.

``\emph{What is happening?}'' screamed Jameia desperately, but her voice seemed faint and remote
against the thunderous noise.

Peno could not answer. They fell to the floor and he threw his trunk around her. The windows
blew out and the walls cracked. In the distance Peno could hear the helpless yowls of the people
in the street. The scorching wind, like a blast from some great furnace, blackened his flesh
with searing pain and darkened the world with smoke and ash. The roof gave way. Stones, tiles,
and beams all crashed into their home. Peno held on to his wife tightly. There was nothing else
to do.

\spacebreak

Colty sat in the back row of the Cogart lecture hall listening to Rachel present on
thermodynamics. She stood on the platform, the board behind her covered in the mathematics of
her last example, arms outstreached with one finger of each hand extending into the empty air.

``So, if the entropy of the universe is always increasing, how is it possible for biological
systems here on Rujar to take disorganized components such as air and soil and arrange them
into, say, a new rigat tree?''

A student in the front row, toward the door, thumped his trunk against his table. Rachel stepped
in his direction and held out her hand in invitation.

``Rujar is not a closed system. Energy is introduced from the sun.''

``So what does that mean about the sun? How does the entropy\ldots''

Suddenly a strange, high pitched beeping filled the air. Rachel paused and looked momentarily
confused. She reached into her clothing and took out a device that she had in a pocket. She
looked at the device with a slightly furrowed brow.

``Ah\ldots\ I'm sorry,'' she said. ``I'm required to respond to this.''

She stepped down from the platform and stood beside the wall between the door and the board,
turned slightly away from the class while everyone watched with fascination. It was rare to see
her using any human technology.

She held the device to her ear and spoke softly into it as if the machine was some kind of
telephone. ``Yes? Hello?'' she said in Argenian. There was a delay while she listened to whoever
was on the other side. The delay grew long. She slowly moved her free hand to her chest,
clutching her clothes. She turned to look at the class but her expression was unreadible.

Finally she turned away again and chittered into the device in her native language. She slid the
device back into the pocket from which it came, but she didn't move and didn't turn around. The
clocked ticked several times, and Rachel put her hand against the wall with her head bowed.
Finally, after long moments, she slowly walked to the center of the floor. She didn't go back
onto the platform but instead stood on the ground beneath the class and looked up over the
seats. Everyone watched her silently.

``I release you.''

She turned and left.

\spacebreak

Even though he was still dozens of karnons away, the cloud rising from the remains of Yartagard
was obvious. The pilot banked the aerocar slightly with his knee control and picked up the radio
with his tentacles.

``S1 to Harken Base,'' he said into the microphone.

The radio crackled slightly. ``Go S1.''

``I'm\ldots\ ah\ldots seeing a lot of smoke. It looks like smoke.''

``Where is it coming from?''

``Yartagard,'' the pilot replied, ``still too far to really see much.''

``What part of Yartagard?''

``I think all of it.''

The pilot climbed higher to see if he could get a better view, pushing the Z-7 reconnaissance
areocar to its altitude limit. There were some natural clouds in the distance that obscured the
scene but the plume was obviously smoke and it seemed to come from an enormous area. He had a
horrible feeling in his gut.

The pilot talked with Harken base regularly. Following the protocol normally reserved for war
missions, he turned on the scramble circuits. It was obvious that Argenia was under attack. Yet
as the pilot approached he started to realize this was no ordinary attack. The breadth and
totality of the destruction was astonishing. He was too young to have experienced the Harkite
War but from his history he knew this was unlike anything ever seen.

``Dind,'' he muttered as the remains of the city came into view. The pilot grabbed at the
microphone.

``Harken,'' he said. ``It's\ldots\ it's gone.''

``What's gone, S1?''

``Yartagard!'' the pilot yelled. ``There's nothing left!''

``What do you mean `nothing'? Report, S1.''

The pilot swallowed hard and clutched at the microphone. Using the knee control he pulled the
areocar into a descending turn toward the base of the smokey plume.

``I'm seeing a circle, maybe three or four karnons across, all destroyed. The outer parts are
burning. Fires everywhere. Lots of smoke. Looks like the middle part isn't burning.''

There was a delay.

``Say again, S1\ldots\ three or four karnons?''

``Yes!''

He turned the Z-7 toward a gap in the smoke and flew over the outskirts of what was once a
thriving city of nearly 100,000. As he pushed past the obscuring curtain the view was, if
anything, more horrifying.

``I'm above the central part now. There is nothing. No buildings. Nothing. All destroyed.'' His
trunk shook as he looked down at the barren landscape.

``Are there survivors?''

The Z-7 flew over what must have been the middle of the blast, right where the center of the
city used to be. There the ground itself had been blown into a crater almost as if Yartagard had
been hit from space.

``I'm descending now,'' the pilot said. ``There is nobody in the city center. Just nothing.''

The aerocar flew through the curtain of smoke at the far side of the city and the pilot turned
to follow the outer part of the circle of destruction. Far below, his sharp dunari eyes could
see people but many of them were sprawled on the ground unmoving. ``Gods of Nar,'' he said
softly to himself.

He flew over the main road leading into the city and he could see some people walking slowly
away from the destruction behind them, carrying nothing, all with nothing left.

``There are some survivors, yes,'' the pilot said. ``They are walking on the Yart Highway. I see
maybe a few hundred.''

Again there was a delay.

``We are sending fire and ambulance services from Tulanor, they should be there in a few
arnets.''

The pilot fluttered. There was nothing he could say.

\spacebreak

The atmosphere in the First Office was tense. They were at war. Fotkey stood at attention beside
the First Table while the Second Citizen lay on his appointed seat at the opposite side. The
First Citizen lay behind the table and leaned his trunk heavily on the mat. In front of him
stood Ragar, the chair of the Parlimentary Defense Committee, and Politia, the ambassador to
Forbin.

``On behalf of my government,'' Politia said, ``I extend our deepest condolences to you and to
the people of Yartagard''

``Save it for the video cameras,'' Ragar interrupted, ``for all we know it was your government
who did it.''

``We know who did it,'' Politia replied sharply, ``no force on Rujar could destroy an entire
city in a single stroke.'' His body was tense but he stood rooted to the floor, his trunk
pressed against his right front leg as protocol required.

``It would suit you to have us believe that. You are little more than barbarians.''

The First Citizen eyed the Forbinite ambassador closely but Politia stood rigidly and did not
respond to Ragar. At that moment there was a knock at the door and the coordination secretary
stepped just inside the room. ``He's here,'' she said. The First Citizen swayed his head. The
secretary withdrew and ambassador Marshall entered the room looking grim.

``Why? Why did you do this? Why?'' demanded Ragar. He pointed his trunk at Marshall angrily and
stepped forward menacingly.

Marshall tried to ignore Ragar and instead addressed the First Citizen. ``Please, FC, we did not
do this thing. We are as horrified by this disaster as you.''

``And you expect us to believe that?'' Ragar shouted. ``An entire city is destroyed with nothing
but a crater left\ldots\ 80,000 people dead\ldots\ \emph{80,000}! Who else has the power to
deliver such destruction?''

Marshall turned to Ragar, his voice raised as well, ``Why would we do something like this? What
would we gain?'' he turned back to the First Citizen. ``It is \emph{not} our doing.''

Everyone waited for the First Citizen to respond. Leaning forward slightly he finally said in a
soft voice, ``Thank you, Ambassador Politia. We will let you know if we need any further help
from Forbin.''

Politia glanced around and then bobbed his head. ``Of course. Again, I convey the condolences of
my nation.'' He then turned and let himself out the door.

``How do we know your intentions\ldots'' Ragar began, facing Marshall. The First Citizen
abruptly raised his trunk silencing the outburst. Nobody spoke or even moved for several long
moments. The Second Citizen tapped his tenticles against his leg. Ragar shifted impatiently.
Marshall watched the First Citizen closely. After what seemed like many arnets, but was probably
not even one, the First Citizen picked up his phone and called the receptionist.

``Has Ambassador Politia left the building? \ldots\ Very good.''

The First Citizen returned the phone to its cradle and looked at Marshall. ``I know you're not
responsible, Tom.'' Fotkey stood unmoving, facing forward like a statue. The Second Citizen
looked very tired.

Ragar strode to the First Table. ``You\ldots\ how do you know?''

The First Citizen paused, looking down slightly as if lost in thought. He tilted his head toward
Fotkey briefly. ``The humans are not responsible.'' He sighed deeply. ``I am.''

``What?'' Ragar said, his voice a rumbling whisper.

Fotkey shifted and then spoke for the first time. ``Sir, the ambassador.''

``Yes, yes,'' the First Citizen replied. He stood up and walked slowly to the window. He looked
out over the lush gardens around the building and across the street to the lawns of Parliament.
The roads were strangely quiet as if everyone was inside, uncertain of what to do or what was
going to happen next, afraid of what the world had become. Without turning around he said,
``We've been doing experiments at our military base in the Rangard Desert. We've been building
special, high energy weapons, all very secret, of course.''

He turned around and walked back toward the table. ``We know the Forbin government has been
doing the same. It was all about national security. We didn't want to be defenseless.
Unfortunately one of the weapons was, well, stolen.''

``Are you saying,'' Ragar growled, ``that Yatagard was destroyed by an experimental weapon
\emph{of our own construction}?''

``Yes, that's what I'm saying,'' the First Citizen replied simply.

``Why wasn't I made aware of this weapon?''

``You would have been. There was a schedule.''

Marshall's alien voice interrupted. ``Who stole it, FC?''

``We think maybe Forbin but\ldots'' the First Citizen glanced at Fotkey, ``General?''

``It was the Junar.''

``The Junar?'' Ragar said with a flutter, ``So the Junar knew about this so-called secret weapon
and I didn't?''

``To be frank, we aren't sure,'' said the First Citizen, ``who stole the farg thing.''

``It was the Junar,'' Fotkey said again confidently. ``If Forbin stole the weapon to study it,
there is no sense in them destroying it, especially since they were unable to secure the human
scientist. Their strategy is sometimes questionable, but they aren't idiots.''

% TODO: This is unconvincing. Perhaps Forbin actually exploded a weapon of their own
% construction. In any case once they learned what they needed to know from the Argenian weapon
% they would be free to use it.

``Why are we having this discussion?'' Ragar said loudly. He leaned against the First Table and
spoke directly to the First Citizen. ``Eighty thousand are dead. Someone must pay!''

``And do you suggest we attack Forbin even if they are innocent?'' the Second Citizen replied
with sudden passion. ``Doing so would be playing into the plan of the radicals, not to mention
cause the deaths of thousands more on both sides. The general seems sure of his conclusion. I,
for one, trust his judgement.''

``The people will demand retribution. They will blame the humans.''

The First Citizen turned toward Marshall. ``It's true. Your people will be in danger until this
is worked out.''

``I have pulled our groups back to the compounds for now,'' Marshall said. ``If necessary we
will retreat to orbit. However, we want to offer what help we can. I've contacted Earth already.
I hope to get medical aid for the thousands of injured. Your medical science is not ready to
deal with the aftermath of an explosion of this nature. Yet it will take many runion for help
to arrive from Earth.''

``In the meantime,'' the First Citizen said softly, ``it's time to tell the people of Argenia
the truth.''

``Are you sure that's wise, Sir?'' Fotkey said, still standing at attention.

The First Citizen rumbled slightly. ``Frankly I don't know what the best course of action is,
but the truth seems like a good default. Don't you agree?''

Fotkey shifted uncomfortably. ``Yes Sir.''

When the meeting finally ended the participants filed out one by one. Marshall was the last to
leave. As he opened the door the First Citizen called out to him.

``Tom,'' he said and Marshall paused. ``Has anything like this ever happened to your people?''

Marhsall put one hand on the door frame and turned slightly. He seemed to think for a moment.
``Yes,'' he said simply.

The First Citizen swayed his head but did not reply.

\spacebreak

A reporter walked slowly down the remains of what was once a road, picking his way around
overturned cars and fallen utility poles.

``This is Following Street Huteria,'' the reporter said, ``or what's left of it. All the bodies
have already been removed by rescue workers. At this location there were bodies, at least.''

The voice of the anchor woman interrupted the reporter's monologue. ``We understand the humans
are recommending you get no closer to the city center than the fifth crossing. Are you at that
point yet?''

``No. It's actually very difficult to get much beyond here, there's so much rubble on the
street. It's more like picking your way through a rock field than walking along a road.''

The picture switched back to the Varnok studio. ``Thank you,'' the anchor woman said, ``we'll
come back to you again soon.'' She turned in her seat slightly and cocked her head toward the
camera. ``Many demonstrations have erupted throughout Argenia demanding that the humans make
amends for what some see as their attack against us. Yet the official word from the First Office
is that the humans are not responsible. In fact, ambassador Marshall has said the humans are
already air lifting people to hospitals all over Argenia in an effort to distribute the enormous
burden of caring for the injured from this horrific event.''

% TODO: Wouldn't people be arming themselves against the humans?

The anchor woman looked momentarly distracted and seemed to touch an earpiece with one of her
tentacles. ``I have on the phone now Sar. Barikoka, an astronomer from Parnon University.
Can you hear me Sar. Barikoka?''

``Yes,'' came a voice over a small speaker on the desk before the anchor woman.

``What can you tell us about a possible cause of this catastrophe?''

``Well,'' the voice continued, ``as incredible as it may seem, the cause might be entirely
natural. The crater at the center of the explosion indicates that a large metorite may have hit
Yartagard.''

``That sounds like an amazing coincidence for it to strike the city center.''

``Yes, true. However, metorites do hit Rujar now and then. There is evidence of much bigger
impacts in the distant past.''

The anchor woman adjusted her earpiece slightly. ``Would that also be the cause of this deadly
radiation we've been hearing about?''

``Well, ah\ldots\ it's possible, yes. We just don't know enough about it. In any case the energy
released in the explosion is far greater than any weapon could produce.''

``Do you think the humans would have been capable of warning us of this approaching meteorite?''

``I don't know what they're capable of doing.''

The anchor woman swayed her head slightly. ``I think a lot of people are asking that same
question right now.'' She lifted her trunk. ``Ah, Sar. Barikoka, I have a light on my panel that
tells me it's time for the First Citizen's speech. We will be going to Jarloz who is standing by
the First Office.''

There was a slight delay and then the camera showed Jarloz in the street with the First Office
building in the background. There were barricades around the building and a number of people
gathered there. Police were on duty making sure things did not get out of control.

``Thank you,'' Jarloz said. She was stylishly dressed as usual. ``In a few moments we will go to
the First Citizen's statement. As you can see there are people here looking for answers. Argenia
is lost, dazed, and confused by what has happened. Everyone is longing for direction. The time
has come for this leader to make his name in history.'' Jarloz paused as if listening. ``They
are about to start. We are going to the First Office now.''

The scene cut again. Now the camera showed the First Citizen of Argenia lying clamly at his
table. There did not appear to be anyone else in the room. Behind and above him the flag of
Argenia was pressed against the wall. The First Citizen looked into the camera with his right
eye confidently. He waited a long moment while the nation stopped to hear him.

``Citizens,'' he began. His voice was firm yet comforting, ``a terrible disaster has befallen
us. The city of Yartagard has been destroyed. I speak with the voice of Argenia when I say that
our hearts and souls go out to the survivors of this clamity and to the families of those who
were lost. This event will forever change the character of our nation.

% TODO: Include a statement about what the government is going to do to help people?

``Yet amidst this dispair there is hope. We will persevere. We will rebuild. It is in our nature
as Argenians and as dunari to push through these dark times to a bright future. Yes, this event
\emph{will} change the character of our nation by reminding us that we are bound together in our
fate and that, together, we will find the way forward.''

The First Citizen paused and the world stood silently.

``It is natural to seek those who are responsible for this horrifying crime and bring them to
swift and proper justice. Even as I speak an investigation is ongoing to find the perpetrators
of this attack.''

He paused again and looked down slighly. He took a breath and then looked back up. ``But now,
citizens, I must share with you a terrible fact. In an effort to protect this great nation from
our enemies, this government has been experimenting with weapons of exceptional power. To ensure
our advantage we kept this experimental program secret.

``Several of these weapons have been made and, to our great sorrow, one was stolen. This
government conducted an aggressive investigation to locate the stolen weapon, and we were
closing on those responsible, when the weapon was exploded by the murderous thieves in the city
of Yartagard.''

The First Citizen looked visibly pained. ``The force that destroyed Yartagard was a force of our
own creation, unwisely built for reasons soon to be obsolete. That force was used by lunatics
bent on undoing all we have done, yet at its heart it was the short sighted arrogance of this
government that was responsible. Eighty thousand of our people paid the ultimate price. For this
reason I am no longer worthy to be your First Citizen. At the end of this broadcast I will step
down from my post. May Jurita forgive me.''

% TODO: It's awkward to have the FC resign here. The reader has come to know and like him.

He paused again, tilted his head slightly, and looked directly at the audience. ``Our role in
this world has changed. We are no longer the simple people we once were. We can wield powers
beyond the imagination of our ancestors. Now we must find compassion beyond their imagination as
well. We must never forget what happened at Yartagard for as long as the sun burns.''
