% Local IspellPersDict: aspell.dunari.pws

\chapter{79}

Colty lay at the table in his apartment. His notebook was open before him with pages slide
across the table top. He reflected, yet again, on the steps of his proof. Pushing the pages
aside, he picked up the copy of his paper and studied the title: \textit{The primality of
  co-sequences with stellating bases}. It was his finest work. The full proofs in Appendix~A
were compelling but, unfortunately, based on human set theory. As suggested by his overseers he
included a sketch of the same proof ideas, but in dunari style, in Appendix~B. In retrospect it
had probably been a mistake to include it.

The reply from the Argenian Journal of Combinatorics had come back in only one runion, an
exceptionally quick turn around. The speed was not due to dazzled enthusiasm. In fact, the paper
hadn't even gone out to formal review. Instead it had been summarily rejected by the journal's
editor. Colty read the conclusion one more time.

\begin{quote}
  It is not reasonable to accept work that uses an approach unsupported in the literature. For
  this reason I dismiss at once the material in Appendix~A, and focus instead on the proof in
  Appendix~B. Unfortunately that proof is deeply flawed\ldots
\end{quote}

Colty fluttered softly. ``Deeply flawed'' might be overstating the case, but he had to agree
with the editor. The proof, such as it was, in Appendix~B had serious problems. It was intended
to only be a sketch to assist readers in their understanding of the true proof in Appendix~A.

He tossed the letter onto his table and leaned back. He wondered what he was going to do next.
He had spent nine terms thinking about stellating co-sequences. If he couldn't get this result
published he'd have to start over with graph extension fields. At that point producing anything
meaningful from scratch in the time remaining would be nearly impossible. He wouldn't be
promoted to full status, and he'd lose his job.

He stood up and went to his small hanaria and got some water to drink. As he sipped from the cup
he leaned against the wall of his apartment and looked out the window at the gathering clouds.
If he had never met Rachel, he would have never spent time learning the human math and he would
have instead used that time working on graph extension fields. Getting to know Rachel had been
an incredible experience on a personal level, but it might have been professional suicide.

Colty clicked his teeth resolutely. Despite what the journal editor thought, he \emph{knew} the
result was true. He understood the human notation, and the proof in Appendix~A was decisive. He
had that satisfaction at least.

He went back to his table and, for the lack of anything else to do, started to tidy up. In doing
so he came across the picture of the Kessler genome sequence that Rachel had given him. He
paused to look at it. Line after line of different colors marched across the page in several
bands. There must have been hundreds of them on the paper.

He put the image down but just as he was about to turn away it caught his attention again. There
was something about the spacing between the bold, dark lines that looked familiar. He picked the
image up again and studied it closely with his right eye. He had definitely seen that pattern
before. Setting the Kessler genome aside, he fished through his pile for the copy of his paper.
Flipping to the diagrams in section three, he held them beside the genome image. There was a
shocking similarity.

Rummaging through the top tray in his table, he found a ruler. He grabbed a fresh pad from the
shelf. Then, very carefully, he measured the distance between the black lines in the genome
image. Using the smallest as a normalizing interval, he computed the stellating correlations of
the other intervals. Even as he wrote the numbers down, he knew what they were. His trunk shook
as he flipped through his paper. The correlation tables where on page nine.

``Jurita's witness,'' Colty whispered.

It was raining when Colty got into the cab. ``The human compound,'' he said simply.

``What?'' the driver said. ``The human compound? Ya sure? You can't get in there now.''

``Yes,'' replied Colty, ``just go.'' He didn't have time to deal with silliness.

Colty looked out the window at the wet streets of the city as the cab drove toward the base. The
driver watched Colty suspiciously in the mirror. ``Ya meeting with a human?'' he said with a
shake of his ears.

``Yes,'' Colty said. That was the end of their conversation.

The entrance to the base was closed with a metal gate but artificial lights glowed softly in the
guard station to ward off the gloom of the storm. The driver pulled into a small parking area
beside the station. ``I guess this is as close as we get,'' he said. Colty paid the fare and got
out of the cab just as a guard was coming out of the station, his trunk on his weapon.

``You don't want to be here,'' the guard called out loudly. The cab pulled away and the guard
fluttered. ``What do you want?''

``I'm Sar. Coltinarly from the university. I'm here to see the human Ra'hel.''

``Good for you.'' The guard drew his weapon and pointed it at Colty.

Colty slowly held out his trunk, tentacles open. This wasn't what he had been expecting.
``I\ldots\ I just want to talk to her about our work together.''

``Inside.'' The guard waved Colty toward the door where the other guard was standing, also with
his weapon drawn. They led Colty into a small room just off the station entrance where there was
a table and nothing more.

``Look, I\ldots\ I don't want any trouble,'' Colty said. ``I'm Ra'hel's host. We are working
together and I need to consult with her.''

The second guard searched Colty and took his pouch, dumping the contents on the table and
sending papers sliding over the smooth surface. He briefly went through them. Eventually he
faced the first guard and swayed his head. The first guard relaxed a little.

``Check out this guy's story,'' the first guard said to the second. Without a word the second
guard headed for the door.

``I'm Sar. Coltinarly,'' Colty called out after him.

The first guard holstered his weapon but stood on the opposite side of the room from Colty
beside the door. He stared at Colty intently first with his left eye, then with his right. Colty
tried not to move.

``I'm sorry\ldots'' Colty began.

``Quiet!''

Finally the second guard returned. He went to the items from Colty's pouch now lying haphazardly
on the table and found his purse. He flipped through it until he came to Colty's university
identification. He studied it carefully with his left eye and then looked at Colty. After a long
moment he bobbed his head. ``It's him,'' he said, ``they say it's okay.''

The tension in the room dissolved. Colty, along with the first guard, both breathed a heavy
sigh. Colty was shaking.

``You can't just show up here,'' the first guard said, ``not now, not this way.''

``I\ldots\ I understand.''

``Do you?'' the guard fluttered.

``I'll escort you to the compound,'' the second guard said.

It was a short trip from the gate to the compound. The road climbed a small rise and then turned
toward a collection of storage buildings one of which had become the home base for humans in the
Varnok area. The rain had mostly stopped and the clouds were starting to break up letting rays
of warm sunshine filter over the base. Neither Colty nor the guard spoke until they stopped
before the last storage building in the group.

The building looked like all the others. It was of single story brick construction with rounded
sand ridges. However, the windows were all covered, and sitting on the ground a short distance
from the corner of the building was a strange device with hoses and cables that fed through the
wall. At one end there was a wide door.

``Just go in that door,'' the guard said, ``they'll talk to you right inside. I'll wait here.''

Colty got out of the vehicle and approached the building. He felt suddenly nervous. Although he
had grown accustomed to Rachel on the Parnon campus, here he was in her world. He felt acutely
aware that despite being friendly and approachable, she was an alien creature with powers
unknown.

Tentatively he tried the latch. It wasn't locked. Inside was a small room with a large table,
several seats, and an inner door on the far wall. A human device of some kind was braced across
the inner door like a bar. A red light on the device flashed ominously.

``Hello? I'm Sar. Coltinarly. I'm here to see Ra'hel. She's not expecting\ldots''

``Just a moment,'' came a voice from a speaker next to the door. The voice sounded synthesized
but otherwise it spoke Argenian very smoothly. ``She's on her way. Please wait.''

Colty paced back and forth. Finally he heard a beep. He turned and saw the flashing red light on
the bar change to green. The door swung open and Rachel stepped out. She was dressed differently
than he had ever seen. Her clothing was plain and resembled a long shroud. Her hair was not tied
back but instead hung loosely about her shoulders. He realized he was seeing her at a time when
she was expecting to be just a human among humans. He felt vaguely embarrassed.

``I'm very sorry to disturb you,'' he said.

``It's fine, Colty. What is it?''

He set his pouch on the table and took out a folder, showing her the genome image. ``You
remember this?'' he asked.

Rachel stepped beside him to take a closer look. ``Yes, of course.''

``Those black lines are part of the seventeenth order stellating co-sequence.''

Rachel blinked her eyes a couple of times. ``What?''

``I have it here,'' Colty said excitedly. He got some other papers from his folder and showed
her the calculations.

She took the papers and sat down on a bench along the side of the room to study them, pulling
the hair back from her face and hooking it over her ears. She looked at the genome image and
then at Colty's calculations. After a moment she looked up, but not at Colty. She seemed lost in
thought. He could see her eyes moving slightly the way they did when she was trying to
understand something difficult.

``How could that be?'' she said finally, more to herself than to him. She touched her face
pensively. Suddenly she stood up. ``Come with me,'' she said as she walked toward the inner
door.

``Inside?'' Colty asked uncertainly.

``Yes, is that fine?''

He folded his ears back. As far as he knew no dunari had ever been inside a human compound.
``Only if it's permitted,'' he replied.

As the inner door closed behind them Colty found himself at the end of a long hall with doors on
either side. It was obviously the same layout as the original building, but the hall was
illuminated with strange, dim, yellow-green lights. The regular lights were off. Colty squinted
slightly in the unnatural glow. ``Don't look at the lights directly,'' Rachel said. ``They emit
a lot of ultra-green.'' She touched Colty on the shoulder as if to reassure him.

Next to where they came in there was a kind of table, obviously added by the humans, with some
electronic devices on it. Rachel walked over to it and touched one of the panels. Colty could
hear a click behind him and he turned to see the red light on the bar flashing once again.
Rachel touched a different panel and the lights brightened some but remained the same ghastly
yellow-green color.

Colty noticed that the walls, floor, and ceiling were all the original materials. The scene was
at the same time other wordly and yet oddly familiar. Rachel motioned for him to follow her and
they started walking down the hall together. ``How many humans are here?'' Colty asked, keeping
his voice just above a whisper.

``There are seven of us right now. We are all here because of what happened.''

``Just seven?'' The way the compound was presented on the news always made it seem like there
were dozens of humans packed inside. Yet as they walked he did not see anyone else.

``It is the time we normally sleep,'' Rachel explained.

They came to a door and Rachel opened it and motioned for Colty to follow her inside. The room
had been one of the storage rooms but the humans had partitioned it into three smaller rooms.
The main door opened into an entry way and there were other doors off that. Rachel turned toward
the door on the right and touched a panel. He heard another electronic click and then she
grabbed knob in her hand and pushed the door open rather than sliding it aside. The lights came
on automatically and Rachel adjusted them brighter right away.

``Is that too dark?'' she asked.

``It's fine,'' he said a bit nervously.

Rachel brightened the lights a little more. ``On Earth we sleep in darkness. The lighting system
is programmed for that.''

``I don't think I would like Earth very much.''

The room itself was small and without windows. It had a high table with a piece of glass mounted
upright on top, a long, low bed, and a piece of furniture that Colty didn't recognize. In the
corner was some kind of storage closet. There was very little there.

``What room is this?'' Colty asked.

``This is my room,'' Rachel said. ``My personal space. I guess you could say it's my apartment
on Rujar. I know it's not much to look at, especially compared to the beautiful buildings at
the university, but we travel light.''

On the wall was a picture of a strange landscape with jagged mountains poking into a dark sky.
In the foreground were tall orange plants with thick, fleshy stalks and broad leaves. The
picture seemed to glow slightly with its own light.

``Is that Earth?''

She glanced at the picture. ``No. That's on a planet officially called \texttt{Groombridge
  1618c} but most people just call it \texttt{Altairia Caledonia}.'' She spoke the names in her
own language making them sound exquisitely exotic. ``It's close to Earth; I've been there.''

She took a tool out of the desk and then stood in front of a mirror on the wall and used the
tool to smooth her hair. She then pulled her hair back and tied it the way he was used to
seeing. ``I shouldn't have imposed,'' he said awkwardly.

She turned toward him as she finished. ``Don't worry. It's fine, really.'' When she was done she
took a deep breath and, bringing her hands together said, ``Now, about Kessler\ldots''

She went to her desk and lifted the glass plate from its mount. It was about the size of two
pieces of paper back to back. It looked like nothing more than a dusky sheet, yet as she held
it, the glass came to life and glowed with symbols and images. Colty realized immediately that
it was actually a human machine, larger than the one Rachel normally carried. She sat on the
bed. ``Is there enough space for you to lie here too?'' she asked.

``I'll just stand if that's okay.''

She touched the surface of the glass a few times and the symbols moved under her touch. Then,
opening a drawer in the desk, she pulled out a small device. She hooked the device around one
ear and pulled her hair aside so it could rest closer to her skin. She touched the glass a few
times and seemed to adjust the device behind her ear as well. After a moment she looked up at
Colty.

``I'm not really fond of this thing,'' she said touching behind her ear. ``But it does make
things faster.''

``What is it?''

``It decodes my brain activity so I can control my machine with my thoughts. Most people have
implants but I really don't like the idea of a computer being able to read my mind.''

Rachel looked at Colty's paper and he could see the formula for calculating stellating
correlations appearing in the dusky glass of her machine. She was transfering the equations by
just reading them. When she was finished she set the paper aside and studied her device. She
took a breath and squinted a little but nothing seemed to happen. She glanced at Colty.

``I'm out of practice,'' she said sheepishly. She closed her eyes for a moment and took a couple
more breaths. Then she opened them and looked at her machine again. The symbols began to move
swiftly. She spoke to herself briefly in her own language and nodded.

``I'm going to do these calculations using the precise distances,'' she said without taking her
gaze off the device, ``right down to the last nucleotide.''

He watched as she worked. Finally, after an arnet or two, she took a deep breath and took off
the headpiece. She watched the machine closely for a moment and then held it up so Colty could
see it.

``Right here,'' she said standing next to him and pointing to a rectangular region in the corner
of the display. ``This table will show the stellating correlations precisely with a comparision
to the seventeenth order correlations. Are you ready?'' She didn't wait for an answer but
instead touched a symbol at the corner of the table. Immediately the first column filled with
values using Argenian script. They matched exactly.

They both looked at the machine in stunned silence. Colty's mind raced. The number of
nucleotides between bases was precisely the correct number; not one more and not one less. It
was beyond belief.

Rachel touched the symbol in the corner of the table again. The values disappeared and then
reappeared the same as before. She rubbed her chin. ``That's really strange.''

``What does it mean?'' Colty asked.

Rachel put her headpiece back on and soon the symbols were moving across the glass again. Colty
waited patiently. Finally she furrowed her brow slightly. ``The other nucleotides in Kessler are
completely random, at least with high confidence. I wonder if that's normal.''

She touched her device again. ``I'm going to leave a message for Brendon. He's probably asleep
right now.'' Suddenly she shifted to her native language and chittered and clicked at the
device. Yet despite the alien sound Colty could sense, in the tone of her voice, an edge of
excitement. When she was done she took off her headpiece and set aside the glass. She drew her
knees up to her chest and leaned back slightly, lost in thought.

``It can't be coincidental,'' Colty said, ``that this sequence appears in both the human and
dunari genome.''

For several moments Rachel said nothing. Then she replied, ``maybe it's artificial.''

``What?''

``Even we have the technology to do genetic engineering. We can insert arbitrary sequences into
the genomes of most living things. If it were done right, such sequences could be passed from
generation to generation for millions or even billions of runion, even as one species evolved
into another.''

A profound silence descended over the room. It was a silence shared by one dunari and one human
together as they absorbed the implications of what they had discovered.
