% Local IspellPersDict: aspell.dunari.pws

\chapter{83}

Jernumia slowly flipped through the land transaction records of the Handani Mining Company.
Despite the catastrophe at Yartagard, work at Aratok continued as usual. She was looking for a
specific sequence of transactions that would show the company in violation of the Land Monopoly
Act. Haratol had a witness that would testify to the violation but unfortunately that witness
was not entirely credible. Hard evidence was necessary.

She heard a soft knock at the kick plate of her office and looked up to see an elderly gentleman
dressed in fine linen. Jernumia stood up at once, stumbling slightly. ``Leemaster Aratok, Sir, I
didn't realize you were there.''

``Please come to my office, Jernumia.''

``Should I\ldots\ bring anything?''

``No.''

Jernumia quickly grabbed a fresh pad from the top tray of her table and then followed Aratok
himself to the office at the end of the hall. As she passed Haratol, he tilted his head
quizzically and looked at her with his left eye.

The ``big office,'' as everyone called it, was about twice the size of anyone else's and was
lavishly furnished. On the floor was a thick rachen hair rug and nanwood paneling ran up the
walls and even across the ceiling. Aratok's table was broad and seemless and made of solid
onnanin wood from the Forbin lake region. It probably cost more than Jernumia made in 10 runion.

When she entered her eye was immediately drawn to a small, slightly disheveled man lying on one
of the onnanin wood seats.

``This is Inspector Narlock,'' Aratok said in a soft voice. A man in his position never needed
to speak loudly.

Narlock got off the seat and Jernumia touched his shoulder briefly. She recognized the IIB look
at once, and her mind focused immediately. ``We are well met,'' she said with a polite click of
her teeth.

Aratok closed the door and then lay down behind his table as Jernumia and Narlock found their
seats as well. Jernumia wasn't sure what to do with her pad so she slipped it under her right
front leg. Narlock watched her closely.

``We have a case,'' Aratok began, ``that has some, shall we say interesting, procedural
issues.''

Jernumia cocked her head.

``Excuse me,'' Narlock interrupted. He held out his trunk slightly in Jernumia's direction,
``what kinds of cases do you prosecute?''

``Lately it's been mostly real estate cases but in the past\ldots''

``Real estate!''

``Yes, but in the past\ldots''

``Are you sure she's right for this?'' Narlock asked Aratok. The Leemaster said nothing.

``Obviously he thinks I'm right for this,'' Jernumia snapped, ``or else he wouldn't have invited
me to his office.'' Narlock jerked his head back with surprise to look at her again. Aratok
clicked his teeth slightly.

``The inspector is gathering evidence,'' Aratok continued, ``in the Yartagard investigation.
Some of this evidence may come from the humans.''

``Oh, I see,'' Jernumia replied. She turned to Narlock. ``What is this evidence?''

Narlock fluttered and swayed his head. ``Frankly I don't fully understand it. Apparently they
have a machine that will allow them to gather surveillance in a way we can not.''

``Survellance, okay.''

Narlock hesitated. ``They say this machine can look into the past and gather surveillance at the
time the weapon was stolen.''

Jernumia bobbed her head slowly. ``I see.''

``So we run afoul of the Rules of Yotal,'' Narlock continued.

``But the surveillance would be gathered now. It would only show us the past.'' Jernumia's voice
trailed off.

``I want you to look into ways of allowing such evidence in the court,'' Aratok said to
Jernumia.

``Yes, Sir. But what about the case I'm on now? I'm looking for land monopoly violations.''

``I'm sure Haratol can take care of that. This is your case now.''

``Yes, Sir.'' Jernumia sat up straight and suppressed a shudder.

``I want you to work with the IIB through Inspector Narlock's office to find out what you can
and put together a strategy. We should assign a page to assist you, perhaps Larno?''

``Yes, that would be fine, Sir,'' Jernumia took a deep breath.

``So she is to be the lead on this?'' Narlock said with a flutter.

``Yes,'' Aratok said simply.

\spacebreak

Jernumia smoothed her sash with her trunk and then turned to look out the window of the express
train. Carefully tended fields rolled past endlessly, dotted by groups of small homes and
crisscrossed by narrow roads. Yet her mind was elsewhere. She put her case on the convenience
table beside her seat and opened it to take out her notes. She felt a need to go over them one
more time. The questions were standard but the situation was far from normal. She is just
another witness, Jernumia reminded herself, just another witness.

An announcement over the speaker brought her back to the moment. ``Parnon up!'' Jernumia slid
her notes back into the case, snapping it shut, and smoothed her sash again. She gripped the
handle tightly in her trunk and waited as the train slowed to the station.

She had been in Parnon only a few times and the last had been long before. As she got off the
train it took her a few moments to get oriented. The station was small, yet busy, and it bustled
with many students. Jernumia glanced at her watch; she still had plenty of time.

She found a taxi along the roadside in front of the station and climbed into the back. The
driver watched her closely. He wasn't used to seeing a woman dressed for business and carrying a
legal case. Jernumia ignored his gawking. ``The human compound,'' she said.

The driver pulled into the street and Jernumia watched the shops and stores go by. Parnon seemed
so quiet and relaxed. Compared to Varnok it was like a small village dropped in the middle of
open farmland. The driver glanced at her frequently in the mirror but she didn't notice.

``You be careful,'' he said finally.

When they got to the gate the guard on duty came out as Jernumia paid her driver. ``I'll call
when I need to get picked up,'' she said.

She walked briskly to the guard trying to look more confident than she felt. She was greeted
with friendly tones. She had her identification ready but the guard barely looked at it. Instead
he directed her to a military car and drove her up the road to the compound.

``I think this is the first time a woman has met with them,'' he said conversationally. Jernumia
bobbed her head briefly. ``I've seen all sorts come here, you know,'' the guard continued, ``but
I never thought I'd see the time when they needed a lawyer.''

When they reached the compound Jernumia sat briefly in the car and just looked at the small,
imposing building. She's just another witness, she reminded herself again.

``Go through that door,'' the guard said, ``they're expecting you.''

She got out of the car and walked purposefully toward the door. She went inside but nobody was
there. It was an empty conference room. She put her pouch down on the table and settled into a
seat relatively far away from the locked inner door and it's blinking red light. She took out
her notes and reviewed them yet again. All was quiet except for the sound of paper as it rustled
in her tentacles. Outside even the Nermella was subdued.

After a short time there was a click and the door swung open. Into the room stepped the human.
She was tall but not as tall as Jernumia had expected. She was wearing grey coverings on her
lower body with red trim and a light grey top. Around one arm she wore a red trunk cord
decorated with small embroidered flowers. She looked like a chortak wearing clothes but her
strange skin and eyes were completely alien.

Jernumia stood up awkwardly, not sure where to look, and haltingly introduced herself. The human
looked at her with a steady gaze, almost childlike yet with a confidence Jernumia found
disconcerting.

``We are well met,'' the human said. ``I am Ra'hel.'' Her voice was light and soft. Jernumia
returned to her seat while the human knelt down on its lower legs diagonally across from her.
The red Rujaran sun highlighted the trim on her clothes and the trunk cord on her arm.

``Um\ldots'' Jernumia felt at a lose and her voice quavered. She reached out and touched her
notes, seeking reassurance from their familiarity. ``I've been assigned to this case,'' she said
finally. ``So that means I need to know more about this surveillance evidence you hope to
gather.''

``Yes,'' Rachel said.

``Before\ldots\ um\ldots\ we get to questions about the evidence, though, I need to explain a
few things.'' She looked at Rachel again but the human hadn't moved. Jernumia took a breath and
reminded herself one more time that she was just another witness. ``The defense will immediately
attempt to have any evidence you provide thrown out because it was collected by\ldots\ well, by
humans. No offense.''

``I understand,'' Rachel said.

``There is a legal principle called \textit{witnearo en tempexia} that allows a non-Argenian
citizen to testify in our courts against Argenians. It has been used in various cases to allow,
for example, Forbin tourists to testify against Argenian businesses.''

Jernumia paused and glanced at the human but Rachel just watched her intently with strange,
circular pupils.

``I believe,'' Jernumia said more confidently, ``if we try to use \textit{en tempexia} the
defense will argue that it doesn't apply because humans are not subject to Argenian law. There
was an obscure case in Arnoxium 21 argued before the National Court where \textit{tempexia}
evidence from a Runar tribesman was disallowed because the witness couldn't be tried for
perjury, even in principle, and so couldn't be trusted to provide a true and complete
accounting. I believe the defense will argue the same in this case, and this is before the
validity of your evidence is even considered.''

Rachel nodded her head slightly and Jernumia wondered briefly if the gesture was something the
human did for her benefit.

``So what can we do?'' Rachel said.

``I'm not sure right now.'' We need to link you personally to Argenia in a legal sense. Once
that link is made, but not before, you can show the court whatever evidence you have.''

There was a long silence while Jernumia pondered the possibilities but nothing came to her. She
fluttered softly. ``Let's talk about that surveillance now,'' she continued, ``have you heard of
the Rules of Yotal?''

``No.''

``They are the result of a number of landmark cases that went through the National Court about
three hundred runion ago when video surveillance became possible. Certain surveillance made
before a formal investigation is opened is not admissible. For example, video recordings made at
the time of a crime, unless they are specifically of the crime itself, are not admissible. Only
recordings made as part of the investigation can be used.''

``But the recordings I propose have not yet been made,'' Rachel said, ``assuming I can convince
my government to allow it, they will be made precisely as part of the Yartagard investigation.''

Jernumia tilted her head and looked at Rachel with her right eye. ``Are you saying you don't
have the support of your government? Are you acting alone?''

Rachel looked off to the side. ``What I want to do is expensive. Tom, the ambassador, feels we
shouldn't get involved in the investigation directly. Earth does not know about my proposal yet
and many people here do not approve.''

Jernumia was uncertain if she should make any notes or not. ``If you are working outside your
authority that could be a problem if it comes out in court.'' She fluttered slightly and shook
her head. ``That creates new issues.''

``I'm trying to get the authority,'' Rachel said. ``I know Tom wants to help somehow. I just
have to convince him what I want to do is the best way.'' Rachel furrowed her brow in an
expression that looked to Jernumia almost like pain.

Jernumia studied her but Rachel said nothing more. ``Well, keep me informed. I'm going to hope
we can avoid talking about your circumstance, but I need to know about it so I can properly
prepare.''

Rachel nodded.

``My understanding,'' Jernumia continued, ``is that you propose to show surveillance of the
crime itself. But the recordings would be made now as part of the investigation, are you talking
about some kind of time travel?''

Rachel nodded again. ``In effect, yes.''

Jernumia fluttered. ``Our legal system has never had to deal with such technology.'' She
scratched her shoulder with a tentacle and then flipped her pad to a fresh page. ``In other
cases involving new technology it is typical for a third party \emph{impartial} expert witness
to testify about the technology's capabilities and limitations. I don't see how we can do that
here.''

``But the technology is not directly related to the evidence\ldots''

``Then there is the problem of evidence tampering. The defense will no doubt call the evidence
into question because it will have spent so much time under human control.''

The two of them fell silent. After a few moments Jernumia wrote some additional notes. She
cocked her head and again looked at Rachel. The human didn't seem so intimidating any more.
``Unless\ldots\ unless we could have an observing witness.''

``What is that?'' Rachel asked.

``An observing witness is someone, an unbiased third party, who observes the investigation and
then testifies to the veracity of the claims made by the investigators. Ra'hel, can you allow a
dunari---an Argenian citizen---to observe you when you make this surveillance?''

Rachel tilted her head back slightly and exhaled sharply. ``No,'' she said looking at Jernumia
again, ``impossible.''

``In that case it may be very difficult to get your evidence before the judges. Don't forget we
are talking about the National Court here; a matter like this would be handled at the highest
level. If there is any doubt or question the defense will latch onto it with great tenacity.''

``You don't understand,'' Rachel said energetically, ``there is a purging process required when
going from one world to the next to prevent biological contamination. It's an elaborate medical
procedure. We don't know dunari physiology well enough to do it safely to one of you. It would
be very dangerous.''

``I see, but it would still be possible?''

``I don't know.''

``Do you know anyone who might be willing to undergo this procedure?''

Rachel looked at her hands. ``There is one person I know who might be fearless enough, or maybe
crazy enough, but it might be sending him to his death. I\ldots\ I can't have that on my
conscience.''

Jernumia tilted her head and looked at the human closely. The creature seemed subdued and tired.
The idea that it might even have a conscience took her by surprise, and she wasn't happy that it
surprised her. ``I understand,'' she said finally.

``He has a family.''

``I understand. I can talk to the IIB. Maybe they can suggest someone else. I do think an
observing witness will be essential, though, if we want this to have much of a chance.''

The two women, from two different worlds, spent many more arnets going over the details of
Rachel's proposal and how they might present whatever evidence it revealed. The fields outside
the window were bathed in the orange, uncaring glow of the ever present Rujaran sun.

Finally when they were done Jernumia slid her pad, along with her other papers, back into her
pouch. The standard questions never came up. There was nothing standard about this case. She
stood up and put her pouch over her back. ``I think that is all I need for now.''

Rachel nodded and then rose as well. She reached out her arm. Jernumia hesitated. Then with a
click of her teeth she stood before her as Rachel lightly touched her shoulder. Just as Rachel
was about to turn away Jernumia said, ``I really like that, ah, trunk cord you're wearing.''

``Yes?'' Rachel replied with a quick smile ``I think it's very pretty.''

``Where did you get it?''

``I bought it at a small shop in Parnon, in the market center.''

``How did you pay for it?''

``Oh,'' Rachel laughed lightly, ``I'm a visiting professor so I just used my paycheck.''

Jernumia cocked her head. ``The university pays you?'' 

\spacebreak

Colty sat staring at his notebook, open to the section on graph extension fields. He had a few
notes and graph diagrams scribbled at the top of the page, but so far nothing had come of those
earlier ponderings.  Open next to his notebook was one of the seminal papers on graph
extensions. He slide his tentacles across the document as if that would somehow extract
inspiration from the work.

A knock at the kick plate to his office distracted him and he turned to see Sar. Molanalor at
his door. With a combination of surprise, pleasure, and forboding Colty invited him to take the
seat beside his table.

Molan closed the door behind him before settling down. ``First, I want to tell you that I think
your paper on co-sequences is brilliant.''

Colty bobbed his head deeply.

``What you are doing by joining dunari and human mathematics is thrilling. I believe in the long
run it will be seen as the most important work in the field in ten Arnoxiums at least\ldots\
far more important than graph extension fields.'' Molan glanced toward Colty's notebook still
open on the table.

``I'm very gratified to hear you say that, Sar. Molanalor. I'm honored.''

Molan shifted slightly in the small seat. ``Unfortunately I understand the difficult position
your work has created for you. I heard the paper was rejected.''

``It's true.''

``I hope you plan to resubmit it promptly.''

Colty suppressed a flutter. ``I don't have much choice. I don't have time to develop anything
else.''

Molan swayed his head. ``I know you tried sending it to a well known journal, but maybe you
should try a journal that's a little more open to publishing innovative work. You might have
better luck.''

Colty looked at his notebook and then reaching out closed it with a soft thud. ``Thanks for the
advice.''

At that moment Colty's phone began to chime. He looked at it uncertainly. Molan got up and
bobbed his head. ``I just wanted to let you know I support you,'' he said as he slid open the
door and left.

Colty picked up the phone, feeling relieved that the conversation with Molan was over. It was
Rachel.

``I'm glad I caught you in your office. We wanted to talk with you about the genome.''

Colty pushed his notebook aside and sat up a little. Rachel's light, alien voice was
surprisingly comforting. ``What do you mean by we?''

``Brendon is here also. He's the biologist I told you about. He's on \textit{Golden Light} right
now but he's on the call.''

``We are well met,'' came another voice. It was very smooth and deeper and obviously
synthesized. Colty was a little disconcerted to hear another person intruding on his
conversation with Rachel.

``Ah, yes,'' Colty answered tentatively, ``we are well met.''

``Brendon, why don't you explain what you've found?'' Rachel said.

``Sure thing, Rach. I tried searching the genome records of other life forms, at least the
records we have on \texttt{Golden Light}, looking for other instances of the Kessler sequence.''

``Yes, I understand'' Colty replied.

``Well, I didn't find any. However, I looked at your paper, Sar. Coltinarly, and saw the
correlation formula you derived.''

``You mean stellating correlation? Or are you talking about the reduction correlation in section
three?'' Colty scolded himself for leaving the copy of his paper at home.

``The stellating correlation. The formula you used to show Kessler is wrapped in the seventeenth
order sequence.''

``Co-sequence,'' Colty corrected him. Suddenly he had the strange feeling that he was talking
with a colleague about some fine point of research or with an advanced student working on a
special project.

``So as I understand it, these co-sequences are infinitely long and Kessler is only a fragment
of one,'' Brendon said.

``Right.''

``Instead of searching for an exact match, I ran your formula over the genome records using
variable sized windows looking for other correlations\ldots\ ah\ldots\ stellating correlations,
I mean.''

``Variable sized windows?''

``I just tried different fragment sizes starting at every possible position in the genomes I
tested.''

Colty's mind reeled. ``I must not be understanding\ldots''

Rachel's voice interrupted. ``We have machines that can do the calculations.''

``Even so,'' Brendon said, ``it does take a while. That formula is very time consuming to
compute when you get to long fragments, as I'm sure you know, and it doesn't help that some of
the life forms I looked at have very large genomes.''

There was a slight delay and Colty tilted his head. ``I understand what you did, I think.''

``Now comes the good part,'' Rachel injected, her voice had a clearly excited tone.

``So I found other fragments of the seventeenth order co-sequence in various life forms on
planets spread over a large region of space. They have different sections of the co-sequence,
but all part of it.''

``Yes,'' Rachel continued, ``and some of them, at least, are ultra-conserved non-coding
sequences.''

``Where we have that information, yes,'' Brendon said.

``It gets better!'' Rachel injected.

``So\ldots'' Brendon paused, ``on one planet the fragment shows up in the genome of
\textit{cancardacs}, they are insect-like creatures. The planet has a lot of them, it's a nasty
place really, but we have good information about their genetics. It seems the fragment appeared
there about 300 million of our years ago, just as Kessler did on Earth.''

Colty tilted his head again, listening carefully as he absorbed the information.

``What's more,'' Rachel interrupted, ``the information between the co-sequence structures
appears entirely random.''

``Yes,'' Brendon said with a certain reluctance.

``What's the significance of that?'' Colty asked.

``It could be a compressed message,'' Rachel replied.

Colty leaned back slightly pulling on the cord to the phone. ``I see. Compression randomizes the
statistics because it removes redundancy. Of course.''

``Rach, I know you think this is artificial,'' Brendon said, ``but we let's not jump to
conclusions. For all we know there is some natural process at work that's causing it. The shell
of the nautilus on Earth contains a logarithmic spiral not because they can do advanced math but
as a consequence of the way they grow. It could just be parallel evolution.''

``Maybe so,'' Rachel said, ``but think about it: what better way would there be to communicate
with intelligent life emerging throughout the galaxy? Wrap your message in a mathematical
structure to draw attention to it. Then inject the message into the genome of life on various
planets, letting biology preserve it while that life evolves into a form that can read it.''

``But why inject just a fragment into each biosphere?'' Brendon said.

``To ensure,'' Colty suddenly blurted out, ``that only life with interstellar travel could
gather the entire message.''

There was a longish pause and Colty wondered if he had spoken somehow out of turn.

``Perhaps,'' Brendon replied. ``I agree this needs to be explained, but I'm not convinced it's
artificial. We need to examine more genomes more deeply and then try to correlate the fragments
with biochemical and environmental factors. Unfortunately the library on \textit{Golden Light}
is limited, and anyway I have my machines analyzing the dunari genome right now working
on\ldots\ um\ldots\ a different project.''

Colty cocked his head. ``What project is that?''

There was silence.

``It's related to the Yartagard investigation,'' Rachel said. ``We've been advised not to
discuss it.''

``I see.''

``Anyway,'' Rachel continued, ``I wanted to let you know what we found. It might be helpful if
you could tell us more about co-sequences; if there are any results you didn't include in your
paper, it might help us understand what we're dealing with.''

After the call ended Colty stood up and paced back and forth in his small office. It was a lot
to consider. He found himself agreeing with Brendon in that cautious analysis was the best
approach. He returned to his seat and pulled his notebook toward him. Opening it to a fresh page
he wrote ``Information Compression'' down one side.

