% Local IspellPersDict: aspell.dunari.pws

\chapter{89}

Jernumia put some wet garjost leaves into the press and, holding it over a mug of warm ranan,
tightly squeezed the juice out of the leaves into her drink. She settled at a table on the
sunward side of her apartment and looked out over the small park behind the building. Lush
crocker with some harsna grew in the shade of the businesses on the far side of the park. It was
a very calm and pleasant view.

Jernumia's apartment was in the Middle City, an older section of Varnok yet very well
maintained. Many young professionals lived there, mainly single men, but a few selecks as well.
Jernumia herself had no seleck, a fact of her life that drove her mother to endless distraction.

``You can't rely on just yourself,'' her mother often said, ``it makes you look like a hick.''

Maybe that was why she had become a lawyer. It was just another way to distinguish herself from
her family. Yet Jernumia was happy with her solitary lifestyle. Her space was small but well
furnished and decorated with tasteful items of good quality. She drank her ranan and valued her
peace.

When she was finished she put her mug away, gathered her papers into her pouch, checked her
clips, and made her way out into the world. She was headed for the legal library in the New
City, only a few following streets from the agency. But first she would meet Larno at the
office. He had offered to meet at the library but Jernumia knew he had things to do, and she
didn't want him wasting his time waiting for her.

When she got to Aratok the office bustled as usual. Haratol was there but didn't look at her.
They had hardly talked since Jernumia started work on the Yartagard case. When they did their
conversations were never more than a few brief words, grudgingly shared.

Larno was in his cubbyhole of an office putting things in order when Jernumia arrived. At once
they left Aratok and took to the streets to make their way to the legal library.

``I can carry your pouch,'' Larno said.

``I'm used to carrying it myself, but thank you anyway.''

They crossed Hillart and started down one of the wide cross streets blocked for vehicle travel.

``I want you to look into employee taxation,'' Jernumia told her assistant as they walked.

Larno tilted his head quizzically.

``And banking laws.''

``Are you looking for something specific?''

``Yes. I'm just not sure what. Look for cases involving foreigners, especially foreigners from
places without treaties, so not Forbinites.''

Larno bobbed his head. ``What about \textit{en tempexia}?''

``It won't work.'' She didn't say anything more.

``What are you going to do?'' Larno asked.

``I'm going to try to find a way to get evidenced gathered by a time machine admitted.''

% TODO: Include some views of the legal library and of Jernumia and Larnon working together?
\textit{Author note: I would like to enhance this scene somewhat to include more descipription
  of the legal library and of Jernumia and Larnon working together. This might also be an
  opportunity to drop some more hints about what she is thinking (without giving too much away,
  of course).}

\spacebreak

Colty lay in the intructor's seat at one side of a hexagonal table in the main Fargon conference
room. To his left and slightly behind him a student stood beside the board. Four other students
lay at the table, all waiting paitently as Colty made a few notes on the brikken pad clipped in
front of him. Finally he looked up.

``Any other questions for Darnart?'' He tilted his head first to one side and then the other.
Hearing no response he tossed his ears and pushed himself back from the table slightly. He
turned toward the board, scanning over the dense notation that covered it. ``I have one.''
Darnart bobbed his head slightly. ``Tangon's Lemma has a side condition,'' Colty continued,
``how is that reconciled?''

``I thought you might ask about that, and I wasn't sure at first myself.'' Darnart turned to the
board and extended his trunk to a formula near the top. ``It's the resequencing done here. If
you look at the proof of Tangon's Lemma you'll see that this accounts for the side condition. In
effect it is equivalent.''

Colty cocked his head slightly as if to get a better look and reflected. ``Interesting,'' he
said finally, ``that's clever.''

``Wait a moment,'' one of the other students called out. He flipped through a couple of pages
open on the table before him. ``I have the paper here, and it looks like the authors are only
conjecturing that the resequencing is equivalent.''

Colty pulled himself up to the table again and opened his copy of the paper. ``Just
conjecturing? Where do they say that?''

At that moment there was a kick at the door just before Dargon slid it partly open. Everyone
turned to look at the intruder with varying mixtures of surprise and annoyance. ``I'm sorry,
Colty, to interrupt while you're in seminar,'' Dargon began, ``but Ra'hel is on campus and wants
to talk with you immediately. She's in her office.''

``Thank you.''

Dargon slid the door closed and Colty looked over the class. ``Well, we are almost out of time
anyway so why don't we stop here. Thank you, Darnart.'' The group briefly stomped their feet as
Darnart turned, with some relief, to erase the board.

Everyone gathered their papers together and returned them to their pouches. ``Taskana? You are
up next, hour 57:025 Parnon time. Yes?''

``That's right, Colty.''

``Good.''

Colty made his way across the Yellow against the blowing Nermella. The rigat leaves whistled
loudly over his head. When the breath of Jurita was strong few lingered on the seats and lawns.
Aside from one or two others moving quickly along the walkways, the campus seemed deserted.

When he got to Cogart he went directly to Rachel's office. The door was wide open and her office
was more empty than usual. The few trinkets she had and even her photo sheets were gone. The air
cooling unit was off. Colty was suddenly gripped with a deep sadness. Her office seemed so
lonely and so vacant. He had a strange feeling that it had all been some kind of extended dream.
Her arrival, the time they spent getting to know each other, her presence at the Juju feast,
their explorations of Varnok\ldots\ had any of that actually happened?

``I thought I saw you come in here.'' Colty turned to see Jarnol standing at the door. ``She
left a message to tell you she's in the hall.''

Colty bobbed his head briefly, and then without a word quickly made his way back toward the
Cogart lecture hall.

The hall itself was empty except for Rachel's bodyguard slumped in one of the seats. He took
notice as Colty opened the door, watching him warily, but quickly returned to a state of
profound boredom. Rachel was standing on the platform, her back to the empty space, writing on
the board with the podium behind her. She turned as Colty came in.

``There you are! I was hoping I'd connect with you.''

Colty climbed up the steps of the platform, his eye drawn to the math sprawled out on the board
written in the thin, curly style Rachel used. It was math he knew well. It was his math.

``What are you doing?''

Rachel cocked her head at him and smiled slightly. ``I'm trying to understand your paper.'' She
put the chalk in its holder and turning to the podium picked up the loose sheets there. ``I've
been working through your proof sketches, but I have to tell you, Colty, some of this is over my
head.''

She held out a particular sheet in her hand and Colty took it from her, gripping the page with
two of his tentacles. ``Right here,'' she said stepping beside Colty and pointing at his
factoring theorem, ``this seems central to the work but I'm having trouble putting it all
together.''

``That is the trickest part. That's the part the journal editor didn't like.''

Rachel walked over to the board and looked at several formula in a notation Colty didn't
recognize at all. He put the sheet back on the podium and joined her.

``What is this?'' Colty asked.

``Remember how I told you all the proofs in our books are checked by machine? Well, to do that
we must first write them in a special way. That's what this is all about. I want to use our
machines to check your proof.''

Colty looked over the notation for several moments. Although he wasn't familiar with it, he
could sense that it had logical structure. He lifted his trunk and touched a formula written in
dunari style, then its isomorphic equivalent written in human style, and finally what he assumed
was it's equivalent in the new notation.

``I understand.''

He glanced over at Rachel and her eyes sparkled. ``I think that you do, indeed,'' she said with
a smile. ``If you translate your entire argument into our form, convert all the dunari parts, it
will be a simple matter to run our automatic checkers over it. The methods are quite standard
and well accepted.''

Colty fluttered. ``Accepted by you maybe, but my reviewers, and my overseers for that matter,
want me to do just the opposite: remove all the human parts. As much as I'd love to see the
proof checked by your machines, I don't see how that will help me to get it published.''

Rachel turned toward Colty and kneeled down to be closer to his level. ``What do you think about
a change of venue?''

``A different journal?''

``I was thinking of, say, the \textit{European Journal of Discrete Mathematics}. It's very
prestigious and noted for publishing unusual papers.''

% TODO: Actually at this point traditional journals might be a thing of the past. Scientific
% research will likely be distributed using some purely electronic "cloud-based" or
% "crowd-sourced" method.

Colty titled his head to look at her with his right eye. ``A human journal?''

``Why not? You can translate everything to our form, yes?''

``That would be simple.''

``So the problem with mixed notation goes away. You would need to write a new section as a
primer to co-sequence math since we've never heard of it before.''

Colty shook his ears. ``It sounds crazy.''

Rachel laughed. ``It is crazy, but I've been suggesting a lot of crazy things lately!''

He took a step back and tilted his head thoughtfully. ``I wouldn't be able to resubmit to a
Argenian journal while the work was being reviewed by your journal, true?''

Rachel sat down on her back legs the way she did. ``Yes, there are some restrictions. But the
review won't take long. Maybe a couple of runion plus another runion in communication time. The
proofs won't be questioned because of the machine checking. It will be all about the signficance
of the work and the clarity of your presentation.''

Colty paused. ``A three runion delay, not to mention preparation time\ldots\ should your journal
reject it, I'd never have time to resubmit here before my full status deadline.''

Rachel didn't reply. Colty slowly walked to the front of the platform and looked out over the
empty seats of the Cogart lecture hall. The Rujaran sun streamed through the back windows as it
always did, filling the space with its warm, orange-red glow. After a moment he walked back to
where Rachel was sitting. Tilting his head he looked closely at her. ``Would my work be reviewed
on its merits?'' he asked seriously.

She looked at him steadily in return. ``That journal has a blind review process, Colty. They
would not even know you are dunari. But, maybe I shouldn't have brought this up. I can't ask you
to throw away your only chance to get it published here in time.''

``It's fine. I won't be able to get the work published here anyway. It was a fantasy. So in that
respect it doesn't matter. In fact, I'm grateful to you that there's even this chance.'' He
clicked his teeth briefly. ``I'll take it.''

The two of them spent many arnets discussing the organization of the new paper. Rachel explained
the details of the notation required for machine proof checking. Colty walked her through
Nulari's proofs and his own isomorphism. They filled and refilled the board several times, and
made many notes on many pages.

``If you write up the new paper,'' Rachel said, ``I'll get it transmitted to Earth right away.
Unfortunately, I'm going to be off-world for a while. Rocolla is taking over my class for the
rest of this term.''

``Where are you going?''

``It's about Yartagard but, like I told you before, I'm not supposed to talk about it for some
legal reason I don't understand.''

\spacebreak

Tusk crossed trunks with Jael and he affectionately caressed the trunk ridges of Jinna and Mart.
There wasn't much to say beyond what they had already said. Behind him, in a hanger of the
Parnon base, was a relatively large human craft waiting for him.

Jael shuddered slightly. ``Don't get sick.'' Tusk just swayed his head.

``Learn as much human language as you can,'' Jinna said with a soft click, ``so you can teach
me.''

Tusk kneeled down to his son who wrapped his trunk around Tusk's shoulder. ``I wish I could tell
my friends,'' he said.

After several moments Rachel came up behind Tusk and stood slightly to the side waiting
patiently. Jael eyed her uncertainly, and Tusk turned to see her. ``I think that is my cue to
get this started.''

As Tusk turned to leave his family, Rachel spoke to them briefly. ``We'll take care of him.''

The hanger was strangely quiet. For the first time in history a dunari would enter a human
spacecraft and leave Rujar. Yet there were no reporters and no crowds of people. Tusk had to
plead to let even his family be there. Eventually ambassador Marshall needed to intervene. The
IIB, and the Aratok lawyer decided discretion, if not outright secrecy, was necessary. This was
strictly a legal matter, not a poltical bonanza nor a media event. If it didn't work, the public
didn't need to know.

Tusk and Rachel walked together toward the rear of the craft where a ramp came down from a tall
narrow doorway in the alien vehicle. Tusk resisted the urge to look back. He avoided looking at
the dark space inside the mysterious door. Instead he focused on the two ambassadors, Marshall
and Lucasa, standing near the bottom of the ramp.

``It's not too late to back out, Tusk,'' Lucas said when they were in ear shot, ``we can likely
find another way.''

``No,'' Tusk replied, glancing up at the door, ``I want this.''

``We will abort the process and pull you out if there are problems,'' Marshall said. Tusk just
swayed his head.

\alien{``I want to thank you again, Tom, for allowing this,''} Rachel said in English.

Marshall shook his head slightly and replied in English. \alien{``You are lucky Earth is so far
  away. I am lucky that it is easier to ask for forgiveness than for permission.''}

Tusk, who understood the exchange perfectly well, clicked his teeth slightly.

Just before ascending the ramp Tusk turned for a last look at his family, standing together by
the wall. He raised his trunk and Mart did the same. Then, turning away again, he followed
Rachel up the ramp toward the door. He was about to go where no dunari had gone before.

% TODO: Add Jernumia going into a panic about his Forbin ancestory.
\textit{Author note: Jernumia discovers here for the first time that Tusk has Forbin ancestory
  and she is upset about that. She worries that he will be seen as biased and thus unacceptable
  as an observing witness. She's also upset that this is the first time she found out about this
  issue. Of course it is too late to change the plan. However, it will hopefully cast
  uncertainty in the reader's mind about the outcome of her plan.}