% Local IspellPersDict: aspell.dunari.pws

\chapter{101}

Jernumia had been in the National Court before, of course, but never at the prosecution desk,
and certainly never as the primary. The seating behind and above the judges' table was packed
shoulder to shoulder with political figures and industrial leaders. The balcony above the
witness seat was overflowing with reporters and other media workers. Yet Jernumia stood steadily
and scanned the crowd slowly. In the back, far above, in a space easily missed, she could see
the human ambassador kneeling with Lucastanonia.

Across the way, at the defense desk, Zarlon was speaking with Pertonal, one of the most well
known---and most expensive---defense advisers in the city. Jernumia caught his left eye briefly
but he pretended not to notice.

Larno lay at the prosecution desk fussing with some papers.

``I can't find the bank statement,'' he said in a hushed but frantic tone.

Jernumia knelt beside him and opened a folder poking unobtrusively from between two others.
``Here it is.'' Larno shook his head slightly and Jernumia rubbed his back.

She lay down next to him to help organize their materials one last time. Across the oval floor,
brightly lite by shafts of sun coming through the sun windows, she could see Aratok's
reservation. Haratol was there, as were several others from the firm. Even the Leemaster himself
was there.

The officer of the court rose and walked to the center of the floor. He stood for a time while
everyone settled. Finally he bent his left front knee and extended his trunk toward the door
behind the raised front table.

``The Eminence.''

The Eminence entered wearing a garish yellow cloak topped with rubinum cords. His trunk was
loosely wrapped in cloth and his legs were covered with richely patterned leggings. He took his
seat at the table without ceremony. Then in formal language he intoned, ``So begins the case of
Zarlonloti v. People. Let the judges enter.''

Plodding slowly down the aisle from the back came the three judges. All wore cloaks of
impartiality, but each was lavishly decorated according to the education and experience of the
wearer. They moved with a practiced shuffle projecting an air of unconcerned attention. Nobody
spoke, moved, or hardly breathed as they took their seats across the oval floor from the
Eminence.

``Zarlonloti, come forward,'' the Eminence commanded.

Rising from behind the defense desk Zarlon walked with his usual confident gait to the middle of
the floor where he stood in the bright light of the Rujaran sun facing the Eminence.

``You are charged with authorizing the destruction of Yartagard, of being a leader of the Junar,
and of ultimate treason against Argenia and its free people. Are you guilty of these crimes?''

``I am not.''

``Then we shall hear evidence related to these claims, both supporting and contrary, so that
judgement can be made. Zarlonloti, you are compelled to take your seat against the charges
leveled upon you. Do so now.''

Zarlon gave the usual formal response, delivered in his usual oratorial style. ``Your Eminence,
I will show these charges are \emph{groundless} and \emph{false}, and by so doing clear the
tarnish upon my name brought to me by People.'' He walked over to the defendant seat and lay
down defiantly.

The Eminence paused and appeared to look over some papers. ``Advisers, stand before me.''

Jernumia and Perton rose and came forward. They stood side by side about a body-length apart,
facing the Eminence with heads tilted to avoid each other's glance. Jernumia did her best to
look tall.

``I have read your \textit{suktinas},'' the Eminence continued, ``and we have an issue related
to \textit{en tempexia}. Perton, state your case.''

While looking directly and only at the Eminence Perton began, ``the prosecution wishes to bring
testimony from a human. Yet the humans are not subject to Argenian law and so the testimony
should be dismissed as irrelevant. I reference Hurtur v. Killkaria where legal precendent for
this decision was handed down from this very court in Arnoxium 21.''

The Eminence turned toward Jernumia. ``I have reviewed the precedent. Do you have additional
comments?''

``Yes, your Eminence.''

She produced some documents from her pouch and passed them to the Eminence.

``These are records of the human's paychecks from Parnon University where she is currently
legally and gainfully employed as a visiting professor.''

Jernumia paused briefly and then continued in a clear voice.

``You will see that she pays taxes to the city of Parnon and to the government of Argenia like
any citizen. Furthermore, here is a record of her bank account at the Bank of Central Argenia
where she has been saving a majority of her income. Notice that she is receiving interest
payments from the bank. In the case of Parlilly v. Yarlat Sporting Goods, a decision made by
this court as well, a Forbin citizen was granted \textit{witnearo en tempexia} based on his
participation in the Argenian economy. I submit that my witness should thus be granted
\textit{en tempexia} status based on her similar participation.''

When Jernumia ended Perton stepped forward. He rubbed his shoulder briefly with his trunk.
``Despite my colleague's points, the key issue in Hurtur v. Killkaria is that of perjury
protection. I submit that regardless of the human's participation in the Argenian economy, this
court is impotent when it comes to charging the human with perjury and thus its testimony, if
allowed, would be worthless.''

The Eminence glanced toward Jernumia. She was ready. ``The First Citizen of Argenia and the
Ambassador of Earth have signed understandment papers whereby humans in Argenian territory agree
to submit to Argenian Law.''

``Understandment papers do not constitute a treaty such as was in place between Forbin and
Argenia at the time of Parlilly v. Yarlat Sporting Goods,'' Perton objected, ``and anyway that
understandment has never been tested.''

``Only because the humans are law abiding\ldots''

``Enough!'' barked the Eminence.

The two advisers stopped talking abruptly. The Eminence scanned over the papers Jernumia
provided. After several long moments he tilted his head toward the two again. ``Anything more?''

Jernumia hesitated. ``I stand here.''

``As do I,'' Perton said.

``Very well. I will review these comments,'' the Eminence said, ``we will resume in six
arnets.''

Jernumia returned to the prosecution desk where she lay down in a slump. The trial had barely
gotten started and already she was shaking.

``Good opening,'' Larno said as he passed the next batch of papers to her.

Jernumia and Larno talked in soft voices while the rest of the court circulated about. Once or
twice she looked up to see the human ambassador and Lucasa still sitting quietly. The arnets
passed steadily\ldots\ first six, then 12. Jernumia checked her watch often. Finally the officer
of the court stood up and everyone returned quickly to their seats.

``The Eminence.''

The Eminience returned through the same door as before and ascended to his center seat. He
motioned with his trunk and Jernumia and Perton went forward.

``I have reviewed the expositions, and I find the argument of the prosecution incomplete.''

Jernumia felt her heart sink.

``However, the defense would have us believe the humans are no more than savages without any
engagement in Argenian society at all. I am moved not so much by the prosecution's arguments but
rather instead by the public fact that the witness in question has suffered personally and yet,
rather than taking counteraction on its own terms, it is here waiting to testify by legal means
before this court. \textit{Witnearo en tempexia} is granted.''

Whispers and a few scattered stomps broke out through the crowd. Jernumia clicked her teeth
softly. ``Thank you, Eminence.''

Perton turned and, as was customary, gave a slight nod toward Jernumia as they went back to
their desks. Zarlon, still on the defendant seat, glared in the general dirction of Perton.

``Please proceed,'' the Eminence said.

Jernumia returned to the center of the floor, standing amid the pool of Rujarn sunlight. As was
traditional when calling witnesses, she turned to face the audience. Zarlon seemed to be
studying her carefully. ``I call to the witness seat the human known as \textit{Rachel}.''

The side door opened and Rachel walked in using the alien bipedal gait that had become well
known. Then, for the first time in the history of the dunari, a creature from another world lay
upon the witness seat of Argenia's highest court. The crowd grew silent.

The Eminence turned toward Rachel. ``Do you swear upon your immortal soul to here put the
absolute truth before all deceptions?''

``Yes I do,'' Rachel said seriously.

Jernumia stepped forward. ``For the court, state your name and position.''

Her voice was soft and light, but loud and clear. ``I am Rachel Spencer. Visiting professor at
Parnon University.''

``Do you bring evidence related to this case?''

``Yes.''

``Please describe the nature of your evidence.''

``It is surveillance showing the accused ordering the destruction of Yartagard.''

Mumors of incredulous surprise swept over the court. Perton stepped forward and raised his
trunk. ``I seek \textit{hutara},'' he said in a loud voice. Jernumia stepped aside without
surprise.

``Yes, of course,'' the Eminence said, ``go ahead.''

Perton snorted slightly in Jernumia's direction. He turned toward the Eminence. ``The
authorization of Yartagard's destruction must obviously have been given before the event, which
in turn occured before the Initial Moment of the investigation. I believe even the prosecution
stipulates this. Clearly the Rules of Yotal are at play. I thus proclaim the evidence out of
order and request the witness withdraw at once.''

``Your Eminence,'' Jernumia said, ``I respond to \textit{hutara} by saying that the Rules of
Yotal do not apply in this case. The evidence you are about to see was gathered after the
Initial Moment. I can provide testimony to that effect.''

``You play at words,'' Perton said directly to Jernumia. ``Your Eminence, please, unless the
humans have technology that allows them to see the past\ldots'' he clicked his teeth with amused
satisfaction.

``Rachel?'' Jernumia said.

Rachel simply replied, ``we have technology that allows us to see the past.''

The courtroom fell silent.

``\ldots\ under certain circumstances.''

Jernumia continued, ``I can and will bring other witnesses to the seat, Argenian witnesses, who
can testify that the evidence we will present was gathered legally and appropriately. If it
pleases the court I can do that now before continuing with this witness.''

Perton stood beside the desk. ``Your Eminence, even if this is true, it was obviously gathered
with a technique beyond our science. Our judges can not evaluate such evidence. It \emph{must}
be thrown out.''

Jernumia immediate pulled a sheet of paper from her pouch and slide it onto the top of the desk
toward the Eminence. She gave a copy to Perton. ``In Hunian vs the Province of Polot evidence
gathered with microscopic techniques was admitted because the method used to gather the evidence
was clearly explained to the judges.''

``I will request a counter-argument, of course,'' Perton said glancing at the paper.

Jernumia did not reply but instead waited patiently for the Eminence to read the document. When
he was finished she took another document from her pouch and passed it to the two of them. ``In
Nariptia vs Taganon scale analysis evidence was also admitted after it was explained. Notice in
that case the defense also argued that the evidence would have been incomprehensible, yet now
scale analysis is a common technique.''

``Scale analysis is one thing,'' Perton said, ``but time travel is entirely different. We are
asking the judges \emph{and} the Argenian people to understand a technology that is far
beyond scale analysis.''

The Eminence reviewed the second document carefully and then put the papers down. Glancing at
Jernumia he added, ``You did your homework, adviser.'' He reflected for a moment and then turned
to Rachel. ``Can you explain how this time travel works in a way that we can understand?''

``Yes, I think so,'' Rachel said.

``Go ahead, I would like to hear this.''

Rachel turned to the court, addressing not just the advisers but everyone. It was as if she was
giving a lecture back at Cogart, and she spoke with the voice of a teacher.

``About five runion ago a nuclear weapon was stolen from the Argenian government. Since then the
light from that event have been steadily moving away from your planet.''

She shifted.

``We are a small band of humans far from home. We don't have many resources. But one thing we do
have is the Nitooli telescope. It is the most advanced optical instrument ever made by my
people, and it is here, waiting to be installed. So we packed Nitooli into \textit{Summer
  Breeze}, and outraced the light rushing from your world to deploy the telescope temporarily at
a point several light runion away. Then, turning back toward Rujar, we captured the images of
the crime as they sailed by.''

Rachel turned toward the Eminence. ``When you see the stars you don't see them as they are now,
but instead as they were in the past. That is all there is to it.''

Perton rubbed his shoulder with his trunk. ``It is still necessary to explain how your faster
than light travel works,'' he said, ``since that technology is critical to your story.''

Jernumia interrupted. ``Your Eminence, it is well known that the humans have faster than light
travel. The precise mechanism of that technology need not be explained in order to make the
method used to gather this legal evidence understandable.''

This time the Eminence was swift. ``I concur. Adviser, present your evidence.''

``Thank you,'' Jernumia said with a quick click of her teeth. There were more murmurs in the
audience. Zarlon remained emotionless. Jernumia looked at Rachel and held her trunk out
slightly. ``Rachel, if you please?''

Rachel hesitated. ``May I rise?'' She glanced toward the Eminence.

``You may,'' he said, ``but you are still bound by your oath until you are formally released.''

Rachel stood up and walked over to a flat, gray screen positioned along one wall. She took her
machine from her pouch and unfolded it. She also turned on the projection system and soon the
screen was filled with a view of Rujar unlike any the dunari had ever seen. It was a view of
Rujar from space. The image of the entire daylight side of the world filled the screen. The tans
and browns of the great deserts faded to deep yellow red toward the Narlar. Several seas were
plainly visible but there was no obvious evidence of dunari activity. There were no cities, no
roads, and no Forbin or Argenia.

``These images are from the date when the weapon was stolen.'' Rachel touched her machine and
the image grew larger as if they were falling toward the world from a great height. As the view
zoomed in the detail increased. There were clouds over the Varsynthia valley and a bloom of
algae in the Argenian Sea. Around the edge of the screen were various human symbols,
incomprehensible to everyone. As the image grew the symbols changed, too quickly to read.

``Notice the Argenian clock in the corner,'' Rachel said pointing out what looked like a drawing
of a clock. The time advanced at a normal rate but was several runion in the past. ``It shows
the time we are seeing,'' she explained.

She focused on the Shunal valley descending lower and lower until roads started to become
visible as thin lines. It felt like they were falling rapidly, and soon the city of Shunalia
started to take shape. Briefly some lines flashed on the screen overlaying the view but they
disappeared as quickly as they came. ``That's just our pattern recognizer,'' Rachel explained.
``It's trying to align the image with the map of the area we got from the IIB before we left.''

The city continued to grow and soon individual buildings started to come into view. The descent
slowed as the buildings of the military base in Shunalia became recognizable. The pattern
overlay flashed on the screen again and a few symbols appeared next to some of the buildings and
roads. Then a circle appeared in the center of the screen that surrounded a single truck.

``That's the truck containing the weapon,'' Rachel said.

The court watched as the truck pulled away from the base and headed out of town. The view
followed the truck as if from high above, looking down on the small vehicle as it crawled along
the road. They watched as the truck was ambushed and as the attackers murdered the drivers and
loaded the weapon into another truck.

``Fortunately the truck stayed under clear skies the entire time,'' Rachel said. ``We were able
to follow it to its destination. It was a long trip but I can turn over the entire sequence to
the court for detailed examination.''

``As you must,'' the Eminence said.

Rachel touched her machine. ``I will skip to the part related to this trial.'' Some alien
symbols appeared in the middle of the screen and Rachel seemed to adjust them. The view shifted
and again they were flying over the truck only this time they were clearly in a very different
part of the world.

The truck stopped at small clearing with another car nearby and a person standing beside the car
waiting. That person stepped forward as the truck parked and two dunari climbed out. Rachel
zoomed in even more closely until the view appeared to be hovering just over the heads of those
who were present. When the men got close to each other they paused and then the three of them
looked up so their faces were all directly visible. The man who had been waiting was clearly
Zarlon. The likeness was unmistakable.

A murmur swept over the court. Perton studied the screen but did not say anything.

Rachel paused the video and pointed at Zarlon on the screen. ``We adapted our pattern matching
system to recognize dunari faces and we believe with 99.97\% certainty that this is the
accused.''

``I don't think human technology is necessary,'' Jernumia said, ``to convince us of who that
is. We have seen him on television often enough.''

Rachel started the video again and the men on the ground began to move. They appeared to
construct a makeshift circle using stones from a nearby gravel pile.

``Explain why they are blurry when they move,'' Jernumia said.

Rachel nodded. ``Light comes in\ldots\ particles that we call \textit{photons}. There aren't
that many photons reflected from small objects and they spread out as they travel. After several
light runion they are very sparse. To get a good image we have to combine information from many
different photons, a process we call photon integration. That process takes time. If the objects
are small or if they move, the integration doesn't have time to work and the view is blurry.''

Finally the men completed their work and Zarlon went to his pouch, laying on top of his car, and
took out what appeared to be a piece of paper. It wasn't easy to see because of the blurring
effect. Zarlon walked toward the circle and, apparently kneeling down, put the paper on the
ground at the center.

``Here we got very lucky,'' Rachel said. She made a few adjustments on her machine and the view
began to tighten up on the small fuzzy square of yellow-white paper.

``They put this paper face up and held down the edges with stones,'' Rachel said, ``and then
they just left it there for a long time.''

Gradually the view zoomed more closely and the fuzzy square grew.

``And so,'' Rachel continued, ``pushing our technology to its very limit, we trained the eye of
Nitooli on that paper, and let photon integration do its work\ldots''

The fuzzy square now filled the screen but it was little more than an amorphous blob at first.
The court was silent as the clock in the corner of the display ticked. Gradually dusky patches
appeared and, as the moments passed, they slowly took form. It was as if someone was focusing a
great lens, bringing the subject into clear view in small increments. The dusky patches took
shape and lines appeared. The lines became curves. The curves became letters. The letters became
words written in a ceremonial script.

Rachel touched her machine and the clock stopped. She turned to the court. ``Now we can read
it,'' she said quietly.

Jernumia stepped forward. This was her moment. ``This is the paper taken by the accused from his
own pouch.'' She turned toward Rachel. ``As I explained to you earlier, in our legal system it
is important, in cases like this, for a witness to utter significant evidence verbally. Rachel,
your Argenian is quite good. Can you read this message?''

``I can.''

``Please do so.''

Rachel turned to look at the screen. She paused briefly and then with her light, alien voice
said, ``I, the Jahar of the Junar, proclaim these words: for the glory of the Seven, the city of
Yartagard and\ldots''

Rachel paused. She shook her head slightly. ``\ldots\ and all the inhabitants therein are to
be\ldots\ erased\ldots from this world.''

Rachel shook slightly and her voice broke. ``So speaks Jurita.'' She rubbed her eyes quickly and
then ran her fingers through her hair.

Jernumia walked over to her. ``Are you all right?'' she said softly.

``Yes,'' Rachel replied, ``I'm fine.''

Jernumia addressed the court as a whole, this time in a loud voice. ``I will bring testimony
that this is the standard Junar assassination note. Except in this case it names the entire city
of Yartagard as the target of the assassination.''

Rachel started the video again and the view pulled back. The court watched as Zarlon and the
other men executed the confirmation ritual. The court watched as Zarlon used the ceremonial
knife to cut between his scales. The court watched as he dripped his own blood onto the yellow
white paper. Finally, the men parted ways and the truck continued on its route while the court
watched. Rachel stopped the video.

``We followed the truck all the way to Yartagard,'' Rachel said, ``where we saw the weapon being
unloaded to a warehouse for a clothing store in the center of the city.''

``Thank you Rachel,'' Jernumia said, ``please return to the witness seat.''

Perton shook his ears slightly and came forward toward the center of the floor. He faced Rachel
and turned his right eye toward her. ``So you are claiming that these pictures were taken from
several \emph{light runion} away?''

``Yes.''

``And you expect us to believe that it is possible to see such detail from that distance?''

Rachel tossed her head slightly and glared at Perton. ``Nitooli can see lakes and rivers on
planets on the other side of the galaxy. So, yes, it can see the detail I've shown here when
it's as close as a few light runion.''

``That is all very well for you to say,'' Perton turned toward the audience and spoke in a loud
voice, ``but we need \emph{proof}.''

He turned back toward Rachel. ``Do you have image processing technology?''

``Yes.''

``Have these images been processed?''

Rachel hesitated. ``Yes, but only\ldots''

``And so how do we know these images haven't been entirely fabricated?''

Jernumia called out suddenly, ``Eminence, the adviser is interrupting the witness's testimony.''

The Eminence turned his right eye briefly toward Perton and then told Rachel to continue.

``I only wished to say that the images have been processed to make them visually accessible. The
data coming off the telescope are not images in the usual sense. But the images I've shown here
are a true accounting of reality.''

Perton addressed the audience again, ``Eminence, despite what we have seen, this evidence is
meaningless. We have no assurance the images are real. I remind you that the witness may have a
vendetta against the accused.''

% TODO: Isn't that an admission that Zarlon had something to do with Rachel's kidnapping?

Rachel replied speaking just as loudly. ``The gathering of this evidence is not my doing alone.
It was gathered with the approval of ambassador Marshall and done by the Nitooli technical
team. I'm merely presenting\ldots''

Jernumia gestured with her trunk and Rachel fell silent.

``My adversary's point is well taken,'' Jernumia continued, ``if his questioning is complete I
am prepared to call an observing witness to add credibility to this evidence.''

``That's the real point,'' Perton said directly to Jernumia, ``your `observing witness'. Then
let us get to it.''

With the Eminence's permission Rachel withdrew and Jernumia once again faced the audience. ``I
call to the witness seat the head linguist at the Office of Alien Affairs, Tuskara.'' The sounds
of murmurs filled the court as Tusk came into the bright oval. He was focused on reaching the
witness seat without looking around.

Immediately Perton slapped a document on the desk before the Eminence and then passed a copy to
Jernumia. ``My counter-argument,'' he said simply.

Jernumia scanned the document quickly. She had already seen the preliminary version. ``I will
point out, that Tuskara is a full bred-and-raised citizen of Argenia in good standing.''

Perton tilted his right eye toward her. ``Do you deny that his father is fully Forbin?''

``Of course not. Tuskara's father was granted citizenship by the First Citizen himself for his
valuable contribution to the Argenian war effort in the Harkite War.'' Jernumia put her rebuttal
on the Eminence's desk beside the counter argument. The Eminence arranged the papers carefully
so they were both in front of him, as if he could read them both simultaneously.

``My rebuttal enumerates many instances where individuals with Forbin ancestory, including some
bred-and-raised Argenians, have testified in Argenian courts.''

``It does not matter,'' Perton said, ``The fact remains that this witness may be motivated to
indict my client in order to protect his homeland from scrutiny. It is no secret that Forbin has
been informally accused of the crime we are here debating.''

``This is his homeland!''

The Eminence raised his trunk and the court fell silent.

``Tuskara,'' the Eminence said, ``recall that you are oathed. Do you owe your allegiance to
Argenia?''

Tusk glanced toward the Eminence and then at Jernumia. He turned his left eye to the audience.

``No.''

Jernumia took a step forward with a flutter. Perton clicked.

``I owe my allegiance to all dunari,'' Tusk said, ``and it is for them that I am here ready to
reveal the truth.''

Perton shook his ears. ``The witness has all but admitted \textit{rustara}. He must withdraw!''

``It does not matter,'' Tusk continued, ``who I am. The destruction of Yartagard is a crime
against us all. It only matters that I am dunari.''

Jernumia gestured with her trunk briefly, but paused.

``Who you are may not matter to you,'' the Eminence said, ``but this is an Argenian court and so
it does matter to us.''

Jernumia lept forward. ``Eminence, in\ldots\ Tagonally v. Rostan, Arnoxium 8, the genocide of
the Endless War was deemed, by this very court, to be a crime against all dunari. Surely this
court could make that determination again.''

``You are desperate now,'' Perton said, ``doing such a thing would amount to saying that war
itelf was a crime against all dunari.''

The Eminence looked tired. He spoke directly to the two advisers. ``I'm interested in the
reliability of this witness, not in making broad philosophical statements.''

``Then please review my rebuttal,'' Jernumia said, ``all the precedent is there.''

The Eminence dismissed the advisers and called for a hiatus. Jernumia returned to the
prosecution desk. Larno was making some notes.

``Tagonally v. Rostan was in Arnoxium 9,'' he said softly to Jernumia.

She didn't reply. Instead she watched Perton pace in front of the defense desk. Zarlon sat
impassively on the defense seat. He could have been a statue. Jernumia always avoided looking at
the accused. It wasn't her business. She constructed her cases based on evidence, facts, and the
rule of law. How the accused looked didn't matter. Yet she found herself drawn to Zarlon. The
quiet stare from his hollow eyes fascinated her.

Finally the Eminence returned. He took his seat as Jernumia and Perton stepped forward again.

``I have read the arguments. I am impressed by the witness's military history. Regardless of
what me might think about Forbin, this witness satisfies all criteria for an observing witness.
I grant him validity.''

There was some stomping of feet. Perton shook his head slightly. Jernumia exhaled. Had she been
holding her breath? She walked over to the witness seat. It would be easy now.

``Tuskara,'' she said, ``you are an observing witness. Tell us what you saw.''

