% Local IspellPersDict: aspell.dunari.pws

\chapter{103}

% TODO: There is a big jump in time here. The trial is over, Nitooli has been deployed
% (properly) and Rachel has gone to do her research. It seems like something more needs to be
% said about all of that (summarizing transition?). Or maybe a scene about Colty's new paper for
% the human journal?

Colty stood outside his office, the door ajar, talking with a student about a recent assignment.
He cradled the student's paper in his trunk and scanned it with his left eye.

``You should use the proof of the Numlis Theorem that I did in class as a guide. If you use the
same technique here it goes through easily.''

``Yes, but this should work, right?''

Colty fluttered slightly. ``I think you're getting into trouble when you reduce the initial
equations. That throws away important information that I'm fairly sure you'll need later.''

The phone in Colty's office started to chime. He returned the paper to the student. ``Look at
Numlis again. That will clarify what I'm talking about.''

``Okay, Sar. C.''

Colty went into his office and closed the door. He quickly stepped over to the phone.

``Yes, hello?''

The line had a strange hiss. The voice on the other end sounded as if it was coming through a
long tube. Yet it was soft and light and unmistakable. ``Colty?''

``Yes\ldots\ is that you Ra'hel? Are you back in Parnon?''

She laughed. ``No, no. I'm actually calling from the observatory. We found something important
and we want to talk to you about it.''

Colty paused as he took in the fact that he was receiving a phone call from another planet.
``What did you find?''

``It's related to your work on stellating co-sequences but it would be easier to show you. Can
you get to Lungast? It would be much better for us to talk over the Link.''

Colty reached his trunk into the top tray of his disk and found his schedule book. ``I could be
there mid-runion.''

``Yes, perfect,'' Rachel said.

Immediately after disconnecting the call, Colty arranged for train tickets and contacted Dargon
for assistance covering a class. ``I can give them a reading assignment if you can't do it,'' he
told Dargon, but Dargon was happy to help.

``What is it about?'' he asked with genuine curiosity.

``Stellating co-sequences.''

The train ride to Lungast was uneventful. It was scenic in some places and dreary in others.
There was rain across The Flats and Colty spent much of that part of the ride reading.
Originally he had thought to review material on co-sequences but there was no point in that.
Colty was the foremost expert on the subject on two worlds. He knew more about co-sequences than
any dunari who had ever lived. Instead he read a book about the Mezanic composers Joleia had
given him just before he left.

``You need a hobby,'' she had told him.

At Lungast Colty was met by a smartly dressed military officer and immediately led to a car
parked directly in front of the station. It was surprisingly luxurous. Colty expected some sort
of combat-ready vehicle but he supposed that was silly. He was just a university instructor
being driven from the Lungast train station to the base, not a general going to a battle front.

Colty had never been in Lungast before. The town was larger than he expected. It was also closer
to the Narlar than he was used to. The sun hung low over the rooftops and the air felt a bit
chilly.

The car turned at a surprisingly unobtrusive sign for the base. Along the road was a collection
of tents, mobile homes, and even some temporary shacks. A few people were outside, walking along
the roadway or laying on the rough grass. They eyed the car suspiciously. Colty could see
several signs saying things like, ``They are not here to help us,'' or, ``Be careful. Jurita is
listening.''

``What's this all about?'' Colty asked the driver.

``People protesting the Link.''

``After all these runion?''

``It was much worse earlier. I used to get an armed escort.''

% TODO: Show, don't tell...

Colty looked out the window at the people. They looked tired. They were dedicated to their
cause, but what cause? Didn't they realize that no matter what they did the humans would be a
part of dunari life from now on? Colty wondered if any of them ever saw a human in the flesh
much less spoke to one.

The ramshackle collection of tents and mobile homes disappeared behind them as they approached
the entrance gate for the base. The car stopped and the guards inspected the driver's papers.

``Transporting Sar. Coltinarly from the train station,'' he said, ``to see Captain
Charkonaless's group.''

The guard asked Colty for identification. He looked over the papers on a clipboard he had
hanging around his neck. He inspected the vehicle. Finally he signaled to the booth and the gate
opened.

When they reached the main facility the driver parked the car immediately by the front door. He
got out and escorted Colty into the building, leading him down the hall and up the ramp. Finally
they came to a door with an engraved plaque saying ``Captain Charkonaless, Director of Link
Communication.'' When they went inside the secretary looked up.

``Hi Jeno,'' the driver said. ``Sar. Coltinarly here for the Captain.''

Colty was told to go right in. He opened the inner door and found a spacious, attractively
furnished office. Two windows were open and a pleasant breeze was gently moving through the
room.

``Sar. Coltinarly!'' Chark got up from his desk and greeted his visitor warmly. The two men
clasped trunks. ``So glad you could come. They seem very intent on talking with you.''

``Thanks,'' Colty said. ``Do you know what it's about?''

``Not really. I'm just an engineer. I don't usually participate in the meetings. Don't you
know?''

``Ra'hel was a little enigmatic when she spoke with me.''

Chark looked at the clock. ``You'll know soon enough. We should go to the conference room to see
how they are coming along with the setup.'' Chark put his trunk on Colty's shoulder and clicked
his teeth. ``It's not like any conference room you've ever seen.''

Chark led Colty back down the hall to the narrow ramp. ``This might be the first time a
non-government official has used the Link,'' Chark said. ``So far it's all been people from the
ambassador's office, Parliamentary delegates, and the like, even the First Citizen. I'm glad to
see them branching out.'' Chark turned to face Colty as he walked, side stepping down the hall.
``There is a similar facility in Forbin and we've used the Link for our own dunari purposes as
well.''

They came to an unassuming door. Inside was a small waiting room. There were windows on one
wall, a counter staffed by a military receptionist, a number of seats, and a desk in the corner.

``Shall we go in?'' Chark asked. He gestured toward a simple door next to the reception counter.
There was a yellow light above it.

When they entered Colty found himself in a triangular room. The floor seemed to be made of dark
gray tiles with the outline of a red square on each tile. The walls and ceiling were also dark
gray and almost completely blank. Along the wall, running all around the room, was a single red
line. In the walls opposite from where they came in were two tall, narrow doors of the sort a
human might find comfortable.

At first Colty did not see anything remarkable about the room, aside from its unusual shape, but
then he noticed that the red lines on the walls and floor were not painted as he had assumed at
first. The color was a glow from the material itself. Even the lights in the ceiling were just
glowing patches and not ordinary artificial light fixtures.

Near the middle of the room stood a young dunari facing a tall human standing slightly to his
left. When Chark and Colty entered the human gestured toward them and the dunari turned around.

``Oh. I didn't hear you come in,'' he said.

``How is it going?'' Chark asked. He introduced the young man as Tarmala.

``Fine on our end. There was a little trouble with the Nermia connection.''

Chark turned toward Colty. ``It's a super-luminal connection to Nermia. They are always a bit
finicky.''

``It's all set now,'' Tarmala said. ``Just a minor synchronization thing. Our equipment is
running fine.''

``Our equipment?'' Colty asked with surprise.

``Not all of it,'' said Chark clicking his teeth. ``The humans provide the end point hardware
and data codecs but we still use good old radio power amplifiers to boost the signal up to their
communications network.''

At that moment the door to the right opened and Rachel walked in. She was wearing green clothing
with black trim that covered her entire body. It was a style Colty had never seen before and he
found it rather striking.

``Ra'hel,'' he said with pleasure. He walked quickly toward her but she froze. Suddenly a grid
of red lines appeared in the air in front of Colty's face and he stopped abruptly.

``Be careful, Colty,'' Rachel said. She put up her hands as if to push him away even as the red
grid faded.

Chark walked up behind Colty and held up his trunk. With a quick motion he swung it in Rachel's
direction and the red grid appeared again, very brightly. His trunk thudded against something
hard even though it appeared to Colty to be stopping in thin air.

``It's a wall,'' Chark explained. ``Just like over there.'' He pointed to the dark gray wall
behind them with the door through which they came in. ``The red lines appear as a warning to
stop people from accidently walking into it.''

Rachel stepped forward so that she was, or seemed to be, standing right next to Colty. ``I'm at
Nitooli,'' she said. She pointed to the other side of the room where the human who had been
speaking to Tarmala was still standing. ``That's \textit{Golden Light}.''

``\alien{Hello, Rachel},'' the other human said with a thick accent.

``This is amazing,'' Colty said. He looked Rachel over closely. ``It's as if you are standing
right here.'' Tentatively he held up his trunk and touched the wall in front of him. It felt
smooth and cool.

The other human chittered something and Rachel turned slightly and spoke a few words in her
native language as well. Colty shivered. To hear her use her strange language still made her
seem very alien.

The other human left. ``Everything is set and your meeting time is about to begin,'' Chark said.
He and Tarmala also left, leaving just Colty and Rachel standing side by side and yet millions
of karnons apart.

``How have you been?'' Colty asked.

``I'm doing fine,''

Colty touched the wall between them again, just to check if it was really there. He considered
telling Rachel that he had missed her but somehow that didn't seem quite appropriate.

``Can you use this technology to communicate with Earth?'' Colty asked.

``It's too far away,'' Rachel said. ``The delay is much too great even with our super-luminal
technology.''

Suddenly the door on the \textit{Golden Light} side opened and in walked a very tall human. At
least he seemed very tall to Colty.

``\alien{Sssorry},'' he said in Argenian. ``\alien{I delayed}.''

``Ah, good,'' Rachel said. ``Colty, let me introduce Brendon. He's the computational biologist
you spoke with before.''

Brendon put a translation unit behind his ear and seemed to adjust it slightly. When he spoke
again he did not move his lips but instead his words were beautifully accented and synthesized
Argenian.

``I know people find this a little strange, but my Argenian isn't good enough yet to do without
it.''

Rachel introduced Colty, and Brendon seemed very pleased.

``I'm honored to finally meet you, Sar. Coltinarly.''

Rachel started the discussion, speaking mostly to Colty. ``I'm going to show you a map, of
sorts. To render it properly I'll have to dim the lights but I won't dim them as much as usual
for your comfort.''

She took a couple of steps back and so did Brendon. Colty followed their lead, uncertain of what
was going to happen.

``Francine,'' Rachel said, ``show the summary map in dark background mode using integrated H-D
visual spectrum. Split across sites and keep ambient light at medium.''

A female dunari voice, speaking Argenian seemed to come out of nowhere. ``Understood. Computing
view\ldots'' After a few moments the voice spoke again, ``Completed.'' The lights faded slightly
and suddenly a galaxy appeared.

Colty gasped from the surprise like a small child. Floating in the air around him was a huge
disk of stars. It filled the room, spanning all three locations smoothly as if it was just one
large space as it appeared to be.

The detail was fantastic. What at first Colty thought was only clouds of light were actually,
upon closer examiniation, millions or even billions of individual points. In the center of the
space was a huge central bulge. A long, thick bar of stars extended from either side of the
center and two, great arms extended from the ends of the bar to wrap in opposite directions
around the combined room. Many smaller spiral arms also could be seen. Colty reached out to
touch the stars but there was nothing solid, only light.

``This is our galaxy,'' Rachel said. ``The map is as accurate as we can make it but the sections
on the far side of the center are much less clear.''

Rachel walked into the center of the space, wading through the stars the way a person might wade
into a lake. She seemed to be in her element, at home both with the technology and with the
galaxy as well.

``There are highly compact spherical clusters that we call globular clusters. They mostly orbit
in a halo around the galaxy as a whole. Francine, brighten all known globular clusters.''

Many bright dots appeared like a swarm of insects buzzing around the galaxy's disk.

``As you know, I came to Rujar to measure the orbits of these clusters. I'm particularly
interested in the orbits of clusters close to the galactic center where the effects of the
central mass are the greatest. Francine, brighten only the globulars in data set 1737-295.''

The swarm of insects faded into the background while a handful of bright dots remained about the
central bar.

``Francine,'' Rachel continued, ``show projected orbits of the 1737-295 globulars using recent
data taken from the Nitooli Observatory.''

The soft voice replied, ``Analysis of observations incomplete. What confidence factor should I
use?''

``Show only orbits at 90\% confidence or higher.''

Graceful arcing green lines appeared circling the central bar. Each line passed through a single
bright dot and Colty assumed it indicated the path that cluster took as it moved about the
galaxy. The lines went up rising out of the disk, swinging over the central bar, and then diving
back down and under. Because the main spiral arms started at the bar's ends, there was a
significant, largely empty space on either side of the bar. It was through that space these
clusters moved.

Rachel waded up to the orbits and pointed at one of them with her finger. ``There is nothing
obviously out of the ordinary here,'' she said. ``Yet these clusters all move in a largely
similar orbit that happens to pass right between the main spiral arms.'' Colty tentatively
walked out into the sea of stars to join her. He wasn't sure exactly where this was going, but
it was fascinating in any case. ``Yet there is something very strange going on here,'' she
continued. ``Francine, show \alien{Sagittarius A-star}.''

A bright green dot appeared at what seemed to be the very center of the galaxy. It was
surrounded by six short green lines that helped to draw attention to its location.

``Okay,'' Rachel said. ``For the first time, thanks to Nitooli's data, combined with data from
several other observatories, we were able to observe precise three dimensional locations of
clusters on the far side of the central black hole.''

``What, exactly, is a `black hole?' '' Colty asked.

``They are highly compressed stars. They are unusual objects, but created by natural means.''

``Most of the time,'' Brendon said.

Rachel didn't react to Brendon's comment. ``Francine,'' she said, ``apply a Larkin normalization
to the semi-major axis distances of the 1737-295 clusters, and show the normalized values.''

Small numbers appeared next to each orbit but Rachel wasn't satisfied by that. ``Francine,
display the previous results in tabular form and show Sar. Coltinarly using Argenian notation.''

A rectangle of light appeared before Colty. He twisted slightly so that he could view the
information with his left eye. He didn't have to look long. He recognized the numbers
immediately. It was the eleventh order stellating co-sequence. He looked up quickly, his ears
spread.

``And now you know,'' Rachel said, ``why you are here.''

``Could it be some kind of coincidence?'' Colty asked.

``Impossible,'' Rachel said. ``Stellar orbits around the galaxy are chaotic. Any pattern like
that would degrade after a few million years. The chances of us being here to see it would be
negligible.''

``But if these orbits are not degrading,'' Brendon said, ``someone is holding them there.''

There was a pause and then Rachel said, ``Francine, remove map and restore lighting.''

Colty paced back and forth across the room. Brendon said, ``I now agree with Rachel. Some
intelligence is responsible for both the Kessler genome fragments and for the orbits of the 295
globulars. Whoever or whatever they are, they seem to be fond of your mathematics, Sar.
Coltinarily. The question is: why?''

Colty hardly heard him. ``Francine,'' he said without even thinking, ``show me again the
previously computed values.'' The illuminated rectangle appeared before him and Colty stopped to
study it more closely.

``This is the TARS variant of the eleventh order co-sequence,'' he said.

``Yes,'' said Rachel, ``definitely. What is significant about that?''

``Well,'' he began. ``You mentioned that the non-coding genome in Kessler might be a compressed
message. I've been reading about information compression, and it occured to me that stellating
co-sequences might form the basis for a very aggressive compression method.''

Rachel and Brendon glanced at each other. ``Go on,'' Brendon said.

``I\ldots\ I haven't worked out the details. I'm not even sure it works at all. If it does, the
behavior depends on the order. Second order co-sequences would produce an identity compression,
third order co-sequences would produce some compression, and so on. The eleventh order
co-sequence\ldots\ well, that could potentially produce a huge compression.''

``Jurita's Witness,'' Rachel said softly.

``If you're right,'' Brendon said, ``then the key to reading the message might be in these
orbital observations Rachel's group is doing.''

``It won't be helpful,'' Colty said. ``By the time you reach the eleventh order, the number of
calculations required to recover the information would be vast. It's totally infeasible.''

Brendon leaned forward and put his hand on the invisible wall, the red grid flashed momentarily.
``Exactly how many calculations are we talking about?''

``I don't know for sure,'' Colty said. ``I'm just starting this line of investigation.''

``Colty,'' Rachel said, ``we have machines on Earth that are capabile of extremely fast
computation. It's not outside the bounds of possibility that we could do the calculations
needed. We would need to know the details.''

Colty paced back and forth again. ``Unfortunately the unclassified dunari literature on
information compression is sparse. I feel like I'm having to derive basic results in that field
in order to make any progress.''

``It might be good if you came to Earth,'' Rachel said.

``What?''

``I agree,'' said Brendon, ``we need to continue this investigation in the Sol system in any
case to have easier access to the master genomic database. Also the computational resources on
Earth make the computer on \textit{Golden Light} look like an abacus.''

``And we need you, Colty,'' Rachel continued, ``we have the technology to gather the message but
it's your mathematics that will allow us to decode it.''

Colty felt overwhelmed. ``I\ldots\ I\ldots\ I'm not sure what to say. There is a lot to\ldots\
consider. Do you think it really is a message, then?''

``It has to be,'' said Brendon.

The two humans and one dunari continued talking about their collaboration and how it might be
arranged. How long they talked Colty wasn't quite sure. There were no clocks and anyway he was
too distracted by the revelations that had come forth. The idea that an incredibly powerful race
might be sending them a message embedded in every living cell of their bodies was overwhelming
enough. The idea of Colty going to Earth to collaborate with two human researchers trying to
read that message was overwhelming on top of that.

When the meeting finally drew to a close Brendon said his goodbyes and Colty was about to leave
as well.

Rachel called out to him. ``If you have a moment, I'd like to show you something.''

``Yes, certainly,'' Colty replied.

``Francine, access my personal storage.''

``Additional authentication required.'' A dark rectangular area appeared before Rachel and she
touched it a few times with her fingers and closed her eyes momementarily in concentration.

``Access granted.''

``Francine, show the image entitled `Cannon Mountain, July 2275.' ''

Suddenly the world changed. Colty almost fell over he was so disoriented. All the walls and even
the ceiling and floor seemed to vanish. Colty found himself standing on, or maybe floating just
above, what looked like green grass beside a lake. Rising up on either side of the lake were
steep slopes covered with green trees. Colty spun around and saw looming behind him a huge cliff
of grey rock, ringed with scruffy green. Overhead white clouds drifted in a dark sky with a
bright yellow sun.

Rachel was there too, standing on the grass beside him and looking out over the water. She
turned toward him. ``This is Earth,'' she said. She pointed toward the water. ``That's Profile
Lake, and that\ldots'' she pointed up at the cliff, ``is Cannon Mountain.''

Colty took a few steps as if to get a better view of the cliff but suddenly a red grid appeared
in the air in front of him. He stopped and clicked his teeth. He turned back toward Rachel.
``This is incredible.''

``I live not far from here,'' Rachel said. She pointed down the valley between the steeply
rising slopes. ``That way 100 karnons or so. This is a kind of park, reserved for its natural
beauty.''

``There is so much green.'' Colty looked at the strange yellow sun, unnaturally high in the sky,
and at the strange green trees. ``It seems so alien.''

``Just as Rujar seemed to me at first.''

``It is beautiful, though.'' Colty turned toward his friend but she was looking across the water
again. She had a strange expression on her face.

``I miss my home,'' she said softly. She sighed. ``Francine, close image.''

The world changed again. Now it was just the two of them in a small triangular room with dark
gray walls and thin red squares on the floor. Colty walked over to Rachel but she seemed very
somber.

``I will see you again when you return to Parnon,'' he said.

``I look forward to that,'' she replied. She sighed and said her final goodbye. The two turned
toward their respective doors. Rachel paused and said, ``Francine, close Link.'' The room
shrank. Where once Rachel stood there was now only a blank, gray wall. Colty stepped into the
waiting room. Once again he was on Rujar, in the city of Lungast.

Chark was laying on a seat waiting for him. ``How did it go?'' he asked. Colty didn't answer
right away. ``It's amazing technology,'' Chark continued.

``Yes,'' Colty said. He was lost in thought. The technology didn't matter. What mattered was
what came next.
