% Local IspellPersDict: aspell.dunari.pws

\chapter{107}

Colty felt strangely relaxed on the trip back to Parnon. He thought about the meeting. He
thought about going to Earth. He thought about Joleia. Going to Earth would be exciting beyond
imagining, yet the idea saddened him as well. Time with Joleia was exciting too in a different
way.

The sun flashed behind the buildings beside the tracks as the train raced by. Colty watched it
carefully. Laying on the seat next to him was a small child in farth-sleep. His mother was on
the other side of the aisle reading. Colty looked at the sleeping boy and clicked his teeth.
Perhaps you will also go to Earth, he thought. Your generation will travel in space.

The mother tapped her foot gently as she read. Perhaps feeling Colty's eye on her, she looked
up. Colty swayed his head slightly and she clicked her teeth politely. He turned again to watch
the world go by outside the window. Everything seemed the same and yet everything was different.
The future had arrived. The world of the dunari had been transformed and most of them didn't
even know.

When Colty got back to Parnon he found a note taped to the door of his apartment. ``Welcome
back,'' it said, ``let's meet for dinner at 80:100, Parnon time. You know the place. Touches...
Joleia.''

Colty just had time to unpack and then stop by the office to check his mail before walking over
to Dolarta's. The air seemed fresh, a little cooler than usual, and the Nermella was little more
than a breeze. The Rigat trees along the Yellow stood nearly in silence with only the occasional
rattle as a reminder of their presence.

Dolarta's was quiet as well. Perhaps the nice weather had encouraged people to spend time
outside. Colty did not have to wait long. ``Your other party is already here,'' the host told
him right away. Colty followed him as he made his way across the restaurant floor, winding
around the various privacy partitions. Eventually they came to a small partition in the back
overlooking the Varsynthia through the sunlight branches of a rigat tree. Joleia was there and
she stood up in greeting.

``So good to see you,'' she said as they crossed trunks. The host gave Colty a menu and then
left them alone. The two wrapped their trunks together and let their tenticles caress their
scales. Joleia was dressed in a lovely orange smock with diamond clips. They briefly touched
their trunk ridges.

``It's good to see you also, \textit{tinka},'' Colty replied. They lay down again and Colty took
a sip of juice from the cup beside his place.

``How was the meeting? What did Rachel want?''

Colty gently flapped his ears. He then started to tell Joleia about orbiting spherical clusters,
TARS co-sequences, and compression methods. Joleia listened attentively but Colty stopped
himself. ``I don't mean to go on about all those details,'' he said finally.

She swayed her head. ``It's fine,'' she said. ``I don't understand the technical stuff, but I
get the general idea. Someone is sending us a message.''

``It appears so.''

``Using your math.''

``The math I discovered, yes.''

``Who is sending the message?'' Joleia asked.

Colty shrugged. ``We have no idea. Whoever it is, they can control the orbits of star clusters
so they must be incredibly powerful.''

Joliea reflected for a moment and then tilted her head slightly. ``Jurita could do it.''

Her comment took Colty by surprise and he wasn't sure how to respond. He always understood
Jurita to be a symbol, a kind of moral guidepost or a source of comfort and inspiration. Despite
the stories, the real physical world was surely just the work of mindless physical laws.
However, Colty never imagined, during his musings as a young adult, that he would ever be faced
with something like this.

``You're right,'' he said finally, ``I'm sure She could.''

Joliea put her trunk against his and they again entwined tenticles. ``And the humans need you
and your math to help them figure this message out.''

Colty folded his ears back and lowered his head slightly.

``What's wrong?'' Joleia asked.

``They want me to go to Earth.''

Joleia spread her ears. ``Oh my, that's wonderful Colty.''

``You think so?''

``I do. You see, they do need you. They want to work with you. This isn't about them teaching us
any more. You are teaching them.''

``So you think I should go?'' Colty asked.

``Absolutely I think you should go,'' Joleia replied. She leaned forward slightly. ``You don't
want to?''

Colty shrugged. ``It would be an incredible experience. Yet\ldots it would take me away from
you.''

Joleia leaned against him slightly. ``We'll be fine,'' she said. ``I know it's a long trip but
I will happily wait for you.''

Colty clicked his teeth and the two wrapped their trunks together.

When the dinner arrived the two continued talking about all manner of things. They talked about
humans and the Earth. Yet they also talked about Parnon, about Joliea's work at the library,
and about Mezanic music. Colty felt completely at home with her. They made each other click. He
found it hard to imagine leaving her.

At the end of their meal they watched a small bird flit among the branches of the rigat tree
outside their window. It seemed to study them through the quartz glass almost as if it
recognized them somehow. Then with a sudden flurry of wings the bird shot into the air and
disappeared over the river in the direction of the university.

The waiter brought some sweet sagnum and fermented russart to drink. Joleia and Colty each took
a square of sagnum and nibbled at the spicy nut mixture.

``So I have a question for you, Sar. Coltinarly,'' Joleia said playfully as they were sipping
their russart.

``Is it about math?''

Joleia leaned toward him and took his trunk in hers. She looked at him with her ears spread and
said, ``Will you marry me?''

Colty spread his ears in surprise. ``I\ldots'' he hesitated and a look of disappointment spread
over her face that pained Colty to the core. ``You said I should go to Earth but how could that
be satisfying for you if we were married? I would be gone for many runion.''

Joleia took a deep breath. She looked at him eye to eye. ``I love you more than anything. Yet
despite that\ldots\ no, because of that\ldots\ I will never hold you back.'' She held his trunk
tightly. ``You are so smart, and yet you still don't understand. You said yourself that it's a
new world for the dunari. You aren't just part of that change, you are the focus of it. I would
be so incredibly proud to stand by your side as your wife as you lead us into our future.''

Colty shook slightly and he wrapped his trunk around Joleia's shoulder and drew her close to
him. ``I love you too, \textit{tinka},'' he said in her ear. ``I want to stand by your side as
well.''

``So is that a yes?'' Joleia said shaking her ears slightly.

``Yes,'' Colty said. ``It absolutely is a yes!''

\spacebreak

Colty stood by the window of the lecture hall in Nark. He looked across the Yellow to Fargon
where his unassuming little office resided. He thought perhaps he could see his office window
but the view was obscured by a large tree.

As Colty waited, faculty and students began to fill the hall. They had come to see his full
status presentation. It was, in some respects, the culmination of his life's work. All the
classes at Marlock College, all the hours studying, all the papers and conferences, and finally
his work at Parnon flowed to this moment. In theory his future career depended on his
presentation being accepted. If he failed, not only would he be forced to leave Parnon, but also
he would most likely never be accepted anywhere else.

Yet strangely Colty did not care. It no longer seemed to matter what Parnon thought of him.
Parnon needed him more than he needed it.

Colty turned away from the window and walked casually toward the board. There was still some
time before his presentation but he didn't want to lay down. Some might say that he should have
been going over his talk in his mind. It wasn't necessary. Instead he scanned the audience.
Almost everyone from the mathematics department was there and he recognized some faculty from
physics and engineering as well. Rumors about the implications of his work had spread and there
were even a couple of faculty from the biology department present. However, Colty noticed that
Rocalla had not come.

He was more interested, though, in the students. There was a surprising number of them, mostly
from his advanced classes, but also many less experienced students whom he did not recognize. He
wasn't sure why the students would be so fascinated by stellating co-sequences, but he supposed
it was his association with the humans that really interested them.

Dargon stepped up to the stage and spoke softly to Colty. ``It's a good showing. Will Ra'hel be
here?''

``No. She's on her way back from the observatory and won't be here until next runion.''

Colty was just as happy that Rachel couldn't make it. While her support had been invaluable he
wanted to make it clear that his work was his own. He didn't want anyone to suppose he had been
prompted by the humans.

Just as the room was reaching capacity Colty noticed Joleia slipping in through one of the rear
doors. She stood against the back wall and he gave her a brief nod. She swayed her head in
response.

``I think it's time to start,'' Dargon said. He walked over to the podium. The audience settled
down and found their seats. Dargon waited a few moments before continuing. ``We usually do these
things here instead of Fargon because Nark has larger lecture halls, but this time we should
have gone to Cogart.'' The audience shook their ears. ``I think there are some more seats in the
next room. Maybe we can move a couple in.''

There was some shifting around as two or three seats were added behind the back row. It wasn't
enough. Dargon glanced over at Colty and Colty bobbed his head slightly. ``I'd like to introduce
the full status presentation of Sar. Coltinarly, a new faculty member in the department of
mathematics, and---now, I've actually checked the records---I believe the youngest faculty
member to ever be part of our department in the history of Parnon.''

The audience stomped their feet.

``The title of his presentation is \textit{The primality of co-sequences with stellating bases
  as shown by isomorphic mapping from Haken collections to Zermelo-Fraenkel set theory.} I hope
I said that right.''

``Yes, just right,'' Colty said.

``In that case\ldots\ you have the stage Sar. Coltinarly.''

Colty stepped forward, hesitated for a moment, and then started to talk. He gave some background
for his work, including a brief literature review. He then launched into his contribution. The
talk was extremely detailed and technical. Colty knew it was the mathematics faculty he needed
to impress.

He presented the human mathematics without apology, describing the basic concepts as needed, and
outlining the nature of his isomorphism to human mathematical foundations. The time did not
allow him to present any detailed proofs but he did work through a proof sketch showing an
important connection between G-reductions and Ranjana's hyper sets. He pointed to the appendix
of his paper for more information.

Colty filled the board with figures and formula. He emphasized the conceptual core of his proofs
and their implications. He was animated yet also deliberate. On the podium were some papers
holding his notes. He never looked at them.

Finally with the background in place, Colty showed the exact form of Nular's theorem and
demonstrated the flaw in Nular's original reasoning. From there it was an easy matter to use
Ranjana's hyper induction on the more precise form of Nular's theorem to derive the final
result: all stellating co-sequences have prime bases.

When he finished he put down the chalk and the hall filled with the sound of enthusiastic
stomps. Dargon returned to the stage. ``Thank you, Sar. Coltinarly. We have time for some
questions.''

The first question came from Molan ``I'm very impressed with this work. Not even considering
your primary result, the isomorphism to human mathematical foundations has the potential to be
an enormous contribution. Congratulations.''

Colty bobbed his head politely as the room again filled with the sound of applause.

``I do have one question. I'm confused about how you collapsed the sub-chains after stellation.
Does Roak's Theorem actually apply there? What about the pre-stokes side condition?''

``Good question,'' Colty said. He erased a section of the board and stepped through the
reasoning in more detail. He fielded several more technical questions from the math faculty,
however Tarnock said nothing.

One of the biologists asked the next question. ``I heard that your work is of interest to the
humans. Could you elaborate on that?''

``The isomorphism between our mathematical structures works both ways, of course. With further
development it could be used by them to access any mathematical results we have and apply those
results to open problems in their domain.''

``I was thinking about\ldots\ something else.''

There was some scattered shaking of ears.

Colty clicked his teeth briefly. ``Yes, I understand. It's true that they've noticed co-sequence
based patterns in the genetic structure of life forms on both Rujar and Earth. The meaning of
those patterns isn't clear right now.''

``Are these patterns they did not know about before?'' another biologist called out.

``Yes. Actually I first noticed them myself while preparing my paper and brought the matter to
their attention.''

``Colty why don't you share with the audience what you told me earlier,'' Dargon said.

Colty bobbed his head. ``I've actually been asked by two human researchers, including Sar.
Ra'hel whom I believe most of you have seen on campus, to collaborate with them investigating
this matter.''

``Sar. Coltinarly has been invited to Earth,'' Dargon added, ``and furthermore his paper has
been accepted by a human mathematical journal.''

The applause broke out again and Colty swayed his head from side to side.

Next an engineer spoke up. ``Congratulations on your exceptional success. I'm concerned,
however, that your access to human technology and human knowledge gives you an unfair advantage
in your research program compared with the rest of us. I guess it's the way of the future. You
are ahead of your time.''

``It is my hope,'' Colty replied, ``that soon both our cultures will be able to share knowledge
freely to the benefit of all.''

There were no further questions so Dargon indicated that the public presentation was over. He
directed Colty's overseers to remain for further questioning while the rest of the audience
filed out of the hall. Colty tried to catch Joleia's eye before she left but she was standing
behind a much taller man.

When the hall was finally emptied the overseers settled down in the front seats while Colty
erased the board. Tarnock started the discussion.

``As you know, my concern,'' he said, ``is that we can't evaluate the novelty of the work since
we don't have access to the relevant literature. It's not that I don't trust Colty, but how can
we know what is original and what was taken from human resources?''

Colty put down the eraser and turned to face the committee. As was traditional in these
situations he would not address any points until explicitly asked to do so.

``I think this is an exceptional circumstance,'' Dargon said, ``and that some flexibility is in
order.''

``I agree,'' said Molan, ``the candidate for full status needs to show that he has become a
`respected and influential' member of the Argenian mathematical community. I think Sar.
Coltinarly has discharged that requirement admirably. It's safe to say the humans didn't create
that isomorphism.''

``Is it?'' Tarnock asked.

Dargon turned to Colty. ``Do you have any comments?''

Colty took a deep breath and stepped forward. ``I agree with Sar. Tarnock,'' he began, ``in that
there is an irregularity here. By its very nature my work can't be properly evaluated in the
context of dunari mathematics alone. I submit, however, that any first attempt at merging dunari
and human mathematical legacies will suffer from this problem. If we are to take advantage of
human knowledge at all then we must bend some rules to make that transition.''

Dargon swayed his head, but Tarnock seemed unconvinced. He fluttered slightly. ``Maybe so, but
should we compromise the integrity of our system doing so?''

Molan leaned forward slightly and changed the subject. ``What of this collaboration? Colty will
be gone for some time, I assume, but he is not yet eligible for leave even supposing we grant
him full status.''

``Yes,'' Dargon replied, ``Colty tells me the humans are looking into a visiting scholarship for
him so technically it would not be a leave but more like a transference. I'm sure we could make
some sort of arrangement in order to take advantage of this unique opportunity.''

``I have a couple of technical questions,'' Tarnock said with a loud flutter.

Tarnock then asked several questions related to Colty's isomorphism proding him about specific
details of notation, equivalence sets, and irreducibility of reductions. He did not ask anything
about co-sequences and was much more interested in the boundary between dunari and human
mathematics. Finally the questions ended and Colty once again erased the board. Tarnock
shrugged.

``Should we convene in private,'' Dargon asked, ``to discuss?''

``I'm content with granting full status right now,'' Molan said.

Dargon swayed his head. ``Tarnock?''

Tarnock tapped his tenticles thoughtfully on his knee. ``I agree,'' he said finally.

Dargon stood up and extended his trunk toward Colty. ``In that case\ldots\ congratulations
Professor Coltinarily!''

\spacebreak

Colty scrambled down the steep embankment along a rough trail forged by adventurous students. At
the bottom he walked a short distance under the shade of several rigats. He climbed over a low
rise and out onto a flat rock that angled down to the river's edge. There, sitting calmly under
a large narnolia tree was Rachel with her back toward him as she looked out over the water.

The Varsynthia was wide and smooth. It flowed by, slowly gurgling now and then with the
occasional swirling eddy. Colty paused and clicked his teeth quietly. The human looked entirely
at peace. She was a creature of the galaxy and as she sat beside the river Colty realized, maybe
for the first time, that she was not an alien.

``They told me I might find you here,'' Colty called out.

Rachel twisted around and smiled broadly when she saw him. She stood up as he came down the
slanted rock to meet her. ``It's great to see you, professor,'' she said. Stepping forward, she
put her arms around his shoulders and hugged him briefly. Colty could feel the warmth of her
body.

``Let's sit.'' Rachel motioned toward the rock at their feet. ``This is such a beautiful spot. I
was just watching the narnolia nuts floating down the river.''

The two of them settled down side by side, Colty stretched his legs out comfortably next to her
while she sat cross-legged. ``It is the season for them, I think,'' he said.

They sat quietly for a time watching the river. The red sun glinted off its waters. Although
they had much to talk about neither felt any particular rush. After a few moments Rachel opened
her pouch and took out a machine. She unfolded it and touched the display a couple of times.
Then she held out the device to Colty.

``For you,'' she said.

``What?''

``You shouldn't have to wait for me to find books. Now you can read anything you like in the
library of \textit{Golden Light}.''

Colty reached out his trunk and took the device in his tentacles. ``I am honored,'' he said.

``It's only fair. I have been reading in your library after all.''

``Does your government know you are doing this?''

``I think I will be able to get it approved.'' She smiled slightly.

Colty held the device up to look at it with his right eye. It was surprisingly rigid considering
how light it was. On the display were several symbols along with writing in Argenian. ``I don't
know how to use it,'' Colty said.

``It's not hard,'' Rachel replied. ``I'll show you.''

``Thank you. I don't know what else to say.''

``That is all you need to say. When you come to Earth you can use this device to access the
Meta-Library with no restrictions. It contains the totality of human knowledge.''

``I guess I have some reading to do,'' Colty said with a click of his teeth.

Rachel chuckled. ``I look forward to finding out what will happen when a mind like yours is
exposed to a library like ours.''

With Rachel's help Colty folded his machine and put it into his pouch. The two of them sat
quietly again for a time. Then Colty said, ``Joleia asked me to marry her.''

``She did? That's wonderful!'' Rachel's eyes were bright and she put her hand on Colty's
shoulder.

``Yes,'' he said with a shake of his ears. ``She's crazy, of course, but I'm glad of it.''

Rachel laughed lightly. ``When is the wedding?''

``We aren't sure yet, but soon. It will be before I leave for Earth.'' He paused. ``We would
both like it if you were there.''

``Colty. I would love to be there! You'll have to tell me how it works, though, so I don't
embarrass myself.''

\spacebreak

A light rain was falling as Colty stepped out of the Varnok train station. He pulled his cloak
tightly around him and then, raising his trunk, he attracted the attention of one of the waiting
cab drivers.

``Where to?'' the driver said as Colty got in.

``The First Office.''

Low clouds covered the sky giving the streets a dark, almost dusk-like feeling. Colty could see
a few artificial lights glowing in some of the buildings.

The First Office building looked the same as ever. There wasn't anything to suggest a special
event was about to occur. Colty did notice a news truck parked in the press parking area across
the street from the building, but of course that wasn't unusual.

``You want me to let you out in front?'' the driver asked.

``I was told there is an entrance on the south side,'' Colty replied.

The driver swayed his head. ``You must be important if you are going in that way.''

The driver rounded the corner and stopped in front of a gate with two police officers standing
beside it. Colty got out and trotted across the street. He slowed to a walk as he approached the
officers.

``The public entrance is around front,'' one officer said.

Colty showed his identification.

``Oh yes, they are expecting you, Professor Coltinarly. This way.''

Colty was shown through an unassuming door and down a hall lined with offices. Raindrops
trickled down the glass of the skylight overhead. After a short distance they came to a large
door and he was shown into a circular room reserved for special events and presentations. The
First Citizen was there wearing diplomatic colors standing with Lucasa and ambassador Marshall.
A short distance from them stood Tusk and Rachel talking together comfortably. Yet Colty was
surprised at how few people where there. He didn't notice anyone from the press.

Rachel's dress was stunning and obviously special. She wore a long garment that covered her
shoulders and arms and that flowed gracefully all the way down to her ankles. It was tan like
the deserts of Rujar, trimmed with yellow the color of harsna grass. The cuffs of the arms and
around the bottom were decorated with a graceful curving design in rubinum, a color Colty knew
she couldn't even see. About her waist she wore a narrow belt of silver, hooked in front. Her
black hair flowed down her neck in a wave, like a bit of night, and was held in place with a
silver clip. On her chest just below her left shoulder she wore a single flower unlike anything
Colty had ever seen. It had an elegant, red bloom and a strange, green stem.

``Colty!'' It was Tusk who first saw him. He greeted everyone and, with honor, clasped trunks
with the First Citizen. Then he stood in front of Rachel and clicked his teeth.

``Do you like it?'' Rachel said with a smile.

``What is that flower?'' Colty asked.

``It's a \textit{rose},'' said Rachel, ``from the gardens of \textit{Golden Light}.''

Tusk spoke to her in her native language and Rachel laughed. She replied in her language as
well, the words rolling off her light, airy voice in a natural way. The sound did not seem so
strange to Colty anymore.

Tusk shook his ears in response to whatever Rachel had said but then she added, ``We should use
Argenian. It's rude to talk when Colty can't understand.''

``We'll have to teach him some \textit{English},'' Tusk said, ``since he is going to Earth.'' He
touched his trunk to Colty's shoulder.

Rachel chuckled, ``Yes, or maybe \textit{French} or \textit{Manderin} or \textit{Hindi}.''

``You are just teasing me now,'' Tusk replied.

At that moment Lucasa stepped onto the platform with the First Citizen behind him. He thumped
the side of the podium with his trunk to get everyone's attention. ``We should get started,'' he
said. He waited until the room grew silent. ``I convene this very special assembly, on this hour
47:060, runion 115, arnoxium 22 since confederation.'' He paused. ``I give you the First Citizen
of Argenia.''

Everyone stomped their feet as Lucasa withdrew and the First Citizen came to the podium. ``It is
a rare and happy event,'' he began, ``when a First Citizen is in a position to bestow the Medal
of Argen. I am privileged to be in such a position. Every moment in history when this honor has
been presented, and every recipient of this honor has been special and unique. Yet I think I can
safely say that Argen himself could never have foreseen the circumstances that bring us here
today.'' He paused and then continued in a commanding voice. ``I call Jahelle to stand before
me.''

Rachel walked to a spot on the floor directly in front of and below the podium.

The First Citizen looked at her and then continued. ``The Medal of Argen cannot be bestowed on
someone who is not Argenian. However, by the Cartana Laws the First Citizen has the power to
grant citizenship to any individual. In this case, by vote of Parliment, it has been deemed that
you are worthy of citizenship.'' He paused and the again continued in a commanding voice.
``Jahelle, do you take onto yourself, willingly and with full knowledge, the rights and
responsibilities of that citizenship?''

Her voice was soft and light. ``I do.''

``Then so be it. From this moment forward, by the power vested in me as First Citizen of
Argenia, you are now Argenian.''

% TODO: The original First Citizen stepped down. I don't think that should happen.

The applause was loud. Colty could see Rachel take a deep breath but she didn't turn around. She
was following custom and faced only the First Citizen during the proceedings.

``Come forward, Jahelle.''

Rachel lightly ascended the steps to the podium where she stood next to the First Citizen. He
reached down and then held up the Medal of Argen in his trunk. ``It is the recommendation of
Parliament, and both the Second Citizen and I concur, that you be given the Medal of Argen.
Your ordeal in the desert and your actions that lead to the arrest and conviction of those
responsible for the disaster at Yartagard have set you apart. What you have done has no doubt
saved countless lives, both Argenian and Forbin alike. Congratulations.''

He then raised the medal and Rachel lowered her head. The First Citizen slipped the medal around
her neck and she adjusted her hair so the cord went underneath it. Again the room was filled
with the sound of stomping. The First Citizen extended his trunk and Rachel reached out her arm.
She grasped his trunk in her hand as he wrapped his tentacles around her elbow.

As the applause faded the First Citizen stepped back from the podium and Rachel took his place.
She adjusted the microphone slightly and then paused. She looked out over the room and then
addressed her country.

``I am deeply moved by this honor,'' she began. ``Never in my life did I imagine that I would be
standing here now, like this.'' She touched the medal briefly. ``I only did what I thought was
right, as I think we all do. I'm just happy that I was at the right place at the right time to
make a difference.

``I haven't lived here very long, but I have come to love this place. I love the trees and the
flowers. I love Parnon. I love the Nermella. Yet it is the people here and the friends I've come
to know that are the most special of all. Really there are too many to name, but I would
especially like to call out my host, Professor Coltinarly, who showed me tolerance and kindness
despite not being too sure about me at first.'' People shook their ears. ``I see goodness
everywhere I look. It makes me proud to be Argenian.''

Once again the room filled with applause. Rachel waited patiently for it to die down.

``That is all. Thank you again.''

With that she once more clasped her arm with the First Citizen's trunk and then climbed down the
steps from the podium.

Lucasa formally ended the proceedings and people crowded around Rachel taking pictures and
asking questions. She finally made her way back to where Colty was standing next to ambassador
Marshall. The ambassador reached out his hand and Rachel shook it.

``Congratulations, Jahelle,'' he said with a smile.

\spacebreak

Colty looked out the cab window while Joleia, his wife, leaned sleepily against him. The streets
of Parnon were nearly deserted. It was the end of the runion and most dunari were at home in
bed. Even the fields surrounding the town seemed subdued. The sun was covered with spots and its
light was softer and redder than usual.

They pulled into the air force base where the driver let them out at the gate. Colty had no
trouble getting inside. In fact, one of the guards, apparently relieved to have someone to talk
to, took extra time showing them the way to the right landing area. There they found a small
alien craft parked some distance from the building. Standing just outside the gate was a group
of three dunari loading some boxes onto a pallet. Standing by and watching the proceedings was
Rachel.

She was dressed in casual clothes. She had on dark tan pants and a light, cream colored top that
looked like something that would be good against the wind. Her hair was tied back, as usual, and
it bounced slightly from side to side as she turned her head.

Colty and Joleia walked out onto the landing pavement toward the people and Rachel soon noticed
them. She raised her hand in greeting and Colty did the same with his trunk. She walked toward
them and soon they were close enough to talk.

``You didn't have to come,'' Rachel said. ``You should be asleep!''

``We wouldn't miss it,'' Joleia replied.

Rachel smiled. ``Well, thank you. I do appreciate it.''

Colty took a deep breath and stood face to face with Rachel. It seemed so long ago when they
first met on that very same landing pavement. Then she was an alien creature. Now she was a
friend. Colty extended his trunk and Rachel clasped it in her hand. Her grip felt firm and yet
warm and gentle. They held each other that way for a time and then Rachel put her other arm
around Colty's shoulders, hugging him.

``Have a safe journey,'' Colty said as they pulled away from each other.

Rachel nodded. ``We will see each other again soon,'' she said, ``and the next time will be in a
city called \textit{London}.''

Colty clicked his teeth. ``I look forward to it.''

``Will you ever come back to Rujar?'' Joleia asked.

``Yes. I hope so,'' Rachel replied, ``after all, Nitooli is here. Anyway there is so much more
of my new country that I'd like to see.''

``You will be welcome here, always,'' Joleia said.

A soft beep came from a device tied to Rachel's wrist. She looked at it briefly. ``That's my
signal. I have to go now.''

``I will miss you,'' Colty said, shaking slightly.

Rachel put her arm around him again, holding him closely for a moment. Her eyes glistened. ``You
keep thinking about data compression,'' she said.

``Oh, he doesn't need any encouragement when it comes to math,'' Joleia replied with a shake of
her ears.

Rachel let go of Colty and stepped back. She smiled and nodded her head. Then she turned and,
without looking back, walked briskly to the craft that would take her to \textit{Summer Breeze}
and, ultimately, to that distant world she called home.

The two of them stood quietly watching her climb into the human aerocar. They watched as it sped
down the runway and lifted into the air. They watched as it climbed steeply upward, moving with
ever increasing speed until it was just a tiny dot. They watched it disappear with a green
streak leaving behind nothing but the orange sky tinted with dust driven from the distant desert
by the Nermella wind.

Colty looked down and then turned to Joleia. He put his trunk over her back as they headed
toward the waiting cab.

``The universe is not as big as it used to be,'' he said.

``Really?'' Joleia replied, ``to me it seems bigger than ever.''

THE END
