% Local IspellPersDict: aspell.dunari.pws

% Suppress the automatic addition of "Epilogue" to the TOC so it doesn't appear under Part III.
% Explicitly add "Epilogue" to the TOC at the same level as the other parts (top level).

\chapter*{Epilogue}
%\markboth{Epilogue}{Epilogue}
\addcontentsline{toc}{part}{Epilogue}

Colty steered the vehicle along the narrow, sandy road that wound beside the Sarnowledge River.
The tops of the mountains rose high above the valley, their snow capped peaks aglow with the
light of the Rujarian sun. Low yola trees grew thickly on their leeward slopes and haganbrush
covered their windward ramparts.

Eventually Colty came to a small town. A circular park of neatly trimmed brikken grass was
surrounded by several old but tidy buildings. At its center the Argenian flag flapped
energetically. He turned the car onto a side road that went past a ramshakle garage and then a
neighborhood consisting of two or three selecks. The houses were simple, but clean, partially
built into the hillside as was typical in the mountain country. The road bent upward toward
higher ground and made its way along the side of the valley following a narrow, gurgling brook.

After several karnons and several steep, twisting turns, the road leveled off just below the top
of a long, broad ridge. From there a grand view stretched out over the Sarnowledge Valley to the
flat lands far beyond. On a clear day one could even sometimes see Lungast. Yet Colty did not
stop to admire the scenery. Instead he continued through open fields until he came to a
distinctive boulder. There he turned onto a narrow lane and followed it for a short distance.
Around the side of a small rise, in the light of the sun and yet protected from the full
strength of the Nermella, he finally came to an unassuming house.

Colty parked in front of the small detached garage and turned off the engine. He got out and
took a deep breath. Her house looked good. It was well kept and the yard was tidy. It had a
large porch on the front, and a small patio at the back, under some rigat trees. Beautiful
gardens surrounded the building and lined the path from lane to door. Flowers in rubinum, red,
yellow, green---all the colors of the rainbow---were layed out before him.

``Colty!''

He looked up and saw Rachel coming out the door and down the step. She had her arms open and a
smile on her face. Colty spread his ears and came up the path. Rachel ran to meet him, throwing
her arms around his neck and shoulders. He wrapped his trunk around her waist.

They embraced for a long moment. ``Let me look at you, Rachel,'' Colty said in English. She
stepped back and smiled but held his trunk affectionately. Colty shook his ears slightly. ``My,
my, your hair is now completely gray. I think the last time I was here there was still a little
black.''

Rachel just laughed. ``Come on, let me show you around. I've got a new shade garden in the
back.''

When she finished the tour she offered Colty something to chew, which he graciously accepted,
and made herself some toast with genic butter. She went into the back room of her house where
unnaturally bright lights illuminated a garden of green plants. She picked a few strawberries
and added them to her plate. Then they settled on the front porch with Colty on a seat and
Rachel sitting cross-legged on a pad. There they could look out at the majestic view of the
valley and the rocky crags of Tomorsa Mountain above.

``How is Joleia?'' Rachel asked as she took a bite of toast.

``Busy,'' replied Colty. ``The National Library is just starting to convert their holdings to
electronic form. She's directing the project now. I don't know where she gets her energy.''

``Wonderful!''

``And would you believe Jarnia is applying to university already?''

``Is she going to Parnon?''

Colty shook his ears. ``She applied for my sake. Her father thinks it would please me, but I
know she'd rather go elsewhere. She's got applications in at several schools including the
University of California. She says she wants an adventure, but I think she really just wants to
get away from the stodgy likes of us.''

Colty clicked his teeth and Rachel chuckled slightly.

``And what about you?'' Colty asked. ``Are you still teaching?''

``Some. Mostly I tend my gardens.''

Colty snorted. ``Yes\ldots\ and\ldots? Come on, I know you better than that.''

Rachel smiled coyly. ``I might also be evaluating the logical consistency of Jenkins
multi-manifold theory.''

Colty shook his ears and Rachel continued, ``I've been using the neticular core at Cambridge
over the hyperluminal link to Earth. It's a bit awkward, but it works.''

The two of them talked for many arnets as they always tended to do. They talked of friends and
of family. They talked of Rujar and of Earth. They talked of the past and of the future,
switching back and forth between English and Argenian as the mood took them. They talked long
after the food ran out, punctuating their conversation with Colty's clicks and Rachel's airy
laughter. And all the while they lay in the light of the unmoving Rujaran sun with the Nermella
whispering in their ears.

Finally their conversation faded and each quietly enjoyed the company of the other. After
several long moments Colty fluttered softly and Rachel glanced his way. ``I'm getting old too,
\textit{tinka},'' he said staring out over the valley. ``I'm planning to retire soon. Joliea and
I are talking about moving into an apartment in Varnok\ldots'' His voice trailed off.

``That sounds nice.''

Colty turned to look at her, clicking his teeth. ``Maybe I'll start thinking about graph
extension fields.''

Rachel laughed out loud.

\spacebreak

On the far side of the galaxy, somewhere near the base of the Perseus sprial arm, there is an
ordinary yellow-orange star surrounded by an ordinary collection of planets. One of those
planets is almost entirely covered by ocean with only a few small, island continents rising
above the waves. In a city on one of those continents, in a brightly lit burrow, a six legged
creature begins to warble with delight. For it has just proved an amazing fact: \textit{all
  stellating co-sequences have prime bases.}
