% Local IspellPersDict: aspell.dunari.pws

\chapter{Prologue}

Crouched beside a small stream was a young girl with tangled black hair and dark eyes. In the
shade of a beech tree she was making a dam of small stones, gradually blocking the trickling
flow. After adding each stone she put a bit of grass in the water and watched it move through
the narrowing gap with ever increasing speed.

``Rachel,'' her mother called from the cottage, ``it's time to get ready for bed.''

She ignored her mother and instead put a flat piece of bark across the gap. She studied the
water as it piled up behind the bark and started flowing over it.

``Rachel!''

She stood up and turned toward the house. ``I'm building a power station, Mumma.''

``Well, you'll have to finish it tomorrow. It's time for bed.''

Kneeling down briefly, she removed the bark and watched the water rush through the opening,
dislodging one of the stones.

``Don't make me call you again.''

The girl turned and ran up the narrow path to the scruffy lawn, her bare feet covered with mud.
The warm summer sun was behind the trees, casting long shadows over the ground as she scampered
up the stone steps to her exasperated parent.

``Straight to the bath with you.''

By the time she was clean and in her night clothes the sun had set and darkness began to fill
the sky. ``Can I stay up and see the stars?'' she asked her father.

``Not tonight. The stars come out too late this time of year. I will read to you, though, for a
little while.''

Her bedroom was at the corner of the cottage, in the back. An orange rug spread over the old
wooden floor. Along the side of the room were shelves with toys, pictures, and even paper books.
Her bed was just below the window, which itself was open slightly to let in a breath of air. She
climbed under the sheets and propped herself up while her father sat next to the bed. He took
the reading glass from the table and touched the screen briefly.

``What will it be today?'' he asked. ``Margarite and the Mouse?''

``Yes\ldots\ please.''

Using only his thoughts her father interfaced with the device and restored their last bookmark.
Immediately the words appeared on the small plastic screen. Leaning back in the chair he
continued the story of the adventures of a girl and her robotic mouse. As he read, he gave each
character a different voice. It made Rachel giggle.

The lights in the room grew gradually brighter as the world outside grew gradually darker.
``That's enough for now,'' her father finally said.

``May I read another chapter on my own?''

Her father smiled. His young daughter, reading already, was well advanced for her age, and he
wanted to encourage her. ``Okay, but just one,'' he replied. He handed the reading glass to her
and was about to get up.

``Daddy?'' she asked, ``Are there really people living around other stars?''

``Probably, but we haven't met any yet so nobody knows for sure.''

``Why not?''

``It's a big universe and very old,'' he replied. ``We just haven't run into anyone else.''

``How many stars have we visited?''

``I don't know. Thousands, I guess.''

Rachel looked pensively at her reading glass. ``I wish I was Margarite so I could talk to people
from other stars.''

Her father chuckled. He leaned over and kissed her on the forehead. ``Just one chapter, then
lights out.''

After her father left Rachel read a little but she was distracted. She turned off the lights and
squirmed up the bed to look out the window. The trees were silhouetted against a deep purple
sky. Through their leaves she could see a bright point of light slowly moving. It was one of the
Arkanen stations, still catching the sun, where almost twenty thousand people lived and worked.
Behind the station, fainter and much farther away, she could see a star glittering in the
deepening dusk. She sighed, put her head on the pillow, and closed her eyes.

\spacebreak

\begin{singlespace}
\noindent Jilfora sings of love. \\
The energy of the universe \\
that builds \\
and destroys.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Jalam sings of strength. \\
The creator of worlds \\
to bear the weight \\
of the meek.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Jinjana sings of wisdom. \\
The knowledge of the void \\
of what is right \\
and wrong.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Jojen sings of spirit. \\
The fire of the soul \\
to drive \\
and to despair.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Jujar sings of home. \\
The gathering of all \\
of some \\
of one.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Jixita sings of patience. \\
The force of the desert wind \\
always blowing \\
in a rush.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Junarit sings of truth. \\
The truth of the sun and sky \\
revealed to all \\
as lies.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent The Seven sing together. \\
They weave their songs to a golden thread \\
the fabric of all that is \\
and is not.
\end{singlespace}
\vspace{1.0em}

\begin{singlespace}
\noindent Jurita sings the Song of Songs. \\
Shaping the world from golden cloth \\
she makes a perfect everything \\
from nothing.
\end{singlespace}
