
% SUMMARY: Narlock talks with the humans about Rachel's kidnapping.

Detective Narlock was at his desk flipping through the report from Parnon when his secretary
knocked on the door.

``They're here, sir,'' she said.

Narlock sighed. He wasn't looking forward to this. He tossed the papers aside, picked up a
notebook, and walked down the short hall to the first interview room. Inside, standing together
beside the table, were two humans. Narlock recognized the tall human from television as
ambassador Marshall. The other human had a name Narlock couldn't begin to pronounce. He got the
impression that it was a female and, unlike Marshall, she had slightly pointed eyes. Narlock
wondered briefly if it was some kind of sexual difference.

After the introductions, Narlock lay down beside the table and put his notepad in front of him.
Lucasa lay down opposite Narlock and the two humans, reluctantly it seemed, lowered themselves
onto their crossed legs on the other side of the table.

``She's been gone for almost two hours,'' Marshall began, anxious to get started. His Argenian
was almost perfect and he spoke with only a slight accent.

``I've read the teletype,'' Narlock said. He hated going over things he already knew. ``We've
already started our search, but my question is, frankly, how can you help us?''

Marshall glanced at the female. She leaned forward slightly. ``We are using high resolution
imaging technology to search for her from space but that will take a long time. Rujaria is a big
place and we aren't sure where to look.''

``Also if she's in a building or in a vehicle we won't be able to see her,'' Marshall added.

Narlock wrote on his pad. He made a mental note to walk under trees from then on while outside.
``I understand that with your help the Paron police located her communications device in the
river,'' he said.

``Yes,'' the female replied. ``the device was smashed but the tracking circuit was far too small
to be destroyed by such simple means\ldots\ not that it helps us.''

``Too bad you don't have one of those tracking circuits attached directly to her,'' Narlock said
with a click of his teeth.

The female shook her head slightly. ``She refuses to wear them.''

Narlock made a few more notes. ``So if we can give you a general idea where to look you'll be
able to just see her as long as she's outside.''

``Exactly,'' Marshall replied.

``In that case, you might start looking at Forbin military bases.''

Marshall sighed and the two humans looked at each other briefly. ``Why?'' he asked.

Narlock tapped his pencil against the table. This was the part he didn't like. Finally he spoke.
``We've received some intelligence that suggests the Forbinites are behind this. However,
something about that information doesn't add up. I specialize in counter-terrorism and I've been
called in to look at\ldots\ other possibilities.''

The female replied. ``If the Forbinites are not behind this, then why should we observe their
bases?''

``We need to rule it out,'' Narlock explained, ``and I can't easily send my agents into Forbin
to do that.''

Marshall nodded his head. ``That makes sense.''

The female was less convinced. ``I don't feel comfortable stopping our current observations. It
makes sense to me that she'd still be in Argenia, probably not far from Parnon.''

``We deal with kidnappings all the time,'' Narlock said. ``We already have people in the field
following standard procedures to locate her.''

``Can we observe both areas?'' Marshall asked the female.

Narlock thought she looked frustrated. She started speaking in the strange human language but
Marshall raised his hand to interrupt her. She took a breath and started again in Argenian. ``We
could but it will take longer.''

Marshall reflected for a while. The female seemed to study him.

``I agree with the Detective,'' he said at last. ``We should use our abilities to complement the
IIB effort, not duplicate it. We'll start observing the Forbin bases and the roads leading to
them, and let the IIB cover Argenia.''

Narlock didn't know much about humans but it was obvious to him that the female was displeased
with the ambassador's decision. However, she said nothing.

When the meeting ended, everyone stood up. Ordinarily that would be when Narlock would grip
trunks with the others but he wasn't sure what to do. Sensing his unease Marshall made no
particular move. ``We will stay in touch,'' he said.

