% Local IspellPersDict: aspell.dunari.pws

\chapter{19}

Chark and Junella found their seats on the train and packed a few items in the drawers beneath
them. They settled down and Junella pulled the booklet about Morgolan Park from her pouch and
started flipping through it. ``This is what I was telling you about,'' she said as she gave the
booklet to Chark, open to a certain page. ``Narleia Falls.''

Chark looked at the picture of a long ribbon of water cascading down a tall rocky cliff into the
shadows below.

``It's just down the road from our hotel,'' Junella said. ``There are tours that go there.''

``It's lovely,'' Chark replied. He passed the booklet back to her and then turned to study the
people on the Lungast platform. Some were running to their trains and some just lay quietly on
public seats watching the world go by. Chark found himself looking up at the sky. The sun was at
his back and the distant horizon seemed dark and mysterious.

``Are you thinking about work again?'' Junella said.

Chark clicked his teeth. ``There's been a lot to think about. I told you there's talk about the
humans setting up a compound at the base, and that's on top of planning for the Link\ldots''

Junella gave a low fluttering growl and Chark stopped abruptly. ``It's our vacation,'' she said,
``leave your work in Lungast. Please?''

``Of course.'' Chark shifted slightly and put his trunk over Junella's back. ``Let me take
another look at that booklet. Maybe there's something else I'd like to see.''

After a time the train started to move. Chark watched as the station slid behind them. The
streets of Lungast quickly followed and soon they were moving swiftly across the countryside.
The trip was long but they were prepared for that. Junella wrote in her diary and Chark read
some technical papers. They chatted now and then with other passengers, and Chark often took
walks down the length of the train to loosen up. Everything seemed entirely normal. It was as if
the humans didn't exist.

The train sped across the Lunarsh flatlands through karnon after karnon of raran fields. They
were heading almost directly east toward the Narlar and behind them the sun crept every so
slowly toward the horizon.

Chark and Junella farth-slept in their seats, getting some rest at least. They wouldn't be able
to paren-sleep until they reached their hotel and by that time they would probably be quite
exhausted. They drank some water provided by the train company and nibbled on some nuts they
brought in their pouches. Yet like their sleep a proper meal would have to wait.

Finally they reached the city of Arbenol. There they made a connection with another train.
Arbenol was smaller than Lungast and the station was correspondingly smaller as well. Despite
the wait it was actually rather pleasant.

Their next train was specifically going to Morgolan Park with destinations at various park
locations. It was much smaller than the first had been but it was very clean and well
maintained. They went mostly south at first, following a small river for a time. Soon, however,
the river disappeared and the route started to climb steadily. They were heading over the Unasta
Highlands. From their window Chark could see the sun hovering just above the horizon. Its large
red disk had numerous spots. Doubtless there would be flashes soon. It was lucky they would be
past the Narlar when they happened.

At the height of land the train turned toward the east again and began to descend into the
valley of Lake Erbum. It was the largest lake in Argenia, even larger than some of the seas. Yet
unlike a sea Lake Erbum was drained, far to the south, by the young Varsynthia River. Long and
narrow, the lake stretched for hundreds of karnons. One end was in the full light of day while
the other was well into the shallow night.

Artificial lights came on in the train as the sun sank below the hills behind them. Junella
looked out the window peering into the dusk. ``I see the lake ahead,'' she said. In the
distance, still many karnons away, Chark could see a large, dark body of water, glinting here
and there with the red glow of the fading sun.

``It's exciting, yes?'' said the voice of an older man in the next booth.

``Yes,'' Chark replied, ``an adventure.''

``We love it,'' the man's wife said. ``You should go all the way to the end of the lake. The
Varnaria River that flows into it is almost as cold as ice. It comes down from glaciers in the
mountains, you know.''

``We won't be going that far up,'' Junella said. ``We have reservations at Morlio.''

Eventually the train came to the lake and stopped at the town of Ebrious. Chark and Junella took
advantage of the break to stretch their legs. The sun was just below the ridge behind the town,
its light glowed about the hills as if they were on fire. Over the lake the sky was filled with
clouds glowing brightly in rubinum. The water, reflecting the color of the sky, seemed to glow
rubinum as well.

Junella gasped at the scenery as she looked around. ``It's gorgeous,'' she said. Chark was
surprised at how warm it was. The Nermella, carrying the heat of the deep day over 6,000
karnons, blew gently over their bodies.

``We should get some heated jackets here,'' he said.

There were plenty to choose from. Catering to tourists many stores bordered the walkways around
the station. Chark and Junella purchased jackets lined with heating pads that could be charged,
for a fee of course, at various places in the Morgolan valley. Along with their new jackets came
a pamphlet describing the signs of lethargy and what to do about them. Junella read every word
carefully.

The two spent time browsing in a small museum at the water's edge. They looked at displays about
brave explorers of the past who dared to travel far up the Varnaria into the night. In more
modern times an expedition using special vehicles with tracks instead of wheels and sprouting
artificial lights powered by on-board electric generators climbed to the very headwaters of the
dark river. There they found a veritable mountain of ice, and high above the valley of the lake
a great, icy plateau beneath a sky with dozens of stars.

``We know so little about even our own world,'' Chark said to his wife.

They returned to the train and were soon traveling again. The train went east and north to
follow, more or less, the shoreline of the lake. Soon they passed a sign beside the tracks that
said, ``Welcome to Morgolan Park.'' The clouds closed in, the sky grew dark and it started to
rain steadily. Artificial lights in the train blazed brightly but outside the window there was
only wet grayness. Junella held Chark's trunk for reassurance.

``It's hard to believe that some people live here all the time,'' Chark said.

``I suppose you get used to it after a while.''

A voice came over the train intercom. ``We're crossing the Narlar,'' it said, ``welcome to the
night.''

Eventually the rain stopped and the clouds parted. The sky grew lighter again. On either side of
the lake the Morgolan Mountains rose up, towering ever higher as the train moved forward. Rocky
crags, still lite by the sun, seemed to almost float above the dark waters. The train started to
slow and Junella asked what was wrong. ``I think we might be approaching the Mount Jaola
platform,'' Chark said.

The train stopped and most passengers got up to take in the view. Chark and Junella donned their
heated jackets, although the warmth from the charge given them in Ebrious was almost gone, and
joined the others. The platform was on a knoll near the lake. Rising above the distant shore was
Mount Jaola, a tall pyramid of rock that reached into the deep green sky. The top of the
mountain was ablaze in sunlight, glowing bright red against the surrounding dusk. On its summit
there was a blanket of fresh snow that stretched far down its flanks.

Soon they were underway again, speeding into the growing darkness. Yet it wasn't much farther to
their destination. The streets of Morlio were lit with artificial lights permanently installed
along the roadways. They had never seen anything like it before. When they got to the hotel it
was also illuminated with more artificial lights than they had ever seen in one place. The lobby
was as bright and warm as that of any hotel in Lungast or even Varnok.

``Good Brightness,'' the woman at the counter greeted them. Chark filled out the paperwork and
got the key to their room. Just as they were about to leave a chart on the wall beside the
counter caught his eye.

``What's that?'' he asked.

``Those are the times when the alien ship goes overhead,'' she replied.

``Really?'' Chark walked over to look more closely.

``There's a path beside the entrance that leads to a small rise. People go up there to see it.''

Chark and Junella went to their room and unpacked their things. As with the rest of the hotel
the room was brightly lite with artificial lights. On the table was a paper with bold print
across the top: ``Please inform management at once if a light burns out.''

``I hope the power doesn't fail,'' Junella said nervously.

``I'm sure they have some kind of backup system. I have to believe every building here does.''

Their room was small and simple, but comfortable. The windows were strangely little but they
didn't let in much light anyway and the extra wall space made the room seem more spacious than
otherwise. They relaxed and had some food together, enjoying a good meal for the first time
since before their travels. It had been a long trip to the edge of the world.

After they ate the two slept for a while, enjoying some much needed paren-sleep. When they awoke
they ate again, feeling decadent on their vacation. Junella wanted to see the sights but Chark
convinced her to start with a walk up the rise to see with their own eyes the human ship.

``According to their chart, it should be going overhead in just a few arnets,'' he added looking
at the clock on the wall.

Junella fluttered slightly but then clicked her teeth and touched his ear. ``Let's do it.''

Carrying portable lights provided by the hotel, the two made their way along the narrow,
unofficial path. They pressed closely together both from the lack of space and to seek
reassurance in the gloom. The sky was a deep rubinum and nearly featureless. The silhouettes of
the mountains could be seen rising above them and reaching toward the darkness of space.

At the top of the rise a group of people were gathered, their artificial lights glowing like a
cluster of stars. ``Here are two more,'' someone said as Chark and Junella joined them. The man
who had spoken was an older gentleman. There was also a young family and a young woman,
apparently alone.

``Is that it?'' Junella asked. She pointed her trunk to a spot just above one of the mountains
while gazing intently with her right eye. Everyone turned to look.

``You have a good eye,'' the older man said, ``but that's not it. That's Aruk, the brightest
star in the sky.''

``I don't see it,'' the young boy said. His father tried to point it out.

``Their ship is much brighter,'' the older man added.

``You've seen it?'' Chark asked.

``Many times.''

The young woman stood quietly off to the side. She seemed to be watching the mountains to the
northwest and not paying any attention to the others. ``There!'' she said suddenly, pointing her
trunk toward the ridge.

Everyone turned to look but nobody spoke. Junella leaned against Chark, holding him tightly.
Silently and surprisingly slowly, a bright red light was moving past the ridge. It was nothing
more than a shining dot crawling against the dark sky, yet the significance of it hung heavily
in the air.

``We are not alone,'' the young woman said simply.

% There are two(?) other places where I say this: the first time is in the scene where Jernumia
% is introduced. I forget exactly when the second time is. Those earlier occurrences detract
% from this utterance.

The light, far brighter than Aruk, gradually ascended, moving in a graceful arc. Everyone turned
to follow it.

``They call it \textit{Summer Breeze},'' Chark said.

``How do you know that?'' the young woman asked.

Chark hesitated. ``I\ldots\ I heard it on the news.''

The group watched as the ship passed nearly overhead and then dipped toward the southeastern
horizon. It was moving into deeper darkness and would soon be passing over a part of Rujar no
dunari had ever seen.

``I think they will be our friends,'' the young woman said.

``What makes you say that?'' the older man asked.

``Because they named their ship \textit{Summer Breeze}.''

