
Chark lay nervously in the luxurious seat in the waiting area outside the First Conference Room.
He checked his papers yet again and, looking up, saw the receptionist eyeing him. He clicked his
teeth slightly and shifted a bit. He was much relieved when the door opened and Lucasa and Tusk
came in together.

``Chark! Good to see you,'' Tusk said. The three of them touched each other's shoulders in
friendly greeting and then Lucasa spoke briefly to the woman at the desk. She nodded and opened
the door to the room to let them enter. Inside was a spectacular carved rosenwood desk
surrounded by several seats each engraved with the seal of Argenia. Large windows on one wall
looked over Tarakia Street.

The three of them selected locations together along the side with the door. ``How was the trip
from Lungast?'' Lucasa asked.

``Very good,'' said Chark.

``Is the Link ready?''

``Nearly.''

They chatted comfortably and Chark was glad to catch up with his friends. Yet he felt
apprehensive as well. ``I don't understand what this meeting is about, exactly,'' he said.

``It will be historic,'' Tusk replied.

As if on cue the door opened and the First Citizen walked in along with the Forbin ambassador to
Argenia and the human ambassador Thomas Marshall. The three men already present stood up and
introductions were made all around. Chark had never met the First Citizen in person before and
he felt overwhelmed to be in the same room with the leader of Argenia and three ambassadors, one
human at that. Yet the First Citizen had a way of making people feel at ease. ``Thanks for
coming Captain Numarky. We really appreciate you taking time to be here.''

``Of course.''

``If I may\ldots\ what is this about, FC?'' the Forbin ambassador asked. He spoke Argenian well
but his accent was as striking as his markings.

The First Citizen took the seat at the head of the table and everyone else lay down as well. He
turned toward Chark. ``Captain Numarky is in charge of the Link effort at Lungast. We know, of
course, that a similar Link is being constructed in Forbin as well.''

``Yes,'' the Forbin ambassador replied, ``at Yatolia.''

% TODO: There is a consistency problem here. In the television show Chark was interviewed for he
% says that they plan to use the Link for Argenian/Forbin communication. Thus the idea isn't new
% at this point.

``What I'm proposing,'' the First Citizen continued, ``is that we use the Link to connect
ourselves not only to the humans, but also to each other.'' He glanced at Marshall who sat
quietly listening intently. ``Tom and I have discussed this and he assures me the humans can
connect the two ground facilities whenever we want.''

The First Citizen looked at Chark. ``Does that sound reasonable to you?''

``Um\ldots,'' he looked at Marshall briefly, ``I'm sure the humans can do the switching as
needed. I don't think it will affect what we are doing.''

``We were thinking of allowing three-way connections,'' Marshall said.

``I see.'' Chark's mind raced. ``We might have to reconfigure a few things. I don't think it
will be too difficult.''

``I'm sure the technical details can be managed,'' the First Citizen said. He turned toward the
Forbin ambassador, ``I'd like you to bring this proposal to your government. I believe if we can
use the Link to hold meetings such as this without flying half way around the world it will help
us understand each other and that can only be good for both our nations.''

``I understand,'' the Forbin ambassador said slowly. ``Yet if ambassador Mar'hall knows of this
idea than surely ambassador LeClaire already knows about it.''

``This idea came from the FC,'' Marshall said. ``I didn't want to speak to my people until
you've had a chance to discuss it with the Prefect. If your government agrees that it's
something you want to do, I'll coordinate with Penelope. We can definitely make it happen and I,
for one, would greatly encourage it.''

During the rest of the meeting they discussed various details of the plan, both political and
technical. The Forbin ambassador worried about expenses and payments, Lucasa worried about
scheduling, and Chark worried about reconfiguring the system for three-way connections. Tusk
mostly listened but, along with Marshall, added words of encouragement.

Finally the group hammered out a general consensus of what the new Link would be like. The First
Citizen consulted his notes and looked at the Forbin ambassador. ``I'll get this written up more
formally and send you a copy within the hour.''

``Yes, thank you First Citizen.''

Just when it seemed the meeting was about to end Marshall said, ``There is one other item of
business I'd like to cover.'' He glanced at the First Citizen.

``Yes, Tom, I know. Please go ahead.''

``The telescope needs a name.''

``A name?'' said the Forbin ambassador.

``It's been provisionally called Project Farsight since it will be used, in part, for exploring
the far side of the galaxy,'' Marshall explained, ``but it is traditional to rename telescopes
when they are installed to something more\ldots\ appropriate to their location.'' He paused.
``Rachel isn't here, but she suggested to me that it be given a dunari name.''

``What about Nermia?'' Chark suggested.

``We don't need to decide right now,'' Marshall said, ``but I do ask you, if it's not too much
trouble, to take the question to your governments for reflection. I think I speak for all my
people when I say it's my hope that giving the telescope a dunari name will be symbolic of
peaceful collaboration among all three of our nations.''

``What about Nitooli?'' Tusk said suddenly.

Everyone turned toward him.

``Long ago,'' Tusk began, ``before Forbin and Argenia existed, there were primitive people
called the Tars who lived in the northern desert at the base of the Haracks Range. Each Tars
herd had a kind of magic man, called a Nitooli, who took care of the spiritual health of his
herd. Every seventeen seasons he would climb the highest mountain in his homeland and look out
over the desert plains searching for a vision from the gods that would instruct him how to best
lead his people.''

``That's very interesting,'' said Marshall. ``Thank you for that suggestion.''
