
% SUMMARY: Chark works on the Link. Here we don't get a clear idea of what the Link is just
% yet.

% TODO: This scene is a bit uninspired. I think I need to work out some conflict for Chark and
% then allude to that conflict in this scene (and others). I also need to understand the role
% played by the Link in the story. Right now it serves as a demonstration of Colty's acceptance
% by the humans and his ascendancy to the new way of the world (when Colty uses the link with
% confidence and comfort in Professor). Is that the only purpose it serves? That's probably not
% enough.

Chark met the crew from \textit{Science News} in the small lobby of the administration building
on the Lungast base. He clasped trunks with Kip, the producer of the show, and also with the
technical crew and an attractive young woman named Jippa who was the face of the show.

``I try to watch your program every year,'' Chark said, ``but it can be hard to find the time.''

``We understand,'' said Kip. ``It's very exciting for us to be here.''

Chark showed the crew into a small conference room just to get started. The cameraman was
already recording some shots of Chark and Jippa together while they clipped a microphone to his
shroud.

``As you are probably aware some parts of the base are off limits to cameras,'' Chark cautioned.

``Yes, yes,'' Kip replied, ``believe me when I say we have been made painfully aware.''

After the lights were set up Kip arranged two seats for Chark and Jippa.

``Ok, we are just going to ask some of the questions we talked about,'' Jippa said, ``and then
you can show us around.''

``That's fine.''

With a camera on its stand and Kip behind the cameraman, Chark and Jippa lay down on the seats.
Chark smoothed his shround and Jippa shook her head slightly. ``Don't worry about making a
mistake. This isn't live television so we can just redo things and let the editors sort it
out.''

Chark clicked his teeth nervously.

``We won't bother with introductions,'' Jippa said, ``the narrator will take care of that. We'll
just cut your answers into the show at appropriate places, okay?''

``Sure, yes,''

Kip swayed his head slightly. ``Let's go,'' he said. Chark tried to sit up and look intelligent.

Jippa looked over her notes and then said, ``Tell us a little about the Link. What is it?''

Chark took a breath. ``The Link is a communications system that will allow us to easily talk
with not only the humans but also our sister installation in Forbin.''

``We understand that the Link uses human technology. Is that right?''

``Yes,'' Chark said, ``the humans have been instrumental in setting it up. However, we are also
using some of our technology as well. The information is processed by human codecs but we boost
the signal to their network using plain old radio.''

``Can you explain what a codec is?'' Jippa asked.

``Oh yes. It's a device for encoding and decoding information. We've been making use of simple
codec-like devices for some time but the human technology is, as I'm sure you can imagine, far
more sophisticated.''

``You said we can use this system to communicate with Forbin. How is it different than the
telephone system we have already?''

``It's much different,'' Chark said. ``When we go over to the other building I'll give you a
demonstration of what's set up now. That's the only way to really appreciate it.''

Kip raised his trunk. ``Why don't you skip the next couple of questions Jip. I think we can go
over them later at the demo.''

Jippa swayed her head and looked over her notes again. ``Ah\ldots\ do you think this Link will
help bring Argenia and Forbin together?''

``I certainly hope so,'' Chark said, ``it can't hurt.''

``Okay, now the question I think everyone in our audience is really curious about: what is it
like to work with the humans?''

Chark paused briefly and clicked his teeth. ``In many ways it's very much like working with
anyone. They each have their own personality, just as we do. I know many people are afraid of
them, but there is no need to be.''

``Do you trust the humans?''

``Yes I do.''
