
Tusk waited patiently in Chark's office. As the official representative of the embassy it was
perhaps an affront to be kept waiting, but Tusk wasn't concerned. He knew Chark probably had
many things to do besides entertain a linguist who was just there to observe. He lay on the seat
quietly and looked out the window to the mountains beyond. Only a few short seasons ago, it
seemed, he was out of a job and shamed. How times change.

``Tusk!'' Chark came into the room with ears spread. Tusk stood up and the two briefly wrapped
their trunks around each other's shoulders.

``We are very well met,'' Tusk said with a click of his teeth. ``It's been a while.''

``Indeed,'' Chark replied, ``and we have been busy. I thought we wouldn't have the second panel
set ready by today but it turns out we do.''

``You are all hard workers,'' Tusk said.

``Not just us. The humans can be rather intent as well. They might have to sleep frequently but
they work through the winter.''

Tusk shook his ears. ``Then I'm sure you got along fine.''

``Come,'' Chark replied, ``let me show you the conference room.'' Chark led Tusk down a flight
of stairs and along a long hallway to an unassuming door. Inside was what appeared to be a small
waiting room. There was a counter staffed by a military receptionist, a number of seats, and a
desk in the corner.

``Shall we go in?'' Chark invited. He gestured toward a simple door next to the reception
counter. There was a yellow light above it.

When they entered Tusk found himself in a triangular room. The floor was made of dark gray tiles
with the outline of a red square on each tile. The walls and ceiling were also dark gray and
almost completely blank. Along the wall, running all around the room, was a single red line. In
the two walls opposite from where they came in were other doors of the same sort as the one they
had entered. Tusk noticed that the red lines on the walls and floor were not painted. Instead
the color was a glow from the material itself. Even the lights in the ceiling were just glowing
patches and not ordinary light fixtures.

Near the middle of the room stood Tarmala facing a tall human standing slightly to his left.
When Chark and Tusk entered the human gestured toward them and the Tarmala turned around.

``Oh. I didn't hear you come in,'' he said.

Tarmala introduced the human and Tusk greeted him in his own language. He replied in synthesized
Argenian.

``How is it going?'' Chark asked.

``All systems are fine,'' Tarmala replied. ``Our equipment is working perfectly.''

``Our equipment?'' Tusk asked.

``Yes,'' explained Chark. ``We are modulating the holographic data onto an RF carrier using
dunari technology and beaming it up to the human's communication network. The high power
connection gives a better fidelity signal.''

``If you say so,'' Tusk replied. He started to walk toward the unoccupied side of the room to
get a feeling for the space when suddenly a grid of red lines appeared in the air in front of
his face and he stopped abruptly.

``Be careful,'' Chark said. He walked up behind Tusk and held up his trunk. With a quick motion
he swung it in the direction of the empty space and the red grid appeared again, very brightly.
His trunk thudded against something hard even though it appeared to be stopping in thin air.

``It's a wall,'' Chark explained. ``Just like over there.'' He pointed to the dark gray wall
behind them with the door through which they came in. ``The red lines appear as a warning to
stop people from accidently walking into it.''

Tusk clicked his teeth. ``Of course,'' he said, ``I should have realized.''

The human spoke, his words obviously being translated by their technology as quickly as he
uttered them. ``I'm at the Forbin facility,'' he said. He pointed toward the empty side of the
room, ``and over there is Nermia.''

``It's a super luminal connection to Nermia,'' Chark explained, ``from \textit{Golden Light} to
the construction facility in orbit there.''

At that moment Marshall came through the door on the Nermia side. He seemed surprised to see so
many people. ``We are well met,'' he said. ``And I thought I was early. Tusk, good to see you
again.''

``And you,'' Tusk replied. ``You are at Nermia? You certainly get around.''

Marshall laughed briefly. ``Yes, and the irony of it is that I don't really like to travel.''

``I think you chose the wrong profession, Tom,'' the other human said.

Tusk studied Marshall closely. It was as if the human was standing right next to him. He reached
out his trunk and touched the invisible wall and the red grid flashed briefly. ``This is
incredible,'' he said, ``it is so clear.''

``That big antenna you have at Lungast helps,'' Marshall explained. ``If you look carefully
you'll see the Forbin connection is blurrier.''

``The holographic displays generate a lot of data,'' Chark said. ``We are sending it on an ultra
high frequency carrier. The Forbinites don't have that technology yet.''

% I think the human on the Forbin side should be the human ambassador to Forbin.

The door on the Forbin side, where the first human stood, opened and Kenellio entered. She was
dressed in a surprisingly casual style. Tusk recognized her from the initial meeting with the
humans at Garnick.

``Greetings, all,'' she said. ``I'm not sure why I need to be here, but here I am.'' Her
Argenian was nearly perfect.

Marshall stepped forward and put his hand on the wall separating the spaces. ``We are all here
to test the system.'' He took a deep breath. ``Shall we start?''

``Yes, go ahead Tom,'' the other human said.

Marshall stood up straight again. ``This hour marks the initiation of the Link, a communications
system that binds together three cultures: human, Argenian, and Forbin. May this link be used to
deepen our understanding of each other and improve the lives of all.''

``Very good,'' the other human said.

``We are preparing a press release,'' Kenellio said, ``that will appear in our news media
shortly. I hope we can convince your First Citizen to talk with the Prefect of Forbin here.''

``Yes,'' Tusk said. ``I've been authorized to say that he looks forward to such a meeting.''

``That is fine. Are we done here?'' Kenellio asked.

``Actually there are some other tests we wanted to run,'' the other human said. ``This is really
the first time we've had the opportunity to run the Link with all three sites engaged.''

He touched the invisible wall and a rectangular display opened in the air in front of him. He
seemed to study it for a moment. ``The translation system appears to be working fine. You are
all speaking your native languages, yes? Can we all understand each other?''

``There's a bit of a delay in the Forbin to English conversion,'' Marshall said. ``It's not
really a problem.''

Tusk swayed his head. ``I don't see how that could be avoided unless you have technology that
can react to words before they are spoken. Forbin word order is almost exactly the opposite of
English word order.''

``He's right,'' the other human said, ``there's nothing we can do except maybe talk faster.''

% Some other test as well?

