
Colty stood on the emblem of Menola before his overseers and nervously reviewed their faces. As
usual Dargon had an expression that seemed like nothing in the world could dampen his mood.
Molan was almost impossible to read. Tarnock, in contrast, was easy.

``Whatever became of graph extension fields,'' Tarnock asked with a cool voice dripping with
displeasure.

Colty was ready for that question. ``I started to investigate them but then I had a breakthrough
with stellating co-sequences and I felt compelled to explore that direction to its conclusion.''

``Quite right,'' Dargon said as he flipped through the manuscript of Colty's paper. ``It's good
to go where inspiration takes you.''

Tarnock was not as easily convinced. ``This paper is unpublishable,'' he said to Dargon as much
as to Colty. ``Your use of human mathematics can't be reviewed by your peers or understood by
the audience of any dunari journal.''

``I understand,'' Colty said, ``that's why\ldots''

``Yes, the appendix. I saw it.'' Tarnock flipped to the back of the manuscript. ``You dedicate
four\ldots\ five pages of the paper to it and yet it is inadequate. How can you reasonably
justify these alien techniques in such a short space?'' He held up the paper to Colty as if to
show him how little was there. ``It lacks rigor.''

``It's a problem,'' said Molan. ``I'm all for what you're trying to do but you can't ask the
reader to digest so much new math just for the sake of understanding your proof. I agree with
Tarnock, you're going to have a lot of trouble getting this published.''

``Colty, is there a way to recast your argument in pure dunari style?'' Dargon asked.

``I've tried. The problem is\ldots'' Colty hesitated. ``The problem is dunari mathematics isn't
powerful enough.''

Silence.

``Are you saying dunari mathematics is incapable of proving this result?'' Molan asked.

``I\ldots\ I don't know. All I know is that I can't prove the result using conventional
techniques, at least not rigorously.''

``Maybe you haven't tried hard enough,'' Tarnock said cooly.

Molan fluttered slightly and glanced at Tarnock. He sat up. ``Colty, I think you're probably the
only dunari in the world who understands this work, but I also think what you're trying to do is
incredibly important.'' He turned to either side to address Dargon and Tarnock. ``We need look
beyond this one paper. Colty is integrating human mathematics into our mathematical culture. The
long term benefits could be profound. It is a very appropriate thing for a professor of Parnon
to be doing. I support him moving to publication.''

``I agree,'' said Dargon, ``I also support publication.''

Tarnock swayed his head. ``It's a violation of our long established academic process,'' he said.
``It's well known that Colty is host to our human visiting professor. How can the reviewers know
the creature didn't just dictate the result to Colty?'' Tarnock raised his trunk as if to hold
back Colty's objection. ``I'm not saying that's what happened, but the review process can't
evaluate this work.'' He fluttered slightly, ``yet if the other overseers feel moving to
publication is appropriate, I won't block it.''

Colty clicked his teeth softly. ``Thank you, all.''

``And good luck, Colty,'' Tarnock added. ``I mean that.''

