
% SUMMARY: Introduce Colty Harnick as a student.

It was the first session of the new year. Students started to gather in the small conference
room at the northwest corner of Denkin Hall. Several seats surrounded three sides of a long, oval
table and at the far end of the room, where the sun illuminated it, was a wide chalkboard.
Despite being just a conference room it was plenty large enough; there were only a few students
in the Advanced Mathematical Topics seminar. The course was reserved for Zarlock College math
majors who were in the final stages of their dictatorium.

Colty Harnick lay quietly by one of the tables as his friend Marnan Fark came in. ``Colty!'' he
said, ``Surprised you made it this early after last winter.''

Colty swayed his head. ``Yeah, well, I stayed up until hour 115 thinking about that proof. I
couldn't get it.''

Marn shook his ears. ``Good thing,'' he said as he slapped his friend's shoulder playfully with
his trunk. ``The conjecture's false. There's a counter example.''

``What?''

``Oh yeah. Read the paper by Rigona and Laracki in the most recent edition of AJAM. It's all
there.''

Colty clicked his teeth with annoyance. ``And you forgot to mention that little detail last
year?''

``I can't help it if you don't keep up with the literature!'' Marn patted Colty on the back
affectionately.

The professor chose that moment to arrive and the group fell silent as he took his position in
front of the chalkboard. He wore the traditional academic shroud. It was black and trimmed with
Zarlock's school colors of yellow and gold. On the left side of the shroud, just behind and
below his ear, was the emblem of Menola, an ancient symbol of mathematical purity and
perfection.

``Since when has Laracki been publishing in combinatorics?'' Colty whispered to Marn. His friend
ignored him.

``Before we get started,'' the professor said in loud tones. ``I want to remind you that the
dictatorium exams will be given in two years. I know a few of you are planning to take them. If
you want to graduate at the end of next term you need to take these exams now.''

The professor looked around the room briefly. ``Bolar, I think you are on the schedule\ldots\
Colty, you too, yes?'' The two swayed their heads. ``How about you Marn?''

``I've got another two terms to go, Sir.''

``You could take the exam now and get it out of the way.''

Colty clicked his teeth quietly. He glanced at Marn.

``Think about it,'' the professor said. He paused briefly and then asked, ``does anyone have
anything interesting to talk about?'' He paused again. ``Bolar? Narlip? Maybe a research
problem? Something you read?''

Marn nodded toward Colty. ``Harnick spent several hours last winter trying to prove an
interesting conjecture.''

``Oh?''

``It's nothing,'' Colty replied. ``Just an annoying distraction.'' He glared briefly at Marn.

``What conjecture is it? Even failed attempts can be enlightening.''

Colty hesitated but Marn didn't. ``It's about multi-partite graphs.''

``That does sound interesting,'' the professor said. ``Colty, can you show us your thoughts?''

The seminar course was not like the lower level courses. It was designed to give the students
practice presenting and defending results, communicating with their peers, and thinking about
the advanced aspects of the entire field. Class participation was a must. Reluctantly Colty
stepped up to the board, with almost no preparation, to go over the various lines of reasoning
he had unsuccessfully tried in his late-year efforts the winter before.

At first he was hesitant and unsure, but as he spoke Colty found his confidence growing. The
result might be false, but there were some interesting aspects to the problem. He had enjoyed
thinking about it.

``Couldn't you use Ticzar's Theorem there?'' asked Narlip.

``Where? Here?'' Colty pointed his trunk at the figure he had just drawn.

``No. In the iteration step right above that.''

``Up here?'' Colty studied the notation. He tried to recall Ticzar. ``Ah\ldots\ I'm not sure.''

``I think so,'' said Bolar. ``Then you don't get the problem in the later reduction step that
you have now. It looks like Ticzar would allow the proof to go through.''

Marn looked displeased.

``I don't really remember Ticzar right now,'' Colty said feeling embarrassed. He had some
studying to do before his exams.

``That's very interesting,'' the professor said as he reviewed the board. ``This looks like the
sort of thing Laracki is working on now. Have you done a literature search?''

Colty glanced at Marn. ``I just found out that Laracki has a paper related to this in the most
recent AJAM. I haven't read it yet.''

``That's excellent,'' said the professor. ``We should all read that paper before next session
and talk about this some more then.''

After Colty lay back down Narlip stood up to talk about his dictatorium project and a few lemmas
he had proved related to it. Colty tried to focus but Narlip's work in divisibility systems was
just not something that Colty knew much about.

As the class was breaking up Marn caught Colty in the hallway. ``Nice proof,'' he said, ``very
clever.''

``There has got to be a flaw,'' Colty replied.

``Maybe so, but the important question is: are you seeing Jingas again?'' Marn clicked his
teeth.

Colty clicked his teeth in reply. ``Actually yes. Amazingly after that disastrous first meet she
agreed to meet again.''

``It wasn't that bad,'' Marn said. ``You only kicked her twice!''
