
There was a knock on Colty's office door. He usually left it ajar when he was there so people
knew he was available. Colty had just gotten in only a few arnets before, returning to the
university after spending almost an hour at home.

``Yes?'' he said. It was Puntara, a specialist in abstract algebra and the closest thing Colty
had to a friend in the math department.

``Your human is apparently over in Physical Sciences checking out its office.''

Colty sighed. He had hoped to get some work done before his Combinatorics class, but he knew he
really should make sure his `guest' was settling in comfortably. ``Thanks Punt,'' he said.

As Colty walked across campus he wondered again why she wasn't being hosted by someone in the
physics department. It was a bit of a walk from the Yellow to the South Campus and it seemed
silly to have her office so far away from his. The Nermela was blowing stiffly that day and the
trees whistled in the wind. Colty wondered if maybe on a nicer day it wouldn't seem so far.

He wasn't familiar with Physical Sciences and it took him a little time to find the stairwell to
the second floor. He stopped at the physics department office and they didn't even ask him what
he wanted. ``It's down the hall that way,'' the receptionist pointed, ``at the end.''

Her office door was closed. It still said ``Conference 3'' on the plaque next to the door. Colty
hesitated. He really didn't want to do it but he knocked on the door anyway. When it opened he
found himself face to face with Dr. Spencer. She seemed a little taken aback by him as well.

``Yes?'' she asked.

Colty got the impression that she didn't recognize him. ``I'm Colty Harnick,'' he said. ``I just
wanted to see how you were doing.''

``Oh, of course,'' she said. ``Come in.''

Colty was shocked by how cold her office was. Instinctively he pressed his ears against his
head. ``You do like it chilly, don't you?'' he said without thinking. It was a rude comment and
he regretted it as soon as he said it.

``I will turn this off,'' she said. She walked over to the air cooling unit and touched a panel
on it. Colty could see that it was of human design; it looked completely out of place in her
Rujaran office. The unit had been humming slightly but in response to her touch it fell silent.

``No, no,'' Colty said. ``Please don't mind my inappropriate comments. You should make yourself
comfortable.''

``It is fine,'' she replied. ``I was feeling chilly also.'' She turned toward him. She was
wearing different clothing than when he first saw her---a surprisingly colorful combination of
tan and yellow with rubinum trim. Around her neck was a fine chain of silver colored metal.
Colty wondered if it was some sort of technology.

She had a standard Parnon University desk, although Colty wasn't sure how she could use it, some
bookshelves, and a file cabinet. The bookshelves were empty except for a clock that was placed
right in the middle of the top self. Dr. Spencer noticed him looking at it.

``My clock,'' she said. She walked over to it and extended one of her front legs. Colty watched
as she curled all of the toes on that foot except for one which she used as a kind of pointer.
She touched the face of the clock gently. ``Hour 61 and\ldots 82 arnets,'' she said. She seemed
pleased and smiled. Colty closed his eyes and took a deep breath. She was so very alien.

``I have a question,'' she said. Colty opened his eyes again. She walked over to her desk and
touched the painted concrete wall over it. ``Is there way to\ldots connect paper to this wall?''

``You could use tape. That's what I do,'' Colty replied.

``How do I get tape?'' She asked.

``There might be some in one of the desk drawers,'' Colty said. He reached out with his trunk
and grabbed hold of the one of the drawer handles to open it. Dr. Spencer stepped quickly
backward and bumped into the file cabinet making a loud banging sound in the process.

Colty froze and saw her looking at him intently. He suddenly realized that he had startled her.
He pulled his trunk back. ``I'm very sorry,'' he said. ``Are you okay?''

She nodded her head and said, ``Yes, yes, I am fine.''

``I only wanted to check the drawer for tape.''

She took a deep breath. ``I know,'' she said. ``I'm just a little nervous.''

Colty clicked his teeth. ``It's okay,'' he said.

The drawers of her desk were quite empty so Colty went down to the physics department office to
get some tape for her. It was nice to get out into the warm hall again. When he returned she had
a box on the desk and was holding a picture up against the wall with her toes.

It was a picture that was both familiar and yet also alien. In the center was a large mountain
capped with what Colty assumed was snow. It looked rocky and barren. Yet in the foreground was a
large field of what, to Colty, looked like flowers. The plants were a strange green color but
the blossoms themselves were yellow, gold, and red.

``The mountain is called \alien{Rainier},'' she said. She looked at the picture and smiled.
``Isn't it nice?''

``Where is that?'' Colty asked.

``It's on Earth,'' she replied. She looked at the wall and then at Colty. ``Maybe I shouldn't
put it up.''

``It is your office,'' Colty said. ``There is nothing wrong with making it feel like home.''

Dr. Spencer put the picture back into the box. ``Maybe later,'' she decided.

Colty explained that he needed to get back to work. Dr. Spencer was very understanding. ``Yes,
yes,'' she said. ``Don't worry about me.'' Then, just as Colty was about to leave she stopped
him with one last question.

``Is there a way for me to read about your mathematics?''

``There are many books in the library,'' Colty replied. ``Are you able to read Argenian?''

She nodded her head and then went to her box again. ``Yes, a little. I'm still learning.'' She
pulled out a map of the campus. It was the standard map given to all new students. ``Where is
library?'' she asked. She held out the map.

Colty hesitated and then reached up to take the map from her. Colty looked at it. ``The library
is the building marked with `J7','' he said.

``Okay,'' she replied. ``Thank you for your help.''

