
Colty waited on the tarmac outside the main terminal building of the Parnon air port. The
landing of a human ship had become fairly routine and so there was little fan fare. Security
kept the public away from this landing especially resulting in even fewer people on hand than
usual.

Lucasa Formand stood beside Colty as they watched the streak approach. ``I'm surprised she's
coming back,'' he said in passing.

``I'm not,'' said Colty.

``She should have at least waited until the matter of a bodyguard is settled,'' Lucasa said.

Colty recalled the first time he met Rachel. He had stood on that very same tarmac and watched
a very similar streak. Yet then he was angry and afraid. Today all he felt was happy.

The vehicle landed uneventfully and taxied to a stop not far from where they stood. A single
figure descended from it and started walking across the open space toward the terminal. It
wasn't an alien monster from beyond the stars. It was Colty's friend and he was glad to see her.

``Colty!'' Rachel called out as she approached. She held out her arms.

Colty walked forward to meet her. He grasped her arm with his trunk and she held him briefly.
``It's good to see you, Rachel,'' he said. ``You are looking well.''

``You too,'' she said as Lucas joined them. She pushed up her sleeve and showed Colty the
bandage there. ``I'm still healing but this should be gone soon enough.''

``Welcome back,'' Lucasa said to her. ``We do have some extra security in place, but I'm afraid
we don't have a bodyguard for you yet.''

She turned to him. ``Thank you. I appreciate that. I'm sure it will be fine,'' she replied
casually. Colty hoped her confidence in the Parnon police force was well placed.

They started walking to the van that would take Rachel to the human compound. ``I get the
impression that some of your people are worried about this too,'' Lucasa said.

Rachel shrugged. ``I think Benedito has the same concerns as you do,'' she said.
