
``Hey, Den, what classes are you taking?''

Denart squinted at his friend with irritation. It was a beautiful day and he was enjoying the
sunshine. Quite a few people were about, going back and forth along the walkways. Some distance
in front of them, standing majestically at the end of the Yellow was Tartuk Hall. Denart
stretched slightly on the seat where he had been basking and then he looked again at Mumalik.

``I don't remember,'' he said. ``Does it matter?''

``Only that they might run out of books at the bookstore,'' Mumalik said. ``That's what I
hear.''

``Good,'' replied Denart. ``Then I can justify not buying them.''

``Do you have your schedule?''

Denart sighed and sat up. He reached over to his pack and rummaged through it a little. He
didn't have much in there but he made a show of the effort anyway just to send a signal to
Mumalik. Eventually he handed him a crumpled piece of paper. Mumalik lay on the grass next to
him and studied it.

``\textit{Zecks},'' he said, ``I think I'm in the same section of Solid Dynamics.'' Denart said
nothing but instead closed his eyes again. ``What's this?'' Mumalik asked. ``What's
Combinatorics?''

``It's some math course,'' Denart replied.

``Math?'' said Mumalik, ``Why are you taking that? I thought you already had all the math you
need.''

Denart opened his eyes again and took the paper from Mumalik's trunk. He was starting to feel
even more irritated. ``I like math, Ok?'' he said. He shoved the paper back into his pack and
closed his eyes again. He hoped Mumalik would go away. His friend was a mechanical engineering
student and tended to take many of the same classes as Denart. They had known each other since
they were first termers taking basic physics together. He was a nice guy even if a little clingy
at times.

``I'm sure you'll slam it,'' Mumalik said. ``You always do.''

The bell of Tartuk Hall sounded the hour and Denart looked at his watch. ``Time to go,'' he
said.

``Where to?'' asked Mumalik getting up to follow him.

``The bookstore.''

Mumalik shook his ears. ``I'm telling you,'' he said, ``they'll be sold out.''

``Let's hope,'' replied Denart.
