
% SUMMARY: Denart consults Rachel about her class.

Reluctantly Denart approached the door to Rachel's office. He checked his watch hoping that it
wasn't yet time but unfortunately the appointment was upon him. Unlike most professors Rachel
kept her door closed. Denart took a breath and then tapped the door lightly with his knee.

After a moment it opened and he was face to face with the alien. ``Yes? Oh, yes,'' Rachel said,
``please come in.''

Denart gasped softly at the cold temperature and Rachel turned off the air cooling unit.

``You like it cold, don't you?'' Denart said undiplomatically.

``It's more comfortable for me,'' Rachel said. ``My body produces heat. That's why I need to eat
so much more than you do.''

Denart clicked his teeth. ``So you're an endotherm,'' he said.

``I'm not sure what that word means.''

``In the shallow night there are animals that heat their bodies with their metabolic activity,''
Denart explained. It felt strange to be teaching something to the human. ``It keeps them from
getting lethargic.''

``Yes,'' said Rachel, ``animals like that are common on Earth\ldots\ and other places.''

Rachel kneeled on the seat in front of her desk and Denart lay on the seat beside her desk. He
had never been so close to a human before and he found himself studying her skin. She didn't
have any scales at all, not even vestigial scales. Nothing was said for several moments and
suddenly Denart relized that Rachel was looking at him. He started.

``I'm sorry,'' he said, ``I\ldots''

``It's fine,'' said Rachel. ``I don't mind and I know you are a biologist.'' She stretched out
her arm and put her hand, palm up, on the desk. She curled her fingers slightly. ``There are
bones in my fingers,'' she said, ``but they are still flexible because of all the joints.''

``That must be nart if you get joint-lock,'' Denart said without thinking. Rachel didn't seem to
care.

``We do get diseases like that sometimes and, yes, it isn't good. Thankfully we have
treatments.''

Denart's grandmother had joint-lock in her back legs. She had to drag them behind her and
eventually she got a wheeled cart to make it easier.

``I feel like the class isn't going well, Denart,'' she was saying. ``Are the students
understanding it?''

Denart shook himself slightly to get his mind back on track. ``I only had about ten come to my
last session. They seemed to be doing fine.''

Rachel studying him closely and it made Denart uncomfortable. ``It's the ones who don't go to
the sessions I'm worried about,'' she said. ``I try to follow the course book, It seems very
well organized.''

``I'm sure it is.''

``What do you mean?'' Rachel asked.

``Well, I just meant that the course book has been used for so many years that all the errors
have long since been worked out of it.''

Rachel looked over at the book sitting on the corner of her desk. ``Maybe everyone is tired of
the book.''

``Do you use course books on Earth?'' Denart asked.

``Not really, no,'' said Rachel.

``How do you teach your classes?''

``We do it over\ldots'' she made a gesture with her hands, ``\ldots\ over the screens. Sometimes
just recorded.''

``But how do you know what to say?''

Rachel thought for a moment. ``I just know,'' she said finally. ``I just talk like I'm talking
now.''

``Why not do that here?'' Denart asked, ``just do what comes naturally to you.''

She ran her fingers through her hair. ``Won't the department be angry if I don't follow the
book?''

``Who cares?''

