
``Did you get the homework done?''

Denart squinted at his friend with irritation. It was a beautiful day and he was enjoying the
sunshine. He looked around briefly. Quite a few people were about, going back and forth along
the walkways. Some distance in front of them, standing majestically at the end of the Yellow was
Tartuk Hall. Denart stretched slightly on the seat where he had been basking and then he looked
again at Mumalik.

``I did it, yes,'' he said.

``\textit{Zecks},'' Mumalik said. ``I'll believe it when I see it.''

``Why does it matter?'' Denart asked closing his eyes again.

``Because I have a bet on it.''

``A bet?''

``Do you have it with you?''

Denart sighed and sat up. He reached over to his pack and rummaged through it a little. He
didn't have much in there but he made a show of the effort anyway just to send a signal to
Mumalik. Eventually he handed him a crumpled piece of paper. Mumalik lay on the grass next to
him and studied it.

``\textit{Zecks},'' he said, ``I think I'm going to have to pay up.'' Denart said nothing but
instead closed his eyes again.

Mumalik was silent for a while, which made Denart happy. ``How did you get this answer for
number three?''

Denart opened his eyes again and snatched the paper from Mumalik's trunk. ``I guessed, Ok?'' he
said. He shoved the paper back into his pack and closed his eyes again. He hoped Mumalik would
go away.

``I'm sure you slammed it,'' Mumalik said.

The bell of Tartuk Hall sounded the hour and Denart looked at his watch. ``Time to go,'' he
said.

``Where to?'' asked Mumalik getting up to follow him.

``The bookstore.''

Mumalik shook his ears. ``You're just buying your books now? The term's already started. I'm
sure they'll be sold out.''

``Let's hope,'' replied Denart.

