
\chapter{Desert}

\textit{Your people are coming!}

\breakpoint

The two of them looked east across the barren landscape, back toward the remote Argenian
heartland. The hot Nermela wind blew against their backs. The road snaked away toward a distant
range of hills. All was sand and rocks and a few scattered clumps of brickken grass. The heat of
the Rujaran sun bore down on them. Colty looked up. The sun was higher than he had ever seen
it; closer to zenith than to the horizon. He felt very far from home as he started to realize
the seriousness of their situation.

``We need to find some shade,'' he said.

``And water,'' Rachel added. ``I will need water.'' She stared into the distance with a hopeless
look on her face.

``I think our best plan is to try to return to their supply depot,'' Colty said pointing his
trunk down the road. ``There may be water there, or a radio. At least there will be shelter from
the sun.''

``It's so far\ldots'' Rachel said softly.

``I don't think too far,'' Colty said hopefully. ``I think maybe three or four karnons.'' He
turned toward her and saw that she was sweating again. Beads of liquid were on her face. ``Can
you make it that far?'' he asked.

Rachel didn't reply at first. She just gazed into the distance. Then she turned to Colty.
``Normally, yes, but it is so hot.''

Colty touched her arm with his trunk and faced her squarely. ``You have deserts on Earth,
right?'' he said.

Rachel took a breath and looked at her friend. ``Yes\ldots but not like this.'' She squinted at
the sun briefly. ``This is not my world, Colty,'' she said, ``my body is not evolved for it.''

``It's only two or three karnons,'' Colty said. ``That's just a few laps around the purple at
Parnon. Then we will look for water.''

Rachel closed her eyes briefly and Colty could almost see her resolve build. ``Right,'' she said
finally. ``There is nothing for us here. Let's go.''

They began trudging down the road. The sand was soft and it made walking difficult. They kept to
the tracks of the sand car as much as they could. There, at least, the sand was packed a little.
The sun beat down on them as they walked. Its bright, red light flooded their eyes. The Nermela,
hot and dry, blew strongly from behind.

Colty felt the heat. He wasn't used to this sort of thing himself. He was an academic who spent
most of his time standing in front of a blackboard or laying in his office. But he barely
noticed his discomfort. Instead he was worried about Rachel.

She walked silently beside him, sweating profusely. Colty could see the water on her skin, even
dripping from her face. Her hair was matted to her forehead and neck. Colty was alarmed. He had
never seen her that way before and it reminded him of just how alien she was. ``You're losing a
lot of water,'' he said. ``Can you\ldots control it at all?''

``Not at all,'' she snapped.

They walked in silence for several arnets. Colty looked at the sky. ``Your people will be
looking for us,'' he said. ``Surely your technology will help them find us quickly.''

Rachel said nothing.

Colty eyed a rock outcrop in the distance. He thought the depot was just behind it. Yet it
approached maddeningly slowly. Rachel was walking much more slowly than usual and she often
stumbled. Colty walked close to her, instinctively huddling with her, and she often reached out
and touched his back with her hand to steady herself. The soft sand made bipedal walking
especially difficult.

After a few more arnets Rachel stopped. ``I have to rest,'' she said. ``I have to\ldots rest.''
She collapsed on the ground, crossing her legs beneath her and sitting where she had stood. She
was panting. ``It's so\ldots hot,'' she said, wheezing slightly.

Colty stood next to her, shading her as much as he could with his body. Yet with the sun so high
his shadow was very short.

``It's way above my body temperature,'' Rachel said. ``I feel like I'm in an oven.''

``Yes it's very hot,'' Colty replied.

Rachel looked up at him and touched the side of his body with her hand. ``How are you?'' she
asked. ``Are you Okay?''

``For now,'' Colty said. ``I might not need water but if my body temperature gets too
high\ldots''

Rachel nodded.

Colty kneeled down slightly and pointed with his trunk. ``But you see\ldots we are more than
half way there, I think.''

Rachel smiled weakly. ``You lie,'' she said, ``but thanks.''

Colty stood up again and looked back to see how far they had come. His heart sank. In the
distance he saw a large yellow cloud, close to the ground. Even as he watched he could see it
moving and billowing. It was coming toward them.

``We need to get going,'' he said. He grabbed Rachel's arm with his trunk and encouraged her to
stand.

``What is it?'' she asked.

``Sandstorm,'' Colty replied. ``It's coming this way.''

They tried walking more quickly now but Rachel didn't have it in her. She labored with every
step, and her breath was like a panting gasp. Colty looked over his shoulder often with rising
fear. Sandstorms on the open desert were extremely dangerous even for a seasoned desert-wise
Rujaran. If they got caught in it they wouldn't be able to see their destination and could
easily wander far off track. In addition the blasting sand could burn the skin off even a
Rujaran body. Colty didn't dare to think what it would do to Rachel's human skin.

The Nermela rose to a strong wind, blowing fiercely. Colty could see the sand starting to move
on the ground around them. The dust was thick in the air. Rachel stumbled and fell and Colty
helped her up. In a few steps she fell again. Colty put his trunk around her waist to help her
and she struggled to stand.

``I can't\ldots'' she said. ``I just can't.''

Colty kneeled down closely beside her, bending all four legs. ``Can you climb on my back?'' he
said. ``I've seen chortak do this sort of thing. Can you do it?'' Colty could feel her small
hands gripping his skin as she pulled herself on top of him. He helped her with his trunk. She
put one leg on either side of his body and layed, face down, against his back. Colty put her
arms around his neck and then wrapped his trunk over her shoulders to help balance her. ``You
okay?'' Colty said finally. Rachel didn't reply. Colty could just hear her ragged breathing in
his ear.

When he stood up Colty was surprised at how heavy the human was. Having her on his back
certainly made walking more difficult. Yet he made better time than they had together. The rock
outcropping was close now.

Yet the sand storm was closer still. The wind rose to a gale and the air darkened. The sun was
visible as a faint red disk high in the sky. Colty could feel the sand stinging into his skin
yet he knew the worst was yet to come. They needed to find shelter right away.

With the Nermela howling around them Colty finally got to the rock outcrop. He intended to find
a place where they could wait out the storm on the leeward side of the rocks but as they turned
the corner he could see the depot just a short distance ahead. It was a small cluster of
temporary metal buildings huddled in the desert.

Colty decided to walk the last short distance to the depot through the raging storm. The sun was
not visible at all now but the sand was biting hard. The buildings ahead were little more than
dark rectangular shapes in the thickening air.

Colty tried the main building but the door was locked. He then went around back to the garage
where the sand car had been. He struggled with the door while keeping Rachel balanced on his
back. Luckily it had been left unlocked and he was able to open it. Inside it was quite hot but
at least there was no sand. The door was on the leeward side of the building so no sand blew
into the garage. Colty just left the door open as he staggered inside.

On the floor was a large oily blanket that had no doubt been used during sand car maintenance.
Colty folded it roughly in half and knelt down again next to it. ``Rachel?'' he said. ``Rachel
are you awake?''

Rachel did not reply. He could no longer hear her breathing. Colty had little choice but to
rather unceremoniously dump her onto the cloth. When he turned to look at her he was aghast.

``Jurita,'' he said softly. Rachel's clothes were drenched in sweat and covered with caked sand
and grit. She also had a large sand burn on her arm running from her wrist up past her elbow.
The sand had abraded away the skin leaving a raw tissue that oozed a thick red liquid that Colty
knew wasn't sweat. The side of her face was burned as well, but not as badly. There the skin
looked red and finely scratched but not as raw as on her arm.

``Rachel? Rachel?'' he said. He positioned her gently so that she would be comfortable.
Afterward he sat back and looked at her, wondering if she was still alive. He shifted a little
and winced. Touching the side of his face he discovered that he'd been sand burned as well. His
Rujaran skin was not as damaged as Rachel's but he knew it would still sting for a while.

Rachel moaned. Colty sat up again. ``Rachel?'' he said. Her eyes fluttered open and she turned
and looked at him.

``Where are we?'' she asked.

``We made it back to the depot,'' Colty said.

Rachel touched her burned arm and then hissed. It was a strange, alien sound but Colty knew
right away it meant pain. She took her hand away and looked at the red liquid on her fingers and
then back at her arm. ``Oh \alien{God},'' she said. ``What happened?''

``It's a sand burn,'' Colty said. ``It looks pretty bad.'' Rachel closed her eyes and started to
shake a little. Colty wasn't sure what it meant but he didn't think it was good.

``I'm going to try to find some water,'' he said, ``and maybe some first aid supplies.''

``You're going out in\ldots that?'' Rachel said pointing weakly out the garage door and into the
blowing storm.

``It's okay,'' Colty reassured her, ``the main building is right there. I won't be in the wind
for long.''

[Do we follow Colty as he explores the depot? How does he get into the locked building?]

When Colty returned he had a can of rather dirty, oily water. There was a layer of sludge on the
bottom. ``This is all I could find,'' he said. ``It's not very clean. I'm not sure it will
help.''

Rachel reached out for the can anxiously but when she saw the water inside she was not happy. It
stank of motor oil and was dark and cloudy. ``I can't drink this,'' she snapped. ``What were you
thinking?'' Rachel put the can on the floor and turned away from it.

Colty layed down but didn't say a word. They listened to the sand blowing against the back of
the garage.

``I'm sorry,'' Rachel said.

``It's okay,'' Colty replied.

``No,'' Rachel said. ``It's not. I shouldn't have snapped at you. I'm sorry.''

Colty slid closer to her and touched her good arm with his trunk. ``Don't worry about it,'' he
said. ``It's not important right now.''

Colty changed the subject. ``Will your people be able to find us out here?'' he asked.

Rachel shrugged. ``I'm not sure,'' she said. ``It's a big world. We do have equipment that could
read a car tag from space, but it won't do any good unless the sky is clear and we are in the
open.''

``Maybe when the storm stops we can find a shady spot outside to wait and hope for the best,''
Colty said. ``We obviously aren't going to walk back to Argenia.''

There wasn't much for them to do so they simply waited. Colty wondered what time it was but his
watch was long gone. He watched Rachel closely. Her arm stopped oozing red but she started to
sweat again. They were out of the sun but the garage was still very hot. Rachel just layed
quietly as if she no longer cared.

``Tell me about Earth,'' Colty said at one point.

Rachel didn't answer right away. ``There are large oceans of water on Earth,'' she said finally.
``Three quarters of our world is covered by them, and they are several karnons deep in places.''

Colty didn't reply but he listened attentively.

``When I was a girl my parents liked to go to the beach,'' she said. She was looking upward,
lost in her memory. ``We spent many hours there over the summer. I loved to swim. There were
these\ldots'' she looked at Colty, ``we call them \alien{horseshoe crabs}. They are really
ugly.'' She laughed a little. ``I liked to watch them crawl around on the sand under the
water.''

``I'm sure you were very cute as a child,'' Colty said.

Rachel chuckled. ``The thing about \alien{horseshoe crabs},'' she continued, ``their species
hasn't changed for, well, billions of your years. They are one of the oldest life forms on
Earth. When my people go extinct there will still be \alien{horseshoe crabs} in Earth's
oceans. I'm glad of that.''

She stopped and the two of them sat silently for a time.

``I wish I could go swimming now,'' Rachel said softly. ``That would sure feel nice.''

\breakpoint

Colty woke up suddenly. He wasn't sure how long he had been asleep. He quickly looked at Rachel.
She seemed to be sleeping too. He studied her closely. Was she asleep or was she unconscious? It
was hard to be sure. He had never seen a sleeping human before but there was something about the
way she was breathing that reassured him. Yet her skin was obviously flushed and she was still
sweating.

It was only after he had checked on Rachel did Colty notice that the storm had abated. He could
no longer hear the sand pelting against the outside wall. When he looked out the open door of
the garage the air was only a little dusty.

Colty stood up and walked to the doorway. The Nermela blew steadily but gently. The sand had
drifted up against the other building and there were no sign of his tracks. Colty realized that
he needed to do something about that. Rachel suggested that her people would be able to find
them from space if they could be seen. Colty didn't want to take Rachel out into the sun again
but he could at least create some kind of sign. Taking a deep breath he headed out into the day.

By the time Colty returned the storm was over and the sand had settled down again into its new
position. The sun was bright in the sky and Colty felt hot. He needed water too. Rachel was
awake and even sitting up when he got back. She was very pleased to see him.

``How is it out there?'' she asked.

``Hot,'' Colty replied as he layed back down next to her. ``I tried to set up something your
people might be able to see. I think that's our only hope.'' Rachel nodded.

They sat silently for a while. Rachel sighed and then reached for the can of oily water Colty
had brought earlier. ``Can you get that for me?'' she asked.

``Are you sure?'' Colty replied as he handed the can to her. Rachel looked into the can and
grimaced. ``Let me find something to hold a little water. We can skim the oil off the top.''

Colty rummaged through some draws and storage areas in the garage. Finally he found a small
metal pan that looked reasonably clean. He took the can of dirty water from Rachel and carefully
poured off the oily layer. He was only partially successful but after a couple of tries he got
most of it. Then, he slowly poured some of the `clean' water into the pan and held it out to
Rachel.

``You are good to me,'' she said. ``Thank you.'' Rachel looked at the water in the pan. It was
cloudy with brown particles. She wrinkled her nose. ``It stinks,'' she said.

``Rachel\ldots'' Colty started.

With one hand she held her nose and with the other she raised the pan to her lips and drank a
little. Colty was fascinated by the way she could do something different with each hand. She
made a face of disgust and then drank some more.

``Ughh,'' she said. Colty understood it wasn't a word. ``That is horrible.''

``Hopefully it will help,'' Colty said.

Rachel put the pan aside. She had finished about half of it. She put one hand over her mouth.
She looked pained. ``I hope so too,'' she said as she layed back down.

``So now we wait,'' Colty said.

``Yes, I guess so,'' Rachel replied. She put her hand on her stomach and looked uncomfortable.

``Are you all right?'' Colty said with concern.

``No,'' Rachel replied. She sat back up with her hand on her stomach. ``I'm not.'' Suddenly she
turned away from Colty and made a horrifying sound that alarmed Colty greatly. He could hear
liquid splashing on the ground. He quickly moved over to her and saw a wet stop on the garage
floor beside the blanket. Rachel was panting slightly. She wiped her mouth.

``Jurita,'' Colty said softly. ``Rachel, what is wrong?''

``It's okay,'' she said. ``My body just didn't like that water very much.''

``Is it normal for you to do that?'' he asked looking at where she had vomited.

``Not normal,'' she said, ``but under the circumstances not surprising. It's a reflex that
protects me from ingesting bad things.''

Colty felt terrible for her. Their situation was serious but for him it was a seriousness that
was still a bit removed. He could last probably several hours, if necessary despite the heat and
lack of water. That would give the searchers time to discover them. Yet Colty was starting to
realize that his human friend might not make it that long.

``You've lost more water,'' Colty observed.

``I know that!'' Rachel said with anger and frustration. Colty moved away from her and layed
back down. The two of them just looked out the garage door into the hot sunshine. There wasn't
much to say or do. They simply waited. After a while Rachel layed back down. She stared at the
ceiling of the garage silently. How much time passed Colty had no way of knowing. It felt like
an hour although it was probably only a few arnets. Rachel was sweating heavily again and
panting slightly.

``I want to go home,'' she said softly.

Colty turned to look at her but she just stared at the ceiling.

``I just want to go home\ldots'' she repeated. Then she made a kind of soft wailing noise and
her face wrinkled. She turned away from Colty and he could hear her making strange choking
sounds. It was obvious to him that she was upset. He slid over to her again, wanting to comfort
her. He reached out with his trunk and caressed her shoulder.

``Don't touch me!'' she yelled. Colty pulled his trunk away immediately. ``Please don't touch
me,'' she repeated softly between cries. Colty watched her but left her alone. After a while her
sobbing stopped and she layed quietly, still facing away from him. Colty got up and stood at the
opening of the garage. He looked anxiously toward the sky. He turned and looked at Rachel's
form, laying motionless on the cloth. Then he went back into the sun to check on his sign.

When Colty returned his eye was drawn to the ceiling of the garage. Heated by the glaring sun,
it was glowing softly in rubinum. As Colty layed down, Rachel stirred and then sat up. She
looked at him and said, ``I'm\ldots late for class.''

Colty noticed that she wasn't sweating anymore. That seemed like a good sign. He shook his ears
slightly, ``Yes, well, we both are.'' he replied.

Rachel started to get up. ``I've got to get my spin dial,'' she said. ``I think I left it in my
office.''

Colty was alarmed. He stood up and held her arm with his trunk. ``Where are you going?'' he
asked. She looked at him and then looked around as if seeing the garage for the first time, and
realization crept over her face. She slumped back down onto the cloth.

They talked for a while but it was obvious that she was having trouble. Often her conversation
made no sense and she seemed to think they were somewhere else. She couldn't remember things he
had just said and the accuracy of her speech started to break down. It was a terrible thing to
behold and Colty could hardly contain his emotion. He held his ears against his head and
shuddered with anguish. His friend was becoming delirious.

``I like how birds fly,'' she said at one point. She lifted her hands and waved them like the
wings of a bird. ``This just flap, flap and away they go.''

She slept now and then, but only for a couple of arnets at a time. Colty knew it wasn't normal.
After one of those short sleeps she woke up and stared at him. Colty could see fear in her eyes.
She shouted at him in her native language, an alien collection of chittering clicks.

``I can't understand,'' Colty said. ``You have to speak Argenian.''

Rachel backed away and started to get up, shouting some more. She pointed frantically out the
garage door, talking rapidly in her language. Colty grabbed her with his trunk and she struggled
to get away from him, hitting him weakly with her injured arm. ``Rachel! Rachel!'' Colty
pleaded. ``It's okay. You're safe. It's okay.''

He was much too strong for her and she feel to floor, defeated. Colty knelt down next to her and
she looked at him. ``I'm going to die, Colty,'' she said clearly. ``I just want you to know that
I consider you my friend.''

He wrapped his trunk around her waist and she leaned against him. He shuddered repeatedly. ``You
are my friend too, Jayelle'' he said finally.

\breakpoint

Colty stood outside the garage and scanned the yellow skies. The air was clear now and the
Nermela nothing more than a gentle breeze. Inside the garage Rachel layed on the floor. She had
slipped into unconsciousness many arnets ago. Colty didn't know if she was even still alive. He
looked down the road, largely covered now in sand, and wondered if he should attempt to walk to
the previous supply depot. It was many karnons away; he doubted if he could make it. Yet Rachel
hadn't been sure if her people even had the ability to locate them. When it was obvious that she
couldn't go farther waiting for a rescue made sense. But now maybe he needed to take matters
by his own trunk.

Colty looked back into the garage. Rachel was nothing more than a pile in the corner, silent and
unmoving. He shuddered. He couldn't just leave her there, of course. He'd have to be sure first
and then bury her properly. He cursed the people who did this to her. How could they snuff out
such a beautiful light.

He turned to head into the garage when he caught a movement out of the corner of his eye.
He looked back and squinted into the sky. There, far in the distance was a short streak. It was
moving with impossible speed. ``Rachel! Rachel!'' he cried out loudly. ``I think they are
coming!'' He ran out into the sand and waved his trunk despite the fact that the streak was
still many karnons away.

Yet his heart sank as he saw the streak move south. It wasn't heading for them. Frantically he
ran back to the garage. He had to do something to get their attention. Perhaps he could light a
fire. He searched for some oil or gas to light or some wood to burn but there was nothing. He
cursed himself for not thinking of this before.

In frustration he ran back into the sun and searched the sky. The streak was there, moving
toward them now. It had circled around and it was descending. ``Rachel!'' he cried out
excitedly. ``You're people are coming!''

The craft approached quickly and then, suddenly, a powerful booming explosion echoed across the
landscape. No Rujaran aerocar made such a noise. It was only human technology that could. It
was the sound of their salvation.

In moments the craft reached the depot and started circling around it, only slightly above the
rocky cliffs. It was a very small craft; one of their fast moving vehicles. Colty waved his
trunk frantically and pointed into the garage, yet the craft made no indication that it was
going to land.

``Come on!'' Colty yelled in frustration.

The second craft was almost upon him before he noticed it. Suddenly appearing, it was much
larger than the first, even bigger than an ordinary aerocar. Unlike the first craft it
approached the ground as if to land, descending almost vertically. Colty ran out of the way. In
short order the second craft touched down and a hatch opened in the back. Four humans came
running from it, carrying with them various cases and materials.

``This way!'' Colty shouted pointing his trunk into the garage. He lead them into the space and
they rushed over to the corner where Rachel layed. Despite his association with Rachel, Colty
had never seen other humans up close before. He was struck by how different they looked from
each other. Their hair was different color and even their skin was different. One human smiled
at him and nodded as it ran by. Colty recognized the gesture.

They gathered around Rachel and started to work on her. They looked in her eyes and they studied
her injured arm, chattering among themselves in their strange language all the while. They
injected some kind of device in her other arm and appeared to put water directly into her body.
Colty watched anxiously.

``Oh!'' he said getting the attention of the one who had smiled at him. ``She tried drinking
some of this water.'' He pointed to the water, but the human didn't understand. Colty mimed the
action of drinking and pointed at Rachel again. The human nodded. It took a small vial from the
case and put some of the water in the vial, sealing it up.

``\alien{Thank yyyou}'' the human said hesitantly in Argenian.

The humans put Rachel on a stretcher and then started to make their way back to their craft.
``Is she alive?'' Colty asked anxiously. But they did not respond. Colty followed them toward
their craft but then hesitated as they approached the alien ship. The human that had spoken to
him before waved him on.

``\alien{Come},'' it said. Colty was being asked to board a human aerocar and, presumably, go to
the human space craft. That was not something he had counted on. Yet there was really little
choice. There was nothing for him at the depot, karnons from anything. Even if the humans told
Argenian officials were to find him, he'd likely be dead by the time they got there.

The human waved him on again impatiently. Colty took a breath and walked up the ramp into the
craft. Quickly the door closed and sealed. Colty was struck immediately by how dark it was.
There were no windows and the room was lit by strange, yellow green lights. The darkness made
Colty unease and he suddenly felt afraid. This was not a Rujaran aerocar. This was an alien
ship filled with alien creatures using an alien technology. Colty wondered if maybe he should
have stayed behind after all.

In addition to the darkness, Colty also noticed how blessedly cool it was. In fact the air was
almost cold; a welcome contrast after spending so much time in the desert. He was reminded of
the way Rachel kept her office.

The humans were mostly focused on Rachel. They attached her stretcher to the wall of the craft
and were continuing to work on her. One of the humans, with darker skin than the others,
appeared to be the lead. It injected several small vials of various liquids into Rachel's skin
while another human cleaned her arm. They cut her clothes from her body and wrapped some kind of
packs around her instead. Colty found her nakedness disconcerting. He had never seen a human
without clothing. Her alien body did not seem like the body of a friend. It seemed more like the
body of a creature.

Colty felt the craft move. Its engines, powered by some means unknown to him, started to whine
and he felt the vehicle shudder slightly. The humans seemed prepared for it and held on to grips
that were obviously designed for the purpose. Colty tried to follow suit but his trunk was the
wrong shape to make that easy.

Suddenly Colty felt acceleration. It was not like anything he'd experienced before and it took
his breath away. His body was pressed downward and toward the back of the craft as the engines
whined loudly. He held on tightly to prevent being thrown, and he was grateful that there were
no windows. The acceleration lasted quite some time. However eventually the whine of the engines
softened and the forces against his body slackened.

The humans were busy strapping Rachel to the stretcher. Then they grabbed hold of the available
hand grips. The human next to Colty touched him and indicated that he should do the same.
Everyone sat quietly for a while. The humans talked between themselves and Colty felt
uncomfortable. He looked at Rachel who was still quite unconscious. The strange, dim lighting
seemed to bear down on him.

The human next to Colty said something to him but it spoke in the human language so Colty had no
idea what it meant. Suddenly Colty had a very strange feeling of lightness. It was a kind of
floating feeling as if the weight on his feet was getting less. The sensation quickly became
more intense and it felt almost as if Colty's internal organs were going to come out of his
mouth. He gasped and grasped frantically at the grips. The human next to Colty spoke to him to
try and calm him just as Colty felt his feet leave the floor of the craft.

It was a strange sensation to be sure, but once he started to float Colty understood. His
physics training told him what was happening. In an orbital trajectory the centrifugal force
exactly cancels the gravitational force resulting in a weightless condition. He knew that
perfectly well. Yet to actually experience it was another matter.

Breathing heavily Colty clutched at the grips and tried to come to terms with his new reality.
His body floated freely and he had trouble maneuvering. The human next to him helped him get his
legs out of the way and showed him how he could push or pull against the wall to help control
the rest of his body. It was basic physics and Colty got the hang of it quickly enough. He
looked over at Rachel but she was securely strapped to the stretcher.

The whine of the engines was different than before. It wasn't as loud yet it seemed closer and
more intimate. After a short time the engines fell almost silent. It didn't feel like they were
moving at all and Colty started to wonder how long they would be on the craft.

One of the humans reached up and touched a panel over its head. Colty saw a collection of
symbols and graphics appear. The human looked at the information, touched a couple of symbols,
and then spoke to the human at its side. The two of them talked for several moments and Colty
watched them. They were so similar to Rachel and yet different. [Say more?]

After several arnets the humans started to move around. One went over to Rachel and made some
adjustments on her stretcher. The others started to prepare in other ways. ``\alien{Almossst
there},'' one of them said to Colty.

Rachel moaned softly and the human next to her moved up to face her. Colty could see her eyes
flutter open and she looked around a little but she did not see him. The human spoke to her in
their language. Colty could recognize the sound of her name. Rachel didn't speak but she nodded.
She looked dazed and groggy, but Colty was very happy to see her alive.

Only moments later, Colty heard the engines again, softly this time. He could feel the walls
shudder a bit. Then, suddenly there was a thud, a few clicking sounds, and then a beep. The
humans waited impatiently beside what appeared to be a kind of door. Two of them disconnected
Rachel's stretcher from the wall and when the door opened they deftly made their way through it,
even in their weightless condition, with Rachel.

One human helped Colty make his way, considerably less deftly, through the door. He found
himself in a kind of short hallway, dimly lit with the same strange yellow green light used in
the aerocar. There were grips along the walls and Colty made his way down the hall, with help
from the human using the grips. It was clear to him that the ship was not designed for a
creature with one trunk. Everything seemed to work a lot better for creatures with two hands.

There was no sign of Rachel. The other group had apparently already moved beyond Colty's sight.
He felt very alone and also frightened. Rachel, at least, was a familiar face, even if a human
one. But she was gone. Here he was the alien and he felt very alien indeed.

As they approached the end of the hall suddenly a door opened before them. Colty did not see any
handles or controls nor did the human touch anything obvious. It was almost as if the door
opened on its own. The human helped Colty through the door into a small, cramped, dark room.
Suddenly the door closed again and Colty held his ears tightly against his head. His heart
pounded. He had no idea what was happening.

``\alien{Thisss move},'' the human said pointing at one wall. Just as he said that Colty could
see the wall coming toward them. He felt panicked and tried to get away but in the weightless
environment he couldn't get any traction. The human put one hand against the approaching wall
and helped Colty get oriented. ``\alien{Feet down},'' it said pointing at the opposite wall.
Colty noticed that the other wall was moving away and he saw that the room was actually moving
as a self contained unit.

In a flash of understanding Colty realized he was in an elevator of some kind. He knew that the
human ship was designed as a rotating cylinder to induce a sense of artificial gravity. The
craft that picked him and Rachel off the desert must have docked on the central axis of the ship
and now they were `descending' toward to the outer edge.

In fact as the elevator moved Colty and his human companion drifted down to the `floor.'
Gradually Colty felt his weight returning and he was much relieved. It was surprising to him
just how disconcerting being weightless was. As the floor pressed ever more firmly against his
feet his anxiety level steadily dropped.

The human put one hand against the wall and Colty felt himself being thrown in that direction as
well. It confused him at first but then he realized it was just the Coriolis effect. Since the
outer rim of the spacecraft was moving laterally as it rotated they had to be accelerated
laterally as well. Colty clicked his teeth as he imagined the examples he could bring to his
next introductory physics class.

Finally, just about when Colty felt his full weight, the elevator stopped and the doors opened.
Waiting for him was a human wearing some sort of formal dress. The human greeted Colty speaking,
much to Colty's immense relief, fluent Argenian.

``Colty Harnick?'' the human said. ``I'm Mike Strauss from our Language Services department.
Welcome aboard the \textit{Golden Light}.''

\breakpoint

Colty stepped out of the elevator and into a small alcove that appeared to be at the end of a
long hall. The human that had accompanied him spoke a few words to Mike and the turned to Colty
and said, ``\alien{I mussst go}.'' Colty thanked him as he left.

``Are you in need of medical attention, Dr. Harnick?'' Mike asked.

Colty took a deep breath. ``No,'' he replied. ``I'm think I'm fine. I\ldots I could use some
water, though.''

``Of course,'' Mike said. ``This way.''

Colty followed Mike down the hall. He was surprised at how little human technology he saw. The
hall was very similar in many ways to halls Colty had walked before, yellow green lighting not
withstanding. There were pictures on the walls, although of strange scenes that Colty wondered
were of Earth. Yet some of the pictures looked like artwork. They were not unlike the art in the
Argenian national gallery except that the figures in the pictures were human. It was an odd
combination of alien and familiar sights.

They passed many humans who looked at Colty with curiosity. Most stopped and stepped aside,
giving him a wide berth. Colty clicked his teeth slightly. Now he knew how Rachel must have
felt on her first day walking the streets of Parnon.

There were doors off the hall but in places there were open spaces where humans worked or
gathered. In one area there was a collection of displays filled with symbols, graphics, and
other information. Two humans were standing there reading those displays.

``I apologize for coming for you without also bringing appropriate language translation
support,'' Mike said. ``It was negligent of us, and I take full responsibility for that
oversight.''

``It's fine,'' Colty said. ``I'm just glad you found us.'' He paused and then added, ``How is
Rachel. I am worried about her.''

Mike stopped and turned toward him. ``She is in hospital,'' he said. ``I don't know anything
else right now, but they are taking care of her. I want to thank you, on behalf of my people,
for helping her.''

They came to an intersection and Colty looked up and down the intersecting hallway. It curved up
slightly in either direction making it seem as if they were at the bottom of a bowl. Along the
walls right at the intersection some plants were growing in pots. They had strange green leaves
and yellow flowers. Colty wondered briefly how they could grow in such dim lighting.

On the ceiling were some small displays with human writing on them. Mike glanced at them and
turned down one of the intersecting halls. A group of three humans walked by. One stopped and
greeted Colty in broken Argenian. ``\alien{Welcome},'' it said with a smile. Colty swayed his
head.

``We are well met,'' he replied.

Finally they came to what looked like a cafeteria of some kind except that it was open and there
were humans eating in public. They stopped and gawked at Colty as he followed his guide into
the space.

``I'm afraid our cups are not ideal for you,'' Mike said. He put a cup under a metal spout and
touched a panel. In short order the cup filled with water. It was much more than Colty usually
drank but he said nothing. ``I understand you prefer privacy for eating and drinking,'' Mike
said.

``That is our usual way,'' Colty replied.

Mike lead him down a side hall to a small room. It was empty at the time. Mike consulted a
display beside the door briefly. ``This room is available now. I will be just outside if you
need me.'' Mike showed Colty how to open the door manually and then left him alone. Colty took a
deep breath. He was nervous but everything seemed fine. He trusted the humans. He had to trust
them.

The purpose of the room was not clear to Colty. It contained several pieces of furniture but
they were not anything Colty could use. On one wall was a large display device. Along the back
were more alien plants, each on a small table with some drawers. Colty laid down on the floor to
sip his water. He felt exhausted. He had no idea what time it was but surely it was almost
winter.

As he drank Colty suddenly realized why it was so dark. The star of Earth was spectral class X5.
He knew that human vision matched that spectral class and so the lighting system no doubt
produced lots of ultra-green that the humans could see. In addition the lights shed no rubinum
at all. It made perfect sense. Why would human engineers waste energy producing a color that
wasn't visible to them? Colty's understanding calmed him. The darkness didn't seem quite so
insidious when he understood the reason for it.

After a couple of arnets Colty heard a soft chime. Then Mike's voice came from what appeared to
be a speaker mounted beside the door. ``Dr. Harnick,'' he said. ``May we enter?''

``Uh, yes,'' Colty replied. ``I was just finishing.''

The door opened and beside Mike stood another human wearing some sort of formal clothing that
Colty had never seen. The two of them stepped into the room and the door closed behind them.

Mike said, ``I'd like to present Olivia Benedito, Captain of \textit{Golden Light}.''

Colty swayed his head solemnly. ``We are well met, Sir,'' he said.

Mike translated Colty's words for the Captain and then she spoke in the human language. Colty
could tell from the expression on her face that she was friendly. He wondered if she was easier
to read because of his familiarity with Rachel, another human female.

``Indeed, yes,'' Captain Benedito replied through Mike. ``We are well met. I am pleased to have
you here, and I must thank you, again, for helping Dr. Spencer in her time of need.''

``I did what I could,'' Colty replied.

``She was lucky to have you.''

There was a pause and then Benedito continued. ``We don't really have good facilities here to
make you comfortable and the authorities on the ground are anxious to have you back. If you are
up to it we'd like to return you to Parnon soon. The shuttle will be ready for launch again in
just twenty arnets or so.''

``I would like to get home,'' Colty said. He hesitated. ``But\ldots''

``You'd like to see Rachel first,'' Benedito finished for him. ``Absolutely. We understand.''

The two humans chattered between themselves briefly and then Mike spoke as himself. ``If you are
ready I can take you up to the hospital and we can find out what is happening there.''

``I am very ready,'' Colty replied. The three of them walked down the small hall to the main
corridor again where Benedito turned in the opposite direction. ``Have a good trip back,'' she
said before parting.

Colty had no idea where he was or where he was going but he followed Mike closely. Colty was
starting to get used to the lighting and he was struck by how comfortable things looked. There
were frequent alcoves along the hall with cushioned chairs, lots of artwork along the walls, and
many plants that decorated the space. It almost felt like he was walking through the Campus
Center in Parnon.

Suddenly they came to a large courtyard, fully three stories high. In the center grew several
full sized trees, their branches reached up into open space and were tipped with flat green
leaves. Around their base grew several smaller plants of various kinds and there were some seats
and tables where a few humans gathered. Bright, yellow green lights shined down from the
ceiling.

% It seems like there should also be some children on Golden Light. Does Colty see any here?

``You have so many plants,'' Colty said as he gazed up at the strange looking trees.

``\textit{Golden Light} has a complex recycling system that allows us to stay in space for
extended missions,'' Mike explained. ``The plants have a role to play in that. Yet their primary
benefit is psychological. We find it difficult to be in a confined space for long periods. The
plants make \textit{Golden Light} feel more open and more like home.''

Mike lead Colty across the courtyard to another elevator. They took the elevator to the second
floor and then walked a short distance down another corridor. They turned into an alcove with a
display on one wall. ``This is the hospital area,'' Mike said. He paused. ``A doctor will be
with us shortly.'' Colty wondered how he knew that.

They didn't have to wait long. Soon a human emerged from a door next to the display and
approached them. He spoke in the human language but Mike translated.

``I'm Doctor Mendon,'' he said through Mike. He addressed Colty when he spoke. ``Rachel is doing
fine. She was suffering from severe heatstroke, dehydration, and the start of organ failure. We
are in the process of rectifying all that. Her arm, as you know, is injured but we should be
able to grow the skin back completely. Fortunately there was no brain damage.''

``That is so good to hear,'' Colty said.

``She was very lucky,'' the doctor said. ``She should make a full recovery.''

``Can\ldots can I see her?'' Colty asked.

The doctor explained that she was groggy due to some of the drugs they were using but that she
was awake. The doctor took Colty and Mike into the back area and through some narrow hallways.
Unlike the spaces Colty had seen so far, human technology was everywhere. They came to a small,
dark room. Behind a partition Colty could see Rachel in a bed. Doctor Mendon pointed and said,
``Just go head. We'll be over there if you need us.''

% The timing isn't quite right here. It hasn't been long enough since their arrival on the
% Golden Light for Rachel to have reached the state she is described as being in below.

Colty stepped into the room. Rachel looked asleep. Beside her was an array of devices unlike
anything Colty had ever seen. A large display showed a series of lines, moving as he watched,
and many alien symbols. Her injured arm was wrapped in in some sort of dressing but there were
several tubes coming from it and connected to a smaller machine. Some more tubes had been
injected into her other arm and appeared to be filled with a dark fluid. Those tubes were
connected to yet another machine.

The room was strangely quiet and Colty shuffled closer to the bed. Rachel had been cleaned up
and was wearing a loose fitting gown of some kind. She opened her eyes. At first she had a
little trouble focusing them but then she saw Colty and smiled.

``Oh,'' she said with a bit of a slur. ``Hello.''

``How are you feeling?'' Colty said. He stood by the bed and looked suspiciously at the machines
around her.

``I'm okay,'' she said. ``I'm good.''

Colty clicked his teeth a little. ``That's good. They tell me you're going to be fine.''

She nodded. Then she looked at him again. ``You\ldots you're on \textit{Golden Light},'' she
said as if she only just realized it.

Colty shook his ears. ``Not for long. They are sending me back to Parnon in twenty arnets.''

``Oh'' Rachel closed her eyes for a moment and then opened them again. ``I feel funny,'' she
said.

``I think I should let you rest,'' Colty said. ``You get better.''

Rachel nodded slowly but her eyes were closed again. Colty quietly left the space and looked for
his guides. He didn't show any emotion but he felt joy in his heart.

% In my original conception of this encounter Rachel is coherent enough to ask Colty to tell her
% students that they should read the next section of the text and that she will be back in time
% for spring. This gives Colty a chance to express concern about the wisdom of her returning at
% all, etc. However, it doesn't make sense for Rachel to be that coherent yet and I don't want
% Colty to be on the ship too long. Perhaps I can slip this issue into Colty's conversation with
% Mike somehow.

\breakpoint

``I want to show you something,'' Mike said to Colty as the two of them walked back down the
corridor. They retraced their steps back to the courtyard and Mike directed him into the
elevator again. They had twenty arnets to kill and Colty had little to do but follow his guide.

They went up to the next level and Mike easily lead Colty through a maze of hallways. Colty
was fascinated by how the doors opened before them and closed behind them all without any touch
or even verbal command.

``How do doors know to open?'' he asked.

Mike smiled. He touched the back of his ear. ``I have an implant here,'' he said. ``It's an
interface to our computer system. It monitors my brain, and I can send it commands with my
thoughts.''

Colty was shocked. ``You mean you have mind reading machines?''

``Yes,'' he replied. ``I also have implants in my eyes that can project images onto my retina. I
can ask the computer questions and read the answers without any other equipment.''

Colty didn't know how to respond. The whole idea was disconcerting to him in the extreme.

``It's convenient,'' Mike was saying, ``but some people don't like the implants.''

``I can understand that,'' Colty said.

``Dr. Spencer feels that way,'' Mike continued. ``She uses wearable interfaces.''

Finally they came to a long room that Colty thought looked something like a conference room.
There were some plants in the corners but one wall was completely made of floor to ceiling
display panels. The door closed after they entered.

``I'm going to turn down the lights a little,'' Mike said. ``Is that okay?''

Colty agreed, a bit reluctantly, and the already dim lights drew even dimmer. Colty was
uncomfortable in the rising darkness but then, suddenly the display panels came to life. Colty
gasped in amazement, stepping back from the panels in a combination of shock and awe.

The wall had apparently dissolved and in its place was a view of Rujaria. A view unlike anything
Colty had ever seen or imagined; the view from \textit{Golden Light}. Cautiously Colty stepped
forward.

``It's all right,'' Mike said. He walked over to the panels and put his hand on them. ``They are
just displays. The wall is still here.''

% I need to be careful about temporal consistency here. They wouldn't launch the Shuttle unless
% Golden Light was in a good position for it. Twenty arnets is close a full orbit. That means if
% the ship is over Argenia now; the shuttle would launch on the next pass. It might be useful to
% calculate more precisely just where Golden Light is located during this scene.

% This entire scene might be gratuitous and could be considered for deletion to save space.
% However, I like the idea of Colty seeing Rujaria "from a distance" and gaining perspective
% about his planet. It would be even better if he could see both Forbin and Argenia, emphasizing
% that both countries are really part of the same world. However, I also like the idea of him
% seeing the sun set over Argenia as Golden Light passes into the night side. He's never seen a
% sunset before and it connects the story line to Earth in a subtle and pleasing way.
% Unfortunately temporal consistency (see above) might make it difficult to show everything.

Colty stood in silence looking down at his world. He could see it moving gradually beneath them
as \textit{Golden Light} orbited. On the left was the deepest day. It was a reddish brown area
bright with rubinum. Directly below he saw [I need to include a lot of place names here, etc]

``It is beautiful,'' Colty said at last.

``Yes,'' Mike replied.

Colty watched the deep yellow of the Argenian sea pass under them. Looking to his right he could
see the mouth of the Varsynthia River. ``Varnok is over that way,'' he told his guide. ``And
then beyond that is Parnon.'' He delighted in pointing out various places to Mike.

``It's too bad we won't get a good view of them on this pass,'' Mike said.

Colty could see clouds over [someplace], thick with rain. He saw the yellow hills of [someplace
else] and the Zagonhost River. Yet the works of his people were almost invisible. He saw a few
roads, maybe, here and there, but from \textit{Golden Light} Rujaria looked entirely wild and
unspoiled.

``I wish Rachel could see this,'' Colty said.

``She has many times,'' Mike said.

Colty clicked his teeth. Of course.

