
The truck halted. Rachel looked up wearily and Colty got on his guard. The big one stood up and
peered through the cracks between the panels on the side of the truck. ``Looks like we're
here,'' he said with a click of his teeth. Outside the little one came around to the back and
swung open the doors letting the bright red sunlight pour in. Colty breathed a sigh of relief.
At least he could see the sun again.

The big one jumped down from the back of the truck and motioned with his gun for the two of them
to get out. Colty climbed down easily but Rachel crawled on her hands and knees to the edge of
the truck's bed.

``Dind,'' the little one said. ``It's a bit worse for wear, I'd say.'' He shook his ears and
clicked his teeth.

Colty used his trunk to help Rachel down. She landed on her feet but stumbled a bit. She gripped
the side of the truck with her hands to steady herself. She did not look at her captors.

``You gonna live?'' the big one asked. Rachel didn't reply so he grabbed a short board from the
truck and poked her in the side with it. ``You gonna live?''

``Yesss,'' she said softly.

``She needs water,'' Colty said. ``If you could\ldots''

``Quiet!'' the little one hit Colty on top of his head with the butt of his gun. A spike of pain
shot down his front legs and he feel to his knees.

The big one watched Colty as he struggled to his feet again. ``Yeah,'' he said slowly. ``Get
them inside. I'll be right there.''

Colty and Rachel were led around the truck toward a small building. Colty looked around trying
to figure out where they were. It was desert. Only a few scrubby patches of brikken grass were
growing here and there. The hills around them were dry and barren but also steep and rocky. The
sun was about half way up the sky.

The building looked like a wooden, two story storage building of some kind. It wasn't large. It
had windows on both floors. There was a small garage next to and slightly behind the main
building. Against the side of the building was a fuel storage tank.

Once inside Colty was struck by how much of a mess it was. The main room on the first floor was
littered with junk. There were empty bookcases, wooden seats and tables, and broken pieces of
other furniture. He noticed some old car parts, some cooking pots, a few storage containers. In
the middle of the room was an open staircase that led to the upper floor.

They were led through a path in the junk to a small side room that was relatively empty. It had
a small table against one wall and a rack for clothing but no other furniture. Above the table
were two windows that faced the road and at the opposite end of the room was a door.

The little one waved his gun toward one of the corners under the windows. ``Make yous selves
comfy,'' he said. Colty and Rachel lay down on the floor there while the little one stood over
them, arms at ready.

After a moment they heard the big one coming into the building. He had his autorifle slung over
his back and he carried with him a small cup. It was water. He offered it to Rachel and she
seemed surprised.

``I don't want you dead before we get to\ldots\ interview you.''

Rachel took the cup and sniffed it suspeciously. 

``Besides,'' the little one said. ``It will be more fun to execute you if yous awake.''

She glared at him and then, wrapping her fingers around the cup, drank the water. She finished
it in moments. The little one shook his ears. Rachel handed the cup back to the big one.

``Thank you,'' she said softly.

The big one kneeled down close to Rachel. ``You understand that you aren't going to leave this
building alive.'' Rachel's breath was ragged. The big one nodded toward Colty. ``But if you
cooperate we will let him go.''

``I don't believe you,'' she said softly.

``We'll have to knock him on the head and leave some place where he can't track us but we don't
care about him. His only value to us is in making you behave.'' The big one stood up and swung
his weapon into his trunk. He pointed it at Colty. ``Otherwise I'll just shoot him now and get
it over with.''

``No!'' Rachel said. ``No\ldots''

``Rachel\ldots'' Colty said softly. She looked at him but didn't reply. Her face and hair was
dirty and matted with sweat and her eyes were filled with water.

The big one stood up. ``I'm going to check the markers and see if I can spot the others. I'll be
back in a bit.'' As he started to walk away Rachel called out to him.

``What have I done to you?''

The big one turned in the doorway between the two rooms. ``You exist,'' he replied.

``Her people\ldots'' Colty began, ``her people will be looking for her. Their technology\ldots''

``Oooo,'' the little one said in mock fear. ``Nobody can find us here. You are naked, human,''
he said looking at her. ``How does it feel to be without your toys?''

Time passed. How long Colty couldn't tell. He wondered if Rachel was going to need to sleep
again. He watched her carefully. She hadn't moved. Her head was down.

The little one had laid down on the floor in front of them and was playing with a knife.
Eventually he slid over toward Rachel and brandished the blade in front of her. ``I'm wondering
what color your blood is,'' he said. ``Can you tell me?'' Rachel didn't say a word but Colty's
heart sank.

``Maybe I needs to find out for meself.'' He grabbed one of Rachel's arms and pulled it toward
him. She let out a yelp and struggled but the dunari was stronger than her. He pressed her down
on the floor with the weight of his body.

``Stop!'' Colty cried out. ``Stop it!''

The little one pinned Rachel's arm to the floor with his foot and then using the knife cut into
the flesh of her arm. Rachel let out a howl of pain that was utterly gut wrenching. ``You
sadistic bastard!'' Colty yelled.

Finally the little one released Rachel and she scrambled away from him clutching at her wound. A
thick red fluid oozed between her fingers and dripped down her arm. She whimpered in pain and
fear. The little one inspected the tip of his knife closely. ``Would ya look at that,'' he said,
``it's red!''

``Was that necessary?'' Colty said.

The little one shrugged. ``What yous worried about? It ain't dunari. It ain't even an animal.''

``\alien{Ick arlomia nestra fornaganio},'' Rachel said in Forbin. It was the first time Colty
had ever heard her speak the language.

The little one glared at her. ``Shut up!'' he said angrily.

``\alien{Ick arlomik garn el munisto}''

``I said shut\ldots\ up! I don't want to hear that.''

Rachel fell silent but Colty could see her eyes moving the way they do when she's thinking.

The little one got up and then went to the door at the back of the room. He opened it and went
inside. It seemed to be some kind of storage room. There were lockers along one wall and some
boxes on the fllor. The little on opened a metal locker and took out a pistol and some
ammunition. He came quickly back into the room and knelt down on the floor in front of Rachel,
holding the weapon in front of her.

``This is what I'm gonna use to execute ya.'' He turned the weapon so she could get a good look.
``It might take a few shots. You probably won't die right away. That way you'll know what's
happening.''

He gripped the weapon in his trunk. ``Do ya want to see how it works? I'll show ya. See, first
you put the ammo here\ldots\ then you point it somewhere important.'' He pointed it directly at
her chest. ``Is that important? I di'nt know. Maybe here?'' He pointed it at her head. ``I guess
I'll just have to use both places.''

Rachel didn't say a word.

``So then ya point it and pull this trigger here, ya?'' With the gun almost directly against her
chest he pulled the trigger. There was a soft click. Rachel jumped and cried out but the gun
didn't fire.

``Oh, my mistake,'' the little one said. ``I forgot the safety.'' He shook his ears as he
disengaged the the safety. He pointed the gun at Rachel's chest again.

``Now if I pull the trigger\ldots'' Suddenly he turned and fired the weapon into the next room.
The sound was deafening and both Rachel and Colty jumped. He fired several times, punching a
couple holes into a metal can sitting on top of a table. Finally the can tipped over the edge
dumping its contents onto the floor.

``Eight shots is all ya get,'' the little one said turning back to Rachel. ``Ya think that'll be
enough to kill ya?'' He shook his ears again.

Rachel didn't reply so the little one stood up and went back into the storage closet. He
reloaded the pistol and put it in the locker, slamming the door closed. The top latch didn't
quite hook.

The little one closed the storage closet door. ``It dark in there,'' he said with a shake. He
came over the Rachel. ``How about I throw ya in there?''

Rachel didn't respond at first but after a moment she shook her head.

``Please, no.''

The little one opened the closet door again and unscrewed the artificial light.

``No, no,'' Rachel said shaking her head.

``Have some mercy for Jurita's sake,'' Colty begged.

``No windows,'' the little one said, ``no lights. Just you and the dark. Total dark.''

``No! Please No!'' Rachel cried out loudly. The little one grabbed her but she struggled with
surprising intensity. Finally he pointed his gun at Colty and said, ``I'll blow his leg off if
ya don't.''

With reluctance Rachel walked toward the open door. She hesitated at the entrance. ``Please,''
she said softly, ``don't make me.''

The little one pushed her into the closet and then closed and locked the door. Rachel screamed
and started pounding on the door. Colty could hear her yelling in the human language. He
struggled against his bonds and felt them weakening. ``Why are you doing this?'' Colty asked but
the little one just shook his ears.

They could hear Rachel thumping around. She moaned. She yowled like an animal. It was the most
alarming thing Colty had ever heard. Suddenly they heard a metalic crash and the little one made
a face of mock surprise. ``I sure hope it di'nt hurt itself.'' After a while the sounds stopped.

``It dark in there?'' the little one asked through the door but Rachel made no reply.

At that moment they heard the big one at the main entrance. He called out and the little one
answered. He came into the room with his weapon ready and looked around.

``What's going on?'' he asked. ``I heard shots.''

``Just a little target practice\ldots''

``Where's the human?''

The little one shook his ears. ``I just give it a little scare.'' He pointed to the closet. ``It
dark in there.''

``You idiot. That's where we keep our weapons.''

``Oh don't worry Gat,'' the little one said. ``theys locked up and she ain't in no condition to
mess with 'em.''

The big one glared at him. ``Get her out of there.'' He leaned his weapon against the wall and
bent down to look at Colty's bonds.

The little one opened the closet to find Rachel kneeling on the floor with the loaded pistol in
her hands. Immediately she fired a shot, at point blank range, into the little one's chest. The
little one reeled and she fired another shot, this time through his trunk and into his shoulder.
The little one fell to the floor.

The big one reached for his autorifle but Rachel bounded over the little one and fired at him.
The bullet lodged itself into a wooden beam uncomfortably close to Colty's head but it was
enough to dissuade the big one from hanging around. He ran into the next room and Rachel
followed him. Colty fought with his bonds and managed to slip his trunk out of them. He heard a
shot in the next room and then the front door slammed.

Rachel ran back into the room and over to where the little one lay on the floor. Orange yellow
blood was pooling around his wounds. The little one moaned and Rachel unceremoniously fired
another bullet into his head, killing him.

Rachel ran over to Colty who had already managed to free himself. ``We have to get out of
here,'' she said. Colty reached for the autorifle the big one had left behind but just then a
rain of bullets started to pour through the window. The two of them scrambled across the floor
to the main room and then ran toward the stairs with Rachel in front. The windows in the main
room shattered as more bullets pulverized them. Colty took a grazing hit to his back leg just as
he reached the top of the steps.

Silence reigned.

Slowly the door opened. Gat crouched low as he rounded the corner. Lifting his autorifle he
fired a few rounds to clear the air and then peeked over a low storage cabinet. ``Look,'' he
said, ``we want you alive but we don't care about the other one. Remember? Give yourself up and
he might live. Right now he will surely die.''

Gat cautiously looked over the room to convince himself that nobody else was there. ``You've
only got three bullets left in that thing. There ain't no way you're gonna get close enough to
me to make 'em count.''

Gat heard a shuffling thumping noise upstairs. He ran to just below where the noise was and
started firing the autorifle upwards right through the second story floorboards. Splinters of
fragmented wood flew through the air. He stopped and listened. He heard another thumping noise
in a slightly different place. Running to that spot he started firing again. This time, his
bullets ricocheted back, blasting large holes in the boards over his head.

Clever human, Gat though. He remembered the metal sheet up there. They must be using it as a
shield. It wasn't very large though. If he shot from an angle he might be able to hit them
anyway. Gat backed up a little and then a little more. He was close to the door as he raised his
autorifle.

Suddenly there was a shot and Gat felt a searing pain on the left side of his back. He spun
around to find the human standing just outside the door, pistol raised. She shot him again in the
face and Gat stumbled backwards falling to the floor. Rachel quickly kicked his autorifle away
and then stood over him, pistol ready.

Gat looked up at her. ``Don't do it,'' he pleaded. His phony Forbin accent was entirely gone.
``I've got two kids. I\ldots''

``I can't risk dealing with you later,'' Rachel said. Without further ado she fired.

Blackness took him.

\breakpoint

The only sound was Rachel's breathing as she lowered the pistol. The deep red Rujaran sun
filtered through the shattered windows and splashed on the floor where Gat's trunk lay
motionless. Rachel stumbled to the side and put the pistol on a table.

``It's fine now,'' she said softly more to herself than to anyone. Then she repeated herself
more loudly. ``It's fine, Colty. You can come down.''

Colty scrambled down the steps and trotted over to the human. He carefully stepped around Gat's
body. ``Are you hurt?''

``Just my arm,'' she said. She held it up and it was covered with red blood still oozing from
the wound the little one had made.

``Jurita,'' Colty said. ``We need to clean that up.''

``There's no time,'' Rachel replied. ``You heard what they said about others coming.''

``I'm still going to look,'' Colty said. ``Very quickly.''

Colty started searching the cabinets and drawers to see what he could find. ``The back room
where they keep the weapons might have something,'' Rachel said.

She knelt down and gingerly felt inside the pouch of Gat's clothing. She found a key and put it
in her pocket. Then she pulled out Gat's wallet and looked through it. ``He has some money,''
she called out. Colty didn't answer but she could hear him opening and closing lockers. Rachel
found a folded piece of paper in Gat's wallet. She opened it carefully and read it.

When Colty returned he was carrying another pistol and some ammo. ``Could this be useful?'' he
said. ``They have autorifles back there but I have no idea what to do with them. I couldn't find
any\ldots''

``Look at this,'' Rachel interrupted. She held out the paper and Colty took it.

``It looks like some kind of assassination order\ldots\ for you.''

``It says 'by order of Lord Junar' ''

``Yes,'' said Colty. ``It might be some kind of Junar thing.''

Rachel looked closely at Gat's body and then reached out and started to rub at the bands on his
lifeless trunk.

``What are you doing?''

She didn't reply but just kept rubbing. The bands started to smudge and when she took her hand
away her fingers were covered with some kind of black ink. ``They aren't Forbinites,'' she
replied. ``It's some kind of deception.''

``Junar.''

``What is that orange symbol on the paper?'' Rachel asked.

``I don't know but it looks like it was written in blood.''

``Blood?''

``I think so,'' Colty said.

Rachel stood up still holding Gat's wallet. ``We should take that note,'' she said. ``My people
have techology that can link a sample of blood with a specific individual. Maybe we can find out
who wrote the order that way.''

``I couldn't find any bandages but there is some tape.''

``It's fine,'' Rachel said. ``I think the bleeding has mostly stopped anyway. Thank Jurita, he
didn't hit anything essential.''

Colty clicked his teeth. ``Maybe we should go.''

Rachel looked around uncertainly. ``I could use more water. I wonder where he got the water he
gave me.''

``I didn't find any,'' Colty said.

The two of them left the building and walked around to the garage in the back. Rachel stopped by
the fuel tank but the pump was locked. She tried the key she had found in Gat's pouch but it
didn't fit. ``You didn't find another key did you?'' she asked.

``No, but we could go back and look some more.''

``Let's check for water first.''

The garage was unlocked but mostly empty. There were only a few tools hanging on the wall and a
couple of boxes filled with some paint cans and brushes in the back. ``I must have missed it
inside,'' Colty said.

``Or maybe it's in the truck.''

Colty turned to head back outside but Rachel paused to get a can of paint and a brush first. ``I
have an idea,'' she explained.

They put the pistols and ammo in the truck and then looked through the accessory compartment but
no water was found. Colty watched the road nervously. ``We should be leaving as soon as we
can,'' he said.

``Hold on.'' Rachel opened a can of paint and used a broken piece of wood to stir it. Then she
climbed up on the truck. ``Give me that,'' she said gesturing to the paint on the ground. Colty
lifted it up to her.

``My people have technology that can see the color of your eyes from space,'' she said as she
worked. ``But it will take time for them to search the whole planet. I want to make it a little
easier by painting my initials on top of this truck.''

``That will make the search faster?'' Colty asked.

``The analysis algorithms will pick it up more quickly, yes.''

Colty looked down the road in both directions hoping against hope that he would see nothing.
``Let's get out of here,'' Rachel said as she climbed back down to the ground. She gave Colty
the key she had found on Gat. ``Hopefully this will work.''

Colty got into the driver seat and Rachel lay down beside him. He slid the key into the
dashboard and then turned on the engine. He clicked his teeth and Rachel nodded. Colty started
to swing the truck around. ``Which way?'' Rachel asked.

``Away from the sun,'' he replied simply.

\breakpoint

Rachel opened the window and let the air blow over her arm. Colty glanced at her and saw that
she was covered with a film of water.

``Are you feeling right?'' he asked.

``I'm hot.''

The road stretched out in front of them, crawling over the dry, barren landscape. Sand was
everywhere with only scattered clumps of brikken grass, mostly in the shady areas beside rocky
outcrops. The world seemed to burn with the red light of the sun.

``I wish I knew where we are,'' Colty said. ``Or where we are going.''

``We must be far from Parnon,'' Rachel said as she stared out the window.

``I'm guessing Tunar province. It's west and a bit north of the Argenian Sea. I've been through
it before and it's a lot like this.''

``It's hard to believe anyone lives here,'' Rachel said.

``Not many do. The people in Tunar are mostly bargat herders or, these days, miners. It's a very
poor area.''

``I hope we don't meet anyone,'' Rachel said.

``We're going to need to meet someone,'' Colty replied as he looked at the dashboard. ``We'll
need fuel, and a phone would be even better.''

After a time they came to an intersection. Colty stopped the truck while they considered their
options. One road turned slightly to the right while the other seemed to continue straight. Both
roads looked equally worn and equally desolate. Colty got out of the truck and looked briefly at
the sun.

``If we are in Tunar then we're going to need to go south so maybe we should take this turn.''
Rachel didn't reply. ``Although we might come up against the Argenian if we do that.'' He walked
over to the window on Rachel's side of the truck. ``What do you think?''

``I don't know, Colty. It's your world. Whatever you think is best.''

Colty returned to the driver's seat. He paused, looking down both roads and then took the
straight route. ``It's just a guess.''

The karnons slowly ticked by without any relief from the barren landscape. Now and then they
passed mining flat trailers parked beside the road. They looked abandoned but Colty explained
that they were probably used as spares in case the regular trailers developed problems on the
long haul back to civilization. It pleased him to see the trailers. ``It means that I might be
right about us being in Tunar.''

They passed two more intersections and Colty just guessed at each of them. They saw no sign of
anyone. He glanced over at Rachel but she was sleeping again. Her clothing was wet and her hair
was matted. ``There's got to be a town around here somewhere,'' he muttered to himself. The fuel
gauge was creeping ever lower.

Rachel woke up maybe a quarter hour later. She looked out the window to see a field of scrubby
brikken grass growing alongside the road. ``I see life,'' she said with optimism.

``You didn't sleep long,'' Colty said.

``It's hard to sleep. It's so hot.''

``You're losing a lot of water.''

``I can't control it,'' she said softly.

They drove in silence for a while and then ahead in the distance they could see a small
building. Rachel was nervous but Colty was hopeful. Gradually it drew closer and soon it was
unmistakably a small, somewhat dilapidated house. ``We'll stop and see if we can get help.''
Rachel nodded but she clutched her seat tightly in her hands.

``Ask for water,'' she said.

Colty parked the truck at the end of the driveway and got out. He walked up to the door and
knocked but there was no answer. ``Is anyone home?'' he called out. ``Do you have a phone we
could use?'' He knocked again. ``Could we buy some fuel?'' He knocked again but eventually gave
up and returned to the truck.

``I think there is someone in there,'' Rachel said. ``I saw some movement at one of the
windows.''

``Maybe so, but they don't want anything to do with us.''

Colty started the truck again and continued down the road. Rachel watched the house recede into
the distance. She sighed softly. ``Don't worry,'' Colty said. ``Where there is one house there
is sure to be another.'' He glanced nervously at the fuel gauge. It was already indicating an
empty tank.

They passed by a couple more trailer stations and then some kind of radio tower. They were
definitely coming into relatively more populated regions. Colty glanced over at Rachel and saw
her taking short, rapid breaths. Her skin glistened with a film of liquid. He didn't say
anything but she noticed him watching her.

``I'm very hot,'' she said. ``The air is so hot.''

They drove around a outcropping of rock. The road stretched out ahead of them across the desert.
Far in front of them, several karnons at least, Colty could see a small black dot. ``Look,'' he
said. ``There's a building of some kind up ahead. We can get help there.''

Rachel leaned forward and peered into the distance. ``I don't see anything. My eyes are too
tired, I guess.''

Colty watched the fuel gauge nervously as the black dot crawled ever so slowly closer. The gauge
was below empty and the building was not approaching quickly enough. Colty glanced over at
Rachel again but she had her eyes closed.

The building was still far away when the engine began to knock. Colty's heart sank. He was
losing power. Rachel opened her eyes. ``What's happening?'' she asked. Before Colty could answer
the engine stalled completely. Colty steered the truck to the side of the road as it rolled to a
stop.

``We are out of fuel,'' he said simply.

Rachel slumped back into her seat. She put her hand on her forehead and wiped the water from her
eyes. She muttered something softly in her language.

Colty looked down the road. ``The building is closer. You see?''

Rachel sat up briefly. ``I see it now, yes,'' she said. ``It looks far to me.''

``We could walk there.''

``No.''

``I think it's less than a karnon.''

``I can't,'' Rachel replied, ``not in the sun.''

Colty shifted in his seat. ``How long will it be before your people find us?''

``I have no idea,'' Rachel said.

Colty sat silently for a moment. ``They might have water,'' he said finally.

``I can't do it!'' Rachel said sharply. She looked out the window and neither of them spoke.
After a pause she added in a softer voice, ``I'm not of this world, Colty, my body can't cope
with this kind of heat.'' She put her arm out the window. ``At least when we were moving there
was a breeze. I don't even feel the Nermela now.''

Colty took a deep breath. ``You could stay here while I go.''

``And leave me here?''

``To get help, Rachel.''

She drew up her legs and wrapped her arms around her knees. Colty had never seen her in such a
position before and he wondered what it meant. She didn't answer right away but then she nodded.
``You're right. You should get help if you can.''

Colty pulled the key from the dashboard and then hesitated. ``I'll leave this with you,'' he
said. Rachel didn't look up. He climbed out of the truck and closed the door behind him. He
looked down the road in both directions and then walked over to where Rachel had her window
open.

``It's less than a karnon,'' he said. ``I shouldn't be gone long.''

Rachel nodded but still didn't look up. ``Okay.''

He reached out his truck toward her but hesitated. Drawing it back he looked toward the house
ahead.

``Colty?''

He turned again to face her. He was struck by how resigned she seemed. She looked exhausted.

``Thank you,'' she said. She reached out and without thinking Colty lifted his trunk. Her hand
held it and his tentacles wrapped around her arm.

``Take care of yourself,'' he said. She nodded weakly.

Colty took a breath and began to walk down the road. He walked for what seemed like ages and yet
the house barely looked any closer. The powerful heat of the sun burned his back as he went. It
was hot. His dunari body protected him from the heat somewhat but even the dunari had limits and
Colty was a college professor not a trained desert survivalist. He was glad it wasn't far.

He paused briefly, flapping his ears to help him cool, and then continued trudging along the
sandy edge of the road. After walking for ages more he turned around. He could see the truck,
now distant and lonely against the sandy expanse. Yet his gaze was drawn above the remote ridge
where a large dark red cloud was looming. It was a dayenstat, a kind of wild sandstorm that
often rolled across the dusty landscape.

``Dind,'' he muttered.

He gaged the distanced back to the truck but it was probably at least as far as the remaining
distance to the house. Yet he wasn't sure what he'd find at the house whereas at the truck there
was at least some semblance of shelter. In the end he decided to push forward. He couldn't bear
returning to the truck with nothing.

He tried to walk more quickly, often glancing over his shoulder at the approaching dayenstat. It
was moving quickly and Colty soon realized that he'd barely be able to reach the house in time.
The wind started to pick up, carrying with it grains of sand. Each grain by itself was nothing,
but by the billions, carried by blasting winds, they would become a formidable force. Colty
started to worry that he had made the wrong choice.

By the time he reached the house the wind was howling strongly and sand scoured his legs and
face. He turned to look back down the road but the dayenstat had filled the air with dust and he
saw very little. He hoped Rachel had the sense to stay inside the truck. He knew she would.

The house was very much like the first one. A small building in need of repair. There was a
garage and a few hatram bushes along the side. The front yard consisted mostly of sand, now
being whipped up by the dayenstat. There was no sign of life.

Colty used his trunk to cover his face as best he could and pushed through the thickening air to
the front door. He knocked. Again. He was going to need to find shelter soon. He looked at the
garage and was wondering if he could hide against the leeward side when suddenly the door
opened.

``Get ya in here, and be quick!''

Colty was startled but he didn't hesitate. Head down he stepped forward into a narrow, dusty
hallway as the door slammed behind him. He turned to find a young man, not much older than Colty
himself standing there regarding him with a cocked head.

``Heya, Jillzy, we got a guest.''

A young woman came around the corner and two small children right behind. ``Oh, good knobs,''
she said with surprise, ``please, sir, come in out of the sand.''

Colty nodded his thanks and introduced himself. He felt so grateful that he had actually found
some people.

``I'm Yuntallah,'' the man said. He gestured toward the woman. ``My wife, Jillziah. The little
ones is Nicknar and Yuntak.''

``We are well met,'' Colty said.

Yuntallah clicked his teeth. ``You ain't from around here. What ya doing in the sands?''

Colty explained about the truck running out of fuel. ``We were kidnapped by Junar terrorists and
escaped,'' he said. Yuntallah waved his trunk at his wife and she quickly herded the childern
into the next room.

``What you talking about? Don't talk zind in front of the kids.''

``I'm sorry,'' Colty said, ``but I have to use your phone. My\ldots\ friend is still in the
truck and she's injured.''

Yuntallah regarded Colty carefully. ``You saying there are Junar around here?''

``It was many karnons west of here. We stole their truck. Look, can I please use your phone. I
have to get word back to Parnon.''

``Parnon?''

``That's where we are from. There are people there who can help her.''

Yuntallah shook his ears violently. ``I think you been in the sands a bit long. That's crazy
zind.''

Colty took a deep breath and touched his forehead with his trunk. ``I know it sounds crazy. But
however it sounds, my friend is hurt. I need to get her help.''

``Let him use the talk box,'' Jillziah said from the end of the hall.

``Talk box?'' Colty asked.

``There ain't no phones out this far,'' Yuntallah said. ``We have a talk box instead. Come.''

Yuntallah led Colty down the hall, through a kitchen filled with dirty dishes to a small alcove
under a window. The sun was blocked by the dayenstat but artifical lights glowed brightly to
ward off the darkness. There on a desk was a two-way radio of the sort Colty had seen only in
old movies.

``It don't got much range but we can get Potan's house on it. They got a phone that can call
into town.''

``Ah, okay,'' Colty said. ``I've\ldots\ never used one of these before.''

Yntallah shook his ears slightly. He lay down in front of the device and powered it on. Once it
had properly warmed up he started adjusting the controls and talking into a tall microphone.
After some time a thin voice could be heard coming from the speaker.

``That you, Jinix?'' Yuntallah said. ``Is your paps there?''

``Ya\ldots''

``Well, be a good girl and get him on the talker. It's important.''

After a few moments an older man's voice replied. Yuntallah briefly described the situation.
``Can you call Darnal and get him to send an ambulance up here?''

``We have to contact the Office of Alien Affairs in Varnok,'' Colty said. ``Can I\ldots\ can I
talk to him?''

Yuntallah looked at Colty closely. While keeping his gaze on the strange visitor he spoke into
the microphone. ``This guy wants to talk to you,'' he said.

Colty lay down in front of the radio and Yuntallah showed him how to use it. Then Colty spoke
directly. ``My name is Colty Harnick. I'm a mathematics professor at Parnon University. My
friend and I were kidnapped by Junar terrorists and we escaped. My friend\ldots.'' Colty paused
and glanced over his shoulder. ``My friend is the human astro-physicist Rachel Spencer. The
kidnappers wanted to kill her. She's been injured but she is alive. You need to get word to the
Office of Alien Affairs in Varnok so they can tell her people where she is.''

``That sounds like zind to me,'' the voice said, ``but I'll tell Darnal. Hold on.''

Colty leaned back. He glanced over and saw Yuntallah and his wife staring at him. ``Are you
saying there's a human just down the road?'' Jillziah asked.

Yuntallah shook his head. ``He's sand burnt, I think.''

The radio crackled. ``What did you say your name was?''

``Colty Harnick.''

There was silence. Colty fidgeted with the cable leading to the microphone. The wind outside
shifted and Colty could hear the sand tinkling against the window.

``No way to get anything out there now,'' Yuntallah said, ``gotta wait until the storm clears at
least.''

Again the radio crackled. ``Ah\ldots\ okay. You there?''

``Yes,'' Colty leaned into the microphone as he spoke.

``Darnal is taking a sand crawler out there now.''

Yuntallah and his wife looked at each other. ``He's crazy,'' Yuntallah said, ``He's crazy!'' he
repeated into the microphone.

``Apparently someone named Harnick and a human have gone missing,'' the voice on the other end
of the radio said. ``Darnel's going up there to check it out. Lay flat 'til he gets there.''

``Dind,'' Yuntallah said softly. He clicked the microphone on. ``Thank ya, Potan.''

Colty stood up again while Yuntallah and his wife looked at him. Colty looked out the window but
the storm was still raging. ``I want to try to bring her some water,'' he said.

``No,'' Yuntallah said flatly. ``It's suicide to go out there now. Wait 'til the storm clears.''

``It won't be long,'' Jillziah added.

``You'd better hope this human is smart enough to stay inside that truck,'' Yuntallah said.

``She's very smart,'' Colty replied. But he felt aggitated. He had told Rachel he would be back
soon and already this was taking far longer than he expected.

Jillziah started going through some cabinets and gathering together some tape, string, cloth,
and assorted things. ``You said she is hurt. You should bring some basic first aid.''

``Jillzy, we can't bandage up a human.''

``They aren't that different than us,'' Colty said. He looked over the items Jillziah had
gathered. ``They cut her. This might help.''

Yuntallah swayed his head slightly. ``Okay. I'll get water to wash out the wound. You don't want
desert mites getting in there. They's bad news.''

``And we'll need water for her to drink too,'' Colty added.

In short order they had everything assembled. Then it was just a matter of waiting for the storm
to clear. They lay in the living room mostly silently while the two boys alternated between
playing together on the floor and staring curiously at the visitor. Colty got up and paced
around some and then lay down again.

``You's worried about her, eh?'' Yuntallah said.

``She's my friend.''

Yuntallah got up and went down the hallway toward the door where Colty had come in. Colty
followed behind. He cracked open the door and looked outside. Colty could feel the hot air
blowing into the hall.

``It's easing up,'' Yuntallah said.

Colty looked over his shoulder. The air was not as thick with dust as before. He could even see
down the road a ways. As he looked he thought he saw a faint orange light. It seemed to flicker
and bounce as it moved along the horizon.

``Jillzy,'' Yuntallah cried out. ``Darnal's coming.''

Yuntallah closed the door while he and his wife made a final check of all the items they had
gathered. Jillziah put them into a box and covered it tightly with a tied cloth. ``You better
wear a sand jacket,'' she said. ``I'll get an extra for our guest.''

``I think it will be okay,'' Yuntallah said. ``It's slacking now.''

``It might start up again.''

Jillziah disappeared and shortly returned with two sand jackets. Yuntallah took one and started
to slid it on. ``I'm sorry this one's so old,'' Jillziah said as she gave the other to Colty.

``It will be fine,'' Colty said. ``I really appreciate your help.''

By the time they were ready the sand crawler pulled into the lane of the house. It was a large
vehicle that looked like an ordinary truck with treads instead of tires. It had large,
artificial lights on the front that glared bright orange and special wipers for the hardened
windows that kept the sand from accumulating against the glass.

When the crawler stopped two men got out, bundled in a sand jackets and made their way toward
the house. Yuntallah let them in right away.

``How ya,'' one said in greeting. Colty could see the symbol of the Tunar Provincial Police on
his jacket. The other man, younger, said nothing. They brushed off the worst of the dust and
then, at Yuntallah's prompting, came into the kitchen. He seemed to know them both.

Colty introduced himself and the older man, Darnal, gave him a quick look. Then he held out his
trunk for Colty to shake. ``Well, well,'' he said. ``The boys at IIB in Tunaria are going to be
surprised to see you. Are you injured.''

``Only slightly,'' Colty said. ``I'm worried about Rachel. She's in the truck down the road and
she's not in good shape.''

``There's really a human just down the road?'' the younger man said.

``Racarta is my deputy,'' Darnal explained.

``Yes, but we should go to her right away.''

``Well,'' said Darnal. ``That's what I got the crawler for.''

``Did you contact the OAA in Varnok?'' Colty asked.

``I called them, yes. They are waiting to hear back.''

Jillziah pushed the box into her husband's trunk and encouraged them to get started. ``Oh,
Jillzy you should come. You know more about bandaging and all that.''

``I got to watch the kids,'' she said.

``Racarta can stay with them,'' Darnal said.

By the time the four of them made their way out to the crawler the storm was almost over. Colty
could even see the green sky again in places and the wind was carrying very little sand. Colty
lay in front with Darnal while Yuntallah and his wife crowded in the narrow back seats. The
radio on the dashboard was on and crackled slightly now and then.

Darnal started the crawler and pulled back into the road. It moved faster than Colty expected
but it still seemed to take much too long. He anxiously scanned the road ahead through the
clearing air looking for the truck.

``How far did you say it was?'' Darnal asked.

``I don't know. A karnon maybe.''

``We've already gone a karnon and a half,'' Darnal said.

Gusts of wind blew occasional clouds of dust across the road, now almost completely covered in
places. The crawler came to a dune, about half the height of a dunari man, flung across the
road. It climbed over it easily. Darnal picked up the mic of his radio. ``This is D. Tell 'em at
the town garage they gonna need to do some clearing out here.''

The crawler climbed over a small rise in the road and then, not far ahead Colty could see the
truck at last. It looked very small and insignificant against the harsh Tunarian desert. The
crawler came to a stop beside the truck and Colty quickly scrambled out.

``Rachel!'' he cried. The front seat where he had left her was empty. The window was still open
and a thick layer of sand covered the seats and dashboard. However, one of the back doors of the
truck was open and inside, laying on the floor he found her.

``Rachel?'' he climbed into the truck but she didn't move. He reached out and touched her arm
but there was no response. ``Oh, Jurita, be alive,'' he said softly. ``Are you sleeping?'' he
said more loudly.

``Good Hosts of Jarnwal,'' Darnal said as he swung open the other door of the truck.

``Dind,'' said Yuntallah.

Jillziah climbed up in into the truck as well and cautiously approached. Colty didn't seem to
notice.

``Rachel? Wake up. We have water for you.'' He gently put his trunk on her head and she opened
her eyes. Colty was ecstatic. ``We have water for you.'' Jillziah got a small container of water
from the box and handed it to Colty. Rachel's eyes fluttered and she moaned but she didn't seem
to know they were there.

``She's covered with water,'' Jillziah said.

``Yeah,'' replied Colty. ``It's something their bodies do when they are overheated. It's not
good.'' Colty opened the bottle but he wasn't sure what to do with it.

Yuntallah had climbed into the truck as well. ``Maybe drip some into it's mouth. That's what
they do with baby uitias that won't drink.''

Just then Rachel coughed and tried to sit up. She looked at them and then cried out, pushing
away from them and muttering in her strange, alien language. Yuntallah backed away but Jillziah
seemed unperturbed. ``She's sand burnt,'' she said. ``It's okay,'' she spoke to her. ``We want
to help.''

Rachel stared at her blankly and then turned to look at Yuntallah and finally Colty. Recognition
came over her face.

``Colty?'' she said weakly.

``Yes! It's me. I found some help.''

``Thank Jurita,'' she said. ``I'm so hot. I\ldots''

``I have water.'' Colty held out the bottle to her and Rachel tried to take it. Her hands shook
and she almost dropped it. Colty helped her bring it to her lips but she only sipped a little.''

Jillziah looked at Rachel's arm, now covered with dried red blood discolored by the orange-brown
sands of the desert. ``Let me look at your hurt,'' she said. Rachel sipped some more water and
then held out her arm. Jillziah hesitated before moving closer. She glanced at Rachel and then
at her injury. She took a cloth from the box, put some water on it and started to wipe away the
desert dirt around the cut.

``I'm sorry this happened to you,'' she said.

Darnal appeared again at the back of the truck. He didn't get inside. ``I got word to the folks
in Varnok,'' he said. Rachel looked at him and he spoke directly to her. ``Your people are
coming for you.''

``Thank you,'' she said softly.

Jillziah took a tube of ointment from the box and uncapped it. Rachel watched her closely.
``What is that?'' she asked.

``It's antiseptic,'' she replied. ``It kills bacteria and mites too.'' Jillziah squeezed some
onto Rachel's arm and started to spread it over the injury. Rachel whinced. It was a gesture
very much like that any dunari might make. ``It stings, huh?'' Rachel nodded.

Rachel reached for the water bottle and Colty handed it to her. Her grip was stronger now and
she drank liberally. She took a deep breath.

``Feeling better?'' Darnal asked.

``Yes. Thank you all.''

Jillziah took a bandage from the box and wrapped it around Rachel's arm. ``You tell me if it's
too tight,'' she said. ``I don't know how tight to make it on you.'' After a few wraps she
closed it off with a bandage clip.

``We have some money,'' Rachel said.

``No money,'' Jillziah replied.

``For the water?''

``No. It wouldn't be righteous.''

``I don't want to take your water without\ldots\ something.''

Jillziah put her trunk on Rachel's arm. ``You pay us back by doing something good for two other
people. Okay?''

Rachel smiled slightly. ``I will.''

Darnal motioned to Colty and Colty climbed back out of the truck. There Darnal and Yuntallah
were apparently having a conversation.

``Can she be moved?'' Darnal asked. ``We should get her back to town where we can wait for her
people.''

``I don't think we need to worry about moving her,'' Colty said. ``The humans will probaby be
here shortly.''

``What do you mean?'' asked Yuntallah.

Colty scanned the now clear and sunny sky. As if on cue he pointed his trunk into the distance
were a thin line was moving toward them. ``Here comes her ambulance,'' Colty said.

``Sweet Jarnwal,'' said Darnal.

Colty put his head in the truck. ``A transport is coming,'' he said.

Rachel turned to Jillziah. ``There's going to be a loud explosion, but it's normal. Don't be
frightened. Colty, tell the others.''

Colty explained to Darnal and Yuntallah about the transition but they seemed mesmerized by the
approaching streak. It descended toward the ground like a meteor with no reduction in speed at
all. Normally, at Parnon, the transports went through the transition at high altitude where the
thin air and great distance dulled the sound. But here the humans were clearly in a hurry. They
obviously intended to make the transition as late as possible.

The three men instinctively stepped back. Even Colty, who knew what was going to happen,
flinched. The transported rocketed toward the ground and then, when it seemed as if it was about
to pulverize itself against the rocks there was a blinding green flash and a powerful,
thundering explosion. They could feel it in their chests and it seemed to shake the very ground.
Yuntallah stumbled backwards and fell against the truck and Jillziah cried out with surprise.

After the transition the transport abruptly, even instantly, was traveling at a more normal
speed right over their heads. It circled around once, twice, as if looking for a place to land.

``I don't know where they are going to put that thing down,'' Darnal said. ``The nearest
areoport is a hundred karnons from here.''

The transport flew down the road and, after making a tight turn, began to fly back toward their
position, dropping closer to the ground as it went.

``Theys gonna land on the road,'' Yuntallah said.

Rachel dragged herself to the edge of the truck and with Jillziah's help managed to climb out.
She leaned against Jillziah's side to help steady herself and walked slowly to where the other
dunari were standing. Her tall, thin body stood out obviously.

The transport touched down on the road, kicking up a huge plume of dust as it did and quickly
braked to a halt only a short distance from them. Colty turned and looked at Rachel as she
watched the arrival. She was covered with sweat and dried blood and looked utterly exhausted.
Yet Colty knew everything would be fine now.

As soon as the transport stopped three humans, completely covered in some kind of protective
suite got out and came running toward them. The three Tunarians held their ground proudly,
apparently unconcerned by the approaching aliens. Colty stepped forward to meet the humans while
Rachel simply leaned on Jillziah.

``I'm so glad you are here,'' Colty said, but the humans didn't really reply. One of them nodded
at him and made a gesture with its hand but Colty didn't understand what it meant.

They walked toward Rachel who now stepped forward as well. They gathered around her and talked
to her in their own language. Rachel replied in kind. They held some devices to Rachel's skin
and eyes and then attached something to her uninjured arm.
