
It was a perfect hour. The Nermelum, never very strong in the Rungard Desert, was almost still.
The dust had fallen out of the cool air leaving it beautifully clear. Paraz looked up at the
tower against the bright green sky, pulling his heated jacket around him. The tower was a modest
structure built for only one purpose. Soon it would be gone.

Several military engineers worked the cable crane, hoisting the Cracker up the side of the tower
to a small platform there. They were going slowly and carefully, mostly out of fear of damaging
the device they had all poured so much time and energy into building.

``We should get back to the observation bunker, Dr. Paraz,'' said a voice at his side. It was
General Fotkey. Paraz got into Fotkey's command car and soon they were being driven across the
barren desert along a temporary road toward a small rise several karnons away. Paraz turned
around to see the tower receding in the distance.

Rungard was at the end of the world. Located far south of any Argenian city, the desert was
bathed in the dry, frigid wind blowing over the Rujaran south pole. Rock lichen was the only
thing that lived there. The landscape was utterly desolate.

When they reached the bunker, Fotkey's driver parked the command car behind a concrete shield
and the men went into the low, thick walled concrete building partially embedded in a rocky
hillside. Inside was one large room punctuated by concrete pillars. Artificial lights gleamed
from the ceiling filling the room with a bright rubinum glow. In one corner was a radio rack
manned by a smartly dressed communications officer. Gathered around were several other military
officers and selected scientists. Dr. Rocalla of Parnon University was there.

``We met at last,'' Rocalla said as Paraz arrived. The two physicists gripped trunks in formal
greeting.

``It's nice to actually see you finally,'' Paraz said. ``I grow weary of all the cloak and
dagger communication.''

The men turned toward the back of the building where narrow slits had been cut into the
concrete. Through those slits, far in the distance, they could see the tower standing like a
thin metal splinter sticking out of the sand.

``It's a shame the human won't be here to see this,'' Rocalla said with a click of his teeth.

``The humans will know soon enough,'' Paraz replied.

``Maybe. Maybe not,'' said Fotkey. He did not seem concerned.

All the observers talked quietly for a time and then Fotkey was called over to the radio rack.
After some time talking on the radio he turned and addressed the group. ``Make ready,'' he said.
There was an announcement over an intercom to warn anyone who might still be outside. The
observers gathered behind the slits and held pieces of welding glass in front of their faces.

Fotkey gave the command and a countdown began. The voice of a young dunari man filled the air.
Paraz shifted slightly. There was a slight chance that this test would start a chain reaction
that would consume Rujaria's atmosphere. The numbers showed that it wouldn't happen. He wasn't
worried. Much.

``Three\ldots\ two\ldots\ one\ldots''

At the location of the tower there was a brilliant flash, for a moment brighter than a thousand
suns. The tower and everything around it was immediately engulfed in an enormous fireball,
glowing with fantastic energy. The mountains all around the site were light with a greenish glow
that temporarily overpowered the sun. The artificial lights in the bunker were pale by
comparison.

``Jurita it works!'' Fotkey said with enthusiasm.

Paraz watched with almost horrified fascination. The fireball spread rapidly coming toward the
base with frightening speed. The calculations all said they would be safe but Paraz couldn't
help but take a step back. Moments later, the roar of the explosion reached them. It was a
deafening howl that shook the very ground. Paraz could see the blast wave moving out ahead of
the main fireball. It kicked up dust in the desert like some kind of hot, angry dayenstat. Above
it all, rising high into the Rujaran air, was a huge mushroom shaped cloud of fire, smoke, and
sand.

\breakpoint

Colty sat with Rachel and Lucasa on the stone porch of the student center. It was a lovely day;
breezy but not windy, warm but not hot. Colty knew it was just the kind of day Rachel liked. She
was sipping some water while they chatted about classes, mathematics, and flowers. Colty wasn't
sure why Lucasa was there. He seemed very intent on spending time with Rachel that day.

``I think your proof of theorem eight in your paper is good,'' Rachel was saying. ``It's clever
how you worked around the renormalization problem.''

``Thanks,'' Colty said. ``It's just a variation of\ldots''

At that moment a strange beeping noise seemed to come from Rachel. She looked surprised to hear
it and opened her pouch. She took out a small rectangular device that fit comfortably in her
hand. It beeped while she held it, and she looked at it with a puzzled expression.

``So sorry,'' she said. ``I have to answer this.'' She held the device to her ear and said,
``Yes, hello?'' in Argenian.

Colty glanced over at Lucasa who was watching the human closely.

Rachel seemed to listen for a moment with her head down. Then, suddenly, she lifted it up with
an expression of surprise. She looked directly at Lucas. She listened a little more and then
spoke into the device in her own language. She nodded her hand, said a few more words, and then
put the device back into her pouch.

``I have to go,'' she said.

``Is everything okay?'' Colty asked.

``It's fine,'' she said, ``they just want me back at the compound.''

Colty swayed his head. ``I need to get ready for my next class anyway.'' The three of them stood
up and Rachel seemed very distracted. ``Are you sure you're okay?'' Colty asked again.

``I'm fine.''

``Why don't I walk you to your van,'' Lucasa said.

Colty left for his office while Lucasa and Rachel made their way back across the Yellow toward
the physical sciences building. They walked in silence for a while. Lucasa wasn't sure what to
say. Finally he spoke. ``It didn't take your people long to find out.''

``You knew about this,'' Rachel said. ``That's why you are here now.''

``I'm sorry,'' Lucasa said. ``I was under strict orders to say nothing.''

Rachel did not reply.

When they reached the access road Rachel raised her arm and waved her driver to come and get
her. Lucasa waited by her side as the van approached.

``Do you know what your people are going to do?'' Lucasa asked.

``I don't know.''

``I fear this will interfere with our relationship.''

``I fear for your world.''

% Would Lucasa know about this? It's possible that in the interest of maximum secrecy the
% Argenian military would not tell him about it. On the other hand as the ambassador to Earth he
% probably should be in the loop (what if the humans started asking questions about the activity
% in Rangard?).
%
% I think it would be good to just drop this scene on the reader with very little build up. That
% way the reader will be as surprised and shocked as Rachel. However the tricky part will be to
% show how Lucasa gets the information without actually revealing what that information is.

\breakpoint
