
``Why do you look like the chortak?'' a voice said from the crowd. A large group had gathered
around them. Colty felt uneasy but Rachel did not seem perturbed at all.

``It's because of parallel evolution,'' she said to the group. ``On Earth there are forests like
here and creatures like the chortak. My species evolved from some of those creatures and my body
reflects the tree climbing heritage of my ancestors.''

``Is Earth dying?'' another voice asked. Colty looked over in the direction from which it came
but he couldn't tell for sure who asked it.

Rachel turned around. ``Oh no,'' she said. ``The total biomass on Earth is several times greater
than on Rujaria. We have huge forests that cover millions of square karnons, teaming with life
of all kinds. We have great, open grasslands with many herds of grazing animals. We have huge
seas some of which are also filled with all sorts of aquatic plants and animals.''

She held up her hands as if to hold an imaginary ball that she pretended to turn. ``Earth
rotates in only about three hours. Each place on the surface alternates between day and night.
This makes the temperature much more uniform than here. On Rujaria life is concentrated in small
regions where it's warm, but not hot, and wet. On Earth life is everywhere. Earth is not dying
at all.''

% What about Jitara and Brutik?

She had everyone's attention. Never before had a human spoken so frankly about their world.
People only had Limzar speculation to satisfy their curiosity about the home of their alien
guests. Now, for the first time, they could hear about Earth from someone who had been there.

``Some people think your world is completely covered by your technology.'' It was an older man
in the front row. Rachel responded to his comment but addressed her words to all.

``We use our technology, yes,'' she said. ``to build cities and similar things. Yet we also love
the natural beauty of Earth. We've designated large areas as wilderness to remain undisturbed by
us. The city of Montreal, not far from where I grew up, is about the size of Varnok, but it has
far more trees and flowers along its streets.

``I hope that one day you will come to see for yourself. I hope that you will visit our museums,
listen to our music, and enjoy the beauty of our world just as I have been enjoying yours.''

Colty could see some people swaying their heads.

``You would have to bring us,'' someone said. ``We could never go to Earth on our own.''

``Oh, but you're wrong,'' Rachel replied, turning to face the direction from which the comment
came. ``You are a highly intelligent people. Look at this great city you built. Look at the
great works of art in the museum just across the park. There is no doubt at all that even
without any help from us, you will work out the details of interstellar travel in a short
time.

``We estimate that our technology is only 15 or 20 arnox ahead of yours. That might seem like a
long time, but it isn't. Rujaria is 60 \emph{billion} years old. The 2000 years that separate us
is only one thirty millionth the age of your world. If things had been ever so slightly
different, you would be the ones reaching across the stars instead of us.

``I know that you will come to visit. It will happen sooner than you think.''

