
\chapter*{Epilogue}
\addcontentsline{toc}{chapter}{Epilogue}

Colty wrapped his trunk around the steering control of the vehicle, his touch activating the
voice recognition system. ``Switch to manual control,'' he said in Argenian. The vehicle
responded with a female dunari voice that sounded to Colty strangely like his mother in law.

``Not recommended.''

It was illegal to operate a car under manual control in urban areas or on high traffic roads.
Yet on the secondary roads of northern Vermont there wasn't much concern, and Colty liked to
drive. He had been warned at the embassy to be careful in any case. The electric motor and
aerodynamic shape of the car made it easy to surprise wildlife. Yet it wasn't Colty's first time
taking this route. He understood the risks.

``Disregard recommendation,'' he commanded.

A red light appeared on the control panel and Colty could feel the steering system engage. He
depressed the accelerator a little to maintain his speed. He drove for some time past fields and
through forests. The road roughly followed along a river that it crossed several times. He drove
by a few farms, some houses, and through the occasional town. Yet between the signs of human
habitation there were large expanses of just trees. The forests of Vermont were lush, green, and
full of life.

Colty recognized the turn without the navigation system's help. The large maple tree at the
corner was a consistent landmark. He slowly drove along the gravel road, following its winding
up hill path. After about a kilometer he turned onto a narrow driveway. It circled around a
small rise and came to an unassuming little house, nestled against the forest and across from a
hayfield.

Colty parked in the driveway in front of the little garage and turned off the car. He got out
and took a deep breath. Her house looked good. It was well kept and clean. The house had a
large, screened in porch on the front, and a small patio at the back, under the trees. The
gardens were beautiful and many flowers were blooming in the late summer sun.

``Colty!''

He looked up and saw Rachel coming out the door and down the step. She had her arms open and a
smile on her face. Colty spread his ears and came up the path. Rachel ran to meet him, throwing
her arms around his neck and shoulders. He wrapped his trunk around her waist.

``It is so good to see you again, my friend,'' she said.

They embraced for a long moment. ``Let me look at you, Rachel,'' Colty said finally in English.
She stepped back and smiled but held his trunk affectionately. Colty shook his ears slightly.
``My, my, your hair has gone completely gray. The last time I was here there was still lots of
black.''

Rachel cocked her head slightly. ``I know. Well, I'm getting old.''

``Eighty four is not that old yet,'' Colty said. ``Yes?''

Rachel just laughed. ``Come on, let me show you the place. I've got a little green house now.''
Colty had been there several times over the years but each time things were a bit different.
Rachel loved to show off the changes and talk about the plants she was growing.

``I had a problem with aphids on my roses last year,'' she said. ``They are the most persistent
little things.''

When she finished the tour she offered Colty some food and drink. He wasn't really hungry but he
graciously accepted a little toast. He knew it was human nature to associate eating with social
events. Rachel got a mat from a closet and unrolled it on the porch. Colty settled down on it
and she sat on the porch beside him with a glass of lemonade in her hand.

``Tell me everything,'' she said with a smile.

Colty shook his ears. ``We have a new First Citizen now.''

``I heard,'' she said.

``He's quite a visionary,'' Colty said. ``He's challenged Rujaria to put a working
communications satellite, of Rujaran design and construction, into orbit by the end of the
arnox.''

Rachel took a sip of her lemonade. ``Wow, that's quick.''

``The Forbinites already have a rocket that can go suborbital, and Argenia has a design for a
prototype satellite. The First Citizen has convinced the Forbin Prefect that we should work
together on this. They hope to have a test satellite in orbit in just 20 years\ldots\ Rujaran
years.''

``That's great!'' Rachel said.

A breeze blew across the field and over Rachel's porch, rattling the wind chimes softly. Colty
took another deep breath, smelling the air of Earth.

``What of Joleia?'' Rachel said.

``She is fine. She's still at the National Library. They are just starting to convert the
holdings to electronic form.''

``That must be exciting for her.''

``Oh!'' said Colty. ``Would you believe Jarnia is applying to university now?''

``How time flies. Is she going to Parnon?''

Colty shook his ears. ``She applied for my sake. Her father thinks it would please me, but I
know she'd rather go elsewhere. She's got applications in at several schools including the
University of California. She wants to see the Pacific\ldots\ and the night.''

Rachel chuckled slightly as she took another sip of her drink.

``And what about you?'' Colty asked. ``Are you doing anything interesting?''

``Oh, you know\ldots,'' Rachel waved her hand nondefinitively. ``I take care of this place, tend
my gardens, look through my telescope.''

Colty clicked his teeth slightly. ``Yes\ldots\ and\ldots? Come on, I know you better than
that.''

Rachel smiled coyly. ``Well, I might also be evaluating the logical consistency of Jenkins
Multi-Manifold Theory.''

Colty shook his ears and Rachel chuckled slightly. ``NEIST has been good about giving me time on
the neticular core for my pet project,'' she added.

``Being Professor Emeritus has its perks,'' Colty said.

Colty had a bite of toast and Rachel smiled at hime. ``Will you be able to stay long? I still
have that room upstairs. You could stay a night or two perhaps.''

Colty sighed. ``Unfortunately no. I have to be at the conference tomorrow and then back to
Rujaria already.''

``Such a long trip for just a conference.''

``To be honest, I came mostly because I wanted to see you again, even if just briefly. I'm
getting old too, Rachel. I'm going to retire soon. This will be my last trip to Earth.''

``Oh.''

Colty looked at her. ``I just wanted to\ldots\ make sure you were going to be all right. I worry
about you here, alone, in the middle of nowhere.''

Rachel gave a little smile. ``I'm fine, Colty. Really. I have my friends, my plants, and my work
to keep me occupied. Besides this place is not as remote as that house I had in the Sarnowledge
Mountains.''

``You were younger then.''

The two friends talked through the afternoon. They spoke of times past and of hopes for the
future. They spoke of Rujaria and of Earth. They spoke of flowers and of physics. Rachel had a
sandwich on the porch for her dinner while Colty sipped some lemonade. All the while the bright
star of Earth drifted slowly across the sky. As it sank toward the west, the shadows grew long
and the sky grew red. Colty checked his watch.

``I should go,'' he said. ``I have a long drive ahead and I need to be there by morning.''

Rachel nodded.

They got up and walked down the path together. Rachel stood by while Colty got into the car. He
lowered the window.

``Take care, \textit{tinka},'' he said.

``I will.''

They paused. Rachel started to cry. ``You should go,'' she said. ``It's a long drive.''

Colty swayed his head. He swallowed hard. He didn't want to leave, yet he raised the window and
powered up the vehicle.

``Destination?'' the system asked, in his mother-in-law's voice.

``Pelican Hotel, Park Street, Philadelphia Pennsylvania.''

The car pulled out and started on its way. Colty looked back at Rachel standing beside the
drive. He waved his trunk but he wasn't sure she could see. In just moments the curve of the
road hide her from view and suddenly he felt a very, very long way from home.
