
The First Citzen hung up and paused, holding his trunk on the earpiece as if to be sure it was
really lodged in its holder properly. He sighed and turned off the secure phone. Getting up he
left his office and strode down the hall to the conference room with the rosenwood table. There
Marshall sat kneeling at the side of the table with his hands clasped on his lap. Across from
him looking uncomfortable was Narlock, the IIB investigator on the case, and Lucasa. All rose as
the First Citizen entered.

``Please take your seats,'' the First Citizen said as he found his way to the head of the table.
Once everyone was settled again he turned to Narlock. ``What news do you have?''

``We've had a few other reports of\ldots\ her being seen but nothing as well confirmed as the
Clickora sighting. All reports point to her being take west, by a circuitous route. It appears
they're taking her to Forbin, although they're changing vehicles frequently making them hard to
track.''

``It doesn't make sense,'' Lucasa blurted out. ``If Forbin wanted Ra'hel why not just ask? I
have to believe she would have gone of her own free will without hesitation.'' Marshall nodded.

``The Forbin government is not entirely in control of everything that goes on in their
country,'' Narlock said. ``This may be the work of some kind of rogue faction.''

Marshall tilted his head slightly in a very chortak-like gesture. ``In any event my counter part
in Forbin tells me the Prefect strongly denies having anything to do with this. She believes
him.''

``The Prefect can be very convincing when he wants to be,'' the First Citizen said. ``Are you
sure your ambassador can always tell the difference between truth and deception?''

Narlock turned to the human. ``You must have some way to track this\ldots\ Ra'hel.''

``I wish we did. Most of us on the ground have an electronic device implanted in our bodies that
can be tracked from space. Unfortunately Rachel has always refused to wear one.''

``What of the devices taken from the canal?''

``Despite the superficial damage they both function normally. They can be used to record visual
and auditory information but apparently Rachel didn't think to activate those features.''

``Or was not able to activate them,'' Narlock said.

``The best we can do is surveillance from space. Yet Rujar is a big world and even knowing the
general area where she might be, we still don't know exactly what we're looking for.''

Narlock swayed his head. ``I can keep you appraised of what we do know. Any information you can
provide would be helpful.''

The group sketched out a general plan of collaboration but before long there was little left to
say. With a nod from the First Citizen everyone got up to leave. Marshall hung back with the
First Citizen and closed the door after the others left. He paused slightly and then turned to
look at the leader of Argenia.

``We've noticed the increased activity of your military.''

``I see.'' The first citizen raised himself off his seat with his trunk.

``I can tell you that Forbin is responding in kind.''

``That does not surprise me.''

Marshall bowed his head slightly and sighed. ``It's not worth it. She's just one person and we
don't yet know what this is all about anyway.''

``We are only preparing ourselves in case Forbin acts aggressively.''

``If you're not careful you might prepare yourself into your next war.''

The First Citizen stepped forward with a fluttering growl. His trunk thudded heavily on the
table. ``I don't need to explain myself to you. We may be simple-minded creatures in your view,
but we've been dealing with Forbin far longer than you have.''

Marshall nodded slightly. ``Yes, very true, and you're right. I don't mean to offend. I just
hope we find her before this gets out of control.''

