
% SUMMARY: The First Citizen admits to the humans that they lost a nuclear weapon.

The atmosphere in the First Office was resigned and yet proud. Fotkey stood tensely beside the
First Desk while his commander chatted with the Second Citizen. Yutaria and Paraz, the head
scientist in charge of the project, both lay on the plush seats that lined the First Office.
Tenar Hutlan, the civilian lead of the Armed Services Group in Parliament, stood uncomfortably
on the rug in the center of the room.

``For the record,'' Hutlan said, ``I disapprove of having the human academic at this meeting.''

``We know you disapprove of this meeting entirely,'' the Second Citizen replied. ``Marshall
wanted her to come.''

``Marshall doesn't know what this meeting is really about,'' Hutlan said. ``I just think it is a
bad idea to have unnecessary people\ldots\ or humans\ldots\ knowing about this.''

Paraz had been listening to the conversation and spoke up. ``I'm looking forward to meeting the
creature,'' he said.

Hutlan checked his watch and paced impatiently.

``Don't worry, Ten,'' the First Citizen said. ``They are very punctual.''

True to his word only moments later the phone on the First Desk rang. The First Citizen picked
it up. ``Yes,'' he said into the receiver, ``please send them in.''

Only the First and Second citizens, and Fotkey, had seen the humans face to face before. The
others shifted nervously and Hutlan took a step back from the door. When it opened, in stepped
Ambassador Marshall and behind him a shorter human with completely black hair.

The First Citizen stood up and came around from behind his desk. He extended his trunk
confidently. Marshall stepped forward and reached out his arm, grasping the First Citizen's trunk
in his hand as the First Citizen wrapped his tenticles around Marshall's arm.

``It's good to see you again, Tom,'' the First Citizen said warmly.

``And you also, FC,'' Marshall replied.

The First Citzen introduced all the dunari present and then Marshall introduced Rachel as well.
``We are well met,'' she said softly. She nodded her head in acknowledgement, dunari-style, and
clicked her teeth slightly.

Marshall looked around the room. ``I was expecting Mr. Formand as well.''

``Unfortunately this meeting is a little more sensitive than the ones we usually have,'' the
First Citizen explained. ``Mr. Formand does not have the right clearances.''

``I see,'' Marshall replied. He glanced toward Yutaria. ``Has Narlock been taken off the
investigation also?''

``That is part of what we are here to talk about,'' the Second Citizen said.

There was a pause and the First Citizen sighed. He glanced briefly around the room and said, ``I
suppose I should begin.'' He returned to the First Desk and lay down behind it, putting his
trunk on the desk. He looked tired.

``In order to protect ourselves from Forbin aggression we have been working on a secret weapons
program at our base in the Rangard Desert.''

Marshall squinted slightly.

``Of course we do weapons research all the time, as do the Forbinites, but the technology we've
been investigating at Rangard is, well, considerably more powerful than anything we've developed
previously.''

``I understand,'' Marshall said slowly. Everyone in the room, including Rachel, watched the two
of them closely.

``We manufactured several devices and were in the process of distributing them to bases where
they could serve as an effective defense. However, well, one of the devices was stolen in
transport.''

``Stolen?'' Marshall said softly.

``Yes, apparently,'' the First Citizen said.

``What are we talking about, exactly?'' Rachel asked.

Marshall turned toward her. ``The Argenians have been building nuclear warheads.''

``Really?'' Rachel said with shock. ``Nuclear warheads?''

The room exploded with commotion as nearly everyone started talking at once. Only Fotkey and
Paraz remained silent. Paraz watched Rachel closely, studying her every move.

``How did you know?'' Hutlan demanded.

``The radioactive debris spread across the Rangard Desert is visible from space,'' Marshall
replied. ``Based on its decay signature we can even estimate how many tests you made and when.''

Paraz shook his ears vigorously. ``You see,'' he said mostly to Fotkey across the room. ``It is
as I told you.''

``I didn't know,'' Rachel said. She turned to Marshall, ``Why didn't you tell me?''

``It's all in the survey reports,'' Marshall said. ``Check the gamma ray map.''

``What I want to know\ldots'' Hutlan began, but Yutaria interrupted him.

``So you have equipment that can detect this kind of radiation from space?''

The room fell silent and everyone waited intently for Marshall's response. He chuckled slightly.
``I see,'' he said, ``you are hoping we can help you find your stolen weapon.''

``Yes,'' the First Citizen said. ``That is what we hope.''

``How many of these weapons have you made?'' Rachel asked.

``That's a detail I think we'd rather keep to ourselves,'' Fotkey injected.

Rachel ran her fingers through her hair in obvious frustration. She turned to Marshall. ``Tom,
are these just fission weapons or\ldots'' Marshall raised his hand abruptly.

``We can discuss this more later,'' he said.

Paraz stared intently at her. ``Fission weapons or\ldots\ what?''

The room fell silent again. Rachel looked at Paraz and then at the First Citizen. She turned to
Marshall. Suddenly she spoke to him in the human language. The alien chittering clicks and
hisses made it abundantly clear that the humans were not of this world.

Marshall did not look at her. ``Argenian, please, Rachel,'' he said.

She tossed her head slighly in a very un-dunari way, and made a groan of frustration. Then she
spoke again. ``These people are playing with fire,'' she said. ``They have to be told\ldots''

``It is not ethical for us to interfer with their choices,'' Marshall said.

``But let them make informed choices,'' Rachel replied, ``let them learn from our experience.''

``You are afraid,'' Hutlan said with satisfaction. ``You are afraid of these weapons we are
building.''

``I fear for your world,'' Rachel said.

She stepped forward and addressed the First Citizen directly. ``I have lived amoung your people
for many years now. I have seen the goodness in their hearts. I have seen the beauty of your
world and of your culture. Yet now you are building a technology that will soon give you the
power to destroy all life on this planet. I implore you, I beg you, turn away from that path
before it is too late.''

The silence was absolute. The First Citizen looked directly at Rachel with a deep intensity. Yet
the tension in his body was obvious. Marshall stepped forward and put his hand on Rachel's
shoulder. She turned toward him. ``Tom, we must get involved. If these people destroy themselves
while we do nothing then the burden of ethical responsibility for that disaster would fall to
us.''

``I agree,'' he said. ``We can start by doing what we can to help them find the missing
weapon.''

The First Citizen sat back again and the room relaxed. Paraz was lost in thought.

``That's good news,'' Yutaria said.

``Don't hope for too much,'' Marshall warned. ``We can't work miracles and we are far from home
with limited resources.''

``Anything you can do would be very much appreciated,'' the First Citizen replied. ``And,
Ra'hel, I will\ldots\ reflect on what you have said.''

``There is something else,'' the Second Citizen said.

``Yes,'' said the First Citizen. ``Yutaria, why don't you take it from here.''

Yutaria stood up. ``Our investigation turned up evidence that the thieves were Forbin military
personnel. However, certain details of that evidence are suspicious. In fact, we believe that
they might actually be Junar terrorists posing as Forbin officers.''

Yutaria looked directly at Rachel as he spoke.

``Do you think the Junar who abducted me were part of the same group?''

``Yes. The Junar believe that there will be a `Last War' between our nations that will herald in
a golden age. They believe it is their mission to start that war. If you were kidnapped and
killed by Forbin military, it would greatly escalate tensions between our nations.''

``Good Jurita,'' Rachel said softly. ``Do you think the Junar might actually try to use the
weapon?''

``Unfortunately we do think that,'' said the First Citizen. ``We know the Forbinites have their
own fission weapons program. We think the Junar might use the weapon against an Argenian target
and try to frame the Forbinites for it.''

``This certainly increases the urgency of our search,'' Marshall said.

``You should talk with Narlock also and combine what you know,'' Rachel said.

``That would not be good for the security of the fission weapons program,'' Fotkey said.

``It would be unthinkable,'' said Hutlan.

Rachel paused and them smiled slightly. ``Maybe you should go public.''

``What? No. The people would panic if they knew such weapons existed.''
