Gartosk Mundon wasn't sure why his day needed this interruption. It's not like he was
doing anything extremely important. However the first hour of the day was always his time to
feed the animals and he didn't like disturbing his routine. It was, perhaps, strange for
someone in his position to take care of such a menial task. Yet he found it relaxing. He often
had his best ideas during that time and, frankly, he enjoyed working with the animals.

Today, however, he had to relegate the feeding to his graduate students. They knew the procedure
well enough. He just hoped they remember Trison's medicine. Gartosk looked at the clock
impatiently. It would be just like government officials to keep him waiting. He tried to
entertain himself by counting the tiles along the wall. He could probably estimate the volume of
the room without even moving from his chair.

He didn't have time to get far with his estimate. These government officials were punctual.
He heard voices in the hall and then in walked Dr. Dargon, chair of the zoology department,
followed by Barnon Lacasa and his personal secretary. Gartosk got up and shook trunks with
Lacasa and with the secretary.

``I'm very pleased to meet you, Mr. Lacasa,'' he said. He even meant it. He didn't
agree always agree with Lacasa's politics, but he seemed competent. At least he did on the
news.

``The pleasure is mine, Dr. Mundon,'' he replied.

``I hope these facilities will be satisfactory,'' Dargon said. ``We could retire to my
office if you'd rather.''

``This is fine,'' Lacasa said. Everyone sat down around the oval table in the middle of the
room. Lacasa's secretary took out some papers and a notepad. Dargon looked nervous. ``It's a
beautiful day today,'' Lacasa said. ``I like the warm weather myself. The campus looks very
nice.''

``Yes,'' Gartosk repled. ``They do a good job with the grounds.''

``I'm not sure,'' Lacasa continued, ``how much Dr. Dargon has told you\ldots''

``Only that this has to do with the humans,'' Gartosk replied.

``Yes,'' said Lacasa. ``We've had an unusual request from them, and we thought you
might be interested.''

``Oh?''

Lacasa paused. He took one of the papers sitting on the table next to his secretary and slid it
over to Gartosk. ``Apparently,'' he began, ``one of their scientists is planning to visit
Rujaria. This individual is an astrophysicist and would like to be on hand during the
construction of the telescope.''

``I see,'' Gartosk said.

``She is a kind of professor as well. I understand that she teaches at the human equivalent of
a university. So that means she would be, in effect, a visiting professor.''

``She?'' Gartosk repeated. ``A female?''

``That's right.''

Gartosk chuckled slightly. ``That's interesting,'' he said softly. Lacasa studied him.
``So she is a visiting astrophysicist. That part sounds rather mundane. What does this have to
do with me?''

Dargon jumped in. ``Of course it is our tradition for a faculty member to host visiting
professors.''

The two of them looked at Gartosk.

``Now wait a minute,'' Gartosk said. ``Are you suggesting that I host this\ldots this human?
Are you suggesting that I bring it into my household\ldots with my family?'' Gartosk didn't
like the sound of this. It was all very well to see the creatures on television but who knows
what kind of bizarre, disgusting practices they engaged in privately.

``That was our thought,'' said Lacasa.

``Well\ldots ah\ldots''

``Look,'' Lacasa said. ``You don't have to decide right now, of course. Give the idea
some consideration and decide in a couple of weeks. But think of the opportunity it gives you to
study these creatures closely.''

``Oh you want me to study it?'' Gartosk said. ``This is because of my work with chortak,
isn't it?''

``We thought you might find the idea exciting,'' Dargon said. ``Who else in the world
better than you could assess their physiology or evaluate their social system?''

``That's crazy,'' Gartosk said with some agitation. ``For one thing\ldots they are not
chortak. What do I know about human physiology, despite any parallel evolution that might have
occurred? And anyway, I'm sure this human astrophysicist is very smart. She will no doubt
guess that I am supposed to spy on her. That's not the way a good host behaves!''

``No, no,'' said Lacasa. ``Don't spy on her. That's not what we are saying. Anyway\ldots
she asked for you.''

``What?''

``It's in the letter.'' Lacasa pointed at the paper sitting in front of Gartosk.

He picked it up and looked at it. ``Is this something she wrote?'' Gartosk asked.

``Yes, apparently.''

Gartosk chuckled again. ``She read my book. I didn't realize I was selling copies on
Earth.''

``Gar\ldots'' Dargon began. ``I know this is a lot to ask, and I'm not saying that you
should accept this. But I do ask you to think about it. Of all the Rujaria I know, I can't
think of anyone who would be a better host. Remember that this human is an academic like us. If
nothing else, you could talk about teaching.''

Gartosk wasn't really paying attention to Dargon. He was reading the letter.

``So when is this\ldots Rachel Spencer\ldots going to be here?'' he asked.

``Our understanding is that she is en-route,'' Lacasa replied. ``But it will take her
several months to get here. It is, after all, a long way.''

``That is for sure,'' Gartosk agreed.

[*****]

Gartosk's office was in the ``animal house'' as his students like to call it. At the edge
of campus it was a small building with a few, crowded offices, a classroom or two, and, of
course, the animals. It was sometimes a dirty, smelly place, but Gartosk found it comforting. He
had always loved animals. He had pets for as long as he could remember. As a child he frustrated
his parents by bringing home every injured animal he encountered only to end up making most of
them pets. Most, that is, except for the brunack. He still remembers the look of horror on his
mother's face when he came home with that 20 pound ball of teeth and claws.

At first Gartosk put the letter on his desk but then he thought better of that and stuffed it in
his case. Then he went down the hall to visit the animals. Narbo was there soaping down Trison.

``Hey, Dr. G,'' Narbo said.

``How's he doing?'' Gartosk asked, pointing at the injured chortak.

``Oh fine. I think it's healing.''

Gartosk took a look at Trison's arm. The skin just above his right wrist was inflamed and
raw.

``Maybe a little,'' Gartosk said.

``The cages upset him,'' Narbo said. ``I caught him this morning still trying to grab the
upper bars.''

It was always difficult when a new chortak arrived at their lab. They couldn't put him with
the others right away until after they were properly introduced. Chortak have an elaborate
social system and won't accept outsiders if they are just thrown together. Yet being on their
own without the support of others was a trying experience for the creatures. It wasn't
unusual for an isolated chortak to compulsively rub their fur off or even chew wounds in their
own skin.

``Maybe we should open the upper section now,'' Gartosk said. ``That might help.'' The
two of them scrubbed down Trison with soapy brushes. It was a process the animal seemed to
enjoy. He grabbed Gartosk's trunk with his hand as if to keep him from leaving.

``He likes you, Dr. G,'' Narbo said. The two finished the bath together and then while
drying off, Narbo asked about Gartosk's meeting.

``So what was that all about, anyway?''

``A human astrophysicist is coming to this university,'' Gartosk said. ``They want me to
host it as a visiting professor.''

``Oh wow!'' Narbo said. ``That's really great!''

``That's easy for you to say,'' Gartosk replied. ``She won't be staying at your
house.''

``Oh, a female! How interesting. Imagine\ldots a chortak you can have a conversation with.''

``Humans are not chortak,'' said Gartosk, ``despite the resemblance.''

