
Colty was almost late when he arrived at the Cogart lecture hall. The room was filled with
students, it seemed like a huge class. Colty scanned the crowd and saw Rocalla in the back with
his trunk raised beckoning to him. Colty found his way to the back of the room and lay down next
to Rocalla.

``You have large sections in the physics department,'' Colty said.

``Not like this,'' Rocalla replied.

A student laying on the other side of Rocalla leaned forward and spoke to Colty. ``A lot of
them are not registered for the course. They come to hear her lecture.''

``This is the class assistant, Denart,'' Rocalla said. Colty and Denart shook trunks.

Colty looked around the room. Many of the students were talking among themselves but the
atmosphere seemed light. He could hear animated voices and see the occasional ear shaking. Some
students were standing beside others and not yet in their seats. The room was filled with the
din of chattering talk. Rocalla did not look particularly pleased.

There was no podium on the platform. Colty could see that it had been moved off the side and
stood pushed up against the wall and out of the way. Instead on the platform there was just a
plain metal table and nothing more.

Colty checked his watch. It was only moments before class. ``She's going to be late.''

``She's never late,'' Denart replied.

At almost exactly the appointed time, the door swung open and in walked Rachel. She did not look
at the class but instead bounded up the platform and stood behind the table. She didn't have the
official course book with her. In fact she didn't have any papers of any kind. She turned and
raised one arm high above her head, curling her fingers into a tight ball. She did not say a
word but instead just waited.

The talking stopped and the students who were standing quickly found their seats. Many students
opened their notebooks. Colty could see Rachel scan the room but she hardly moved. She held her
fist above her head steadily until everyone was seated and the room was nearly silent.

Then with a swift motion she brought her hand down, slamming it hard against the table and
sending a explosive, reverberating bang throughout the hall. She put her whole body into it,
bending her legs slightly to increase the power of the blow. The sound startled everyone. Colty
jumped slightly and stretched his legs instinctively. There were several short cries of
surprise.

Rachel stood up straight again, letting her arms relax at her sides.

``Why didn't my hand go through the table?'' she asked.

A voice blurted out from the right side of the hall. ``Because you didn't hit it hard enough!''

Everyone in the room shook their ears vigorously and Rachel made the strange, short breathy
sounds that she did when amused. She walked to the edge of the platform and held out her hand in
the direction of the voice. ``That's good, Mr. Vernodar,'' she said, ``but not the answer I was
looking for.''

She returned to the table and faced the class. Putting on hand firmly on the table she pressed
down. ``Think about what is happening now between the surface of this table and my hand. If you
could see at an atomic scale what would you see?''

Rocalla leaned toward Colty and spoke softly. ``I thought this was a lecture about
electrostatics.'' His ears were firmly against his head.

Several students raised their trunks and Rachel picked one of them. ``Go ahead Yartak,'' she
said.

``You would see the atoms all smooshed together,'' the student replied.

``I don't know this word, \textit{smooshed},'' she said.

Several students shook their ears and one called out, ``he means pressed together tightly.''

Colty leaned over and spoke softly to Rocalla. ``Does she always talk with her students like
this in class?''

``All the time,'' Denart replied.

``Yes,'' Rachel was saying, ``the atoms would be pressed together. Yet atoms are mostly empty
space. Why don't they just pass right through each other?''

There was a moment of silence and then one student raised his trunk. Even before Rachel could
properly call on him he blurted out, ``The negatively charged electrons repel each other.''

Another voice called out, ``If it were that simple, materials would just fall apart. We wouldn't
have chemical bonds.''

``But there are postively charged protons in atoms too.''

``They are shielded by the negative electrons. Two atoms approaching each other would just see
the electrons and be repelled.''

``Then why do we have solid objects at all?''

The conversation flowed back and forth between a handful of students. Rachel simple held her
arms in front of her body and listened attentively without saying a word. She did nothing to
correct or clarify but instead let the students try to work it out. Finally as the discussion
started to die down she spoke again. ``When things are confusing it is often best to consider
simple cases first. Hydrogen is simple. It forms molecules of two atoms but if a third atom
comes along it does not bind to the two. What is happening there?''

Rocalla was busy making some notes on a pad he had in front of him. ``This is outside the
prescribed sequence,'' he said more to himself than to anyone.

``She hasn't brought that book to class in ages,'' Denart said, ``thank Jurita.''

A different student was speaking. ``The electron from one atom is attracted to the proton in the
other but when a third atom arrives the two electrons in the molecule create too much
shielding.''

``That's an interesting idea,'' Rachel said.

``Is it correct?'' the student asked.

``It's not important right now if it is correct or not,'' she replied. ``Let's take a step back
for a moment and ask: what fundamental force of nature are we talking about here?''

There was a pause and then a student in the front row blurted out, ``Electric!''

Rachel seemed pleased. ``Yes. The details of how chemical bonds work are complicated but
ultimately it is the electric force that defines their strength and structure. It is the
electric force that gives material objects their rigidity, that acounts for the pressure of
gases, and that explains the energy that comes out of chemical reactions.''

