
% SUMMARY: Karn tells Gat that they have a job to do.

The car drove along the road between row after row of langolia plants. A cloud of dust rose
behind it and dissipated in the warm Nermela. At the end of the road was a modest farm house
like an island in a sea of langolia. The house was surrounded by a small lawn with some toys
strewn about the front yard under the shade of a large rigat tree. It was reasonably well kept
with freshly painted trim and new screens in the windows. Yet the shingles on the roof were
crumbling and the porch boards were dry and cracked. Behind the house rose a large barn, full of
drying langolia and a few simple pieces of farm equipment.

The car came to a stop beside the lawn, the dust of the road enshrouding it momentarily. Karn
got out. He was wearing expedition clothing that was cool, comfortable, yet eminently
functional. He walked confidently up the steps to the porch and knocked on the front door with
his trunk. The door opened almost at once. A older man wearing rough farm clothes looked at the
arrival and then down the road at the settling cloud of dust.

Karn just swayed his head briefly. ``I'm clean.''

The older man swayed his head in reply and ushered his guest into the kitchen where another,
much younger man was laying beside the table. Two children played with toy animals on the floor.
They looked up and the younger man's wife who had been standing at the counter called to them.

``Let's go outside,'' she said. ``Time to let the men talk.''

Karn lay down and helped himself to a cup of water from a pitcher in the center of the table.
The men waited while the young wife scooted her children out the front door. Once they were out
of the way Karn reached into his pouch and slapped a small piece of paper on the table.

``The time has come,'' he said. ``We have word from the seleck of Junarit.''

The old man slid the paper to him and eyed it carefully. ``This is crazy,'' he said at last.

``What is it Pap?'' the younger man asked. His father didn't reply directly but let him take the
paper to see for himself.

``It is the word of the Junar,'' Karn said. ``It's not our place to question why.''

``Maybe. But the creature ain't got nothing to do with it. Jurita only knows what they'll do.''

``It's a strategic move,'' Karn replied. ``I understand it.''

``We ain't done no training for this.''

The younger man let his trunk fall to the table in a gesture of annoyance, the paper slipping
from his tentacles. ``Paps, we've trained plenty. We're ready for this. We're ready for
anything.''

``You only think so because ya ain't ready for anything.''

Karn put his empty cup on the table. ``I agree with Gat,'' he said. ``We might not be ready for
anything but we are ready enough.'' The old man sputtered and waved his trunk in frustration,
but Gat just nodded.

``The humans a wildcard,'' the old man said. He pointed his trunk at Gat to emphasize his words.
``Theys the anything ya ain't ready for.''

``It's alright, Paps. We can do it.''

The old man banged the table with his trunk and grunted.

Karn took the paper and returned it to his pouch and the pulled out a packet tied off with a
string. reached into his vest and pulled out an envelop. He handed it to Gat. ``Read this and
then burn it. You should find the information on our target useful.'' Gat started looking at the
papers. Karn put his trunk on the old man's shoulders. ``This is our moment,'' he said. ``This
is when we finally set the world on the path to righteousness.''

The old man was not consoled. ``You take my boy on a suicide mission. I ain't never forgive
ya.''

``It'll be fine, Paps,'' Gat said. ``Doesn't look hard. We'll be home for the next harvest,'' he
added with a click of his teeth.

Karn clicked his teeth and then stood up. ``I must be getting back..'' Gat escorted him to the
door and stood on the porch as Karn said good bye to his wife and children. He climbed back into
his car and soon was nothing more than a dusty plume receeding into the distance.

The old man grunted and pushed his way through the door, walked down the creaking steps, and
then around back toward the barn. Gat's wife went to her husband. ``What's going on?'' she asked
in a hushed voice.

``We're gonna act,'' Gat said. He continued looking down the road, staring into the distance.

``Good Jurita,'' his wife said with a gasp.

``It's alright, \textit{tinka},'' Gat said. He wrapped his trunk around her shoulders and pulled
her close to him. ``This is the beginning of a better world for our children. A world were
Jurita matters again.''
