
Narlock entered the interview room and threw a pad of paper onto the table. Laying on the other
side was a young agent he had never seen before.

``You are the lip reading specialist?'' Narlock asked.

``Yes, Sir,'' the agrent replied nervously. ``Tolanor Haglost.''

The two agents clasped trunks and then Narlock lay down across from him. Tolanor shifted
uneasily and looked at the clock. Narlock studied him carefully. ``They are usually very
punctual,'' he said.

``Yes, Sir.''

``Relax. They're trying to help us.''

``Yes, Sir.''

At that moment the phone on the table rang. Narlock picked it up and listened briefly. ``Yes,
fine,'' he said into the device. He hung up and turned to Tolanor, ``They're here.''

Narlock stood up and Tolanor followed his lead just as two humans entered the room. One was
ambassador Marshall. Narlock did not recognize the other but he could tell it was a female. She
had much darker skin than the ambassador and very dark, almost black eyes.

Marshall introduced himself to Tolanor who awkwardly muttered a reply. He then introduced his
companion as Jaya. ``She is here for technical support but she doesn't speak much Argenian.''

``Jaya is a dunari name,'' Narlock said with surprise.

She nodded her head. ``Yesss,'' she replied with a heavy accent. ``Alssso name mine.''

``Shall we get started?'' Marshall said. He nodded toward Jaya and spoke to her briefly in their
language. Narlock lay back down and Tolanor, who had been standing stifly through the
introductions followed.

Jaya unfastened a long tube from a strap over her sholder and then, opening one end, pulled out
a thin rolled up sheet of some dark material. She unrolled the sheet and layed it out flat on
the table. It seemed to stiffen on its own and when she lifted it back off the table it had
become a hard, clear rectangular piece about the size of a small window. From her pack Jaya also
took some soft material, rather like modeling clay, and used it to make a kind of stand. She put
the stiffened sheet into the clay so that it stood upright on the end of the table before them.

Jaya kneeled down next to the strange device and stared at it. Suddenly it came to life. After a
few moments it showed an image no dunari had ever seen before: Rujaria from space.

``This device is just a display,'' Marshall explained. ``It gets information from \textit{Golden
  Light}. Jaya here knows how to operate the device and get it to show us what needs to be
shown.''

``She's not even touching it,'' Tolanor objected.

``She communicates with the device using an electronic interface implanted in her brain,'' said
Marshall. ``It's quite convenient.''

``Yes, yes,'' Narlock said. ``That's very fascinating, I'm sure. What information do you have
for us?''

Once again Marshall spoke to Jaya in their language. She nodded and turned her attention back to
the device. The group watched as the image of Rujaria started to grow. Quickly at first and then
more slowly they zoomed toward the surface. As they approached the ground more and more details
could be seen. Narlock recognized the Argenian and Organite seas. He recognized (some other
things).

``This is recorded image information,'' Marshall was saying. ``We located T1 getting into a car
in X1 and heading west.''

When the zoom stopped they could see a small vehicle moving down a dusty road. The view panned
along the road, following the vehicle. Jaya watched the display intently but didn't move.

Narlock was unimpressed. ``How do you know T1 is in that car?''

``You'll see,'' Marshall said. He spoke to Jaya again. The action seemed to speed up. The car
was now rocketing down the road at an unrealistic speed. After a few moments it came to its
destination and the action returned to its normal pace. They could see a small figure get out of
the car and walk across the desert sands to a group of several other small figures gathered
around some kind of stone ring.

``Jaya?'' Marshall said. She nodded her head again. The action stopped and the view again
started to zoom in. It was as if they were descending on the gathering from above, falling from
the sky upon the backs of the people standing there. As they approached even more detail became
visible. The ring resolved in a collection of stones. The color of the clothing could be
discerned. Their footprints in the sand could be seen as well.

``The optical array has multiple receivers,'' Marshall said. ``So we can see the same scene from
several angles. Using information from all receivers allows our computing devices to reconstruct
a three dimensional view.''

It was now as if they were hovering right over the heads of the dunari, all blissfully unaware
that every detail of their movements could be seen by aliens in space. Suddenly the view lowered
even more. The camera appeared to turn and soon it was as if they were standing there themselves
right next to the men on the sand. There was no doubt about it now. One of them was T1. It was,
in fact, a better picture of him than one normally saw on television.

Suddenly the action continued and the characters all started to move. There were occassionally
strange hesitations and several times sections of the display were briefly blanked out. ``Seeing
this much detail from 100,000 karnons isn't easy,'' Marshall explained. ``The optical array was
designed for astronomical targets so we had to make a lot of changes to our systems to do this.
Canceling the atomospheric turbulance in the image data was a challenge, right Jaya?'' She
didn't seem to notice. She watched the screen intently.

The image focused on the leader (who?). When he started to speak the view of his lips was clear.
Tolanor sat up.

``Now I'm impressed,'' said Narlock.
