
Narlock and his officer walked into the Parliament East Wing where various political observers
kept their offices. They drew a few eyes as they went down the hall but for the most part they
were inconspicuous. They soon came to the door of the Limzar party office.

Narlock took a breath and glanced toward his officer. ``Be alert,'' he said, ``don't
underestimate him.''

They entered and two interns watched them carefully as they approached the reception desk.

``It's important that I talk with Zarlon right away,'' Narlock said.

``Is he expecting you?''

``No.''

Narlock waited for a moment and when it was obvious that he wasn't going to leave the
receptionist picked up her phone and called into the inner office. In short order Narlock and
his officer were invited inside.

``Narlock!'' Zarlon greeted him as he stood up and came around to the front of his desk. ``I do
hope this is good news about the investigation.''

``I think so,'' Narlock said.

``Would you like something to chew while we talk?''

``Not now. This is strictly an official visit.''

``I understand.''

Zarlon stretched his trunk toward some seats where all three of them could lay but Narlock did
not move. Instead he regarded Zarlon closely. ``By the authority vested in me by the Internal
Investigation Bureau, I hereby place you under arrest for ordering the destruction of
Yartagard.'' He nodded to his officer who stepped forward with a trunk binding ready.

``What?'' Zarlon clicked his teeth. ``Do you have any evidence to support such an outlandish
claim?''

``All evidence will be presented according to due legal process,'' Narlock replied formally. He
nodded to his officer again.

``If you please, sir,'' the officer said holding the trunk binding toward Zarlon.

``This\ldots\ this is outrageous,'' Zarlon said. ``I'm sure there is a misunderstanding. I'm not
guilty of even a parking violation let alone this\ldots\ this horrifying crime.''

The officer took Zarlon's trunk and begin to slip it into the binding. Narlock opened his arrest
order and held it up so Zarlon could see it.

``Are you sure you know what you're doing, Narlock?'' Zarlon said. ``You know I could make your
life miserable once this is all cleared up.''

Narlock shook his ears slightly. ``I'm not concerned.''

The officer tightened the strap of the binding around Zarlon's left front leg and hooked the
loop through his own trunk.

``Listen, this is an unnecessary embarrassment to my office,'' Zarlon continued. ``I'm happy to
accompany you to the IIB center where we can get this cleared up but for the sake of my staff
can we do it without the binding?''

The officer looked at Narlock.

``No.''
 
Zarlon folded his ears back tightly and snorted as the officer led him back to the outer office.
The interns and the receptionist all stood up in alarmed surprise to see the leader of the
Limzar party in bindings.

``Call my legal staff,'' Zarlon barked.

