
The campus security office at Parnon was a small, round building at the edge of the new campus
just off Blakart Road. It had a radio tower on top and a garage in the back for the security
cars. Inside were a few offices, the call dispatch center, and some interview rooms. It had all
the usual amenities of a big city police office, only less of them.

Ordinarily Narlock would not have agreed to doing debriefing interviews in such a place. He would
have much preferred the IIB offices in Varnok. Under the circumstances, though, the less the
human had to move the better. Parnon police seemed to be doing a good job at setting up
protection, at least based on what Narlock saw in their reports.

Narlock was greeted politely, even if a bit coolly, by the Parnon and University authorities. They
knew this matter was outside their jurisdiction but that didn't mean they liked it. He was shown
into the room where the human was waiting along with the human ambassador.

The two creatures were sitting quietly beside the table in a strange, upright posture. Their
legs were folded underneath their bodies and their backs were aligned vertically. They looked up
when Narlock entered the room.

``Hello,'' Narlock said.

``Hello,'' the female replied. It's voice was surprisingly soft and airy.

``Hello,'' the ambassador said. It's voice was deeper but had a similar airy quality. ``I'm
ambassador Marshall and this is Dr. Spencer,'' it said.

Narlock lay down beside the table and spread out some papers from his pack in front of him. ``I'm
the investigator from the IIB,'' he said after introducing himself. ``I specialize in matters
related to terrorism.''

``I understand,'' Dr. Spencer said.

Narlock flipped through his papers briefly and found the one with the questions on it. He found
it very uncomfortable being the only dunari in the room. ``We have a statement from Dr. Harnick,
but we are glad to be able to interview you as well, Dr. 'pen'er.''

``I want to help.''

``I have a general idea of what happened, but can you go over it again in your words?'' Narlock
took a small tape recorder from his pack. ``It's ok if I record this, yes?''

Rachel went over the events as best she could remember them starting from when they were
kidnapped to their eventual rescue. Narlock made some notes as she talked but relied on the tape
recorder to catch most of the details. The ambassador listened carefully to the proceeding but
said nothing.

``That sounds pretty much the same as what Dr. Harnick remembers as well,'' Narlock said when
Rachel had finished. ``I wanted to ask about a few details to help me understand the situation.
First how were you able to locate the pistol in the darkness of the storage closet?''

``I come from a world with alternating day and night. In the dark my eyes become 100 times more
sensitive. There was enough light for me to see.''

Narlock clicked his teeth. ``So you were not really frightened of the darkness, that was a
rouse.''

``Yes.''

Narlock made a few notes. ``Dr. Harnick says you jumped out of the second story window so that
you could attack your captors from behind. Is that true?''

``Yes.''

``You didn't sustain any injury doing that?''

``No,'' Rachel said. ``My species is descendant from tree climbers so a certain amount of
jumping is normal for us. Also the gravity on Rujaria is only about 7/8 of that on Earth and I
was jumping into sand. It wasn't a problem.''

Narlock paused. ``It was lucky for you they didn't know any of that.''

``Yes.''

Narlock continued. ``Dr. Harnick's statement says that you disabled both of your captors but then
went out of your way to kill them despite that. Why did you do that?'' The ambassador shifted
slightly in his seat and then spoke to Rachel briefly in the human language. Rachel didn't seem
to pay any attention to what he said.

``I\ldots\ I was frightened,'' she said finally, lowering her head. ``I was worried that they
would maybe sneak up on us again. I wasn't sure how serious their injuries were.''

Suddenly Narlock felt his heart go out to the strange creature before him. ``It's fine,'' he said
reassuringly. ``You did what you had to do. I'm just trying to understand what happened.'' He
looked at the ambassador but he couldn't read him at all.

``Did that second one really have two children?'' Rachel asked.

Narlock studied her but he couldn't read her either. ``He did, yes.'' Rachel sighed.

Narlock flipped to another sheet of his notepad. ``I want to ask you about the note you found.
What prompted you to go through their pouches?''

``I knew they were not Forbinites,'' Rachel said.

``How did you know?''

``One of them, the really evil one, didn't seem to understand Forbin. He didn't talk like a
Forbinite and I tried speaking some Forbin to him but his reaction didn't make any sense.''

``What did you say?'' Narlock asked.

``I just said 'The grass is very tall.' I don't really know much Forbin. Yet he got angry and
acted as if I had insulted him.''

Narlock made some more notes. ``So you wanted to find out more and that's why you searched
them?''

``Yes.''

The ambassador leaned forward. ``Are you going to be able to use that note in any legal
proceedings that might come from this?''

``I doubt it,'' Narlock said. ``It's gone through your trunks and the defense would surely ask
for it to be declared inadmissible. I don't know if that would work, but I think probably.''

``We did a genetic analysis of the blood on the note,'' the ambassador said. Narlock folded his
ears back against his head. ``If we get some biological material from a suspect we can tell you
if it's the same person.''

``What kind of biological material?''

``It doesn't matter. It could be a bit of blood or a scale or a toe shield.''

Narlock made a few more notes but mostly for show. He wasn't comfortable with the turn the
conversation had taken. ``I can tell you that any evidence derived from human technology will
never be accepted by our courts.''

``I understand that,'' the ambassador said. ``Yet the information would still be correct and you
might find it useful in guiding your investigation.''

Narlock admitted to himself that made a certain amount of sense. Despite his reservations about
the humans, if their technology helped the IIB track down the Junar leaders he couldn't afford
to ignore it. ``I'll keep it in mind,'' he said finally.

Narlock had several more questions about the weapons the terrorists had used. He asked mostly
about how they held them and how they wore them. He also wanted to know more about why Rachel
felt they hadn't talked like Forbinites. Rachel did her best to answer all the questions fully.

As he finished up the interview, Narlock gathered his papers together. The ambassador spoke. ``I
want you to know that we completely respect your legal process, but that we are willing to help
as we can.''

``Sure,'' said Narlock. He turned to Rachel ``In the meantime you might want to stay out of
Parnon, perhaps you should remain in the compound where it is safer for you.''

``Oh no,'' Rachel objected. ``I can't do that. I have classes to teach.''

The ambassador turned to her and spoke in Argenian. ``I'm sure under the circumstances they can
find someone else to cover the class.''

``No,'' said Rachel. ``I\ldots\ I should finish my classes.''

Narlock leaned forward. ``It is very possible these terrorists, or even Forbinites for real, will
try again. You are in danger being here. That is obvious.''

``The Parnon police department has dispatched extra officers,'' Rachel objected. ``I will be
fine.''

Narlock looked at the ambassador but the human said nothing. ``At least let us give you an IIB
body guard. No disrespect to the Parnon police force but they aren't trained for this sort of
thing.''

``I think that's a good idea, Rachel,'' the ambassador replied.

Reluctantly she agreed to a body guard. ``As long as he stays out of the way.''

Narlock stood up and the ambassador said that he wanted a few private words with Rachel before
they left. Narlock nodded his head and took his leave.
