
``Why do you have to go Tusk?'' Jael cried. ``Narlock I can understand, but you?''

Tusk wrapped his trunk around his wife's shoulders. ``They need a second witness. Besides Lucasa
wants me to find out as much as I can about them and I know \alien{English} better than any
other dunari.''

``So you are a spy now? And what about this medical procedure? How do they know it's not going
to kill you?'' Jael swung her trunk and moved away from Tusk to emphasize her displeasure.

``I will be fine, don't worry.''

Tusk embraced his wife again and the two of them stood quietly for several long moments in their
hanaria. Then Tusk gathered his bag, just a few personal items mostly, and crossing trunks with
Jael, left on the most adventurous journey of his life. Yet there were no news cameras, and no
reporters. Not even his neighbors realized what was happening. For security reasons the trip had
been kept secret. Tusk himself had been informed only a few hours before. Just his immediate
family, Jael's seleck, and a few people in the government knew that Tusk would be one of two
dunari to set foot on the alien starship \textit{Summer Breeze}.

For the next three hours Tusk, Narlock, and Rachel were sealed into a small air tight chamber
somewhere inside the human compound were they received injections and pills designed to purge
their bodies of microbial life. Although the biological incompatibility between human and dunari
chemistry made mutual infections virtually impossible, the humans insisted on avoiding cross
contaimination between their worlds. ``We've had some bad experiences in the past,'' Rachel
explained.

% Jernumia should come and be upset to discover that Tusk is half Forbin. This might hurt his
% credibility should the perpetrators of the crime be Forbinites.

``Thisss\ldots\ maaay make you\ldots\ feel\ldots\ sssick,'' a young human said to Tusk in
halting Argenian as he gave him a bottle of small yellow tablets.

``\alien{Thank you, I'll let you know},'' he replied in English.

The three of them talked for a while but as the drugs took effect their conversation died down.
Narlock feel asleep and Rachel took to sitting in the corner and rocking slightly. Tusk tried to
stand up to get comfortable but just felt dizzy instead.

``Relax,'' came the voice of the dunari doctor monitoring them from the other side of the glass.

Tusk moaned slightly.

Rachel smiled a little. ``It doesn't last too long.''

Time passed and Tusk tried to read. When he looked up Rachel was laying on a small bed and
appeared to be asleep. He watched her in fascination for he had never seen a sleeping human
before. He looked at his watch. Just one hour to go\ldots\ eight human hours he reminded
himself. He had to get used to thinking in human terms.

Finally after 24 human hours the four of them were given clearance to leave the life lock. From
the dunari point of view it wasn't soon enough. Even Rachel looked impatient as she stood beside
the inner door. Yet when it finally opened with a slight hiss she invited the dunari guests to
go ahead before her.

The space they entered was not much larger than the space they came from. There were two
``windows'' along the side at about human height and beneath them was some metal framework with
straps and other attachments that Tusk did not recognize. Toward the front of the space there
was a doorway without a door and beyond that point what looked like a kind of cockpit bristling
with electronic devices.

% The ship would have no human pilot. It should be flown entirely automatically.

There were just two other humans, the doctor and someone Tusk assumed was the pilot. Based on
her size and shape Tusk recognized her as a female\ldots\ a woman Tusk reminded himself. Rachel
did the introductions although the humans already knew the dunari. Tusk didn't quite catch the
pilot's name but it wasn't like any of the other human names he knew. He wondered if she had a
different ethnic background from the others.

``\e{Rachel,}'' the pilot said in English, ``\e{can you help them get ready?}''

``\e{Yes, of course,}'' Rachel replied in her own language. As the pilot stepped toward the
front of the aerocar Rachel turned to the dunari. ``When we are traveling inside the
displacement field the ride will be very smooth and you'll feel the gravity of Rujar normally,''
she said. ``But we'll be climbing at a steep angle so you'll feel pulled toward the back of the
shuttle. When we reach altitude we'll switch back to this inertial reference frame and then,
because we'll be in orbit, you'll feel weightless.''

``Weightless?'' Narlock asked.

``Yes, you will float. It's a little strange at first and you might find it alarming because it
feels like you're falling.''

Narlock grumbled slightly. ``I'm looking forward to that.''

``That's where these straps come in,'' Rachel continued. ``You can lay here and hold on to this
framing. We can strap you down which is often a good thing. Also be sure your clothes are
buttoned and you aren't wearing anything loose.''

``I help,'' the doctor said in Argenian. The two humans got the dunari into position with
Narlock at the back and Tusk toward the front.

``\e{Will we be able to see the ground?}'' Tusk asked the doctor in English as he connected
Tusk's straps to the framework.

``\e{Through the window displays,}'' the doctor replied pointing toward the wide displays above
their heads, ``\e{although you might not see much from down here.}''

``Don't worry, Tusk,'' Rachel said in Argenian. ``The view is magnificent from \textit{Summer
  Breeze}.''

When the dunari were situated the doctor and Rachel stood at either side of the space with
their feet in a kind of slot in the floor and their hands wrapped around some grips that were
obviously designed for the purpose. Tusk realized the dunari were strapped in more like some
kind of living cargo. He clicked his teeth slowly and leaned over to Narlock. ``I don't think
they designed this craft for dunari passengers,'' he said softly.

The pilot chittered something but Tusk didn't catch it. Rachel turned and said, ``We are on our
way.''

``How long will this trip last?'' Narlock said out loud.

``Not very long,'' Rachel said. ``The longest part will be aligning with \textit{Summer Breeze}
when we get there.''

There was a soft whine as the engines, or whatever it was that powered the shuttle, came on.
Tusk could feel them starting to move. He looked up at the windows but all he could see was the
green Rujaran sky. It was a reassuring view.

The shuttle was obviously moving toward a runway. The experience was surprisingly ordinary. Tusk
could feel the periodic bumps from the grooves between the concrete slabs that formed the
landing pavement. Narlock tightened his grip on the framework.

Rachel was obviously looking out the window and Tusk envied her height. He was tempted to try
standing up when the craft suddenly stopped. He could hear the pilot talking some but again he
couldn't catch her words. They were probably outside his vocabulary anyway.

``This is it,'' Rachel said. She tightened her grip on the holds and Tusk followed her lead.

The whine of the engines increased suddenly to a kind of ethereal scream. It was nothing like
the sound of an aerocar engine. Yet the effect was dramatic. Tusk felt himself being dragged
backward by the force of the acceleration. Rachel and the doctor leaned against the post that
was conveniently placed and held on with both hands.

The bumps in the landing pavement came faster and faster and in only moments were little more
than rapid clicking. Then, suddenly, they were gone as Tusk could feel the craft lift from the
ground. It banked slightly and he caught a glimpse of the base through one of the windows. Then
the shuttle angled upward steeply.

``Jurita\ldots'' Narlock said softly. Tusk could feel himself leaning heavily into the straps
and he hung onto the frame tightly. Between the angle and the acceleration the experience was
quite intense. Tusk found it rather exhilarating. The shuttle banked again and this time Tusk
could see out across Parnon and even, briefly, the spire of Tartuk Hall.

They continued this way for at least an arnet, maybe two. Then, suddenly, Tusk felt a strange
sensation. It was like his body momentarily expanded a small amount and then contracted back to
its normal size. It was like a wave that seemed to come from everywhere. At that same moment the
whine of the engines stopped and all was strangely quiet. The steep incline of the craft
remained, but it felt more like they were perched on the side of a fixed, sloping wall rather
than in a moving ship.

``Spacial displacement,'' Rachel said. ``We've left our original reference frame. In effect, we
are no longer moving through the universe but instead the universe is moving around us.''

Tusk looked out the windows to see the sky growing gradually darker. Slowly the floor tilted
back toward a more comfortable flatness and everyone, dunari and human alike, relaxed their
grip. By the time they leveled out completely the sky out the windows was utterly dark. Tusk
shuddered slightly. Narlock breathed steadily but seemed more relaxed.

Rachel looked over her shoulder and seemed to be watching through the forward windows. Tusk
could hear the pilot this time. ``\e{Tell them to get ready for transition,}'' she said.

``We're about to get weightless,'' Rachel told the group. ``It's best to hang on and for us, at
least, it helps to breath deeply.''

The change was not nearly as dramatic as Tusk expected. He again felt the curious sensation of
his whole body changing size momentarily, this time getting slightly smaller before returning to
its normal size. Afterward he felt like he was lifting off the floor although it didn't seem
like he had moved at all. The whine of the engines returned, soft and faint, as if coming from a
distance. Tusk felt his clothing ripple over his body in an unusual way.

He pushed slightly against the floor and started to drift upward. Feeling a wave of panic he
tucked in his legs in tightly and pulled himself down again with his trunk.

``It's easy to get disoriented,'' Rachel said. Her hair, tied behind her head, was floating
above her shoulder. ``It takes training to be able to work under these conditions. We're just
going for a ride. Focus on the floor and remind yourself that it is the floor.''

She turned again and looked out the front window but Tusk couldn't see from his vantage point.
``\textit{Summer Breeze} approaches,'' she said.

As she predicted it took a surprisingly long time to connect with \textit{Summer Breeze}.
Rachel explained that the process was largely automatic but to guarantee lock integrity tight
synchronization was needed and certain tests were done. Tusk didn't mind. He actually found
being weightless rather enjoyable once he got used to it. He couldn't wait to tell Jinna about
it.

Finally they heard a soft click and felt a gentle shudder. Uncertain what to do the dunari
waited. Rachel didn't move either although the doctor moved toward a storage unit of some
kind...

% TODO: They would have put on traction boots before the trip.

The pilot came out of the cockpit area, pulling herself across the ceiling with her arms. It was
a very creepy sight and Tusk tried not to watch. Fortunately when she reached the far end of the
space she thrust off the ceiling, spun around in the air and landed, traction boots first, on
the floor where she stuck firmly.

Rachel much more tentatively stepped forward, carefully placing each foot one at a time. She and
the doctor than helped the dunari to stand and get their four feet firmly attached to the floor.
Through the windows Tusk could see that they were surrounded by a great structure with arms
stretching out and connecting with a large arc overhead. But he couldn't get a proper view and
he didn't have time to really think about it. The humans were soon herding them toward the door
at the back through when they had come originally.

When the door opened, again with a slight hiss, there was a short narrow passage to a similar
door at the other end. Narlock checked his watch. ``Twenty more arnets before I can take another
one of those pills,'' he said to Tusk softly.

``I don't think opening a pill bottle right now would be a good idea anyway,'' Tusk said.

It felt strange to be weightless and Tusk had to constantly remind himself to hold his trunk in
an appropriate way. Yet the traction pads on their shoes helped greatly. They plodded along one
step at a time, in a single file, down the short passage.

The pilot opened the other door first and then deftly hopped through to what looked like a
small, windowless room. The doctor followed and the two of them helped Rachel and the dunari
into the space as well. Tusk tried to steady his breathing. The enclosed space was filled with
strange greenish lighting. It wasn't as dark as he feared but the lack of windows bothered him
more than he realized.

The pilot closed the door and then turned toward them. ``\e{I hope they are okay,}'' she said
mostly to Rachel.

``This is an elevator of sorts,'' Rachel explained to the dunari. ``It will take us down to the
rim of the ship where things will be more comfortable.''

Without touching any controls, for indeed Tusk could not see any controls, the elevator began to
move. Aside from the weightlessness and lack of windows the experience wasn't much different
than any other elevator ride Tusk had taken. He could feel the movement quite distinctly.

As they descended the sensation of weight gradually returned. It was almost imperceptible at
first but soon it was very noticeable. Narlock seemed especially relieved. The elevator clicked
slightly and a panel near the door showed some changing lights. By the time they reached the
bottom they felt their normal weight.

There was a slight delay and then the opposite door opened. Standing there were two more humans.
One, a male, had black hair like Rachel's but darker skin. The other, female, had long,
extremely light hair unlike anything Tusk had seen before. It seemed like the humans had a lot
more genetic variety in their species than the dunari did.

``I Captain \alien{Hector Rodriguez},'' the male said in thickly accented Argenian. ``Welcome to
\textit{Summer Breeze}.''

\breakpoint

After a quick introduction Captain Rodriguez excused himself. He spoke to Rachel in English,
``\e{Please convey my apologies. We have a tight schedule and there is much to do. If we have to
  wait I'd rather wait there than here.}''

Rachel nodded and explained to the dunari. Then she turned to the other woman and the two put
their arms around each other briefly, a gesture Tusk had never seen before.

``This is Jaya,'' Rachel said. ``She is one of my students, actually, here to observe with me
once this is all over.''

``\e{Let me show you their space},'' Jaya said. She led the way as the three of them followed
down the hallway.

Tusk wasn't sure what he expected but somehow the reality didn't match whatever it was. It was
cool, that much he had expected, but the lighting was very different. It was brighter and
greener than normal lighting on Rujar. It didn't have the same warmth. The halls were narrow and
tall, suitable for humans, perhaps, but very cramped from a dunari point of view. They passed
the occasional open area. Two were empty but at one a human was sitting upright in a kind of
raised seat looking at a display panel. He seemed happy to see Rachel and he spoke with her and
Jaya briefly. Overall, though, there were far fewer humans around than Tusk expected.

``Where is everyone?'' he asked Rachel as they walked.

``Most are probably getting ready,'' she replied. ``But understand there are only about ten or
twelve people on board.''

``So few people to run all of this?'' Narlock said.

``The ship is largely run by machine,'' Rachel said. ``It can even navigate its way back to
Earth on its own if it has to. There are really very few of us here.''

Rachel reached out and touched Jaya's arm. She turned but kept walking. ``\e{Where is Kevin?}''
Rachel asked.

``\e{I think he's reviewing the assembly sequence again. To be honest he's a bit nervous about
  meeting our dunari friends.}''

Rachel glanced at Tusk. ``\e{I understand. It's okay.}''

``\e{Who is Kevin?}'' Tusk asked.

Jaya paused and gave an embarrassed chuckle.

``Kevin is my other student,'' Rachel said switching to Argenian for Narlock's sake.

They finally arrived and Jaya slid open a door and invited the others inside. ``\c{This is
  it?}'' Rachel said and Jaya nodded.

``This is the space we set aside for you,'' Rachel said to the two dunari. ``I realize it's
small. We actually combined two adjacent spaces; the walls are partially adjustable. I hope it
is suitable.''

% TODO: Describe space

``If you don't mind I think I would like to just rest here for a while,'' Narlock said. ``It's
been exhausting and I might even sleep a little.''

``Of course,'' said Rachel.

As they were leaving Narlock settled down on the bed to just stretch out. Rachel, Jaya, and Tusk
went back out into the hall. ``\e{Let's go to the tin-tin,}'' Rachel said, ``\e{Tusk will enjoy
  the view.}''

Jaya led the way again while Rachel walked, more or less, beside Tusk in the narrow hallway. She
explained that the tin-tin was a kind of cafeteria, meeting place, and observation area all
rolled into one. ``We don't have a lot of space on board so the spaces we have get used for
multiple purposes.''

``\e{Can you use English, please? I want to practice,}'' Tusk replied.

``\e{Yes, of course. Jaya would prefer it too, I bet,}'' Rachel said.

``\e{I'm practicing my Argenian,}'' she said, ``\e{but I won't embarrass myself by trying to
  speak it now.}''

When they reached the tin-tin there was nobody there...

% TODO: Describe the tin-tin

On the far side of the room was one large, blank gray wall.

``\e{Francine,}'' Rachel said loudly as if she was addressing the room itself. ``\e{Activate the
  forward view.}''

The gray wall changed. Suddenly before them and below them was the huge, curving shape of Rujar.
Tusk gasped slightly and stepped forward, cautiously at first, and then more boldly as he
reminded himself that it was only a display.

Rachel came up beside him and outstretched her hand. ``\e{I give you Rujar,}'' she said,
``\e{the home world of the dunari.}''

Tusk didn't recognize much. The ground was so far away, starkly red and brown. The display
hadn't been color corrected for dunari eyes and so it lacked rubinum making the view seem alien
even to Tusk. Then he realized the view was changing. As \textit{Summer Breeze} flew in its
orbit the world seemed to rotate beneath it.

Suddenly he realized they were actually over Forbin. He recognized the confluence of the Tariori
and Tolopias rivers, where two of the largest rivers in the land joined to form the Tulari. From
there the waters flowed for nearly two thousand karnons through rich forest lands until finally,
succumbing to the heat, the river just ended in the middle of the burning desert.

``\e{It is beautiful, isn't it?}'' Rachel said.

Tusk swayed his head but didn't reply.

They watched as the Rujaran deep day passed slowly beneath them. It seemed slow but of course
they were moving far faster than any aerocar could ever go. In the distance Tusk could see the
mighty Tukonook Mountains, as little more than a crease on the land. Argenian civilization
started beyond that point and Tusk watched eagerly for Sumar Sea.

``\e{You can't really see any evidence of us,}'' Tusk said mostly to himself.

``\e{You can,}'' Rachel said, ``\e{you just have to know where to look. Varnok is easy to see
  from this altitude and the farming country in Bulartria Provence stands out as well.}''

``\e{Still it makes us seem rather small.}''

``\e{There is nothing wrong with that,}'' Rachel said.

As they spoke the Narlar approached and Tusk started to feel uneasy. Rachel was unperturbed and
watched as the border of night and day passed under them. The red Rujaran sun shone steadily
above the edge of the world but it sank quickly. Rachel watched it closely with the red glow on
her face. Tusk turned to see something no dunari had ever seen before: the sun setting over the
edge of his world.

% TODO: Describe the sunset.

``\e{Let me show you something else,}'' Rachel said. Again she spoke a command to the room.
``\e{Francine, switch to central view.}''

Now instead of looking down at the cold night of Rujar, they appeared to be on the inside of the
ring looking toward the central shaft. ``\e{This gives the best view of the ship,}'' Rachel
explained.

% TODO: Describe the ship.

``\e{What are those projections along the side?}'' Tusk asked. ``\e{Are they part of your
  engines?}''

``\e{No,}'' said Rachel. ``\e{Those are actually shipping containers. They contain the Nitooli
  telescope.}''

% Do we also want to look out onto the galaxy? That would be a good opportunity to remind the
% reader about Rachel's real purpose there and to hook the main plot to this subplot, even if
% only indirectly (Rachel could talk about how anxious she is to observe far side spherical
% clusters, etc, etc).
%
% On the other hand it's nice to end the scene with a powerful reference to Nitooli.
%
% Should this scene end with them leaving the Rujaran system? I probably need to describe that
% somewhere.

\breakpoint

% I should have some other scene here, probably, to break up the Summer Breeze scenes. I also
% need to make it clear that Rachel is the driving force behind this undertaking and that it
% most likely wouldn't be happening if it wasn't for her.

% SUMMARY: The deployment of Nitooli

There was a lot of activity in the makeshift Nitooli control room. Kevin, Jaya, and Rachel where
there and three other humans who served as engineers on the project. Rodriguez was there as
well, mostly just as an observer. Tusk and Narlock stood to the side with Narlock scribbling
notes on a pad as he watched.

``I need you tell me if they say anything interesting,'' he told Tusk quietly.

The humans sat in chairs in front of a long narrow desk that seemed to be partially built into
the wall. In front of them were various electronic displays and other devices. Tusk could see
images of \textit{Summer Breeze} and of the Nitooli storage containers. He could also see images
that made no sense to him.

Rachel turned and motioned for them to come and take a closer look. Both Tusk and Narlock
stepped closer to her station. On the display in front of her was a picture that showed the
entire ship. The Nitooli containers were disconnected and seemed to be floating nearby.

``\e{This is taken from a dingus some distance from the ship?}'' Rachel explained.

``\e{What's a dingus?}'' Tusk asked.

``\e{It's just what we call the small robotic units that we use on construction projects. I'm
  really not sure where the word comes from.}''

As they watched the container doors opened and they could see, barely on the small display,
several more dinguses going inside and taking various components out. There was nothing that was
recognizable to Tusk as a telescope but he expected that.

``\e{I don't have much to do right now. This step is mostly in the hands of the engineers.}''
Rachel gestured down the table to where David was studying his display intently. He seemed to be
looking at some sort of three dimensional graphic but Tusk had no idea what it meant.
``\e{Afterward my students and I will do the calibration and testing.}''

Things seem to happen in slow motion but also several things happened at once. It was like
watching an intricately choreographed dance. As components were connected to other components
the entire assembly moved away from the ship. The picture backed up as well so that it could
take in as much as it could.

It appeared to Tusk that frames were being built. It also seemed like there were several of
them. ``\e{Why so many?}'' he asked.

``\e{Nitooli isn't just one telescope,}'' Rachel said. ``\e{It is really an array of telescopes.
  Once the frames are done they will be covered with linare, it's a high reflective
  microscopically thin layer that will form the mirrors. Then the individual telescopes will be
  moved apart and linked electronically to synthesize a single large mirror.}''

Tusk roughly described the process to Narlock but he seemed more interested in what the other
humans were doing. ``They are mostly just monitoring the assembly process,'' Rachel explained.
``Everything is automated and has been tested. This should be routine.''

Several of the mirrors seemed complete and using a mechanism that Tusk could only guess about
they started to rotate in a stately, perfectly synchronized way. The completed mirrors started
to move away, in all directions, outside the field of view. As they did so, more mirrors were
assembled.

``\e{Locks are established,}'' said David. ``\e{Branches 1A, 1B, and 1C are deploying normally.
  All systems seem nominal.}''

Rachel leaned back in her chair. ``We wait,'' she said. ``It will take time to finish the
assembly.'' She touched the display in front of her and an image of a dunari clock appeared
displaying the current time. The subarnets ticked by steadily.

``Are we on schedule?'' Tusk asked.

``We are so far.''

\breakpoint

% SUMMARY: First light!

