
Chark's office was on the leeward side of the Advanced Technologies Center, on the second floor
and near the rounded northeast corner of the sprawling structure. On clear days, when the sand
was out of the air, he could stand beside the window and see the Sarnowledge Mountains rising in
the distance as a darkening against the yellow-green sky. Sometimes he liked to look at the
mountains and reflect on the relaxed days when Junella's old seleck took camping trips there. Of
course that was all before the project.

His office was not small but not large, reflecting in its size the significance of his position.
He was lucky to have two windows. On the hottest days he opened both and even with the Nermella
blowing steadily, as it always did, only a gentle breeze circulated through the room. His table
was functional, his seat was comfortable, and his phone worked. In the hall outside was an
engraved plaque: ``Captain Sar. Charkonaless, Director of Radio Technology Research.''

Chark was entering notes into the project management book. It was one of the regular drudgeries
that came with his job. He didn't mind it, at least not much. He understood the RRRF project was
closely watched and needed detailed documentation despite, or perhaps because of, its secret
nature. He had a small collection of skilled engineers under his direction. Some were recent
university graduates, excited to be a part of such an important and challenging project. Some
were members of the fledgling commercial aerocar industry. All were carefully vetted by military
intelligence.

The phone started to chime in its usual, insistent manner. Distracted, Chark reached his trunk
out clumsily to grab the transceiver, wrapping four tentacles around the handle and lifting it
to hold under his ear.

``Captain,'' came the voice on the other end, ``could you come down here?''

``Why? What's happening?''

``We were calibrating the power amplifier and we seem to be receiving a signal. We can't figure
out why. We're hoping you might have some ideas.''

Normally Chark enjoyed working with the younger engineers. They had enthusiasm even if they
lacked experience. Yet he was tired and at that moment he mostly just wanted to get his notes
done and go home.

``I'm sure you'll get it. Is Brogan there?''

``The target seems to be moving.''

Chark flapped his ears slowly in annoyance.

``Okay, fine,'' he replied, ``I'll come down and take a look.''

He put the phone transceiver back on its stand and finished writing the comment he had started.
Then with a soft, fluttering sound of impatience that came from deep in his chest, he wrapped
his trunk around the table grip and rolled his seat back. He slid his rear legs to the floor
and, unfolding his front legs, stretched slightly. He straightened his military cloak to tidy
himself, checking the buttons to ensure all were fastened properly.

Jeno was at home. Her table, as always, was perfectly arranged with orderly trays filled with
papers and tools. Chark made his way through the broad opening separating Jeno's space from his
own, and from there into the main passage that ran the length of the building.

As he walked along the tiled hallway splashes of blotchy sun came through the windward windows
and spread across the floor. They filled the passage with deep, red-rubinum light. At Lungast
the humid, tropical air made the sun seem darker than normal, covering it as it did with a
slight haze. On the Northern Plains where Chark lived before, the air was dry and clear and
everything seemed so much brighter.

