
Tusk poked at the last remaining brag root lying in a dejected pile on the copper serving tray.
It didn't look very appetizing although he couldn't fault the hospitality of the First Staff for
providing the tray. After all, they hadn't been invited. He picked up the root gingerly with two
tentacles and popped it into his mouth. Chewing slowly he walked over to Lucasa who was standing
beside a shelving unit of books and looking out the window at a large tree. The street was a
full story below and the window was at the level of the tree's branches, swaying as they did in
the steady Nermella.

``I'm impatient,'' Lucasa said without turning around.

``Does it seem odd to you that Marshall is in there and you're not?''

Lucasa shrugged, ``Not really. Politia is there too. It's natural for Argenia to consult with
other nations and I'm glad the humans are seen that way.''

Tusk leaned his trunk against the shelf and looked around at the fine furniture and rare books
that decorated the Argenian Reading Room. ``You are one step closer to the First Table now, my
friend. It seems to me you should have insisted.''

Lucasa didn't answer right away. Instead he walked to the middle of the room and pulled the
serving tray toward him slightly, eyeing it's empty contents. ``This is not the time,'' he said
finally, ``to be insisting about such things.''

Suddenly the phone chimed. Tusk lifted the old-style handle and awkwardly held the small speaker
under his ear. He listened for a moment as Lucasa watched him and then he leaned down and spoke
directly into the microphone on the base. ``Yes, thank you.'' He hung up the phone and turned to
Lucasa. ``They're finishing up.''

The two men left the private confines of the Reading Room and headed down the tiled hallway with
the sun at their backs toward the conference circle. They passed by several offices, most with
their doors closed but some opening inward to bustling staff working on the business of the
country. They rounded a corner into a large foyer flanked by tall sun windows. Before them the
broad doors of the conference circle were open, with the seal of Argenia perched above the
entrance. A throng of reporters gathered there, held back by a low partition and the watchful
eyes of the First Guard.

The new First Citizen stood a short distance behind the partition and was taking questions.

``How close are we to identifying the criminals responsible?''

``We are making progress every hour. I am confident we will apprend the perpetrators before
long,'' the First Citizen said.

``What does your government intend to do for the survivors of Yartagard?''

``I will have a full statement in an hour and can answer more questions at that time.'' The
First Citizen said. With that he turned his head and, with his bodyguard close at by, walked
quickly toward his new office.

Tusk snorted softly and leaned toward Lucasa. ``You are perhaps even closer to the First Table
than I thought.''

Fotkey and Politia were next to endure the interrogation of the press. Tusk thought they made an
odd even if symbolic pair.

``What is our stance toward the Forbin government?'' a reporter asked. Fotkey glanced at Politia
and the Forbin ambassador stepped forward.

``We intend to offer as much aid as possible,'' he began. ``The Prefect has sent word to me
saying that Forbin intends to send at least 5000 construction workers to help with the clean up
and reconstruction of Yartagard.''

``What about the investigation. Do you believe the Junar are responsible?''

Fotkey took that question. ``There is little doubt of it.''

``Can they be brought to justice?''

``That is our intention.''

``General, there have been open investigations against the Junar for many years and none have
come to fruition. How is this going to be any different?''

``It's different because now the full resources of Argenia will be brought to bear. It's as
simple as that.''

With that the two moved past the reporters and turned down the hall away from the partitions and
from the crowd. Now it was Marshall's turn and the reporters swung back to him as if he was any
other political figure.

``Ambassador, what assistance can you offer us?''

Marshall stood beside the Argenian flag, his hands clasped in front of him. ``Unfortunately
there isn't much we can do. We will continue to help move injured and we will consult with
doctors and engineers about the many problems you'll face on the way to recovery.''

``Surely your technology can help in the investigation.''

Marshall took a deep breath. ``We don't have the tools here to be very useful. Even our
construction tools on Nermia are ill-suited to operation on the ground. We are looking into
moving a digger here to help with the clean-up but it will take time to reconfigure it.''

``Ambassador\ldots'' but Marshall raised his hand slightly.

``I'm sorry but I need to return to my tasks. I'm sure the First Citizen will say more when he
gives his statement.''

Marshall walked along the partition and Lucasa and Tusk stepped up to intercept him. The
reporters quickly gathered around Lucasa.

``Lucastanonia! Are you planning to announce your interest in the position of First Citizen?''

Lucasa paused slightly and tilted his head. ``I have no plans of that nature. My plan is to do
what my office has always done: coordinate with the humans and represent Argenia to them.''

``Lucastanonia\ldots''

Lucasa waved his trunk and pressed slightly against Marshall's leg to indicate they should move
on. The three of them quickly walked down the hall and Tusk could hear the sound of the
reporters receding.

They rounded the corner and Marshall turned his head the way only a human could to look back
briefly. He leaned down slightly toward the two dunari. ``Rachel has a plan,'' he said softly,
``but we need to get approval from Earth.''

``How long will that take?'' Lucasa asked.

``Too long. Rachel wants to start before we hear back.''

``What if they disapprove?''

Marshall stopped and turned toward them. He pursed his lips slightly. ``Let's just say,
gentlemen, that it's a long way from Earth and signals are often noisy.''
