
After the introductions, the humans who accompanied Dr. Spencer returned to their craft. She
walked forward with Lucasa on one side and Colty on the other into the tent area where the
reports waited behind a rope. Dr. Spencer did not seem surprised to see the reporters. She had
no doubt been briefed on what would happen. The humans who had been on the ground waited off to
the side.

The human was shown the way to an impromptu podium where she could say a few words and answer a
few questions. She seemed shy and somewhat reluctant to speak. She walked slowly to the podium
and then paused while everyone waited.

``Greetings,'' she said in her strange voice. ``My name is \alien{Rachel Spencer}. I am a
physicist and mathematician. I am here to be on hand during the construction and testing of the
telescope on Fresnik. I am also here to get to know you better and to begin a process of
intellectual sharing between our peoples that I hope will continue far into the future.'' She
turned slightly toward the greeters. ``I'd like to thank you all for your welcome. I'd like to
especially thank Dr. Harnick for being my host. I know that it will be strange to have me
around. I will now answer a few questions if you have any.''

Colty was impressed with her mastery of the Argenian language. Despite what he had been told,
she did not once use the electronic translation devices the other humans usually relied on to
communicate. Yet despite her alien voice she spoke very clearly.

There were plenty of questions. Lucasa took the role of moderator and selected the first
reporter for a question.

``You spoke of `intellectual sharing.' What do you mean, exactly?''

She paused for a moment as if to gather her thoughts. ``I am a researcher and a teacher,'' she
said. ``I hope to learn about your physics and mathematics and, in due course, I hope to teach.
I also hope to do some astronomy once our facilities on Fresnik are complete.''

``What will you teach?'' the reporter asked.

``I don't know yet.''

There was some jostling for the next question. Lucasa pointed to a young reporter in the
front.

``All humans who have come to the surface have so far kept mostly to themselves. Why did you
want to be a part of our community?''

``I do not want to know your world only from the windows of a spacecraft or of our compound on
the outskirts of Parnon. I want to know your world as you do. If we are to one day work
together, we need to understand each other. Let this be a step toward that understanding.''

``Just one more question,'' Lucasa said. ``There\ldots''

``Are you religious?''

Colty was taken aback by the question. It seemed inappropriate. Lucasa turned questioningly
toward Dr. Spencer but she didn't seem to notice.

``I am religious, yes,'' she said. ``Although the way I express my spiritual feelings might seem
strange to you. Yet in many ways we are not that different. We also struggle with questions of
morality, the nature of divinity, and the relationship between rational thought and faith. Our
knowledge of the universe is more detailed than yours and that creates both challenges and
opportunities for religious thinkers. I am sure you will discover these challenges and
opportunities in time.''

``Okay, that's it,'' said Lucasa. Dr. Spencer got down from the podium and security escorted the
reporters back to the main building.

Colty spoke to his guest now for the first time free of the protocols of formal introduction.
``You are very eloquent,'' he said.

Dr. Spencer smiled slightly. Colty recognized the gesture from the picture Tusk has shown him.
``Thank you,'' she said. ``I've spent a lot of time studying your language while coming here. It
is difficult for us.''

``You speak better than me,'' Colty said.

Finally, once most others had left, the other humans approached. Colty knew that Dr. Spencer
would first settle in at the human compound before visiting the university. The humans had a
small vehicle [mention this earlier] to take her there [mention also that the human compound is
near the air base].

The humans spoke to Dr. Spencer in their language, an eerie collection of hisses and clicks that
seemed utterly alien. Colty had never heard it spoken in real life and it sent shivers down his
spine. He couldn't help but recall the blood sucking space monster stories of his youth. Humans
might be able to learn Argenian with difficulty but Colty seriously doubted if typical dunari
could learn the human language.

Dr. Spencer turned to him and said, ``I understand there is a reception later at the
University.''

Yes, Lucasa interjected, ``at hour 62. We were planning to send a van for you.''

``I will see you then,'' she said. It seemed like a strange comment. Dr. Spencer followed the
other humans to their vehicle and they left toward the compound.
