
\noindent Rachel tied her black hair into a short ponytail at the base of her neck using an
elastic. She did a quick check in the mirror to see if there were any horrible blemishes on her
fair skin. It wouldn't have mattered if there were. She preferred using her real image anyway.
She put on a light green tunic with embroidered white trim that was both simple and elegant,
along with white pants. Then, staying barefoot, she went downstairs.

In the corner of the finished lower level of her home she sat in a standard office padded chair
in front of a large, curved gray screen. The screen stood about a half meter above the floor and
was almost a meter high. It wrapped around the chair, at arms length, so that if she looked
straight ahead she could only just barely see the two sides in her peripheral vision. The rest
of the room was comfortably furnished. Behind her was Sarah's desk, still covered with her
papers, all in neat piles. To her right was a comfortable sofa facing the back windows with some
house plants decorating the well lite space. The blinds were mostly closed, but diffuse sunlight
filled the room.

Rachel deftly slipped the device behind her left ear and hooked the delicate headpiece over her
scalp. She turned the device on and stared at the screen in front of her as a simple test
pattern appeared. With a few adjustments of her hair she moved the sensors close enough to her
skin to activate the green symbols beside each sensor icon. She moved her hands to her lap and
concentrated.

The test pattern first swung to the left, then to the right. It spun clockwise and then
counterclockwise. Rachel nodded slightly and took a breath. The test pattern disappeared and,
using only her thoughts, she commanded the conference dialog to appear. There she connected with
the university conference system and selected the appropriate meeting. After alpha wave
authentication the image of Dr. Toombs, the physics department chair filled the left half of
Rachel's field of vision. He had dark, piercing eyes, was dressed in an expensive suit, and had
a full head of brown hair. Of course, the real Dr. Toombs was nearly bald, and probably didn't
own any suits.

``Good afternoon, Rachel,'' he said with a British accent, ``oh, I guess it's still morning
there.''

``Hello, James.''

``We are waiting for Rajagopal. He's the representative from the Xnn planning committee. How are
your classes going?''

Rachel tossed her head slightly despite her hair being pulled back. ``They're fine.''

``The students like you,'' Toombs said. Rachel shrugged.

The other side of the screen activated and the image of a short, somewhat paunchy man appeared.
He had dark skin, slightly greasy hair, and was dressed in an inappropriately colorful shirt.
Rachel assumed it was a real image since nobody in their right mind would create an avatar like
that.

Introductions were made and Rachel waited patiently. The holographic imaging system gave the
attendees a three dimensional appearance and presented the illusion that they were all sitting
around a table together in the same room.

``Rachel, as you know,'' Toombs said finally, ``your proposal for telescope time on the
Jordan-Fair instrument hasn't been approved.''

Rachel nodded slightly.

``But I believe your work with globular clusters is important so I took the liberty of
contacting the Xnn planners to see if there was any opportunities there.'' Toombs gestured
toward Rajagopal.

Rajagopal leaned forward and coughed slightly. ``Yeah. So it turns out X17 has been
funded\ldots''

``What?'' Rachel said. She looked intently at Rajagopal and then glanced at Toombs who grinned
back.

``Yeah. So it's going out to the top of the galactic disk at a latitude of as close to ninety
degrees as we can make it. So we've already got probes out that way looking around.''

``That's perfect,'' Rachel said, squirming a bit in her chair. ``I never thought I'd see X17
become real.''

``So I want you to consider applying for time on it. The position is very well suited to your
project, making it a good fit for you. I want to encourage you here.''

``Yes, of course.''

``I'll send you the application materials, but they are still in draft form right now. I want to
fast track this a bit, okay? So we might do things a bit off the book. I personally am very
excited about your work. Yeah, it's fascinating that our galaxy is the only one in the vicinity
with anomalous globular orbits.''

When the meeting was over Rachel paced around the room. Without thinking she went into Sarah's
studio and then drew up short when she realized that Sarah wasn't there. Instead Rachel looked
again at the painting Sarah had been working on, still unfinished. The news about X17 was
exciting, but Rachel knew all too well that such projects often never make it beyond the early
stages. It's too easy to cancel a project before the big money is spent.

In the weeks that followed Rachel spent many hours just walking the fields and forest above the
house. With her younger sister gone there seemed little else to do. There was her research, of
course, but that didn't hold her interest for long. There were also her classes, but they were
routine.

She stopped to sit on her favorite log, with a mature maple-beech forest at her back. A long
grassy field sloped away before her to the house below. Across the valley the Adirondack High
Peaks rose, veiled slightly by the muggy summer air. She brushed away a couple of flies that
buzzed about her head.

The house looked so peaceful and inviting, surrounded as it was by the gardens she and Sarah had
put in. It was the house their parents bought for their summers in the mountains, steeped in
memory and tradition. The accident took their parents, and now disease had taken Sarah. Despite
the genetically engineered viruses, and despite the nanite treatments, she could not be saved.
At least the drugs spared her the pain.

``Do something good,'' she told Rachel one spring afternoon. She was in her room in the house
below, on her bed and surrounded by machines. She seemed so sincere, but Rachel wondered if it
were the drugs talking.

A hawk caught Rachel's eye and she followed it as it circled over the mountain slopes.

Then there was David. He tried to jam Rachel several times, but she avoided him. She set her
avatar to apologize on her behalf. ``I am very busy with my work right now, please jam me
later.''

``I'm worried about you lassie,'' he said in a message, ``why don't I come up and we can do a
weekend on the river. You'll feel better.''

That was a month before. Rachel sighed and stood up to walk back to the house.

The house was surrounded by winter snow when the message from Toombs came through. ``A location
for the X17 observatory has been found. Your application for observing time has already been
granted. See attached. It looks like it will be about two years to first light. You should apply
for a sabbatical and for deep space training now if you want to be ready in time.''

With a thought Rachel flicked the message aside and opened the attachment. It was a formal
note from the X17 schedule committee. Her proposal would be among the first to observe. Rachel
smiled and nodded slightly. It was about time. The note also said something about a complication
at the optimal site location causing possible delays. Rachel wondered what that meant.

That evening she walked down the road a little. Ice crystals crunched under her feet, and her
breath was a fog in front of her face. Above her head, and through the trees, the sky was
spectacularly clear. The stars twinkled against the dark background of space. Rachel looked up
and took a deep breath, taking in the constellations with a glance. It had been a long time
since she'd been off Earth. The idea of going off Earth again seemed to her like something good.
