
Darkot arrived with his closest sibling, as was traditional. They parked the car in the spot
reserved for them at the head of a path leading up a small rise. Darkot climbed out and swept
his trunk over his formal shroud to smooth the wrinkles, tugging on the fabric with his three,
short tentacles. He swayed his head so that he could see all the guests, friends and family,
standing at either side. He kicked at the ground with his back leg and then slowly raised one
front foot in the air. Before he put it down his brother, now out of the car himself, started
the traditional banter.

``Don't do it!'' he said loudly. He raised his head slightly and pointed his trunk at Darkot.
``You'll regret it. There's no going back.''

Darkot stepped forward and walked deliberately up the path between the guests as his brother ran
around him, sometimes trotting, sometimes sidestepping, constantly talking. ``You're making the
mistake of your life---look at our parents!'' He pointed toward an older, well dressed couple
standing to the side. Many of the guests shook their ears with amusement but Darkot just clicked
his teeth and walked on.

In a short distance the path came to the top of a large, flat grassy knoll. There opposite
Darkot standing in an arc along the far edge of the knoll stood seven women. In the center was
Jiata, his bride. At her sides were the six others of her seleck. Darkot stepped forward and his
brother fell silent and withdrew.

The knoll was the only feature in an enormous plain of harsna grass. The tall, lush yellow
plants surrounded them like a sea surrounds a boat, stretching into the vast distance as far as
even dunari eyes could see. The sky above was clear and green with a touch of golden sand driven
from the distant desert by the Nermela wind, the breath of Jurita Herself. Hovering above his
bride the unmoving sun shone upon his face, its deep red-rubinum light washed over the world.

Darkot moved forward and Jiata also. They approached each other slowly and solemnly. She was the
most beautiful woman he'd ever seen. Her wedding shroud was embroidered with the Harnick family
emblem and trimmed with orange and gold. Her large ears were clipped back with silver chains,
and around her legs and trunk she wore a fine mesh of brikken web. She was the love of his life.

When the two met at the center of the knoll Jiata gave Darkot a single naria flower, its green
bloom symbolic of long life. Darkot clipped the flower to his shroud and then they wrapped their
trunks together and clicked their teeth with happiness.

The women of Jiata's seleck raised their trunks and paused. All was silent for a moment save the
rustling of the harsna grass in the tireless Nermela wind.

``We are the proxies of Jinalia, the keeper of passion and love,'' the women called out in
unison. Everyone knew the words by heart. ``We are here to seal the union of Jiata Harnick and
Darkot Talinan. We are here to bind them before our eyes and before Jurita\ldots''

When the ceremony was over Jiata formally introduced her new husband to the members of her
seleck. Darkot knew them all, of course, but as traditional he pretended he was meeting them for
the first time.

``I'm so excited for you,'' Jiata's best friend said. ``Now if only I could find someone half as
nice.''

The couple mingled with the guests talking comfortably with people from both families. Darkot
overheard one of Jiata's older relatives, perhaps an aunt, speaking softly with another elderly
woman.

``I don't understand,'' she said in a whisper. ``Jiata's a doctor and he's just an accountant.
She could have done so much better.''

Darkot pretended not to notice. He didn't care. Nothing could dull his happiness on that day.
Instead he walked with his bride to the edge of the knoll where a wizened old man stood quietly.
He had braces on three of his legs and thick glasses wrapped over his trunk ridge.

``Pap, I certainly hope you didn't stand through the whole ceremony,'' Darkot said with concern.

``It was fine,'' the man said. ``It was only a few arnets and I have these steel legs now.'' He
tapped the brace on his front leg with his trunk.

``We are so happy you could make it,'' Jiata said. She rubbed his shoulder affectionately.

``I wouldn't miss it for the world.''

Jiata got the attention of her friend who had been temporarily distracted. ``I want to introduce
you to Darkot's grandfather, Colty Lanaran.''

``We are well met, Sir,'' the young woman said. She held out her trunk and Colty touched it
briefly.

``We are well met,'' he replied.

Colty glanced over the crowd. ``Now if you will excuse me. I want to find out what my little
girl has been doing.'' With short, difficult steps he walked across the grass toward a small
group where Darkot's mother appeared to be telling jokes.

``I worry about him,'' Darkot said.

``He's a wonderful man,'' Jiata said, leaning against her husband affectionately. ``I love the
name Colty also. I think\ldots\ if we ever have a son we should name him Colty.''
