
% SUMMARY: Rachel and Colty continue their Varnok visit... this time at the museum of Natural
% History.

It was a short walk across the park to the National Museum of Natural History. Yet the
experience was, in some ways, worlds apart. The natural history museum was in a much newer and
more modern building with unique revolving doors. Directly in front of the entrance was a large
statue of a gregonerar. The fearsome beast was reared up on its hind legs with its front claws
extended as if ready for battle. Its eyes were polished flopstone and seemed to sparkle with
rage in the red Rujaran sun.

Rachel paused and looked at the animal frozen in stone. ``That's probably close to actual
size,'' Colty said.

``It looks dangerous,'' Rachel replied. ``Do they live around here?''

``Actually, they're extinct,'' Colty said with a shake of his ears.

In the central hall of the museum was growing a full size manago tree. Windows along the side of
the hall let the sun in so the tree could grow. Climbing its trunk was a thick vine with wide,
yellow leaves and dark red flowers. Rachel read the plaque with interest.

``It says there is a symbiotic relationship between the tree and the vine,'' Rachel said. ``The
vine extracts water vapor from the air that it shares with the tree, while the tree provides
support and access to the sun for the vine.''

There were quite a few more people in the natural history museum than there were in the art
museum, including many children. Yet despite that, it did not feel crowded and Rachel seemed
comfortable. As they were looking at the museum map, they were approached, as before, by the
museum director. He was a young man, hardly any older than Colty. He greeted Rachel warmly and
very openly. He seemed remarkably at ease with her.

``Is there any particular area you'd like to see?'' he asked.

``The animals,'' she said.

They were directed down a large hallway with exhibit rooms on either side and large doors at the
end. In the exhibit rooms were small animals in various elaborate cages. It was obvious that an
attempt was made to give the animals a reasonable habitat while at the same time allowing them
be seen by the visitors. There were figzats, lolirars, and termicosts. Rachel was particularly
interested in the termicosts.

``They look like little \alien{cats},'' she said.

``They eat insects,'' Colty read from the plaque beside the cages.

The people were amazingly tolerant of Rachel and several spoke to her, always in friendly tones.
Rachel was very polite, saying ``thank you,'' or ``we are well met,'' in response to each
greeting. Colty was happy to see the people so accepting of her.

% It might be nice to have Rachel look at some animals of the night.

The doors at the end of the hall led outside to an open area with displays containing living
animals from all over Rujaria. It was hard to believe they were in Argenia's largest city; the
paths and walkways took them through all sorts of wild habitats.

Rachel was very impressed. ``This is really great,'' she said looking into the rorack space.
Colty pointed to a grassy knoll where a large rorak was sleeping.

``They are herbivores,'' he said. Colty was fascinated with Rachel's reaction to the animals.
She was as excited as the children to see them. What a curious people the humans were. What a
curious human Rachel was.

``Do you have many different kinds of animals on Earth?'' Colty asked.

``Oh yes,'' said Rachel, ``but I've never seen these animals before.''

They came to a large display filled with exotic trees. A group of children and some adults stood
next to the railing looking inside and pointing with their trunks. It was the chortak habitat.
Colty felt uneasy. The human resemblance to chortak had turned the word into a kind of insult to
his ears.

As they approached the other people fell back but Rachel stopped at the end of the railing so as
not to impose herself on the dunari already there.

``What's in here?'' she asked Colty.

``This is the chortak display,'' he said.

At that moment there was a loud screech from inside the habitat. A group of four chortak
bounded from a nest in some high branches and scurried down the tree. They leaped from branch to
branch with amazing ease, clutching at them with their flexible toes. As they move they howled
and cried at each other with chittering voices.

Rachel watched with fascination. ``I don't think they like that one,'' she said with a nod
toward the animal now in the lower branches. The cast out climbed toward the railing and
screeched at the onlookers who stomped their feet in response. Then, jumping to the next branch
it hung upside down just opposite Colty and Rachel and howled at them.

Rachel stepped back with alarm but the animal was more noise than action. She moved forward
again and the chortak climbed to the end of the branch. It chittered to her and made clicking
noises. Rachel leaned over the railing a bit. ``You don't know what to make of me, do you?'' she
said in fluent Argenian.

A crowd had gathered around the two of them. People stood by to watch the meeting between the
human and the chortak. Colty could hear the sound of camera shutters as many pictures were
taken. Rachel had become as much of a exhibit as the animals behind the railings.

The chortak chittered some more and then reached its front leg out toward Rachel almost as if he
wanted to touch her. Rachel pulled back, well behind the railing. ``Oh no,'' she said. ``I don't
trust you.''

A uniformed woman who had been hanging back with the crowd took that opportunity to step forward
and introduce herself. ``Dr. 'pencer'' she said. ``I'm Jitara. That's Brutik. He's got quite a
personality as you can see.''

``He certainly has a lot to say,'' Rachel replied.

``Would you like to meet him?''

Rachel wasn't sure what she meant so Jitara walked to the end of the railing and opened a locked
gate into the chortak habitat. She climbed down some steps and started making some chittering
noises. All the visitors looked over the railing and watched as several chortak approached her,
chittering back with loud voices. Brutik watched for a moment from above and then with a screech
jumped down from the branches to the ground. Jitara picked him up with her trunk and swung him
onto her shoulders. She then made her way back up the steps to the waiting crowd above.

The children crowded around Jitara as she came through the gate again. She told them to stay
calm and not move too quickly. Brutik obviously knew the procedure. He swayed from side to side
with obvious glee and reached out his hands to touch the children's uplifted trunks. Rachel and
Colty hung back and watched. Rachel smiled.

After a few moments Jitara walked over toward Rachel and Colty, still with Brutik on her back.
The animal seemed both curious about and fearful of Rachel. She definitely affected him. Jitara
stopped a short distance from them and Rachel stood quietly holding her hands together in front
of her. Brutik bobbed his head up and down looking at her as the dunari visitors gathered
around.

Brutik chittered questioningly and then Rachel spoke to him, this time in her own language. It
was the first time Colty ever heard her speak her native tongue and the sound of it was
shockingly alien. Her voice had a similar quality as the chortak vocalizations, yet it was
obviously a much more complex utterance. Brutik feel silent and stared at her. Rachel laughed.

``I don't think I've ever seen him at a lose for words,'' Jitara said with a shake of her ears.
``What did you say to him?''

``I just told him I was friendly and that he has nothing to fear from me.''

Jitara asked Colty to reach into the front pocket of her shroud and take out the container
there. Inside were some rigbar roots. ``Go ahead and give him one,'' Jitara said.

Rachel took one of the long, black roots and tentatively held it out. Brutik backed away at
first but then reached for the root. He lost his nerve and backed away again. Rachel stepped a
little closer and held the root steadily. Eventually Brutik got brave and reached out again. He
grabbed the root from Rachel's hand and retreated to the safety of Jitara's back to eat it.

Rachel held out another root and this time Brutik took it right away. Rachel moved a little
closer and gave him another. Colty could hear the cameras clicking again.

``There you see?'' Jitara said. ``You are friends now.''

Brutik clung to Jitara's shoulders and leaned toward Rachel. She held up her empty hand. ``I
don't have any more,'' she said. Brutik reached out and Rachel didn't move. He took her hand in
his and seemed to study it. Rachel relaxed her fingers and let him explore. The chortak had
three long, curved, parallel fingers that he used to hook around branches. Rachel's hand was
quite different. Her four fingers were shorter but, as Colty knew, far more dexterous. Her fifth
finger, set apart from the others, gave her as much ability to manipulate small objects as the
dunari trunk. No chortak could compete with that.

Colty watched Rachel closely. She studied the chortak and smiled slight as Brutik touched her.
It was as if she felt a kind of kinship with the creature. At some level they understood each
other.

Brutik chittered suddenly and then jumped back onto Jitara. Rachel lowered her hand.
