
Denart had never been in Fargon before but he found that he liked it. Unlike the newer buildings
Fargon was ornately decorated with carved moldings and intricate ceiling panels. The floors were
polished wood, not the tile covered concrete to which he was accustomed.

His class was at the back corner of the building overlooking the hagnolia bushes that grew
around Fargon. As usual Denart got there early. He liked having his pick of the seats. Yet when
he pushed open the door he stopped short. The human was sitting at the back of the otherwise
empty room, using a very choice seat next to one of the tall windows.

Instinctively Denart stepped back into the hallway and looked at the number on the door. He had
the right place. ``Zecks,'' he said softly to himself. He paused briefly and then with a shrug
went back inside.

The human regarded him with its strange forward facing eyes but Denart pretended not to notice
it. Instead he boldly walked to the back of the room and lay down on the seat right next to the
creature. It was, after all, the second best seat in the class. Denart glanced at the human and
saw that it was looking at him.

``I am Ra'hel,'' it said. The human spoke with an odd accent but was entirely understandable.
That meant he couldn't just ignore it.

``I'm Denart,'' he replied. ``People just call me Den.''

``We are well met,'' the human said formally.

``Yeah,'' replied Denart. He curled his trunk and tried not to look at it.

The two sat quietly for a moment and then Denart turned toward the human. ``Are you taking this
class?''

``Yes.''

``Your people don't have combinatorics?''

``I want to learn how you do your classes,'' it explained. ``I'm supposed to give lectures next
terms.''

At that moment a small group of other students walked in. Denart recognized one but the others
he had never seen before. They were all advanced math students, probably in their fifth or sixth
term. One of them was a rather cute girl.

As Denart had done the group stopped abruptly when they saw Rachel. The human sat quietly with
its rear legs folded underneath its body and its front legs resting lightly on the table. It had
a notebook there as well, complete with the Parnon University seal on the cover.

``Oh wow,'' one of the students said. ``I guess this is an advanced math class.'' The other
students clicked their teeth but the girl shook her ears vigorously. They mostly all found seats
some distance away from the human but the girl walked right up to Rachel.

``I'm Jirano,'' she said. She extended her trunk as one dunari might to another.

The human paused. Then, slowly, it raised one of its front legs and held it out, bent slightly,
with the toes curled downward. Jirano touched the human with her trunk and gently wrapped her
tenticles around its leg.

``I am Ra'hel,'' it said.

``Welcome to Parnon,'' Jirano said.

The human looked at her intently and made an odd expression with its mouth. ``Thank you,'' it
said.

Jirano lay down next to Rachel on the opposite side from Denart. ``I'm glad I'm not going to be
the only girl in this class,'' she said.

``You're a girl?'' Denart blurted out.

Rachel turned toward him and spoke slowly. ``Our genders are not exactly like yours but
essentially, yes, I'm a girl.''

``Are you married?'' Jirano asked bluntly.

``No.''

``Do your people even get married?'' Denart asked.

Rachel turned back toward him. ``Yes. We have social customs like your marriage.''

The door to the class swung open again and in walked another student. He took one look at Rachel
and squinted. ``What is that doing here?'' he asked undiplomatically.

``What are you doing here, Tor?'' replied Jirano. ``I thought you failed Abstract Algebra.''

``I managed to pull it out,'' he replied. He stroked his leg with his trunk in an exaggerated
gesture of self approval.

As the start of class approached, Denart realized it would be a small group. Besides Rachel
there were only seven students seated when the professor walked in. He was surprisingly young.
Denart thought he was hardly any older than the students. That was a good thing. The old faculty
were typically very boring. The young ones usually told more jokes.

``Spring well to you all,'' he said. He put a large notebook on the lectern and stood facing the
class. ``I'm Colty Harnick, and this is math F-12, Combinatorics.'' He looked directly at
Rachel. ``I'd like to start by introducing Dr. 'pen'er. She is a visiting professor who will be
observing this class.''

Everyone looked at her and Rachel just nodded her head slightly the way a dunari might to
acknowledge an introduction. The professor then went around the room and had everyone else
briefly introduce themselves. Finally he opened the course book and the students, including
Rachel, all opened their notebooks and got ready to write. Denart didn't bother with that. He
didn't make it his habit to record every word the professor said. In fact, he usually didn't
write any notes at all.

Despite being young the professor obviously meant business. He dived right into the material,
defining terms and proving lemmas. There weren't any jokes. The material wasn't hard but Denart
decided he might need a notebook for this class after all. In the meantime he studied the board
as Harnick wrote and let the ideas fill his mind.

Now and then he glanced over toward Rachel. She also seemed to spend more time listening and
less time writing. Denart figured she probably already knew the answers. Yet now and then she
did write a few notes, sometimes using Argenian words and notation, and sometimes using a script
that was unlike anything Denart had ever seen before.

% TODO: Need to wrap this scene up.
