
``Your Eminence,'' she said while continuing to look at Perton, ``this is an exceptional case
and exceptional cases require exceptional evidence. It is my intention to show that the signing
blood on the assissination decree naming Rachel Spencer, and the signing blood on the Junar
declaration of juista against Yartagard, are both blood from the one and only Zarlon.''

Perton swayed, obviously anxious to rebut, but Jernumia held the floor.

``To do this I will bring testimony from two experts in genetics: a Dr. Warlic Tulinar from
Zarlock College and Brendon Fischer, a human molecular biologist aboard \textit{Golden Light}.''
Jernumia took a breath. ``To justify Fischer's testimony I invoke the principle of
\textit{huras} allowing one expert, Dr. Tulinar in this case, to support another. I also refer
specifically to the case of Bartoka vs Paliglia where a Forbin dignitary allowed himself to be
subject to Argenian law and by so doing was able to testify as a foreign observer.''

Jernumia stopped and the Eminence glanced toward Perton.

``\textit{Huras} does not apply here. The human technology is beyond our means and no testimony
of Dr. Tulinar, or any dunari, could support it. Let me remind the court that the principal of
\textit{huras} allows a more authorative expert to support evidence gathered by a less
authorative expert\ldots''

``Then do you stipulate,'' Jernumia interrupted, ``that Fischer's testimony is \emph{more}
authoritative than Dr. Tulinar's?''

Perton hesitated, but only briefly. ``I do stipulate that. Hence \textit{huras} does not apply
and Dr. Tulinar's testimony is irrelevant. I further stipulate, however, that Bartoka vs
Paliglia, and the other cases you mention in your \textit{suk-tina} are equally irrelevant. Even
if Fischer claims to be subject to Argenian law there are no consequences that result from that
assertion and so the claim is vacuous. The humans are not foreigners, they are aliens. That puts
them outside our law and their testimony is inherently invalid in our courts.''

Jernumia responded at once. ``Your Eminence, I intend to invoke the ancient right of
\textit{kolopta}. The human Rachel Spencer, who has lived among us for many years, is willing to
assume the legal responsibility for Fischer's testimony. If it is found perjurious she will
accept the consequences according to Argenian law.''

``The human 'pen'er may have lived among us,'' Perton continued, ``but it is still an alien
creature on this world. The right of \textit{kolopta} was only intended to apply between the
members of noble families in the age of the Regents so lesser family members could accept
punishment that might otherwise disrupt family structure. It is an affront to our history and to
the honor of those ancient families to apply it here.''

``Your Eminence,'' Jernumia said, ``the humans are indeed a family of sorts. They are a small
group very far from home and rely on each other as a family might. Furthermore Spencer is a
visiting professor at Parnon University where she is paid in Argenian julas for her services.
She's even paid the standard tax to the Argenian government and if that doesn't make her one of
us, I don't know what does.''

The audience shook their ears.

``More seriously,'' Jernumia continued, ``Spencer's employment by Parnon University clearly
indicates, according to Argenian tax law, that is has the legal status of a foreign observer and
can thus, legally, exercise the right of \textit{kolopta}.''

The Eminence clicked his teeth slowly and swayed his head slightly from side to side.
