
Colty met Rachel at her usual spot on the patio behind the Student Center. There, around the
corner, was a small table under the shade of a rigat tree yet still reachable by the cooling
Nermela. When Colty arrived Rachel was already there, sitting as she did in an upright posture.
A few paces away stood a well dressed but obviously atheletic man. He watched Colty carefully as
he approached. Colty nodded a greeting to him but the man didn't react.

``He doesn't say much, does he?'' Colty said to Rachel as he lay down beside her.

``Not much,'' she said. ``The other one is more talkative.''

``I'm glad you have some protection,'' Colty said.

``Yes, but I also have this.'' She lifted her arm and showed Colty a small bandage just below
her shoulder. ``It's an implant. They can now locate me anyone on Rujaria, even if I'm naked.''

Colty shook his ears slightly. ``That would have been useful.''

``Yeah. I should have had one before, but I don't really like the idea of them tracking me all
the time.'' Rachel took a piece of paper from her pouch. ``Anyway I wanted to talk to you about
something interesting.''

She laid the paper on the table and Colty could see that it was a picture of some kind. It
consisted of several bands of colored stripes. The stripes were very narrow and were in red,
yellow, gold, and green. Here and there were also some black stripes distributed randomly among
the others.

``This is a representation of a DNA fragment from the blood on that note we found,'' Rachel
said. ``Each colored strip is a nucleotide.''

``This is a picture of dunari DNA?'' Colty asked

``Well, just a fragment,'' Rachel said. ``there are only about 15,000 nucleotides here.''

Colty took the picture in his trunk and looked at it carefully. ``That's incredible,'' he said.
``Your technology is just amazing.''

Rachel touched the picture briefly. ``This is probably a junk fragment. It doesn't actually do
anything.''

``Junk?''

``Yeah, there are long sections in the DNA of every creature that don't encode any proteins. We
assume they are leftovers from the evolutionary process.''

Colty shrugged. ``So why do we care about this junk fragment?''

Rachel put her hands on the table. ``Because,'' she began. ``This junk fragment is Kessler J-7.
It appears in the genome of every vertebrate creature on Earth.'' Her eyes shone.

``Wait. Are you saying that the blood on that note is human blood?''

``No, no,'' Rachel objected. ``We decoded the entire genome. There's nothing human about it.
Except for this.''

``How could dunari and humans share the same junk DNA?'' Colty asked.

Rachel smiled with obvious excitement. ``I have no idea,'' she said. ``But there's got to be
some reason for it.'' Rachel paused. ``There's a biologist on \textit{Golden Light} named Brent.
He's very interested in exploring this more. Our ambassador is going to talk to Mr. Formand
about maybe doing a genetic examination of other Rujaran life forms.''

Rachel shifted in her seat and clasped her hands together. ``This is totally unexpected,'' she
said. Her voice was animated and she spoke quickly. ``This could be an important discovery.''

``Maybe it's parallel evolution,'' Colty said.

``Brent thought of that too but he says he's never heard of it manifesting like this.'' Rachel
sighed. ``It's funny how things work out sometimes. I came here to observe the far side of the
galaxy with the new telescope and yet maybe here\ldots'' She held her arm up against Colty's
trunk. ``in our blood is where the real discovery is to be made.''

``Maybe we're related,'' Colty said with a click of his teeth.
