
``It's just in here,'' Colty said as he pointed toward several doorways. Rachel was at his side
and next to her was Joleia. The three of them walked along the street toward the entrance to the
center district. They went between the tall sheet walls that protected the interior space from
the Nermela. The path zig-zagged several times to break up the wind. Posters and drawings
decorated the walls. Rachel seemed fascinated with it all. She stopped to look more closely at a
sign about a lost tranchen; someone's pet that had disappeared in the shuffle several hours
before.

``It's sad,'' she said.

Finally they came into the center district: a large area surrounded by shops and stores of all
kinds. Only foot traffic was allowed there. In the middle of the space was a sand park with
brushtik grass and rargen flowers. Around the park seats were sprinkled about where people could
relax and enjoy the sun.

Dolarta's was on the south side. It had a large entrance with a tall sign beside the doors
spelling out the name. Colty led the way with Joleia and Rachel behind. As they approached they
could see many dunari inside, some standing beside the wall tables and others laying on the
lower seats. Rachel hesitated but Joleia encouraged her.

``It will be fine,'' she said. ``We'll get a wall table near the back out of the way.''

As they entered many people turned to look at them. Some didn't seem to notice but kept talking
as usual but others fell silent as they passed. Rachel kept her head down and stayed behind her
hosts. Yet her height made her stand out.

Colty scanned the crowed and saw Tusk standing beside a wall table. He raised his trunk slightly
and the three of them made their way to his table.

``I thought this would be a good spot,'' Tusk said as they arrived. ``It's close enough to the
front and yet not in the middle of things.''

Rachel seemed relieved when they all knelt on the ground around the table. She took a position
next to the wall and tried to be as inconspicuous as she could. Joleia was next to her while
Colty and Tusk were across from her.

Tusk spoke to her softly in her own language and Rachel smiled briefly. ``Yes, thank you,'' she
replied in Argenian.

``You have such an interesting sounding language,'' Joleia said to Rachel, ``almost musical.''

The seesung arrived and then froze when he saw Rachel. Yet he quickly regained his composure.
``Welcome to Dolarta's,'' he said to everyone. He then put knock boxes on the table for all,
complete with knock sticks. ``Charara?'' he asked, ``Toran root?''

``Nothing for me,'' Tusk said.

``I'll have some toran root,'' Joleia said. Colty swayed his head also.

As the dunari took some money out of their pouches, the seesung looked at Rachel. ``Would you
like something?''

``Um,'' Rachel said uncertainly.

``He's just asking if you want anything to chew,'' Joleia explained.

``No, nothing,'' Rachel said, ``but you need some money?''

``Just the price of admission,'' the seesung said.

Tusk slid a few more coins across the table. ``I'll pay for her. She's my guest.''

After the seesung left Rachel spent time just looking around the room. The ceiling was open with
a fine mesh to break up the Nermela and to let the sun in. Around the sides there were several
doors through which people came and left. Not far from their table, close to the center of the
room was a raised open space on which stood an empty seat and a small stand. Joleia could see
Rachel looking at it.

``For the performers,'' she said.

``How is your research going?'' Tusk asked Colty.

``Uninspired,'' Colty said. He turned toward Rachel. ``I have almost finished that book you gave
me.''

``Already?'' Rachel said with surprise.

``It's fascinating.''

``What book is that?'' Tusk asked.

``Rachel gave me a human math book,'' Colty said.

Tusk cocked his head and clicked his teeth. ``Oh yes?''

``The approach is so meticulous,'' Colty said, ``such exquisite attention to detail. It's
wonderful.''

Rachel smiled. ``I'm glad you liked it, Colty. I'll get you another one.''

``You don't need to do that, Ra'hel.''

``I want to.''

At that moment the seesung returned with a bowl of charara and a bowl of toran root. He put them
on the table and then pulled a porlap feather from his belt and gave it to Rachel. She hesitated
but then reached out to take it. ``For our special guest this hour,'' the seesung said.

``Thank\ldots\ thank you,'' Rachel said.

Joleia clicked her teeth. ``It's an honor to get the feather,'' she said.

``What do I do with it?''

``Just keep it with you,'' the seesung explained, ``and let the players see it so they know they
play for you.''

After the seesung left Joleia took a bit of toran root and put it in her mouth. Colty did the
same with the charara. Tusk pushed the charara bowl toward the far end of the table. ``Keep that
away from me,'' he said, ``but I will try the toran.''

``It's a little intoxicating,'' Joleia said to Rachel as she chewed. ``I suppose your people
don't do anything like this.''

Rachel laughed and smiled. It was a strange gesture but the three dunari with her had come to
know it. She held up her knock box. ``What are these?'' She had to speak in loud tones to be
heard over the increasing noise from all the patrons.

``Knock boxes,'' Colty said. ``When the time comes you hit them with the knock stick.''

``What time?''

``You'll know,'' he said with a click of his teeth.

``Do you have music on Earth?'' Joleia asked. ``You know, sounds that are nice to hear?''

``Oh yes,'' Rachel said. Suddenly the crowd settled down and everyone looked toward the raised
platform where the seesung was standing with his trunk pointed almost straight up. He waited for
several long moments until the all the talking stopped and the room was strangely quiet. Then he
brought his trunk down and held it out to the crowd.

``Tonight,'' he began, ``our feathered guest is visiting from afar.'' He pointed toward Rachel
and everyone turned to look at her. ``Raise up your feather for all to see!''

Rachel picked up the feather from the table in front of her and briefly held it up.

``Does it crow?'' a voice from the back said. Several people shook their ears.

``That is not for me to say,'' the seesung replied. ``What I can say, though, is that in this
hour we will hear the crowing of many. Who shall start?''

A young woman on the opposite side of the room stood up. ``I will,'' she said.

``Then come forward,'' the seesung replied.

The woman took a long gurt from the table where she had been standing and walked onto the
platform. The seesung stepped down while the woman lay on the seat with the gurt across her
knees. She adjusted the strings slightly. She turned toward Rachel. ``Feathered one,'' she
began, ``this music is from the hills of Basuptia in the north. It is called \textit{Shallow
  Speech}.''

The woman then began to play. It was a bright, lyrical piece with a moderate tempo and notes
that rose and fell like the gusts of the Nermela. As she played people began to join her using
their knock boxes, tapping out the rhythm in time with her music. At first only a few joined in,
but then more and more as the intensity of the sound built. Joleia joined and so did Colty and
even Rachel used her knock box. Tusk just watched clicking his teeth in time with the music.
When the woman finished everyone stomped their feet and Rachel hit the floor with her hand.

``Very nice Hills of Basuptia,'' the seesung said as he again took the platform while the woman
returned to her seat. ``Who is next to crow?''

More and more people came forward to play. They had all manner of instruments and played all
manner of pieces. Some where slow and sad. Some where light and happy. Some called out for
motion and often people got up to sway and stomp. The knock boxes were used often and Rachel
even clapped her hands together at times. It was an unexpected gesture that seemed to please the
crowd.

After many players had come to the platform the seesung started having trouble getting people to
come forward. Tradition barred those who had gone from going again and fewer and fewer players
remained.

``Eventually we run out of players,'' Joleia explained, ``and then it's over until next time.''

``Who is next to crow?'' the seesung called out but there was no response. ``Nobody?''

Suddenly Rachel raised her arm. ``I will crow,'' she said loudly.

The seesung nodded her way and clicked his teeth. ``Then come forward,'' he said.

The room grew quiet as Rachel stood up and walked toward the platform as the seesung withdrew.
She knelt down on the seat and looked around the nervously. Everyone waited.

``Um\ldots'' she began. ``This music is from a country called \alien{New Wales}, on Earth.
I\ldots\ I don't have any instruments, but I will do my best without. I'm sorry that I have to
do it in my own language so you won't understand the words, but the music won't work
otherwise.''

She closed her eyes and took a few deep breaths. Then she opened her mouth and spoke. Yet it
wasn't like ordinary speech. Instead her voice formed notes, changing in pitch and intensity as
much as the notes from a gurt or any instrument. Yet among those notes were also words. They
were words in her own language but they resonated together, connecting to each other in a way
normal speech did not. The audience was transfixed.

Her song was slow and filled with longing sadness. It was a song of a broken heart and
resignation. Yet buried under the pain there was also a sense of resolve and a feeling of
strength. It was also a song of hope. The flavor of the music was evident to all even if the
words where not.

At the end her voice taper off and stopped. She paused for a moment and opened her eyes. She
smiled briefly and the room was filled with stomping.

The seesung stepped back up to the platform. ``That was a lovely crowing from the feathered
one,'' he said with a nod, ``lovely and unique.'' He flapped his ears.

``What do the words mean?'' someone in the crowd called out.

``It's about a woman who lost her husband at sea,'' she explained, ``she is wishing him well
where ever he might go, and sending him her love.''

``Do you have seas on Earth, then?'' another voice called out.

``Oh yes,'' said Rachel. ``In fact, almost three quarters of my world is covered by a great sea
that is larger than all the seas of Rujaria put together. We have a special word for it. We call
it the \alien{ocean}.''

``I heard that Earth is very cold, is that really true?'' It was someone at the table right
behind Colty. Rachel answered the question, and many others, as the people in Dolarta's
expressed their curiosity about Earth openly. The seesung did nothing to discourage it, and
Rachel stayed in the raised seat for many arnets talking about her world in ways no other human
ever had.

She spoke of tall, snow covered mountains, wide plains of grass, forests of green trees, and of
great rivers. She spoke of buildings hundreds of stories tall, and of cities orbiting Earth in
space. She spoke of day and night, sunsets and sunrises, and of seeing thousands of stars in the
sky.

``I hope one day you are able to come to Earth and see for yourself,'' she said. ``My world is a
beautiful place\ldots\ just as yours is.''

% I like the connection between Rachel talking about the plants on Earth and her interest in the
% flowers around Parnon. I like how this makes Rachel seem less alien to Colty. Perhaps this is
% where their friendship really... er... blooms?

There was a hush in the room as she spoke, and her words seemed to transport everyone present
across the gulf of space to a place that seemed both alien and yet familiar. Colty watched
Joleia as she listened with rapt attention. When he turned back toward the platform Rachel was
talking about the wild flowers that grew along the roadside not far from her home. It made him
think of her fascination with the red topolia flowers on the Parnon campus and he felt a strange
kinship with her. She was not, perhaps, quite so alien after all.

