
``There is no way we can evacuate the city in time,'' Porganen said.

Fotkey looked grim.

``We can get the First Citizen out,'' Porganen continued, ``and some government officials. If we
make a general announcement there will be mass panic and nobody will get out.''

The First Citizen said. ``If I impose military law perhaps we can force an orderly evacuation.''

``It might work for a while,'' Porganen replied. ``But as people get more desperate it will
break down.''

``We must do what we can.''

Rachel reached into her pouch and took out her machine and unfolded it. She touched the screen
lightly and started at it intently. Marshall watched her closely.

``Fotkey, use whatever forces you have,'' the First Citizen commanded. ``Get my people out of
Varnok.''

``Yes, Sir.'' Fotkey left immediately.

``Are you sure all this is necessary?'' Hutlan asked. ``The weapon is sealed in concrete well
below ground level. It might destroy the Toman building but that hardly justifies evacuating the
entire city.''

Paraz snorted. ``Did you see the video of Yartagard? There was nothing left. The weapon here is
more powerful. I estimate it will completely destroy the entire downtown district and cause
extensive destruction throughout the rest of the city.''

The First Citizen turned toward Marshall. ``Tom, would your people be able to help with the
evacuation?''

``Yes, probably,'' Marshall replied. ``We could maybe move some people by air. I'll work on
setting it up.''

``Hold on,'' Rachel said. Everyone turned toward her. She was still looking intently at her
machine. After a moment she looked up.

``Tom, I think we could use a spacial displacement field around the device to contain the
explosion.''

``Really?'' Marshall replied. ``Won't the explosion disrupt the field?''

``I just did a quick calculation,'' Rachel replied. ``I think we can support the field with
enough energy, if we make the volume large enough, to make it work. I'll know more once I model
it better.''

Paraz interrupted. ``What is this `spacial displacement field?' ''

Rachel folded up her machine. ``The idea is to enclose a region of space and, in effect,
separate it from the rest of this universe. If the explosion happens inside such a region, it
will have no effect in this universe.''

``That's utterly incredible,'' Paraz said.

``We do this sort of thing to build our faster-than-light engines.''

``Yes,'' Marshall said, ``but you are talking about enclosing a major explosion. If the field
loses integrity the explosion will escape and we'll be no further ahead.''

``It might even be worse,'' Rachel said. ``We'd basically be creating a constrained reaction
chamber so the reaction will be more complete and more energetic than normal. Yet if it works
the damage to Varnok would be minimal.''

``You said the explosion would have no effect,'' Paraz said.

``The explosion, yes, but raising a spacial displacement field here on the ground will cut
through any object in its way. It will damage the Toman building significant.''

``That is a small price,'' the First Citizen said. ``Will it work?''

``I'll have to do some more calculations. If Dr. Perez helps the results will be more accurate.
I'll need to know all the details of the composition and design of the weapon.''

``Do it,'' Marshall said, ``with the permission of the FC, of course.''

``I'll need to borrow the computers on \textit{Golden Light}. The modeling computations will be
significant.''

% Someone, probably Perez, should ask about using the spatial displacement field to slice
% through the weapon, making it inoperable. For example perhaps the field could separate the
% subcritical masses in the weapon so they can't be combined. The characters should probably
% discuss this but ultimate decide that it won't work because they don't know the exact location
% of the weapon. If they miss they could end up with an explosion in their universe after all.
