
% SUMMARY: They raise the field.

% Dunari... Fotkey (and some military personnel), Tusk, Paraz.
% Human... Rachel, three others.

% I probably need some preliminaries to warm the scene up. I'm not sure how that should work.

The humans unrolled some flexible sheets of a thin material. It was slightly cloudy but still
easy to see through. ``We need to hang these in front of the windows and walls,'' Rachel said.
``They will block the radiation.''

``I thought you said nothing would escape from the field,'' Paraz replied.

``I could be wrong,'' Rachel said with a wry smile. ``It's normally considered a bad idea to
raise the field through solid objects. As it cuts through the atoms in the concrete there will
be a veritable stew of radiation produced.''

Fotkey held the thin sheet in his tentacles. ``Can we hammer nails through this?''

``I should think so,'' Rachel said.

Fotkey ordered his men to start hanging the sheets along the side wall of the cafe while Rachel
helped set up some of the equipment brought down from \textit{Golden Light}. Parez and Tusk
watched while Rachel and the two other humans chittered to each other rapidly. After a moment
Parez touched Rachel's shoulder and asked, ``what should we be doing?''

One of the other humans made a gesture with its hand. ``Emit,'' he said in broken Argenian.

``Yes,'' Rachel replied. She turned toward the dunari. ``This is were we need your help.'' She
walked over to one of the tables and called for Fotkey to join them. She unfolded her machine,
and activated it. At once an image of the street layout appeared on the display.

``We need to set up field emitters around the space we are trying to enclose.'' She pointed at
certain spots on the display that were marked with small circles. ``We can mount some on poles
here and here, but we also need to get emitters below the space here, here, and here.''

``This spot is probably in the basement of the building,'' Fotkey observed touching the display
with one tentacle. ``But we'll have to drill holes in the street for these other two.''

``Yes, exactly,'' Rachel said. ``They need to be as close to these locations as possible and to
the depth I've calculated here.''

``I'll get the drill equipment in position and get that work started,'' Fotkey said.

Rachel turned toward Tusk. ``You speak English, yes?'' she asked.

``I try,'' Tusk said.

``Then let me introduce you to Yang,'' Rachel said. ``She could use some help getting the cables
ready. We need to connect the energy supply to the emitters and we'll need cables on the street
for that.''

Yang chittered a greeting and Tusk clicked his teeth briefly before replying in kind.

``What can I do?'' Parez asked.

``We'll need to synchronize the emitters to this reference frame,'' Rachel said. ``We're on a
rotating planet moving around your star. When we raise the field we have to keep it locked in
our reference frame or else it will move relative to us.'' She paused. ``That's bad.''

The other human pointed toward several cylindrical metallic devices laying on the floor beside
their equipment. Rachel walked over to one of them motioning to Parez. ``Let's get this out on
the street where we have some room,'' she said. ``John here is an engineer and will do the
calibration, but we'll have to move the gyros manually I think.''

John grabbed two machines, one in each hand, as Rachel and Parez moved the first emitter to the
sidewalk just outside the cafe.

``It this really part of your faster than light engine?'' Parez asked.

``Yeah,'' said Rachel. ``In fact right now \textit{Golden Light} can't move on its own power
until we get these returned.''

``What if they never get returned?'' Parez asked.

Rachel shrugged. ``They will have to call Earth and wait for more to be sent.''

After connecting his machines to the device, John fussed with the display on one of the machines
for a while. He chittered to Rachel and the two of them talked back and forth animatedly. Rachel
shook her head slightly and turned to Parez. ``We usually do this sort of thing in a
laboratory,'' she said. ``John's a bit unhappy with the noise level here.''

Parez clicked his teeth. ``Your engineers sound a lot like ours.''

``These are definitely not ideal conditions,'' Rachel said. ``If we had more time we could
have automated the whole thing and been a thousand karnons from here when the critical moment
came.''

Fotkey's men started drilling into the street filling the air with loud rattling noise. The
ground vibrated around them.

% blah, blah...

Back inside Rachel sat down at the machines set up on the table. It was a strange contrast. The
cafe's simple decor surrounded several electronic devices the likes of which no dunari had ever
seen before. The displays were covered with human symbols and many mysterious graphic images.
Rachel put a head listener on, securing it behind her ear. As she did the display in front of
her came alive. Staring at it intently she manipulated the symbols and images with the power of
her thoughts. After a moment she reached up with one hand and touched the screen of the other
machine to manipulate some information displayed there as well.

Parez stood behind her watching intently. There was so much he wanted to ask but he also didn't
want to distract her. Presently Rachel stopped and turned toward him. ``I've modeled the
explosion,'' she said as she tapped the screen in front of her. ``That's what this is. As you
know there are a lot of unknowns that I had to guess about. Now that we've got the calibration
information on the emitters I can get a better idea of how it's going to work.''

``These machines do all these calculations?'' Parez asked.

``Yes,'' Rachel replied. She seemed to be copying information from one place to another. As she
did so, the colors on the image before her shifted. She studied the image carefully. ``It should
work,'' she said, ``but it will be a bit close.''

Parez kneeled down next to her. ``I'd love to use this technology in my research,'' he said.

Rachel smiled slightly. ``It is very useful,'' she said. ``Your people will be building machines
like this in no time, I'm sure.'' She leaned back a bit and made a few more adjustments on the
image. ``Normally we can use our machines to control as well as calculate. But these emitters
were not designed to be used this way and we haven't had time to adapt our control systems.''

Rachel turned toward Parez. ``We humans have a saying that describes what we're doing here.
We're winging it.''

Finally after many arnets all the emitters were in place and all the cables were connected. The
atmosphere in the cafe, now control room, was tense. Fotkey looked at the clock.

``Will the Junar be on time?'' Rachel asked.

``Yes,'' Fotkey replied. ``We have only about 20 arnets to go.''

``Then it's time to raise the field. We need to get every non-essential person out of here,''
Rachel said. ``General you should send your men away. They've done everything they can.'' Rachel
turned toward Parez. ``You should go too. You've been very helpful but you don't want to die
here if it comes to that.''

``No way,'' Parez said. ``I need to see this if it's the last thing I do. I promise I'll stay
out of your way.''

``I'm staying also,'' Tusk said. ``You might need me to translate some more.''

``Tusk, no,'' Rachel replied. ``You have a wife and children. Go to them. Anyway if this fails
your language skills will be needed more than ever by your people.''

``I don't want to abandon you,'' Tusk said.

``It's not about that,'' Rachel said. ``I can translate if necessary. There is no need for you
to risk your life here any more.''

Tusk hesitated. He swayed his head. ``Okay,'' he said finally. ``But I better see you again,
Rachel, or I'm going to be very angry with you.''

Rachel laughed. ``It's a deal, then.''

The dunari packed into a military truck and drove off leaving a cloud of dust in the street.
Parez wondered if they could get far enough away in just 20 arnets, but there was nothing to be
done about that. In the end the only people left where Parez, Fotkey, Rachel, and two other
humans. Just five small individuals to stop what promised to be the most powerful explosion ever
seen. Yet they had something the Junar hadn't anticipated. They had human technology.

Rachel and the two other humans knelt down before their machines. They chittered among
themselves intently as they worked the information displayed there. Fotkey and Parez stood back
some distance and just waited.

After a moment there was a kind of hum followed by a faint hiss. Parez looked out the window of
the cafe and saw a greenish glow surrounding the end of the emitter positioned there. As he
watched the glow seemed to extend from the emitter, like ghostly tentacles, reaching toward the
Tomar building and toward the ground. It flashed and flickered slight as it moved but this
seemed to displease the humans.

Yang stood up and started inspecting the mass of cabling that connected their control machines
to the power supply to the emitters. She chittered to Rachel and John who replied back in kind.
Finally she seemed to find what she was looking for. She made a few adjustments and the field
stablized.

``Bad connection,'' Rachel called out. ``Fixed now.''

Parez leaned over the Fotkey. ``They are winging it,'' he said.

``Let's pray whatever they are doing works,'' Fotkey said simply.

The green glow grew steadily. Parez could see it coming out of the ground, no doubt from the
emitters that were buried below the street level. The glow seemed to thicken and where it
intersected with the building the view through the greenish haze grew blurry.

The humming hiss grew louder and the humans seemed content. Rachel made some adjustments on her
machine and watched the evolving field. ``Everything is fine,'' she called out. ``The field will
start cutting through matter soon. Let's hope the building doesn't fall on us!''

As Parez watched the glow became more of a cloud. He could still see the building through the
haze but with difficulty. The hum became a kind of long, screeching scream, rising in pitch
toward the edge of hearing. Parez heard the sound of glass breaking and saw some windows on the
side of the building, above where the field was forming, shatter. Fotkey took a deep breath next
to him.

Suddenly a bright yellow glowing line appeared on the side of the building and along the ground.
It was just inside the green haze and formed a circular path. Rachel watched the effect as well.
The screeching sound went ultrasonic and a strange quiet descended on the scene. Yet shortly
after that there was a gut wrenching crack and seemed to shake the ground. Parez saw a crack in
the wall of the Tomar building stretching up the side at least two stories above the field.
Then, to his amazement, the very stones from which the building was made seemed to bend like
some kind of strange taffy. A lound groan of twisted metal and cracking rock filled the air.
Even the road in front of the building, outlined by the bright yellow line started to twist in
an impossible way.

Fotkey braced himself against the wall rigidly and watched uneasily. Strangely Parez did not
feel afraid. He was witnessing something beyond his wildest dreams. He was actually seeing a
piece of our universe been torn away.

The humans seemed happy with how things were going. The yellow green fog had become nearly
inpentrable and started to look more like a wall. Parez could barely see through it and the
sounds of breaking stone seemed to receed into the distance as if moving away through a long
tunnel.

The field continued to strengthen and soon became a ball of shimmering yellow light. It rose out
of the ground and curved into the wall of the Tomar building, wrapping itself around the spot
where the weapon was embedded.

``We will reach pinch-off soon,'' Rachel called out. As she spoke the glittering ball seemed to
brighten and then turn reddish. The surface grew smooth and solid until it looked almost like a
large marble, nearly a full story high, buried partway into street and into the Tomar building.

Suddenly, almost like a switch was thrown, the surface of the ball become nearly perfectly
reflecting. It had become a large spherical mirror. Parez could easily see the cafe in its
curved surface and the buildings down the street as well. The light from the Rujaran sun
glinted off the field and flooded the cafe with its ruddy glow.

``That's it,'' Rachel said excitedly. The humans all seemed very pleased and chittered between
themselves enthusiastically.

Rachel stood up and walked over to the two dunari. Fotkey seemed to relax a little. ``We've
pinched off that volume of space from this universe,'' Rachel said. ``If all goes well, when the
explosion happens we'll see no effect here whatsoever.''

Parez stepped forward staring at the sphere intently. ``That's incredible,'' he said. ``Does the
light reflect from it because it can't penetrate it and can't be absorbed by it?''

Rachel smiled slightly. ``Yes. That boundary is essentially the edge of the universe. For us
there is nothing inside, not even space.''

``That doesn't make any sense to me,'' Parez said.

Rachel laughed. ``It isn't very intuitive. Maybe another time I can show you the math.''

``Oh you are teasing me now,'' Parez said with a shake of his ears.

``So what happens now?'' Fotkey asked.

``We wait,'' said Rachel. ``Well know when the explosion occurs by our instruments. In the
meantime, I guess we can just relax.''
