
When the van arrived at the human compound Dr. Spencer was waiting outside. She climbed into the
back and sat in the special seat they had constructed for her. She sat in a mostly upright
position, her head just touching the roof of the van, with her legs tucked underneath the seat.
Colty laid down on the seat next to her.

As the van made its way through Parnon the human looked out the window with curiosity. She
wasn't monstrous at all. In fact to Colty she seemed almost like a small child seeing wondrous
sights for the first time. Colty thought that Rujaria must be very strange to her.

``Dr\ldots ah,'' Colty began. ``Dr. 'pen'er\ldots''

She turned toward him. ``You can call me \alien{Rachel},'' she said.

``That is easier for me to say,'' Colty admitted, ``but I don't want to address you
inappropriately.''

``It is fine,'' she said. ``Rachel is my familiar name, but I know it is easier for you. I don't
mind.''

[I need a different question here. Rachel is asked this at the reception.]
Colty took a breath and started over. ``Well\ldots Ra'hel,'' he began. ``I hope your trip here
was a good one.''

``It was very good, thank you,'' she said.

At the university the van parked beside the campus center. Parnon University's tradition was to
welcome all visiting professors with a small reception. Due to the special circumstances it was
suggested by some to skip the reception in Rachel's case. However, most department faculty felt
that the tradition should be honored.

The reception included a collection of faculty, mostly from the Physics or Mathematics
departments, a few advanced graduate students, some university officials, a couple of invited
members of the press, and Lucasa. In total perhaps twenty or thirty people were there. As was
traditional at these events several faculty and their students had erected posters describing
their research.

When she arrived Rachel took on a more formal posture than she had while riding in the van. She
walked slowly and kept her arms close to her body. Most people at the reception kept their
distance from her as well. People were nervous around her. She didn't appear hideous to dunari
eyes, but she wasn't dunari and most people didn't know how to respond to her. Some simply
avoided her.

Rachel looked at the posters and asked a few questions but mostly she didn't say very much.
Colty got the impression that she was as nervous as everyone else. He pointed out the names and
positions of various people, and like any good host might, explained to her a few things about
Argenian social customs.

``The students won't initiate a conversation with you, but it is considered appropriate to
greet and acknowledge them,'' Colty told her.

``Thank you,'' she replied. Colty was taken by how polite she was.

The people who did talk to her mostly asked her simple questions but some of them took a turn
toward the surreal because of the circumstances.

``How was your trip?'' asked Dr. Marley. He was a specialist in solid state theory. She
hesitated and then said, ``It was long but comfortable enough.''

``It sounds like it wasn't as comfortable as you would have liked,'' Dr. Marley continued.

``That's true,'' Rachel admitted. ``The accommodations were very small to save mass.''

The physicists all nodded. ``I can imagine that would be a concern,'' Dr. Marley said. ``I
recently took a trip to a conference in Burlusket and the train cars were exceptionally small.
It wasn't very nice.''

None of the established scientists asked her about anything of a technical or personal nature.
The students, however, were not so inhibited.

When she talked with a group of graduate students one of them quite directly asked ``I'm not
clear about what, exactly, this telescope your people are building will observe. Can you explain
that?''

Colty thought Rachel looked distinctly relieved to be talking science. ``We are interested in
observing the universe when it was only a few seconds old,'' she said.

``Whoa, what?'' said the student. The group looked stunned. ``So\ldots are you saying that
the universe is not infinitely old?''

Rachel hesitated. ``Well\ldots it depends on how you define your terms,'' she said
enigmatically.

Before the first student could follow up on her comment, another student jumped in. ``I
understand that Earth rotates quite rapidly,'' he said.

``Yes,'' replied Rachel. ``In only about three of your hours. Earth is not locked to our sun so
it alternates between day and night\ldots like Fresnik''

Everyone paused to reflect on her comment. ``That\ldots must be strange,'' the student said.

The first student spoke up again, ``It has been speculated that life can't evolve on a planet
with alternating day and night.''

``Oh yes?'' said Rachel with interest.

``Yes, the idea is that during the night the compounds built up in the day would just break down
again and life would never get started.''

Although Colty wasn't quite sure how he could tell, Rachel seemed much more at ease. She held
her body less stiffly and seemed more animated than before. ``Well,'' she said. ``I'm here to
tell you that it can happen.''

Rachel ended up speaking with the students for a long time, a fact that somewhat annoyed the
regular faculty. With Colty's encouragement she finally broke off from them and mingled with the
rest of the group.

As the reception was winding down, Dr. Dargon invited Rachel to see the office space they
created for her. Dargon, Colty, and Lucasa showed her the way across campus to the mathematics
building.

``You are a physicist,'' Colty said as they walked, ``yet you asked to be a part of the math
department. Why?''

``Where I come from,'' Rachel said, ``there isn't much difference between mathematics and
physics.''

Her office was located at the end of the first floor hall. It had been a small office with a
storage closet but now it was a somewhat more comfortable office with some toilet facilities.
The duncolans renovated the space, but the humans installed the facilities. Apparently the
humans were very sensitive about ``losing organic matter'' from the closed loop processing
aboard \textit{Golden Light}. Rachel ate food from Earth and her biological by products had to
enter into the recycling system that would eventually return it all to Earth in one form or
another.

``So in other words, they eat their own waste,'' Rutbarn had ungraciously said during one of the
department meetings leading up to Rachel's visit.

``We all do,'' Colty reminded him.

Rachel seemed pleased with her office. ``I appreciate all the effort you have gone through on my
behalf.'' The humans also installed a air cooling device so that Rachel could lower the
temperature in her office to the chilly level that humans seem to prefer. Compared to a duncolan
unit the human device was smaller, quieter, consumed less power, and was more effective.

After reviewing her space, Rachel said that she wanted to return to the compound for a while.
``I will need to sleep soon,'' she said. ``I would like to return at hour 63 to settle in here,
if that is acceptable.''

``Yes, yes,'' said Dargon. ``But you won't get much sleep before hour 63.''

``I only sleep an hour at a time,'' she said. ``My biochemistry is synchronized to Earth's
rotation. At home we sleep at night and are awake during the day\ldots all in only three
hours.''

``Of course,'' Dargon replied. ``I should have realized.''
