
% SUMMARY: Rachel gives a Ronbar lecture

As usual Denart sat in the back. He was one of the few students invited to attend Rachel's
Ronbar Lecture thanks to her explicit recommendation. He was surprised when she spoke to him
about it. He didn't understand why she would want a biology student in a lecture about physics.

Denart watched as some of the greatest scientific minds in Argenia gathered in the hall. Some he
knew as faculty at Parnon. Many he didn't recognize. The ambassador to Earth was there, standing
beside one of the doors greeting people who entered. Next to him stood a large man he had also
seen on television. The scientists mulled about, standing in the aisles or on the floor
chattering among themselves. Gathered to one side was a small group of Forbin scientists who had
traveled from the other side of the world by aerocar to attend.

The Ronbar Lecture Hall was the biggest hall on campus with steeply arranged seating surrounding
in a semi-circular arc a large, open floor. A small podium had been placed in the middle of that
floor and behind and above it hung a large screen. The projector sitting on the projection wedge
several rows back looked like nothing Denart, or anyone else on Rujaria, had ever seen. It was
much smaller than an ordinary projector and had no obvious controls.

Rachel sat quietly by herself in the front row. She seemed lost in thought. Perhaps she was
reflecting on what she was about to say, reviewing complex equations in her mind. Perhaps she
was meditating in some way, following some sort of exotic human ritual. Perhaps she was asleep.

Finally Rocalla moved to the podium and stood quietly as the audience took their seats and
settled down. Then he spoke.

``It is my honor,'' he said, ``to open the 237th Ronbar Distinguished Lecture.'' There was some
scattered stomping. ``Today we have a very special guest who has traveled an exceptionally long
distance to be with us.'' There was some scattered ear shaking.

``Our speaker for this year's lecture,'' Rocalla continued, ``is Dr\ldots\ \alien{Rachel
  'pencer}.'' Rocalla did his best to pronouce her name properly but he still glanced her way as
if to check if he got it right. She didn't make a sign.

``She is a professor of Physics at \ldots\ \alien{Sunapee College}\ldots'' Rocalla paused. ``I
hope that's right,'' he added.

``Yes.'' Rachel's thin, high pitched voice was soft but still clear in the hall's fine
acoustics.

``Yes?''

``Yes, just fine,'' she said.

With the tricky part over, Rocalla continued with more confidence. ``Dr. 'pencer has many awards
including a \alien{Nobel} Prize in Physics\ldots\ I understand that to be similar to our Brulark
Medallion\ldots\ and several best paper awards. She has authored two books, one of which is
considered a landmark work in its field, and she has also been awarded the designation of
`Master Teacher.' She has published in diverse areas including physics, mathematics, logic,
and\ldots\ philosophy. However, her primary research interest is in the logical consistency of
physics. Dr. 'pencer\ldots''

Rocalla stepped down and the hall filled with respectful stomping. Rachel walked toward the
podium. She did not rush but, in her usual style, moved in a light, unassuming way. She wore a
kind of tunic-like garment that covered most of her body with leggings and high boot-like foot
covers. Her clothes were a subdued color but trimmed with green. Her black hair was tied back to
keep it out of her face. In many ways she looked like a well dressed chornak but otherwise her
clothing and her manners were surprisingly ordinary.

She didn't stand behind the podium but instead stood next to it, surveying the hall briefly
before starting to speak. ``Thank you, Dr. Rocalla, for your introduction,'' she began. She
appeared to be wearing a wireless microphone of some sort\ldots\ another bit of human
technology. ``I am honored to be here. It is my hope that this marks the beginning of a long and
fruitful collaboration between our peoples.''

She paused and then stepped forward out into the middle of the floor and well in front of the
podium. Denart clicked his teeth slightly. He suspected the podium was not her idea.

``Why does the spectrum of a radiating black body have its distinctive shape? Why do protons
have the mass that they have? Why does anything exist?'' She took a few steps toward the left as
if she was speaking directly to the people on that side of the hall. ``As the questions get
simpler the problems get harder. It is the questions a child might ask that challenge us the
most.'' She took some steps toward the right and spoke to the other side of the hall. ``Yet it
is the answer to the simplest of questions that lead to the most profound insights.''

She returned to the center of the floor. ``Light is waves, but waves in what? The effort to
answer that question ultimately led to our first truly successful theory of gravity. A theory
so profound that it predicted results that, in the words of the physicist who initially
developed it, were too weird to be true. Yet they are true.''

Rachel walked to the podium and picked up a small headset. She slipped the thin band over her
hair and positioned each end just behind her ears. She paused and seemed to concentrate
momentarily and then continued. ``I don't normally wear these,'' she explained, ``but they do
make things convenient.'' Suddenly the projector came to life filling the screen with an
artistic design of curved lines and yellow-tan colors.

She turned to look at the screen and a white rectangular region appeared with lines of alien
text. Briefly one of the lines flashed and then the screen was filled with an astounding image.
On the left a white hot star glowed brightly. Yet it wasn't round. Instead it was distorted into
the shape of a glowing water drop. The sharp tip of the drop pointed to the right where, a short
distance away, there was a bright twisted disk of matter.

Some more text flashed briefly on the screen and then a scale appeared along the bottom. As if
by its own will a pointer slid along that scale a short distance before vanishing. Then the
image started to move. Globs of glowing gas were gathered from the space around the star and
swirled into the distorted disk.

Rachel turned back to the hall. ``This is \alien{Cygnus X-1},'' she said. A small arrow appeared
on the screen pointing at the disk. ``It's a major source of X radiation and has long been a
subject of study. These pictures were taken by an automated space craft we sent to the system to
investigate.'' The picture zoomed in on the disk until it overflowed the edge of the screen. At
its center where the view was highly twisted was an unexpected silhouette: a tiny dark spot in
the midst of a swirling maelstrom.

``That dark object at the center,'' Rachel said, ``has 100 times the mass of your sun yet is
only 20 karnons across.''

``Why is the rotation around it so irregular?'' a voice came from the middle of the room.

``The gravitational field from the object is distorting the structure of space in its vicinity.
The disk is actually fairly uniform but the light from the far side curves around the central
object yielding this strange appearance.''

A silent hush passed over the room.

% Actually there is a point to this. Rachel is foreshadowing Colty's discovery. I need to make
% that connection more clearly. Also this scene ends a bit abruptly (but at least it's far
% better than any of my earlier versions!).

``At least that is the simplest way to look at it,'' Rachel said with a slight smile. ``The
important point is that the theory that first predicted such things arose from a simple problem
phrased as a simple question.''

Rachel continued with several other examples. She talked about experiments, both human and
dunari with strange results and the basic motivations for exploring them.

Finally she paused and looked at the clock in the back of the room. ``I'm out of time for
today,'' she said. ``But before you go remember to never let yourself become too complacent
about your work. Let yourself be filled with wonder. The structure of the universe is revealed
in the simplest questions about the simplest systems.''

Rachel paused and said, ``That is all for now. Be well and learn well.''
