
% SUMMARY: Paraz is introduced. There is a threshold accident at the Rangard research station.

Lendon Paraz reviewed his notes as he rode in the back seat of the car. He didn't notice the
bumps as they jostled down the rough road. He had been that way many times before. The great,
dusty expanse of the Rangard desert stretched away in every direction about them.

The driver glanced at Paraz in the rear-view mirror but said nothing. His military training was
clear. It was not his place to engage someone such as Paraz in idle conversation. Even though
Paraz was a civilian he was, quite probably, the most brilliant physicist in the world.

Paraz leaned back and closed his eyes. He was tired and the season was still young. He knew he
worked too hard, but it was necessary and important work. At least that's what he told himself.

The car rounded a familiar corner and Paraz opened his eyes. He started to organize his notes
and close them back into their case, complete with tri-locks. Ahead, out the window, he could
see the base coming into view around a low rise. It wasn't exactly home, but after spending so
much time there it felt that way. He was glad his long trip from Varnok was nearly at an end. As
much as he respected the First Citizen, the regular face to face meetings took away too much
time from his work.

Without a word, the driver approached the gate. He stopped his car while several guards came
forward, weapons at the ready. It was routine.

``Dr. Paraz is here, returning from Varnok,'' the driver said. The guards inspected the driver's
identification. Paraz had his ready as well. The first guard looked at him suspiciously while
the other guard circled the car, taping on the sides and looking underneath.

``When did you last leave Rangard?'' the guard asked.

``Hour 50:080 last year,'' Paraz said.

``I don't believe you had that case with you when you left,'' the guard said. It wasn't true.

``I did have it,'' he said simply.

``Please step out of the car,'' the guard said.

The driver and Paraz got out. Paraz was unconcerned. The security at Rangard was greater than at
any other military base in Argenia. This ritual, slightly different every time, was designed to
confuse and intimidate would-be infiltrators.

``Open the case,'' the guard demanded.

``That is forbidden by Order 197 issued directly by General Fotkey,'' Paraz replied. The guard
said nothing. Paraz knew that if he had opened the case he would have been arrested immediately.

They stood by while the second guard pulled the seats out of the car. The first guard asked some
questions about their trip and took photographs of them. Finally after several arnets they
assembled the car again. The first guard clicked his teeth. ``Welcome back, Dr. Paraz.''

``Thank you.''

Paraz and the driver got back in the car and headed toward the main building. The base at
Rangard wasn't anything special. It mostly contained garages for experimental tanks that could
withstand the scorching heat of the deep day. It was a real research program, actively pursued.
However, Paraz wasn't involved with it.

When they arrived at the main building, two more guards stood by the door. Paraz got out, with
his case, and headed inside where he was greeted by a receptionist. ``Good Summer, Doctor,'' she
said, ``they want you in the underworld right away.''

He walked briskly down the hall to a plain looking elevator door at the end. From there he
descended down three levels to the underworld. He had grown used to the confined space. It
wasn't natural but at least it was well lite. Many years ago there had been a power failure in
the underworld surrounding those unfortunate enough to be there with the darkness of night.
Brave solders wailed with desperate fear until someone managed to activate the secondary
generator. Now, with multiple backup systems routinely tested Paraz didn't worry about that
happening again. He had much bigger things to worry about.

The door at the bottom opened and he was immediately greeted by Commander Barnol, the military
head of the underworld. Barnol looked grim.

``What's wrong?'' Paraz said.

``There's been a threshold accident.''

The breath caught in Paraz's throat. ``What?'' he gasped softly. ``What happened?''

``Harteria was making measurements with the bricks and one of the supports collapsed. The bricks
touched.''

``Good Jurita,'' Paraz said. He looked around as if to verify that the underworld was really
still there.

``Harteria pushed the top brick away,'' Barnol said, ``with his trunk.''

Paraz leaned against the wall. Harteria was a good friend of his. A fellow physicist. ``Is he
alive?'' Paraz asked finally.

``For now,'' said Barnol, ``but probably not for long. He's in the infirmary.''

``I must speak with him at once,'' Paraz said.

The two went directly to the infirmary. It was a small space more suited to patching cuts than
tending to a dying man. Yet there was little else they could do. Bringing him to a regular
hospital was out of the question as it would break security. They found Harteria on a bed, eyes
closed, wheezing slightly. He had a terrible burn on his trunk but it was really the yellow
orange color of his scales that alarmed Paraz.

Harteria opened his eyes as they approached and weakly clicked his teeth. Paraz kneeled down
beside the bed and put his trunk around his friend's shoulders. ``You aren't supposed to do
that,'' Paraz said.

Harteria tried to click his teeth again. Then with a voice that could hardly be heard said,
``Now you tell me.''

``You are a crazy fool to be grabbing at the bricks with your trunk,'' Paraz said.

Harteria turned slightly toward him and tried to lift his head. ``Tell\ldots\ tell my
wife\ldots,'' he said softly, laboring to get out the words. ``\ldots\ that I love her.''

Paraz nodded. ``I'll be sure she knows.''

Harteria lay back onto the bed. He tried to move his trunk but the pain kept him from doing so.
He closed his eyes.

``You saved everyone here, you know,'' Paraz said, ``with what you did.'' Harteria didn't
answer. His breath came in short gasps. Paraz looked at him sadly. ``I will let you rest for a
while.'' Harteria swayed his head slightly. ``I will come back to check on you after I get
settled.''

Paraz stood up and joined Barnol who had been waiting by the door. The two went out into the
hallway.

``The doctor doesn't think he'll last another hour,'' Barnol said softly.

Paraz didn't answer right away. Instead he turned and looked at the closed door that separated
him from his friend. ``He's a hero,'' Paraz said, ``but nobody will know.''
