
Colty heard a knock. ``Dr. Colty Harnick?'' Colty turned from his desk and was surprised to see
a large man with Forbinite markings standing at the door to his office.

``Ah\ldots yes?'' he said tentatively.

``I'm Tusk Mulnermin,'' the man said as he swayed his head slightly. He stepped into Colty's
office and extended his trunk.

Colty stood up awkwardly. ``Oh. I wasn't\ldots expecting you so soon.'' He shook Tusk's trunk.
Tusk was right on time but Colty wasn't sure what else to say. He hadn't realized Tusk was a
Forbinite and it caught him off guard.

Tusk clicked his teeth slightly. ``I don't mean to disturb you,'' he said.

``No, no,'' Colty replied. ``Please come in.''

Colty's office was little more than a closet off a side hall in Fargon. As a junior faculty
without full status he did not get one of the spacious offices with a window, a blackboard, and
a conference table. Instead his office was just barely large enough for a small bookshelf, his
desk, and a seat beside his desk for the occasional student. Tusk was a large man and seemed to
fill the small space to bursting.

``Maybe we should find a conference room,'' Colty said.

``This is fine with me,'' said Tusk as he settled down on the seat. ``It's cozy,'' he added with
a shake of his ears.

Colty laid back down and Tusk took a folder of papers from his pack. He put it on the desk.
``Take a look,'' he said. ``It's an orientation package I put together for you.''

Colty flipped open the folder and there, stapled to the inside cover was a picture of
\emph{her}. Colty stopped and stared. He was struck by how like a chortak she looked. She had
forward facing eyes, no trunk, and a prominent chin. He had seen pictures of humans before on
television or in magazines but this was different. This was a human he was going to meet.

The human's mouth was partly open and her teeth were just visible. Tusk noticed what had drawn
Colty's attention. ``It looks fearsome doesn't it?'' Tusk said. ``It's an expression they make
when they are happy or pleased with something.''

``So she is happy here?'' Colty asked pointing at the picture.

``That's right,'' Tusk replied. He shook his ears, ``or maybe just posing for the picture.''

``What is the black stuff?'' Colty asked pointing at the top of her head.

``That's fur,'' Tusk explained. ``They have fur very much like chortak except it only covers a
part of their body. It's mostly on top of their head or here\ldots'' Tusk pointed at the
picture. ``\ldots above their eyes.''

Colty studied the picture some more. ``What a strange looking beast,'' he said at last.

Tusk shrugged. ``She is an animal of Earth just as we are animals of Rujaria,'' he said. ``Once
you get used to them they are not really all that strange looking.''

Colty paused and then asked, ``So what about these\ldots hands.''

``Ah yes,'' said Tusk. ``The picture doesn't show it but as you know her front legs serve the
same purpose as your trunk. They call them \alien{arms}. Her front feet, the \alien{hands}, are
bony but amazingly dexterous. I bet she'll be able to pick up anything with them that you can.''

``You've worked with these humans closely, I understand,'' Colty said.

Tusk swayed his head. ``Not with Dr. \alien{Spencer}, but with other humans, yes.''

``Everyone calls her `doctor' but that can't really be right,'' Colty said.

``Their academic titles are different than ours, yes, but that's the best translation we can
come up with. She has the highest title their culture can confer,'' Tusk explained. ``She is
apparently exceptionally smart, even by human standards. She has all kinds of awards and has
published ground breaking results. This is what we've been told.''

Colty sat back and looked around his cramped little office. Once again he felt annoyed about
being pressed into this job. ``She's not going to be impressed by me, I'm afraid.''

``You have a nice space here,'' Tusk said looking around. He shook his ears.

Tusk lifted his trunk briefly in an attention getting gesture. ``Oh, speaking of offices. The
plan is to renovate Fargon 215 to make it into an office for her.'' Colty looked surprised. Tusk
clicked his teeth slightly. ``She'll live at the compound, but if she's going to be part of the
university community she should have her own office, don't you think?''

``I suppose so,'' Colty said, ``but we use 215 for our department meetings. I guess we'll have
to find someplace else.''

Tusk shrugged. ``That's the room your department chair came up with when we asked him.''

Tusk flipped over a couple of sheets of paper in the folder. ``There are some things you need to
know,'' he said. ``First, she won't like the heat. The humans prefer cooler temperatures. Their
bodies generate their own warmth so they can live comfortably in temperatures that would quickly
make us lethargic. I'm sure it's great on Earth, but here it works against them.

``We've been told that the temperatures at Parnon are tolerable for them. You should know,
though, that when she gets hot her skin will seep water. Their metabolism requires a closely
regulated body temperature, and as the water evaporates it will cool her.''

Colty looked disgusted. Tusk clicked his teeth. ``She is from another planet, Dr. Harnick. Don't
forget that. It's a natural reflex for them. She won't be able to control it.''

``What about all the water lose?'' Colty asked.

``She'll drink a lot. It's what they do.''

Tusk flipped to another page. ``She'll sleep at the compound\ldots\ yes, humans do sleep. In
fact, she'll need to sleep every two hours, but for only an hour or so at a time. Humans have a
very short sleep cycle, but its the same for them year round\ldots\ even in winter.''

Colty read over the notes while Tusk talked. He never really considered all of these details
before. Like everyone else he took the annual rhythms of life on Rujaria for granted. Yet these
humans were not following the same rhythms.

``Socially eye contact is important for them,'' Tusk was saying. ``When she talks to you she'll
want to look in your eyes. You should try to do the same when you talk to her. However, don't
stare. That comes across as threatening.''

Colty took a breath. ``So if I'm not careful I could start our first interstellar war?''

Tusk shook his ears. ``Don't worry about it too much. You won't be dealing with a caged animal.
She will be a highly sophisticated individual. She understands that we are getting to know each
other. I'm sure my counterpart, Mike, is going to be talking to her about all this as well. It
wouldn't surprise me if she is just as worried about offending you as you are of offending her.
They are a polite people. I wouldn't expect Dr. \alien{Spencer} to be any different.''

Colty had turned back to her picture again and was looking at it. He thought he could perhaps
get used to her appearance. The way Tusk described the humans was both shocking and intriguing.
It might be a distraction having her around but it would be a unique distraction at least.

``What about language?'' Colty said. ``Will she speak Argenian?''

Tusk clicked his teeth. ``Well,'' he began, ``she will most likely speak through a translation
device. It's a small box that she'll wear on her clothing. The box translates her thoughts into
Argenian and speaks using a synthesized voice. When you speak the box translates your words into
their language, I guess, and injects them directly into her mind.''

``What?'' Colty was shocked.

Tusk swayed his head. ``Yes, they have machines that can read their minds and interface with
their thoughts,'' he said. ``It's likely that you will never hear her speak directly. She'll
talk to you through the translation device without actually uttering a sound.''

Colty scratched his forehead slowly with his trunk. He looked at Tusk, unsure what to say.

``It works well,'' Tusk said. ``Anyway their natural language is very strange. It's organized in
an unusual way, and they use sounds we can't easily make with our mouths. It's hard for us to
understand and extremely hard for us to speak.''

Tusk leaned back a little in his seat and stretched out his front legs a bit. ``It goes both
ways, though,'' he said. ``They have a lot of trouble with our language too. Although it seems
to be somewhat easier for them to speak Argenian than for us to speak \alien{English}.''

``Can I learn to speak her name?'' Colty asked. It seemed to him that he should at least show
enough respect to call her by name when appropriate.

``Ah yes,'' Tusk said leaning forward again. ``I have a page on that.'' He flipped to another
sheet of paper in the folder. ``Her name is \alien{Rachel Spencer}.'' As a trained linguist who
had been working with humans for some time, Tusk was able to pronounce the hard sounds reasonably
well.

``Jurita,'' Colty said. ``I could never say that.''

``I know it's difficult,'' Tusk said. ``I'm still not very good. Try dropping a few sounds:
Ra'hel 'pencer.''

Colty struggled with it as Tusk gave him advice on how to hold his tongue and how to force air
through his teeth. He had reasonable success saying `Ra'hel' but much more trouble with her last
name. ``How important is it that I pronounce her name?'' Colty asked.

``I'm not entirely sure,'' Tusk said honestly.

Colty tried a few more times.

``Oh and one other thing,'' Tusk said. ``She will have some sort of communication device with
her. I gather it's like a portable telephone that she can use literally anywhere on Rujaria. If
she starts having trouble, from the heat maybe, she can call her people for help. You are not
responsible for taking care of her. She can take care of herself.''

``I understand,'' Colty said.

Colty found himself liking Tusk. He had a casual, yet confident approach that made the whole
idea of hosting the human seem like a natural thing. He consistently referred to the humans as
``people.'' Perhaps having the human as a visiting professor wouldn't be any different than
having, say, a Forbinite visiting professor.

Finally, Tusk got up to return to Varnok. ``My phone number is in the folder. Don't hesitate to
call me if you think of more questions. I'm sure we will meet again before Dr. \alien{Spencer}
arrives.''

``Thank you,'' Colty said. He shook trunks with Tusk. After his visitor left, Colty lay back
down at his desk. He picked up the folder again and organized the pages a little. He looked at
Rachel's picture. ``Ra'hel 'pencer,'' he said softly.

