
The door opened into a tall, narrow room. It wasn't as dark as Tusk had feared. There were
strange yellow-green lights, as he had been told to expect, but also more natural lighting the
humans had installed for his benefit. A human was there dressed in a suit that covered its
entire body. One wall was covered with displays, glowing with human symbols Tusk did not have
time to read.

``This way please,'' the suited human said in synthesized Argenian. He or she---Tusk had no idea
which---directed them toward one end of the room and to a small, very narrow open door. With
some awkwardness and trepidation Tusk made his way into a smaller room. Rachel followed.

The room had some furniture that could only be comfortable to humans and at the far end, near a
second door similar to the one they had come in, was a metal seat of the sort that was common on
military bases. It wasn't very comfortable, but that was the least of his concerns.

Rachel touched some small panels on the walls and, to Tusk's surprise, several larger panels
came to life. They must have been display panels but they showed views as if they were windows.
On one side they could look out at the hanger, a familiar and reassuring sight. On the other
side they appeared to look into another room, larger and with many consoles and display panels.
It looked to Tusk like a control room. Another human, not wearing protective clothing, was there
studying a tray filled with vials. Tusk judged it to be a female, although it was hard for him
to be sure.

She looked up from her work and smiled at the display. \alien{``Make yourself at home,''} she
said, \alien{``we'll get started with phase one in a moment.''}

Rachel didn't say much but simply waited patiently. Tusk fretted about not having anything to
write on. The lawyer had told him to take careful notes and briefed him at length about what to
record. One was a timeline of events. Unfortunately the humans didn't allow him to bring
anything into the purging process except for the disposable clothes on his back; even they would
be destroyed in phase two. Or was it in phase three, Tusk wondered. He couldn't remember. His
notepads and other personal items were being handled separately. He wasn't sure when they would
be reunited.

After a few arnets the door by Tusk's seat clicked open and a fully suited human, perhaps the
one who had been in the control room, stepped in. It, or she, had a transparent box filled with
various fearful items.

First the human went to Rachel and gave her something to swallow with some water. She then
picked up a small electronic device and held near Rachel's torso. She looked at the display on
the device for some time before putting it back with satisfaction. Rachel then held out her arm
and the human strapped a small white band around her wrist. She picked up the electronic device
again, touched it a few times, and then waved it over Rachel's wristband.

\alien{``Welcome home,''} the human said. Tusk could tell from the voice that it was the person
who had been in the control room before.

The human next went to Tusk and gingerly held out a large tablet with a cup of water using both
her hands at once. Tusk took the tablet from her and swallowed it. He didn't need any of the
water. The human then held the electronic device near Tusk's side.

\alien{``We can monitor what's happening in your gut with that instrument,''} she explained,
\alien{``I'm just making sure it's working.''}

After putting the device away she pulled one of the human seats over and sat on it, keeping her
body upright as the humans generally did.

\alien{``We don't have an infuser for you.''}

\alien{``Infuser?''}

\alien{``Yes, like Rachel has.''} The human nodded toward Rachel and she held up her wristband.
\alien{``It injects through the skin but there is no way our infusers could get through your
  skin. I'll have to do the injections the old fashion way with a needle.''}

% There should be a dunari doctor in house. We should meet him before Tusk starts this process
% and he should be guiding the human here.

Tusk swayed his head. The human took a long needle from the box, already filled with a green
tinted liquid.

\alien{``Where do they usually do this?''} she asked, \alien{``I was told just behind your right
  shoulder, yes?''}

Tusk used his trunk to point to the area.

\alien{``Between scales, yes?''} the human asked.

\alien{``That's right.''}

\alien{``And really this whole thing? Almost all the length?''}

\alien{``Yes, it has to go through a fat layer.''}

She seemed nervous and tentative. Tusk understand she had probably never done this before, at
least not on a real person. After a moment he felt the sting of the needle as it burrowed into
his flesh. He winced slightly. She was going too slowly, but he didn't say anything. After a
moment he felt a cool tingle as the strange concoction poured directly into his veins.

\alien{``You did it just right,''} Tusk told her afterward.

% NEXT: The human leaves. Tusk and Rachel have a conversation of interst.
%       Jurnumia comes and is upset about something (can't be that Tusk is a Forbinite since the
%       text above makes it clear they met before).
%       Tusk has a bad reaction to the treatment and there is talk about him backing out.
