
Colty waited just outside the lecture hall with his pack slung over his shoulder. Rachel was
finishing a session on basic thermodynamics. The door was open and Colty listened as she
explained Grogan's Laws. Her voice was animated and filled with excitment. It was obvious that
she loved her work. The students seemed engaged as well. They asked her questions and she spoke
with them, talking back and forth the way colleagues might in a bresken shop. Colty clicked his
teeth. She was a talented instructor. It was a privilege to have her at Parnon. It was a
privilege to be her friend.

When the class ended Colty waited as the students filed out of the hall. Finally he went inside
to find Rachel speaking with a student. She leaned back against the table comfortably as she
talked. She seemed completely at ease and the student seemed at ease with her.

``I'm just having trouble with the algebra,'' the student was saying. ``I can set up the
problem, but I can't get the answer.''

``Come to my office,'' Rachel said. ``It will be easier to look at your work there.''

``I have another class right now.''

``Well, I'll be there until 32:100. Maybe after\ldots'' Rachel noticed Colty standing to the
side. ``Oh, hello.''

``Don't mind me,'' Colty said. ``It can wait.''

``I have to go anyway,'' the student said. ``I'll try to stop by your office when I'm done with
my next class.''

The student left and Rachel started to gather up her papers. ``What brings you to the Physical
Sciences building?'' she asked. ``Did you want a break from the abstract world of pure
mathematics?'' Rachel chuckled slightly.

``Actually I have something I want to show you,'' Colty said. ``Can we go to your office?''

Rachel put her spin dial into its pouch and put everything into her pack. The two of them walked
down the hall toward Rachel's office at the end. They passed several students who greeted them
warmly. They also came across Dr. Narse coming out of his office.

``Oh Ra'hel,'' he said. ``We should talk later. My students are making progress on those low
temperature resistivity measurements you suggested.''

``That's great!'' she said. ``Let's definitely talk.''

Dr. Narse rushed off to a meeting and Rachel smiled to herself. She turned to Colty. ``They
were recently successful in liquifying helium,'' she explained. ``I suggested they measure the
resistivity of lead at that temperature.'' She laughed.

``You sound very impish,'' Colty said clicking his teeth.

``I think they will find the results\ldots\ interesting.''

When they got to Rachel's office she unloaded her pack while Colty looked around. Next to her
picture of \textit{Rainier} she had put up a picture of Lake Erbum with Mount Nizala towering
above it. Nizala's rocky slopes were afire with the deep rubinum light of the sun, reflected in
the dark, cool waters of the lake.

``Isn't it nice?'' Rachel said. She was suddenly standing next to Colty. She reached up and
touched the picture with her fingers. ``I'd love to visit.''

``It's ironic,'' Colty said, ``that you can travel over 5000 light years to get to Rujaria and
yet can't go 300 karnons up the road to Lake Erbum.''

Rachel shrugged. ``Come,'' she said, ``you wanted to talk.'' She sat down on a small pad beside
her desk, folding her legs under her. She gestured toward a seat next to her and Colty lay down
there.

